### 2010-10-07 cvprkit2009/egpaper_for_review.tex
Initial import.

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for CVPR Proceedings}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\small\url{http://www.author.org/~second}}
}

\maketitle
% \thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, at the top
   of the left-hand column, below the author and affiliation
   information. Use the word ``Abstract'' as the title, in 12-point
   Times, boldface type, centered relative to the column, initially
   capitalized. The abstract is to be in 10-point, single-spaced type.
   Leave two blank lines after the Abstract, then begin the main text.
   Look at previous CVPR abstracts to get a feel for style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Please follow the steps outlined below when submitting your manuscript to
the IEEE Computer Society Press.  This style guide now has several
important modifications (for example, you are no longer warned against the
use of sticky tape to attach your artwork to the paper), so all authors
should read this new version.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

By submitting a manuscript to CVPR, the authors assert that it has not been
previously published in substantially similar form. Furthermore, no paper which
contains significant overlap with the contributions of this paper either has
 been or will be submitted during the CVPR 2009 review period to {\bf either
a journal} or a conference (including CVPR 2009).
{\bf Note that this is a strengthening of previous CVPR policy,}
and papers violating this condition will be rejected.

If there are papers that may appear to the reviewers
to violate this condition, then it is your responsibility to: (1)~cite
these papers (preserving anonymity as described in Section 1.6 below),
(2)~argue in the body of your paper why your CVPR paper is non-trivially
different from these concurrent submissions, and (3)~include anonymized
versions of those papers in the supplemental material.

%By submitting this manuscript to CVPR, the authors assert that it has not
%been previously published in substantially similar form, and no paper
%currently under submission to a conference contains significant overlap
%with this one.  If you are in doubt about the amount of overlap, cite the
%dual submission (as described below), and argue in the body of your paper
%why this is a nontrivial advance.  A simultaneous journal submission would
%be expected to have significant additional material not in the conference
%paper, and should not be previously published, nor in the final acceptance
%stages. The conference chairs reserve the right to cancel
%submission of any paper which is found to violate these conditions.  In
%particular, {\em uncited} dual submissions will be summarily dealt with.

\subsection{Paper length}
Consult the call for papers for page-length limits.  Overlength papers will
simply not be reviewed.  This includes papers where the margins and
formatting are deemed to have been significantly altered from those laid
down by this style guide.  Note that this \LaTeX\ guide already sets figure
captions and references in a smaller font.  The reason such papers will not
be reviewed is that there is no provision for supervised revisions of
manuscripts.  The reviewing process cannot determine the suitability of the
paper for presentation in eight pages if it is reviewed in eleven.

%-------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
note that the ruler measurements do not align well with lines in the paper
--- this turns out to be very difficult to do well when the paper contains
many figures and equations, and, when done, looks ugly.  Just use fractional
references (e.g.\ this line is $095.5$), although in most cases one would
expect that the approximate location will be adequate.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  It is
important for readers to be able to refer to any particular equation.  Just
because you didn't refer to it in the text doesn't mean some future reader
might not need to refer to it.  It is cumbersome to have to use
circumlocutions like ``the equation second from the top of page 3 column
1''.  (Note that the ruler will not be present in the final copy, so is not
an alternative to equation numbers).  All authors will benefit from reading
Mermin's description of how to write mathematics: \url{http://www.cvpr.org/doc/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove
citations to one's own work---in fact it is often impossible to
review a paper unless the previous citations are known and
available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for
techreports)

Saying ``this builds on the work of Lucy Smith [1]'' does not say
that you are Lucy Smith, it says that you are building on her
work.  If you are Smith and Jones, do not say ``as we show in
[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
paper, include reference 7 as you would any other cited work.

An example of a bad paper:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of our
   previous paper [1], and show it to be inferior to all
   previously known methods.  Why the previous paper was
   accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an excellent paper:

\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of the
   paper of Smith \etal [1], and show it to be inferior to
   all previously known methods.  Why the previous paper
   was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo
   filter, a fundamental contribution to human knowledge''.
   Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to that
submission in order to explain the differences, just as you would if you
had previously published related work.  In such cases, include the
anonymized parallel submission~\cite{Authors06} as additional material and
cite it as
\begin{quote}
[1] Authors. ``The frobnicatable foo filter'', ECCV 2006 Submission ID 324,
Supplied as additional material {\tt eccv06.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be
found elsewhere, and refer them to a technical report.  For conference
submissions, the paper must stand on its own, and not {\em require} the
reviewer to go to a techreport for further details.  Thus, you may say in
the body of the paper ``further details may be found
in~\cite{Authors06b}''.  Then submit the techreport as additional material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool which
is widely known to be restricted to a single institution.  For example,
let's say it's 1969, you have solved a key problem on the Apollo lander,
and you believe that the CVPR70 audience would like to hear about your
solution.  The work is a development of your celebrated 1968 paper entitled
``Zero-g frobnication: How being the only people in the world with access to
the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.  Don't write ``We show how to
improve our previous work [Anonymous, 1968].  This time we tested the
algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors. Instead
write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.  This
   system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] didn't
   handle case B properly.  Ours handles it by including
   a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo
   lunar lander, and went all the way to the moon, don't
   you know.  It displayed the following behaviours
   which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you as
the authors.  A reviewer might think it likely that the new paper was
written by Zeus \etal, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been
contracted to solve problem B.

FAQ: Are acknowledgements OK?  No.  Leave them for the final copy.

\begin{figure}[t]
\begin{center}
\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
\end{center}
   \caption{Example of caption.  It is set in Roman so that mathematics
   (always set in Roman: $B \sin A = A \sin B$) may be included without an
   ugly clash.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
 \verb'$conf_a$' &  $conf_a$ \\
 \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a
sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
\verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et alia'',
shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
However, use it only when there are three or more authors.  Thus, the
following is correct: ``
   Frobnication has been trendy lately.
   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
because reference~\cite{Alpher03} has just two authors.  If you use the
\verb'\etal' macro provided, then you need not worry about double periods
when used at the end of a sentence as in Alpher \etal.

For this citation style, keep multiple citations in numerical (not
chronological) order, so prefer \cite{Alpher03,Alpher02,Authors06} to
\cite{Alpher02,Alpher03,Authors06}.

\begin{figure*}
\begin{center}
\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
\end{center}
   \caption{Example of a short caption, which should be centered.}
\label{fig:short}
\end{figure*}

%------------------------------------------------------------------------
\section{Formatting your paper}

All text must be in a two-column format. The total allowable width of the
text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
first page) should begin 1.0 inch (2.54 cm) from the top edge of the
page. The second and following pages should begin 1.0 inch (2.54 cm) from
the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
page.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be
kept within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches
(22.54 cm) high.

%-------------------------------------------------------------------------
\subsection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If neither is
available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
the first page. The title should be in Times 14-point, boldface type.
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
adverbs; do not capitalize articles, coordinate conjunctions, or
prepositions (unless the title begins with such a word). Leave two blank
lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type. This information is to
be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
inch or 0.422 cm). Make sure your text is fully justified---that is,
flush left and flush right. Please do not place any additional blank
lines between paragraphs.

Figure and table captions should be 9-point Roman type as in
Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-,
second-, and third-order headings.

FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
should be Times 12-point boldface, initially capitalized, flush left,
with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
should be Times 11-point boldface, initially capitalized, flush left,
with one blank line before, and one after. If you require a third-order
heading (we discourage it), use 10-point Times, boldface, initially
capitalized, flush left, preceded by one blank line, followed by a period
and your text on the same line.

%-------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote {This is what a footnote looks like.  It
often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral
observations in
the text (within parentheses, if you prefer, as in this sentence).  If you
wish to use a footnote, place it at the bottom of the column on the page on
which it is referenced. Use Times 8-point type, single-spaced.

%-------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times,
single-spaced, at the end of your paper. When referenced in the text,
enclose the citation number in square brackets, for
example~\cite{Authors06}.  Where appropriate, include the name(s) of
editors of referenced books.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Method & Frobnability \\
\hline\hline
Theirs & Frumpy \\
Yours & Frobbly \\
Ours & Makes one's heart Frob\\
\hline
\end{tabular}
\end{center}
\caption{Results.   Ours is better.}
\end{table}

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.  Please ensure that any point you wish to
make is resolvable in a printed copy of the paper.  Resize fonts in figures
to match the font in the body text, and choose line widths which render
effectively in print.  Many readers (and reviewers), even of an electronic
copy, will choose to print your paper in order to read it.  You cannot
insist that they do otherwise, and therefore must not assume that they can
zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use
\verb+\includegraphics+, and to specify the  figure width as a multiple of
the line width as in the example below
{\small\begin{verbatim}
   \usepackage[dvips]{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.eps}
\end{verbatim}
}

%-------------------------------------------------------------------------
\subsection{Color}

Color is valuable, and will be visible to readers of the electronic copy.
However ensure that, when printed on a monochrome printer, no important
information is lost by the conversion to grayscale.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

Please direct any questions to the production editor in charge of these
proceedings at the IEEE Computer Society Press: Phone (714) 821-8380, or
Fax (714) 761-1784.

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage
\rule{0pt}{1pt}\newpage

\end{document}

### 2010-10-20 authorkit/egpaper_final.tex
Stripped out template stuff, added face recognition bibliography

--- 2010-10-07 cvprkit2009/egpaper_for_review.tex
@@
 % run, let it finish, and you should be clear).
-\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+%\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
 
-% \cvprfinalcopy % *** Uncomment this line for the final submission
+\cvprfinalcopy % *** Uncomment this line for the final submission
 
@@
 First line of institution2 address\\
-{\small\url{http://www.author.org/~second}}
+{\tt\small secondauthor@i2.org}
 }
@@
 \maketitle
-% \thispagestyle{empty}
+\thispagestyle{empty}
 
@@
 By submitting a manuscript to CVPR, the authors assert that it has not been
-previously published in substantially similar form. Furthermore, no paper which
-contains significant overlap with the contributions of this paper either has
- been or will be submitted during the CVPR 2009 review period to {\bf either
-a journal} or a conference (including CVPR 2009).
-{\bf Note that this is a strengthening of previous CVPR policy,}
-and papers violating this condition will be rejected.
+previously published in substantially similar form. Furthermore, no paper
+which contains significant overlap with the contributions of this paper
+either has been or will be submitted during the CVPR 2011 review period to
+{\bf either a journal} or any conference (including CVPR 2011) or any
+workshop (including CVPR2011 workshops)
+ {\bf Note that
+  this is consistent with CVPR2010 but a strengthening from some previous CVPR
+  policy}.  Papers violating this condition will be rejected and a list of violating authors may be included in the proceedings.
 
@@
 
-%By submitting this manuscript to CVPR, the authors assert that it has not
-%been previously published in substantially similar form, and no paper
-%currently under submission to a conference contains significant overlap
-%with this one.  If you are in doubt about the amount of overlap, cite the
-%dual submission (as described below), and argue in the body of your paper
-%why this is a nontrivial advance.  A simultaneous journal submission would
-%be expected to have significant additional material not in the conference
-%paper, and should not be previously published, nor in the final acceptance
-%stages. The conference chairs reserve the right to cancel
-%submission of any paper which is found to violate these conditions.  In
-%particular, {\em uncited} dual submissions will be summarily dealt with.
-
 \subsection{Paper length}
-Consult the call for papers for page-length limits.  Overlength papers will
-simply not be reviewed.  This includes papers where the margins and
-formatting are deemed to have been significantly altered from those laid
-down by this style guide.  Note that this \LaTeX\ guide already sets figure
-captions and references in a smaller font.  The reason such papers will not
-be reviewed is that there is no provision for supervised revisions of
-manuscripts.  The reviewing process cannot determine the suitability of the
-paper for presentation in eight pages if it is reviewed in eleven.
+CVPR papers may be between 6 pages and 8 pages, with a \$100 per page added
+fee.  Overlength papers will simply not be reviewed.  This includes papers
+where the margins and formatting are deemed to have been significantly
+altered from those laid down by this style guide.  Note that this
+\LaTeX\ guide already sets figure captions and references in a smaller font.
+The reason such papers will not be reviewed is that there is no provision for
+supervised revisions of manuscripts.  The reviewing process cannot determine
+the suitability of the paper for presentation in eight pages if it is
+reviewed in eleven.  If you submit 8 for review expect to pay the added page
+charges for them.
 
@@
 an alternative to equation numbers).  All authors will benefit from reading
-Mermin's description of how to write mathematics: \url{http://www.cvpr.org/doc/mermin.pdf}.
+Mermin's description of how to write mathematics.%: \url{http://www.cvpr.org/doc/mermin.pdf}.
 
@@
 
-An example of a bad paper:
+An example of a bad paper just asking to be rejected:
 \begin{quote}
@@
 
-An example of an excellent paper:
+An example of an acceptable paper:
 
@@
 had previously published related work.  In such cases, include the
-anonymized parallel submission~\cite{Authors06} as additional material and
+anonymized parallel submission~\cite{Authors11} as additional material and
 cite it as
 \begin{quote}
-[1] Authors. ``The frobnicatable foo filter'', ECCV 2006 Submission ID 324,
-Supplied as additional material {\tt eccv06.pdf}.
+[1] Authors. ``The frobnicatable foo filter'', F\&G 2011 Submission ID 324,
+Supplied as additional material {\tt fg324.pdf}.
 \end{quote}
@@
 the body of the paper ``further details may be found
-in~\cite{Authors06b}''.  Then submit the techreport as additional material.
+in~\cite{Authors11b}''.  Then submit the techreport as additional material.
 Again, you may not assume the reviewers will read this material.
@@
 let's say it's 1969, you have solved a key problem on the Apollo lander,
-and you believe that the CVPR70 audience would like to hear about your
+and you believe that the CVPR11 audience would like to hear about your
 solution.  The work is a development of your celebrated 1968 paper entitled
@@
 should be Times 12-point boldface, initially capitalized, flush left,
-with one blank line before, and one blank line after.
+with one blank line before.
 
@@
 
-Please direct any questions to the production editor in charge of these
-proceedings at the IEEE Computer Society Press: Phone (714) 821-8380, or
-Fax (714) 761-1784.
-
 {\small
@@
 
-\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-\rule{0pt}{1pt}\newpage
-
 \end{document}

### 2010-10-20 authorkit/egpaper_for_review.tex
Stripped out template stuff, added face recognition bibliography

--- 2010-10-07 cvprkit2009/egpaper_for_review.tex
@@
 By submitting a manuscript to CVPR, the authors assert that it has not been
-previously published in substantially similar form. Furthermore, no paper which
-contains significant overlap with the contributions of this paper either has
- been or will be submitted during the CVPR 2009 review period to {\bf either
-a journal} or a conference (including CVPR 2009).
-{\bf Note that this is a strengthening of previous CVPR policy,}
-and papers violating this condition will be rejected.
+previously published in substantially similar form. Furthermore, no paper
+which contains significant overlap with the contributions of this paper
+either has been or will be submitted during the CVPR 2011 review period to
+{\bf either a journal} or any conference (including CVPR 2011) or any
+workshop (including CVPR2011 workshops).  {\bf Note that
+  this is consistent with CVPR2010 but a strengthening from some previous CVPR
+  policy,} and papers violating this condition will be rejected.
 
@@
 many figures and equations, and, when done, looks ugly.  Just use fractional
-references (e.g.\ this line is $095.5$), although in most cases one would
+references (e.g.\ this line is $097.15$), although in most cases one would
 expect that the approximate location will be adequate.
@@
 an alternative to equation numbers).  All authors will benefit from reading
-Mermin's description of how to write mathematics: \url{http://www.cvpr.org/doc/mermin.pdf}.
+Mermin's description of how to write mathematics:
+\url{http://www.cvpr.org/doc/mermin.pdf}.
 
@@
 \begin{quote}
-[1] Authors. ``The frobnicatable foo filter'', ECCV 2006 Submission ID 324,
-Supplied as additional material {\tt eccv06.pdf}.
+[1] Authors. ``The frobnicatable foo filter'', F\&G 2011 Submission ID 324,
+Supplied as additional material {\tt fg324.pdf}.
 \end{quote}

### 2010-10-20 paper/main.tex
Stripped out template stuff, added face recognition bibliography

--- 2010-10-20 authorkit/egpaper_for_review.tex
@@
 
-\usepackage{cvpr}
+\usepackage{../authorkit/cvpr}
 \usepackage{times}
@@
 
-% Include other packages here, before hyperref.
+% Packages added by us
+\usepackage{mydefs}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
-% If you comment hyperref and then uncomment it, you should delete
-% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
-% run, let it finish, and you should be clear).
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
@@
 
-%%%%%%%%% TITLE
-\title{\LaTeX\ Author Guidelines for CVPR Proceedings}
+\title{Efficient Parallelization of Sparse Representation for Face Recognition}
 
@@
 \maketitle
-% \thispagestyle{empty}
 
-%%%%%%%%% ABSTRACT
 \begin{abstract}
@@
 
-%%%%%%%%% BODY TEXT
 \section{Introduction}
 
-Please follow the steps outlined below when submitting your manuscript to
-the IEEE Computer Society Press.  This style guide now has several
-important modifications (for example, you are no longer warned against the
-use of sticky tape to attach your artwork to the paper), so all authors
-should read this new version.
-
-%-------------------------------------------------------------------------
-\subsection{Language}
-
-All manuscripts must be in English.
-
-\subsection{Dual submission}
-
-By submitting a manuscript to CVPR, the authors assert that it has not been
-previously published in substantially similar form. Furthermore, no paper
-which contains significant overlap with the contributions of this paper
-either has been or will be submitted during the CVPR 2011 review period to
-{\bf either a journal} or any conference (including CVPR 2011) or any
-workshop (including CVPR2011 workshops).  {\bf Note that
-  this is consistent with CVPR2010 but a strengthening from some previous CVPR
-  policy,} and papers violating this condition will be rejected.
-
-If there are papers that may appear to the reviewers
-to violate this condition, then it is your responsibility to: (1)~cite
-these papers (preserving anonymity as described in Section 1.6 below),
-(2)~argue in the body of your paper why your CVPR paper is non-trivially
-different from these concurrent submissions, and (3)~include anonymized
-versions of those papers in the supplemental material.
-
-\subsection{Paper length}
-CVPR papers may be between 6 pages and 8 pages, with a \$100 per page added
-fee.  Overlength papers will simply not be reviewed.  This includes papers
-where the margins and formatting are deemed to have been significantly
-altered from those laid down by this style guide.  Note that this
-\LaTeX\ guide already sets figure captions and references in a smaller font.
-The reason such papers will not be reviewed is that there is no provision for
-supervised revisions of manuscripts.  The reviewing process cannot determine
-the suitability of the paper for presentation in eight pages if it is
-reviewed in eleven.  If you submit 8 for review expect to pay the added page
-charges for them.
-
-%-------------------------------------------------------------------------
-\subsection{The ruler}
-The \LaTeX\ style defines a printed ruler which should be present in the
-version submitted for review.  The ruler is provided in order that
-reviewers may comment on particular lines in the paper without
-circumlocution.  If you are preparing a document using a non-\LaTeX\
-document preparation system, please arrange for an equivalent ruler to
-appear on the final output pages.  The presence or absence of the ruler
-should not change the appearance of any other content on the page.  The
-camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
-the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
-note that the ruler measurements do not align well with lines in the paper
---- this turns out to be very difficult to do well when the paper contains
-many figures and equations, and, when done, looks ugly.  Just use fractional
-references (e.g.\ this line is $097.15$), although in most cases one would
-expect that the approximate location will be adequate.
-
-\subsection{Mathematics}
-
-Please number all of your sections and displayed equations.  It is
-important for readers to be able to refer to any particular equation.  Just
-because you didn't refer to it in the text doesn't mean some future reader
-might not need to refer to it.  It is cumbersome to have to use
-circumlocutions like ``the equation second from the top of page 3 column
-1''.  (Note that the ruler will not be present in the final copy, so is not
-an alternative to equation numbers).  All authors will benefit from reading
-Mermin's description of how to write mathematics:
-\url{http://www.cvpr.org/doc/mermin.pdf}.
-
-\subsection{Blind review}
-
-Many authors misunderstand the concept of anonymizing for blind
-review.  Blind review does not mean that one must remove
-citations to one's own work---in fact it is often impossible to
-review a paper unless the previous citations are known and
-available.
-
-Blind review means that you do not use the words ``my'' or ``our''
-when citing previous work.  That is all.  (But see below for
-techreports)
-
-Saying ``this builds on the work of Lucy Smith [1]'' does not say
-that you are Lucy Smith, it says that you are building on her
-work.  If you are Smith and Jones, do not say ``as we show in
-[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
-paper, include reference 7 as you would any other cited work.
-
-An example of a bad paper just asking to be rejected:
-\begin{quote}
-\begin{center}
-    An analysis of the frobnicatable foo filter.
-\end{center}
-
-   In this paper we present a performance analysis of our
-   previous paper [1], and show it to be inferior to all
-   previously known methods.  Why the previous paper was
-   accepted without this analysis is beyond me.
-
-   [1] Removed for blind review
-\end{quote}
-
-An example of an acceptable paper:
-
-\begin{quote}
-\begin{center}
-     An analysis of the frobnicatable foo filter.
-\end{center}
-
-   In this paper we present a performance analysis of the
-   paper of Smith \etal [1], and show it to be inferior to
-   all previously known methods.  Why the previous paper
-   was accepted without this analysis is beyond me.
-
-   [1] Smith, L and Jones, C. ``The frobnicatable foo
-   filter, a fundamental contribution to human knowledge''.
-   Nature 381(12), 1-213.
-\end{quote}
-
-If you are making a submission to another conference at the same time,
-which covers similar or overlapping material, you may need to refer to that
-submission in order to explain the differences, just as you would if you
-had previously published related work.  In such cases, include the
-anonymized parallel submission~\cite{Authors06} as additional material and
-cite it as
-\begin{quote}
-[1] Authors. ``The frobnicatable foo filter'', F\&G 2011 Submission ID 324,
-Supplied as additional material {\tt fg324.pdf}.
-\end{quote}
-
-Finally, you may feel you need to tell the reader that more details can be
-found elsewhere, and refer them to a technical report.  For conference
-submissions, the paper must stand on its own, and not {\em require} the
-reviewer to go to a techreport for further details.  Thus, you may say in
-the body of the paper ``further details may be found
-in~\cite{Authors06b}''.  Then submit the techreport as additional material.
-Again, you may not assume the reviewers will read this material.
-
-Sometimes your paper is about a problem which you tested using a tool which
-is widely known to be restricted to a single institution.  For example,
-let's say it's 1969, you have solved a key problem on the Apollo lander,
-and you believe that the CVPR70 audience would like to hear about your
-solution.  The work is a development of your celebrated 1968 paper entitled
-``Zero-g frobnication: How being the only people in the world with access to
-the Apollo lander source code makes us a wow at parties'', by Zeus \etal.
-
-You can handle this paper like any other.  Don't write ``We show how to
-improve our previous work [Anonymous, 1968].  This time we tested the
-algorithm on a lunar lander [name of lander removed for blind review]''.
-That would be silly, and would immediately identify the authors. Instead
-write the following:
-\begin{quotation}
-\noindent
-   We describe a system for zero-g frobnication.  This
-   system is new because it handles the following cases:
-   A, B.  Previous systems [Zeus et al. 1968] didn't
-   handle case B properly.  Ours handles it by including
-   a foo term in the bar integral.
-
-   ...
-
-   The proposed system was integrated with the Apollo
-   lunar lander, and went all the way to the moon, don't
-   you know.  It displayed the following behaviours
-   which show how well we solved cases A and B: ...
-\end{quotation}
-As you can see, the above text follows standard scientific convention,
-reads better than the first version, and does not explicitly name you as
-the authors.  A reviewer might think it likely that the new paper was
-written by Zeus \etal, but cannot make any decision based on that guess.
-He or she would have to be sure that no other authors could have been
-contracted to solve problem B.
-
-FAQ: Are acknowledgements OK?  No.  Leave them for the final copy.
-
-\begin{figure}[t]
-\begin{center}
-\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
-   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
-\end{center}
-   \caption{Example of caption.  It is set in Roman so that mathematics
-   (always set in Roman: $B \sin A = A \sin B$) may be included without an
-   ugly clash.}
-\label{fig:long}
-\label{fig:onecol}
-\end{figure}
-
-\subsection{Miscellaneous}
-
-\noindent
-Compare the following:\\
-\begin{tabular}{ll}
- \verb'$conf_a$' &  $conf_a$ \\
- \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
-\end{tabular}\\
-See The \TeX book, p165.
-
-The space after \eg, meaning ``for example'', should not be a
-sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
-\verb'\eg' macro takes care of this.
-
-When citing a multi-author paper, you may save space by using ``et alia'',
-shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
-However, use it only when there are three or more authors.  Thus, the
-following is correct: ``
-   Frobnication has been trendy lately.
-   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
-   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''
-
-This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
-because reference~\cite{Alpher03} has just two authors.  If you use the
-\verb'\etal' macro provided, then you need not worry about double periods
-when used at the end of a sentence as in Alpher \etal.
-
-For this citation style, keep multiple citations in numerical (not
-chronological) order, so prefer \cite{Alpher03,Alpher02,Authors06} to
-\cite{Alpher02,Alpher03,Authors06}.
-
-\begin{figure*}
-\begin{center}
-\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-\end{center}
-   \caption{Example of a short caption, which should be centered.}
-\label{fig:short}
-\end{figure*}
-
-%------------------------------------------------------------------------
-\section{Formatting your paper}
-
-All text must be in a two-column format. The total allowable width of the
-text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
-cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
-$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
-first page) should begin 1.0 inch (2.54 cm) from the top edge of the
-page. The second and following pages should begin 1.0 inch (2.54 cm) from
-the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
-cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
-paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
-page.
-
-%-------------------------------------------------------------------------
-\subsection{Margins and page numbering}
-
-All printed material, including text, illustrations, and charts, must be
-kept within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches
-(22.54 cm) high.
-
-%-------------------------------------------------------------------------
-\subsection{Type-style and fonts}
-
-Wherever Times is specified, Times Roman may also be used. If neither is
-available on your word processor, please use the font closest in
-appearance to Times to which you have access.
-
-MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
-the first page. The title should be in Times 14-point, boldface type.
-Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
-adverbs; do not capitalize articles, coordinate conjunctions, or
-prepositions (unless the title begins with such a word). Leave two blank
-lines after the title.
-
-AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
-and printed in Times 12-point, non-boldface type. This information is to
-be followed by two blank lines.
-
-The ABSTRACT and MAIN TEXT are to be in a two-column format.
-
-MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
-double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
-inch or 0.422 cm). Make sure your text is fully justified---that is,
-flush left and flush right. Please do not place any additional blank
-lines between paragraphs.
-
-Figure and table captions should be 9-point Roman type as in
-Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.
-
-\noindent Callouts should be 9-point Helvetica, non-boldface type.
-Initially capitalize only the first word of section titles and first-,
-second-, and third-order headings.
-
-FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
-should be Times 12-point boldface, initially capitalized, flush left,
-with one blank line before, and one blank line after.
-
-SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
-should be Times 11-point boldface, initially capitalized, flush left,
-with one blank line before, and one after. If you require a third-order
-heading (we discourage it), use 10-point Times, boldface, initially
-capitalized, flush left, preceded by one blank line, followed by a period
-and your text on the same line.
-
-%-------------------------------------------------------------------------
-\subsection{Footnotes}
-
-Please use footnotes\footnote {This is what a footnote looks like.  It
-often distracts the reader from the main flow of the argument.} sparingly.
-Indeed, try to avoid footnotes altogether and include necessary peripheral
-observations in
-the text (within parentheses, if you prefer, as in this sentence).  If you
-wish to use a footnote, place it at the bottom of the column on the page on
-which it is referenced. Use Times 8-point type, single-spaced.
-
-%-------------------------------------------------------------------------
-\subsection{References}
-
-List and number all bibliographical references in 9-point Times,
-single-spaced, at the end of your paper. When referenced in the text,
-enclose the citation number in square brackets, for
-example~\cite{Authors06}.  Where appropriate, include the name(s) of
-editors of referenced books.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|}
-\hline
-Method & Frobnability \\
-\hline\hline
-Theirs & Frumpy \\
-Yours & Frobbly \\
-Ours & Makes one's heart Frob\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Results.   Ours is better.}
-\end{table}
-
-%-------------------------------------------------------------------------
-\subsection{Illustrations, graphs, and photographs}
-
-All graphics should be centered.  Please ensure that any point you wish to
-make is resolvable in a printed copy of the paper.  Resize fonts in figures
-to match the font in the body text, and choose line widths which render
-effectively in print.  Many readers (and reviewers), even of an electronic
-copy, will choose to print your paper in order to read it.  You cannot
-insist that they do otherwise, and therefore must not assume that they can
-zoom in to see tiny details on a graphic.
-
-When placing figures in \LaTeX, it's almost always best to use
-\verb+\includegraphics+, and to specify the  figure width as a multiple of
-the line width as in the example below
-{\small\begin{verbatim}
-   \usepackage[dvips]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]
-                   {myfile.eps}
-\end{verbatim}
-}
-
-%-------------------------------------------------------------------------
-\subsection{Color}
-
-Color is valuable, and will be visible to readers of the electronic copy.
-However ensure that, when printed on a monochrome printer, no important
-information is lost by the conversion to grayscale.
-
-%------------------------------------------------------------------------
-\section{Final copy}
-
-You must include your signed IEEE copyright release form when you submit
-your finished paper. We MUST have this form before your paper can be
-published in the proceedings.
-
-Please direct any questions to the production editor in charge of these
-proceedings at the IEEE Computer Society Press: Phone (714) 821-8380, or
-Fax (714) 761-1784.
+\section{Application to Face Recognition}
+\begin{algorithm}[thb]
+\caption{\bf (Deformable Sparse Recovery and Classification for
+Face Recognition)} \label{alg:deformable-src}
+\begin{algorithmic}[1]
+\STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
+$\y\in\Re^m$ and a deformation group $T$. 
+\STATE
+{\bf for} each subject $i$, 
+\STATE \hspace{3mm} $\tau^{(0)}
+\leftarrow I$. 
+\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
+\STATE \hspace{6mm}
+$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
+\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
+\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
+%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
+\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
+\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$ 
+\STATE
+\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
+\tau$; 
+\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
+the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
+residuals $\|\e\|_1$. \STATE Compute an average transformation
+$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
+\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
+\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
+k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
+\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
+\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
+$\ell^1$-minimization problem:
+$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$$
+\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
+\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
+\end{algorithmic}
+\end{algorithm}
 
@@
 \bibliographystyle{ieee}
-\bibliography{egbib}
+\bibliography{faces}
 }

### 2010-10-20 paper/main.tex
Added missing style files
Writing the outline...

@@
 
+\section{Characteristics of the ALM Algorithm}
+Describe the two variants of the ALM algorithm
+Show Pseudocode for one variant
+\subsection{Computational Complexity}
+
+\section{Efficient Implementation on the CPU}
+% With a reshuffling of the order of operations and a change of the termination
+% condition, the inner loop takes the form of two matrix-vector multiplications
+% and a sequence of vector-vector operations.
+\subsection{CPU Hardware Parallelism}
+High-level discussion of CPU parallelism, i.e.
+thread-level, vector-level, cache size, cache locality
+\subsection{Tuning the GEMV operation}
+When matrix-vector operations are called sequentially on the same data,
+the operation can be blocked so that each thread operates on the same block of data for each operation.
+
+For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
+
+For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
+
+For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
+
+\subsection{Tuning the Soft-Thresholding Operator}
+...Make sure there are no branches
+
+...Make sure there are no function calls
+
+...Make sure the loop gets automatically vectorized
+
+...Block the computation into work for different cpu cores
+
+\section{Efficient Implementation on the CPU}
+\subsection{GPU Hardware Parallelism}
+High-level description of GPU parallelism, i.e.
+
+cuda cores, SIMD width, scheduling restrictions, bandwidth
+
+\subsection{Solving multiple L1 minimizations in parallel}
+This section is about Drew's kernels for solving multiple L1 minimizations in parllel
+
+How to efficiently perform an L1 minimization on a single core, quote performance on a large batch.
+
+How to solve multiple instances in smaller batches without losing concurrency, using CUDA streams.
+
+...Mention effect of compiler verion on streams implementation
+
 \section{Application to Face Recognition}

### 2010-10-20 paper/main.tex
Fleshing out the outline...

@@
 \begin{abstract}
-   The ABSTRACT is to be in fully-justified italicized text, at the top
-   of the left-hand column, below the author and affiliation
-   information. Use the word ``Abstract'' as the title, in 12-point
-   Times, boldface type, centered relative to the column, initially
-   capitalized. The abstract is to be in 10-point, single-spaced type.
-   Leave two blank lines after the Abstract, then begin the main text.
-   Look at previous CVPR abstracts to get a feel for style and length.
+Abstract
 \end{abstract}
@@
 \section{Introduction}
+
+\cite{Wagner2009-CVPR}
+\cite{Wright2009-PAMI}
 
@@
 \subsection{Computational Complexity}
+The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$), 
+in terms of both computation cost and memory-bandwidth.
+Therefore, assuming the iterative solver always takes a constant number of iterations,
+the computational complexity of the overall algorithm is linear in the problem size.
 
@@
 {\small
-\bibliographystyle{ieee}
+\bibliographystyle{../authorkit/ieee}
 \bibliography{faces}

### 2010-10-20 paper/main.tex
Working on introduction.  Added some pseudocode.

@@
 \begin{abstract}
-Abstract
+
 \end{abstract}
@@
 
+Significant strides have been made in automatic face recognition by re-casting
+it as a sparse representation problem.  This core of this algorithm consists of
+stacking the test image into a vector $\y$, stacking the training images for
+all of the users into the columns of a matrix $A$, and solving the following
+minimization problem:
+\begin{equation}
+\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
+%The solution we are looking for should be sparse in $\x$ because
+
 \cite{Wagner2009-CVPR}
@@
 
+For iterative alignment, we need to perform the following optimization problem:
+\begin{equation}
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+
 \section{Characteristics of the ALM Algorithm}
 Describe the two variants of the ALM algorithm
+
+We need to define the following soft-thresholding operator for a
+scalar $x$ and a scalar $\alpha \geq 0$:
+\begin{equation}
+\textup{shrink}(x,\alpha) = \textup{sign}(x)\cdot \max \{|x| - \alpha, 0\},
+\end{equation}
+In practice, this operator will be applied elementwise to a vector $\x$ with a single $\lambda$,
+and will be notated as $\texup{shrink}(\x,\alpha).
+
 Show Pseudocode for one variant
@@
 cuda cores, SIMD width, scheduling restrictions, bandwidth
+\subsection{Solving a Single L1 minimization}
+This section is about Victor's CUBLAS implementations.
+Plot Solver Bandwidth in GB/s vs. Problem size in MB
 
@@
 
+We summarize the entire ALM
+algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
+largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
+in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
+\begin{algorithm}[h]
+\caption{\bf (Augmented Lagrange Multiplier Method for Global
+Recognition)}
+\begin{algorithmic}[1]
+\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
+\mathbf{0}$.
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
+\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
+\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
+\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
+\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
+\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
+\sqrt{1+4t_l^2}\right)$;
+\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
+\ENDWHILE
+\STATE $\x_{k+1} \leftarrow \w_{l}$;
+\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
+\ENDWHILE \STATE
+{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+\end{algorithmic}
+\label{alg:alm}
+\end{algorithm}
+
+\section{Conclusion}
+...How lessons from this paper extend (or don't extend) to other similar problems in computer vision
+
+...Reiterate how the appropriate hardware choice depends on the problem size.
+
 {\small

### 2010-10-21 allen_outline/ALM.tex
Added Allen's outline

\documentclass[times, 10pt,twocolumn, conference]{article}
\usepackage{cvpr}
\usepackage{times}
\usepackage{balance}
\usepackage{cite}
\usepackage{times}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{times}
\usepackage{color}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{latexsym, amsmath, amsfonts, amscd}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
%\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{principle}{Principle}
\newtheorem{problem}{Problem}
\newtheorem{property}{Property}
\newtheorem{example}{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{proposition}[theorem]{Proposition}
\def\arr{{\mathcal{A}}}
\newcommand{\Null}{{\mbox{Null}}}
\newcommand{\rank}{{\mbox{rank}}}
\renewcommand{\mathbf}{\boldsymbol}
\newcommand{\p}{{p}}
\renewcommand{\Re}{{\mathbb{R}}}
\newcommand{\Ce}{{\mathbb C}}
\newcommand{\Ze}{{\mathbb Z}}
\newcommand{\Pe}{{\mathbb P}}
\newcommand{\I}{\mathcal{I}}
\def\tt{{\boldsymbol{t}}}
\def\ee{{\boldsymbol{e}}}
\def\aa{{\boldsymbol{a}}}
\def\bb{{\boldsymbol{b}}}
\def\ss{{\boldsymbol{s}}}
\def\cc{{\boldsymbol{c}}}
\def\pp{{\boldsymbol{p}}}
\def\chat{{\boldsymbol{\hat{\c}}}}
\def\nn{{\boldsymbol{n}}}
\def\m{{\boldsymbol{m}}}
\def\X{{\boldsymbol{X}}}
\def\L{{\boldsymbol{L}}}
\def\l{{\boldsymbol{l}}}
\def\Y{{\boldsymbol{Y}}}
\def\V{{\boldsymbol{V}}}
\def\uu{{\boldsymbol{u}}}
\def\vv{{\boldsymbol{v}}}
\def\yy{{\boldsymbol{y}}}
\def\xx{{\boldsymbol{x}}}
\def\ww{{\boldsymbol{w}}}
\def\zz{{\boldsymbol{z}}}
\def\uu{{\boldsymbol{u}}}
\def\F{{\boldsymbol{F}}}
\def\veronese{{\nu}}
\def\Qveronese{{\mu}}
\def\H{{\mathcal{H}}}
\def\Q{{\mathcal{Q}}}
\def\E{{\mathcal{E}}}
\def\L{{\mathcal{L}}}
\def\M{{\mathcal{M}}}
\def\A{{\mathcal{A}}}
\def\B{{\mathcal{B}}}
\def\F{{\mathcal{F}}}
\def\field{\mathbb{F}}
\def\QED{~\rule[-1pt]{5pt}{5pt}\par\medskip}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{10 \vspace{-0.2in}} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

\twocolumn

%%%%%%%%% TITLE
\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}

\author{Anonymous\\
Paper ID: XXX
}

\maketitle
% \thispagestyle{empty}
%%%%%%%%% ABSTRACT
\begin{abstract}

\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:introduction}
{\bf Drew, please draft this section.}

[Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization].

[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.]

[Zero in and justify why we choose ALM for our implementation.]

[Further zero in and limit the application to face recognition. What are the special data structure. What are the selected hardware architecture. Why it is important to consider both CPU and GPU implementation (difference in cores and caches)].

\subsection{Literature review}
[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]

\subsection{Contributions of the paper}
[Can leave empty until later.]

\section{Parallelization using Augmented Lagrange Multiplier Methods}
{\bf Victor and Mark, please draft this section.}

[An discussion of the original ALM algorithm should go first]

[Parallelization steps go here]

[A discussion about computational complexity and memory cost should go here]

\section{Implementation in Many-Core CPU/GPU Architectures}
\label{sec:algorithm}

{\bf Victor and Drew, please draft this section}

\subsection{CPU Implementation}

\subsection{GPU Implementation}

\section{Experiment}
\label{sec:experiment}

{\bf Drew, please draft this section}

\section{Conclusion}
\label{sec:conclusion}

{\small
\bibliographystyle{ieee}
\bibliography{paper}
}

\end{document}

### 2010-10-21 paper/main.tex
Fixed some missing textup macro
Merged in most of Allen's outline.

@@
 % Packages added by us
+\usepackage{amsfonts}
 \usepackage{mydefs}
@@
 
-\title{Efficient Parallelization of Sparse Representation for Face Recognition}
+%\title{Efficient Parallelization of Sparse Representation for Face Recognition}
+%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
+%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
+\title{Efficient Parallelization of $\ell_1$-Minimization for Face Recognition
+\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
 
@@
 \section{Introduction}
+
+[Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization].
+
+[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.]
+
+[Zero in and justify why we choose ALM for our implementation.]
+
+\subsection{Literature review}
+[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
+
+\subsection{Contributions of the paper}
+[Can leave empty until later.]
 
@@
 In practice, this operator will be applied elementwise to a vector $\x$ with a single $\lambda$,
-and will be notated as $\texup{shrink}(\x,\alpha).
+and will be notated as $\textup{shrink}(\x,\alpha)$.
 

### 2010-10-21 paper/main.tex
Pulled in my notes on results from a synthetic benchmark we did.

@@
 
+\section{Benchmarks on Synthetic Data}
+\subsection{PALM L1 minimization benchmark}
+
+Optimization:
+\begin{equation}
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+
+$A$ is $mxn$
+$m = 5120$
+$n = 32$
+
+One of these minimizations is performed per 
+alignment iteration in our face recognition pipeline.
+
+All of the following are implementations of primal ALM.
+Primal ALM contains outer and inner iterations.
+To simplify the benchmark, the number of outer iterations
+is clamped to 50, and the number of inner iterations is
+clamped to 50 iterations per outer loop.  Running with
+a convergence criterion will typically result in fewer
+iterations, depending on the data, so these numbers are
+for comparison across hardware, and may look worse than
+with other benchmarks on face data.
+
+To generate a meaningful comparison between different hardware
+architectures, we compare on a per-board basis:
+
+For CPU implementations, the benchmark makes use of
+as all of the cores in as many cpu's are present.
+
+For GPU implementations, the benchmark makes the best use
+of the entire GPU chip (most GPU boards have a single GPU chip).
+
+Since different platforms may perform most efficiently when
+executing multiple L1 minimizations simultaneously,
+we quote the average time required to solve a single L1
+minimization problem instance.
+
+schur:  Mac Pro, dual socket, quad core, hyperthreading intel nehalem @ 2.66 GHz
+
+Nvidia GTX480: This is the current state of the art consumer card.  
+It is clocked slightly faster than the pro version of the card, so it 
+should have slightly faster single precision float performance. (double precision is crippled, though)
+GTX480 has 15 cores, each of which has 2x32-wide floating point pipelines.
+All Fermi GPU's have 768KB L2 cache.
+
+Zihan's OpenCV implementation: *** 190 ms per L1 minimization ***
+amortized over 16 instances,
+four of which run simultaneously.  Tuning note:  the cpus
+can support up to 16 logical threads, but 4 instances seem
+sufficient to saturate the memory architecture.  Using more
+than 5 instances results in lower performance, perhaps due
+to less efficient usage of the cpu caches.
+
+Zihan's MATLAB implementation: *** 60 ms per L1 minimization ***
+This is amortized over 100 sequential instances.  All 16 logical
+cores were used, i.e. showed 100% usage in htop.  Pararllelism
+is a result of MATLAB's usage of Intel's MKL, a parallel blas
+implementation. I believe performance is very high to
+efficient cache usage:
+
+The A matrix takes up 5120*32*8 bytes = 1.25 MB
+The two processors have 16MB L3 Cache in total.
+The two processors have 2MB L2 Cache in total.
+
+Drew's pure CUDA, single kernel implementation: *** 36 ms per L1 minimization ***
+This implementation performs L1 minimization on 60 instances of L1.
+They are all run as a single cuda kernel.  
+Each instance of L1 is mapped onto a cuda block. (tuned to 64 threads / block)
+This means that the GPU hardware will schedule the multiple instances to run
+on the (15 for GTX480) cuda cores as concurrently (15? 30?) as possible.  
+For this to work, all of the data has to be ready at the same time.
+
+Drew's pure CUDA, streams implementation: *** 70 ms per L1 minimization ***
+This implementation performs L1 minimization on 60 instances of L1.
+Each instance of L1 is mapped onto a cuda block. (tuned to 128 threads / block)
+Each instance is run in a separate single-block CUDA kernel.
+The 60 kernels are added to 15 cuda streams.
+The GPU (hardware? driver?) then runs the kernels, up to 15 concurrently.
+
+Victor's CUDA/cublas mixed implementation: *** 302 ms per L1 minimization ***
+This implementation makes use of hand-written kernels for vector-vector operations,
+and cublas for matrix-vector operations.  There are two implementations under development:
+1. Perform a single L1 minimization using the entire GPU.
+2. Use cuda streams to perform multiple L1 minimizations concurrently on the GPU.
+Benchmarked on Drew's GTX480
+
+Drew's OpenMP implementation: *** 200 ms per L1 minimization***
+Benchmarked on schur(Dual Socket Quad Core Nehalem), all compiler optimizations turned on.
+This divides the matrix operations into 8 chunks, each of which is handled by a thread
+on a different CPU core.  Fine-grained parallelism is achieved through SSE2 vector
+floating point operations.  This implementation should ultimately beat the MATLAB/MKL
+implementation, especially for small problem sizes, since the primary advantage
+over MKL is a reduction in the number of times image-sized temporary variables
+are written and read from memory.
+
+Drew's Single-Threaded MKL implementation: *** 226 ms per L1 minimization ***
+Benchmarked on schur(Dual Socket Quad Core Nehalem), all compiler optimizations turned on.
+Uses MKL for single-threaded matrix-vector multiplies.  
+Saves ~almost as many writes of temporary vector variables as the previous OpenMP implementation.
+Shrinkage operator improved; at least with g++ there is now no branching or function call overhead.
+This implementation uses a single CPU core...
+... so if linear speedup can be achieved, could possibly get *** 28 ms per minimization ***
+
 \section{Application to Face Recognition}

### 2010-10-21 paper/main.tex
Working on intro...

@@
 This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
-%The solution we are looking for should be sparse in $\x$ because
+Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
+correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
+that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
+The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.
 
-\cite{Wagner2009-CVPR}
-\cite{Wright2009-PAMI}
-
+For this idea to work, all of the images must be aligned. \cite{Wagner2009-CVPR}
 For iterative alignment, we need to perform the following optimization problem:

### 2010-10-26 paper/main.tex
More on face recognition...

@@
 [Can leave empty until later.]
-
-Significant strides have been made in automatic face recognition by re-casting
-it as a sparse representation problem.  This core of this algorithm consists of
-stacking the test image into a vector $\y$, stacking the training images for
-all of the users into the columns of a matrix $A$, and solving the following
-minimization problem:
-\begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y = A \x + \e.
-\end{equation}
-This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
-Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
-correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
-that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
-The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.
-
-For this idea to work, all of the images must be aligned. \cite{Wagner2009-CVPR}
-For iterative alignment, we need to perform the following optimization problem:
-\begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
-\end{equation}
 
@@
 \section{Application to Face Recognition}
+
+Significant strides have been made in automatic face recognition by re-casting
+it as a sparse representation problem.  This core of this algorithm consists of
+Significant strides have been made in automatic face recognition by re-casting
+it as a sparse representation problem.  This core of this algorithm consists of
+stacking the test image into a vector $\y$, stacking the training images for
+all of the users into the columns of a matrix $A$, and solving the following
+minimization problem:
+\begin{equation}
+\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
+Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
+correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
+that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
+The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.
+
+For this idea to work, all of the images must be aligned. \cite{Wagner2009-CVPR}
+For iterative alignment, we need to perform the following optimization problem:
+\begin{equation}
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+stacking the test image into a vector $\y$, stacking the training images for
+all of the users into the columns of a matrix $A$, and solving the following
+minimization problem:
+\begin{equation}
+\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
+Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
+correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
+that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
+The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.
+
+For this idea to work, all of the images must be aligned. \cite{Wagner2009-CVPR}
+For iterative alignment, we need to perform the following optimization problem:
+\begin{equation}
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\end{equation}
+
 \begin{algorithm}[thb]

### 2010-10-29 results/bench_PALM.m
adding PALM GPU benchmark data and temporary SVG files bench_PALM.m outputs to SVG now, but an external library is required from http://www.mathworks.com/matlabcentral/fileexchange/7401-scalable-vector-graphics-svg-export-of-figures

@@ -1,0 +1,69 @@
+function rtn = bench_PALM()
+
+type = 'PALM';
+n  = 10; % denominator for ratio
+ks = 0.1;
+tols = [1e-4 1e-5 1e-6 1e-7];
+maxIter = 5000;
+iterations = 10;
+dimsPowers = [500:100:900 1000:1000:10000];
+lambda = 10;
+ratio = 1:5;
+
+h = figure;
+
+for i=ratio
+    name = [type '_' num2str(n) '_' num2str(i) '_'];
+    
+    gpuName = [name 'GPU.mat'];
+    
+    gpuData = load(gpuName);
+    
+    cpuName = [name 'CPU.mat'];
+    
+    cpuData = load(cpuName);
+    
+    dimsPowers = [500:100:900 1000:1000:2000];
+    lenPlot = numel(dimsPowers);
+    
+    for j=1:numel(tols)
+        
+        subplot(numel(tols), 3, (j-1)*3+1);
+        plotA = gpuData.averageRunTime(j,j,:,2,2);
+        plotB = cpuData.averageRunTime(j,j,:,1,2);
+        plot(dimsPowers, squeeze(plotA(1:lenPlot)), 'b');
+        hold on
+        plot(dimsPowers, squeeze(plotB(1:lenPlot)), 'r');
+        title(['Average runtime - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        hold off
+        
+        subplot(numel(tols), 3, (j-1)*3+2);
+        plotA = gpuData.averageL2error(j,j,:,2);
+        plotB = cpuData.averageL2error(j,j,:,1);
+        plot(dimsPowers, squeeze(plotA(1:lenPlot)), 'b');
+        hold on
+        plot(dimsPowers, squeeze(plotB(1:lenPlot)), 'r');
+        title(['Average L2 error - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        hold off
+        
+        subplot(numel(tols), 3, (j-1)*3+3);
+        plotA = gpuData.averageIteration(j,j,:,2);
+        plotB = cpuData.averageIteration(j,j,:,1);
+        plot(dimsPowers, squeeze(plotA(1:lenPlot)), 'b');
+        hold on
+        plot(dimsPowers, squeeze(plotB(1:lenPlot)), 'r');
+        title(['Average iterations - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        hold off
+    end
+    
[…]

### 2010-11-01 paper/main.tex
Intro drafted... working on ALM description.

--- 2010-10-21 paper/main.tex
@@
 \usepackage{mydefs}
+\usepackage{algorithmic}
 \usepackage{algorithm}
-\usepackage{algorithmic}
 
@@
 
+%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
 %\title{Efficient Parallelization of Sparse Representation for Face Recognition}
-%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
 %\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
@@
 
-[Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization].
+%Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization.
+$\ell_1$-minimization is one of the most powerful tools to be added to the engineer's
+toolbox in years.  It has strong theoretical motivation for both for use in recovering sparse
+solutions to systems of linear equations, and has been shown to perform well even in the
+presence of a combination of sparse and dense error.  Since solutions of linear systems
+arise in practically every branch of engineering and there are very few application which
+do {\em not} benefit from a robust error function, the applications are far to numerous to
+itemize here.  
+For many applications of $\ell_1$-minimization solving a system $b=Ax$, $A$ often has a
+special structure that can be leveraged to compute portions of $A$ on the fly
+as they are needed, speeding up perfomance of the minimization dramatically.
+Unfortunately, there are many import applications where this does not hold true.
+This paper targets applications where $\b$ and $A$ contain
+image data, and thus must be loaded every time they are used, as is the 
+case for many problems in computer vision.
+The techniques presented in this paper target these applications, as well
+as others where the dictionary is dense and unstructured. $A$ must be loaded every time
+it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
+that arithmetic logic.
 
-[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.]
-
-[Zero in and justify why we choose ALM for our implementation.]
-
-\subsection{Literature review}
-[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
-
-\subsection{Contributions of the paper}
-[Can leave empty until later.]
-
-Significant strides have been made in automatic face recognition by re-casting
-it as a sparse representation problem.  This core of this algorithm consists of
-stacking the test image into a vector $\y$, stacking the training images for
+% Narrow focus for applications to Recognition Application
+The primary motivating application for this paper is face recognition for access control applications.
+Recently, advances in automatic face recognition have been made by re-casting
+it as a sparse representation problem.  This core of this technique consists of
+re-casting recognition as a sparse representation problem. This is done by  
+stacking the test image into a vector $\b$, stacking the training images for
 all of the users into the columns of a matrix $A$, and solving the following
@@
 \begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \b = A \x + \e.
 \end{equation}
@@
 
-For this idea to work, all of the images must be aligned. \cite{Wagner2009-CVPR}
-For iterative alignment, we need to perform the following optimization problem:
+For this idea to work, all of the images must be aligned. An iterative alignment
+algorithm for this purpose was described as part of the \cite{Wagner2009-CVPR}, which was
+the first paper to expand \cite{Wright2009-PAMI} into a functioning face recognition system.
+This iterative alignment routine is based on repeatedly solving the following optimization problem:
 \begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \b = A_k \x + \e.
 \end{equation}
+Where $A_k$ contains the training images for user $k$ as well as several vectors consisting of the
+Jacobian of the test image $\b$ with respect to the transformation parameters, and $\x$ contains
+both the current representation coefficients, as well as an update to the transformation parameters.
+
+\subsection{Literature review}
+[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
+
+[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.]
+[Allen]
+
+[Zero in and justify why we choose ALM for our implementation.]
+
+\subsection{Contributions of the paper}
 
@@
 Describe the two variants of the ALM algorithm
+
+	%Major steps:
+	%while:
+		%mu *= rho
+		%lambda += mu(y - Ax - e)
+		%while:
+			%e <- shrink(y-Ax+lambda/mu, 1/mu)
+			%x <- Apinvt'*(y-e+lamba/mu)
+We summarize the entire ALM
+algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
+largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
+in \cite{YangJ2009-pp} and set $\mu = 2m / \|\b\|_1$.
+\begin{algorithm}[h]
+\caption{\bf (Augmented Lagrange Multiplier Method for Global
+Recognition)}
+\begin{algorithmic}[1]
+\STATE {\bf Input:} $\b \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\e_1 = \b$, $\blamda_1 =
+\mathbf{0}$.
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $\e_{k+1} = \textup{shrink}\left(\b - A\x_k + \frac{\blamda_k}{\mu}, \frac{1}{\mu}\right)$;
+\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
+\frac{1}{\gamma}A^T\left(\b - A\z_l - \e_{k+1} +
+\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
+\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
+\sqrt{1+4t_l^2}\right)$;
+\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
+\ENDWHILE
+\STATE $\x_{k+1} \leftarrow \w_{l}$;
+\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\b - A\x_{k+1} - \e_{k+1})$;
+\ENDWHILE \STATE
+{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+\end{algorithmic}
+\label{alg:alm}
+\end{algorithm}
 
@@
 \begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \b = A \x + \e.
 \end{equation}
@@
 is clamped to 50, and the number of inner iterations is
-clamped to 50 iterations per outer loop.  Running with
+clamped to 50 iterations per outer loopsretrsitrsi.  Running with
 a convergence criterion will typically result in fewer
@@
 
-\section{Application to Face Recognition}
 \begin{algorithm}[thb]
@@
 \STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
-$\y\in\Re^m$ and a deformation group $T$. 
+$\b\in\Re^m$ and a deformation group $T$. 
 \STATE
@@
 \STATE \hspace{6mm}
-$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
+$\tilde \b(\tau) \leftarrow \frac{\b \circ \tau}{\|\b \circ
 \tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
-%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
+\tau} \tilde\b(\tau)  \bigr|_{\tau^{(j)}} $;
+%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \b + J \Delta \tau = A_k \x + \e \end{array}\right.$
 \STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$ 
+\subj \; \tilde \b + J \Delta \tau = A_i \x + \e.$ 
 \STATE
@@
 $\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
+\STATE Update $\b \leftarrow \b \circ \bar{\tau}$ and $\tau_i
 \leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
@@
 $\ell^1$-minimization problem:
-$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$$
-\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
+$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \b = A \x + \e.$$
+\STATE Compute residuals $r_i(\b) = \| {\b} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
+\STATE {\bf Output:} $\mbox{identity}(\b) = \arg\min_i r_i(\b)$.
 \end{algorithmic}
@@
 largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
+in \cite{YangJ2009-pp} and set $\mu = 2m / \|\b\|_1$.
 \begin{algorithm}[h]
@@
 \begin{algorithmic}[1]
-\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
+\STATE {\bf Input:} $\b \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\e_1 = \b$, $\blamda_1 =
 \mathbf{0}$.
 \WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
+\STATE $\e_{k+1} = \textup{shrink}\left(\b - A\x_k +
 \frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
@@
 \STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
+\frac{1}{\gamma}A^T\left(\b - A\z_l - \e_{k+1} +
 \frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
@@
 \STATE $\x_{k+1} \leftarrow \w_{l}$;
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
+\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\b - A\x_{k+1} - \e_{k+1})$;
 \ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+{\bf Output:} $\x^* \leAftarrow \x_k, \e^* \leftarrow \e_k$.
 \end{algorithmic}

### 2010-11-01 results/bench_PALM.m
accepts different dimsPowers arrays

@@ -23,35 +23,37 @@ for i=ratio
     
     cpuData = load(cpuName);
     
-    dimsPowers = [500:100:900 1000:1000:2000];
-    lenPlot = numel(dimsPowers);
+    gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 2000);
+    gpuDimLen     = numel(gpuDimsPowers);
+    cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 2000);
+    cpuDimLen     = numel(cpuDimsPowers);
     
     for j=1:numel(tols)
         
         subplot(numel(tols), 3, (j-1)*3+1);
         plotA = gpuData.averageRunTime(j,j,:,2,2);
         plotB = cpuData.averageRunTime(j,j,:,1,2);
-        plot(dimsPowers, squeeze(plotA(1:lenPlot)), 'b');
+        plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
         hold on
-        plot(dimsPowers, squeeze(plotB(1:lenPlot)), 'r');
+        plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
         title(['Average runtime - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
         hold off
         
         subplot(numel(tols), 3, (j-1)*3+2);
         plotA = gpuData.averageL2error(j,j,:,2);
         plotB = cpuData.averageL2error(j,j,:,1);
-        plot(dimsPowers, squeeze(plotA(1:lenPlot)), 'b');
+        plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
         hold on
-        plot(dimsPowers, squeeze(plotB(1:lenPlot)), 'r');
+        plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
         title(['Average L2 error - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
         hold off
         
         subplot(numel(tols), 3, (j-1)*3+3);
         plotA = gpuData.averageIteration(j,j,:,2);
         plotB = cpuData.averageIteration(j,j,:,1);
-        plot(dimsPowers, squeeze(plotA(1:lenPlot)), 'b');
+        plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
         hold on
[…]

### 2010-11-01 paper/main.tex
Fixed algorithm pseudocode. also, y->b

@@
 Unfortunately, there are many import applications where this does not hold true.
-This paper targets applications where $\b$ and $A$ contain
+This paper targets applications where $\bb$ and $A$ contain
 image data, and thus must be loaded every time they are used, as is the 
@@
 re-casting recognition as a sparse representation problem. This is done by  
-stacking the test image into a vector $\b$, stacking the training images for
+stacking the test image into a vector $\bb$, stacking the training images for
 all of the users into the columns of a matrix $A$, and solving the following
@@
 \begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \b = A \x + \e.
+\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
 \end{equation}
@@
 \begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \b = A_k \x + \e.
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_k \x + \e.
 \end{equation}
 Where $A_k$ contains the training images for user $k$ as well as several vectors consisting of the
-Jacobian of the test image $\b$ with respect to the transformation parameters, and $\x$ contains
+Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
 both the current representation coefficients, as well as an update to the transformation parameters.
@@
 largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu = 2m / \|\b\|_1$.
+in \cite{YangJ2009-pp} and set $\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
 \begin{algorithm}[h]
-\caption{\bf (Augmented Lagrange Multiplier Method for Global
-Recognition)}
+\caption{\bf (Augmented Lagrange Multiplier Method Used in Alignment Inner Loop)}
 \begin{algorithmic}[1]
-\STATE {\bf Input:} $\b \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \b$, $\blamda_1 =
-\mathbf{0}$.
+\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
 \WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}\left(\b - A\x_k + \frac{\blamda_k}{\mu}, \frac{1}{\mu}\right)$;
-\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
 \WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\b - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$;
-\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
+\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
+\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
 \ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$;
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\b - A\x_{k+1} - \e_{k+1})$;
+\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
+\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
 \ENDWHILE \STATE
@@
 \begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \b = A \x + \e.
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
 \end{equation}
 
-$A$ is $mxn$
+$A$ is $m \times n$
 $m = 5120$
@@
 
-\begin{algorithm}[thb]
-\caption{\bf (Deformable Sparse Recovery and Classification for
-Face Recognition)} \label{alg:deformable-src}
-\begin{algorithmic}[1]
-\STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
-$\b\in\Re^m$ and a deformation group $T$. 
-\STATE
-{\bf for} each subject $i$, 
-\STATE \hspace{3mm} $\tau^{(0)}
-\leftarrow I$. 
-\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
-\STATE \hspace{6mm}
-$\tilde \b(\tau) \leftarrow \frac{\b \circ \tau}{\|\b \circ
-\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-\tau} \tilde\b(\tau)  \bigr|_{\tau^{(j)}} $;
-%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \b + J \Delta \tau = A_k \x + \e \end{array}\right.$
-\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-\subj \; \tilde \b + J \Delta \tau = A_i \x + \e.$ 
-\STATE
-\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$; 
-\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
-the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
-residuals $\|\e\|_1$. \STATE Compute an average transformation
-$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-\STATE Update $\b \leftarrow \b \circ \bar{\tau}$ and $\tau_i
-\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
-k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
-\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
-\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-$\ell^1$-minimization problem:
-$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \b = A \x + \e.$$
-\STATE Compute residuals $r_i(\b) = \| {\b} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-\STATE {\bf Output:} $\mbox{identity}(\b) = \arg\min_i r_i(\b)$.
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}[thb]
+%\caption{\bf (Deformable Sparse Recovery and Classification for
+%Face Recognition)} \label{alg:deformable-src}
+%\begin{algorithmic}[1]
+%\STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
+%$\bb\in\Re^m$ and a deformation group $T$. 
+%\STATE
+%{\bf for} each subject $i$, 
+%\STATE \hspace{3mm} $\tau^{(0)}
+%\leftarrow I$. 
+%\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
+%\STATE \hspace{6mm}
+%$\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ
+%\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
+%\tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
+%%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \bb + J \Delta \tau = A_k \x + \e \end{array}\right.$
+%\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
+%\subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
+%\STATE
+%\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
+%\tau$; 
+%\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
+%the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
+%residuals $\|\e\|_1$. \STATE Compute an average transformation
+%$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
+%\STATE Update $\bb \leftarrow \bb \circ \bar{\tau}$ and $\tau_i
+%\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
+%k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
+%\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
+%\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
+%$\ell^1$-minimization problem:
+%$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \bb = A \x + \e.$$
+%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
+%\STATE {\bf Output:} $\mbox{identity}(\bb) = \arg\min_i r_i(\bb)$.
+%\end{algorithmic}
+%\end{algorithm}
 
-We summarize the entire ALM
-algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
-largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu = 2m / \|\b\|_1$.
-\begin{algorithm}[h]
-\caption{\bf (Augmented Lagrange Multiplier Method for Global
-Recognition)}
-\begin{algorithmic}[1]
-\STATE {\bf Input:} $\b \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \b$, $\blamda_1 =
-\mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}\left(\b - A\x_k +
-\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
-\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\b - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$;
-\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
-\ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$;
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\b - A\x_{k+1} - \e_{k+1})$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leAftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{algorithmic}
-\label{alg:alm}
-\end{algorithm}
+%We summarize the entire ALM
+%algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
+%largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
+%in \cite{YangJ2009-pp} and set $\mu = 2m / \|\bb\|_1$.
+%\begin{algorithm}[h]
+%\caption{\bf (Augmented Lagrange Multiplier Method for Global
+%Recognition)}
+%\begin{algorithmic}[1]
+%\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
+%$\x_1 = \mathbf{0}$, $\e_1 = \bb$, $\blamda_1 =
+%\mathbf{0}$.
+%\WHILE{not converged ($k = 1,2,\ldots$)}
+%\STATE $\e_{k+1} = \textup{shrink}\left(\bb - A\x_k +
+%\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
+%\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
+%\WHILE{not converged ($l = 1,2,\ldots$)}
+%\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
+%\frac{1}{\gamma}A^T\left(\bb - A\z_l - \e_{k+1} +
+%\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
+%\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
+%\sqrt{1+4t_l^2}\right)$;
+%\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
+%\ENDWHILE
+%\STATE $\x_{k+1} \leftarrow \w_{l}$;
+%\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
+%\ENDWHILE \STATE
+%{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+%\end{algorithmic}
+%\label{alg:alm}
+%\end{algorithm}
 

### 2010-11-01 results/bench_PALM.m
new results

@@ -3,7 +3,7 @@ function rtn = bench_PALM()
 type = 'PALM';
 n  = 10; % denominator for ratio
 ks = 0.1;
-tols = [1e-4 1e-5 1e-6 1e-7];
+tols = [1e-4 1e-5 1e-6];
 maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];

### 2010-11-02 results/bench_PALM.m
new GPU data plotting 2000:250:8000

@@ -3,7 +3,7 @@ function rtn = bench_PALM()
 type = 'PALM';
 n  = 10; % denominator for ratio
 ks = 0.1;
-tols = [1e-4 1e-5 1e-6];
+tols = [1e-4 1e-5];
 maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];
@@ -23,9 +23,9 @@ for i=ratio
     
     cpuData = load(cpuName);
     
-    gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 5000);
+    gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 8000);
     gpuDimLen     = numel(gpuDimsPowers);
-    cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 5000);
+    cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 8000);
     cpuDimLen     = numel(cpuDimsPowers);
     
     for j=1:numel(tols)

### 2010-11-02 results/bench_PALM.m
new plots

@@ -23,9 +23,9 @@ for i=ratio
     
     cpuData = load(cpuName);
     
-    gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 8000);
+    gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 6000);
     gpuDimLen     = numel(gpuDimsPowers);
-    cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 8000);
+    cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 6000);
     cpuDimLen     = numel(cpuDimsPowers);
     
     for j=1:numel(tols)

### 2010-11-04 results/recognition/bench_rec.m
reorganizing results. added recognition benchmark results

@@ -1,0 +1,53 @@
+function rtn = bench_rec()
+
+type = 'PALM';
+tols = [1e-3 1e-4 1e-5];
+
+h = figure;
+
+name = [type '_rec_'];
+
+gpuName = [name 'GPU.mat'];
+
+gpuData = load(gpuName);
+
+cpuName = [name 'CPU.mat'];
+
+cpuData = load(cpuName);
+
+gpuNumUsers = gpuData.numUsers(gpuData.numUsers < 50);
+gpuNUlen    = numel(gpuNumUsers);
+cpuNumUsers = cpuData.numUsers(cpuData.numUsers < 50);
+cpuNUlen    = numel(cpuNumUsers);
+
+for j=1:numel(tols)
+    
+    subplot(numel(tols), 2, (j-1)*3+1);
+    plotA = gpuData.averageRunTime(j,:,2,2);
+    plotB = cpuData.averageRunTime(j,:,1,2);
+    plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
+    hold on
+    plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
+    title(['Average runtime - tolerance ' num2str(tols(j))]);
+    hold off
+    
+    subplot(numel(tols), 2, (j-1)*3+2);
+    plotA = gpuData.averageIteration(j,:,2);
+    plotB = cpuData.averageIteration(j,:,1);
+    plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
+    hold on
+    plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
+    title(['Average iterations - tolerance ' num2str(tols(j))]);
+    hold off
+end
+
+set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
+set (gcf,'Units','Inches');                     %   Using units of inches
+pos=get(gcf,'position');                        %   Save position values
+set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
+set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
+
+plot2svg([type '_benchmark.svg'], h, 'png');
+%saveas(h, [type '_benchmark_ratio_' num2str(i)], 'jpg');
+
+end

### 2010-11-04 results/recognition/bench_rec.m
updated plots

@@ -22,22 +22,24 @@ cpuNUlen    = numel(cpuNumUsers);
 
 for j=1:numel(tols)
     
-    subplot(numel(tols), 2, (j-1)*3+1);
+    subplot(numel(tols), 2, (j-1)*2+1);
     plotA = gpuData.averageRunTime(j,:,2,2);
     plotB = cpuData.averageRunTime(j,:,1,2);
     plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
     hold on
     plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
     title(['Average runtime - tolerance ' num2str(tols(j))]);
+    xlabel('Number of Users');
     hold off
     
-    subplot(numel(tols), 2, (j-1)*3+2);
+    subplot(numel(tols), 2, (j-1)*2+2);
     plotA = gpuData.averageIteration(j,:,2);
     plotB = cpuData.averageIteration(j,:,1);
     plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
     hold on
     plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
     title(['Average iterations - tolerance ' num2str(tols(j))]);
+    xlabel('Number of Users');
     hold off
 end
 

### 2010-11-06 results/random_data/bench_PALM.m
added BLAS implementation of PALM in ratio 1o to 3 plots

@@ -8,7 +8,7 @@ maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];
 lambda = 10;
-ratio = 1:5;
+ratio = 1:3;
 
 h = figure;
 
@@ -23,37 +23,49 @@ for i=ratio
     
     cpuData = load(cpuName);
     
+    cpuName = [name 'BLAS.mat'];
+    
+    blasData = load(cpuName);
+    
     gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 6000);
     gpuDimLen     = numel(gpuDimsPowers);
     cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 6000);
     cpuDimLen     = numel(cpuDimsPowers);
+    blasDimsPowers = blasData.dimsPowers(blasData.dimsPowers <= 6000);
+    blasDimLen     = numel(blasDimsPowers);
     
     for j=1:numel(tols)
         
         subplot(numel(tols), 3, (j-1)*3+1);
         plotA = gpuData.averageRunTime(j,j,:,2,2);
         plotB = cpuData.averageRunTime(j,j,:,1,2);
+        plotC = blasData.averageRunTime(j,j,:,1,3);
         plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
         hold on
         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
+        plot(cpuDimsPowers, squeeze(plotC(1:blasDimLen)), 'g+-');
         title(['Average runtime - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
         hold off
         
         subplot(numel(tols), 3, (j-1)*3+2);
         plotA = gpuData.averageL2error(j,j,:,2);
         plotB = cpuData.averageL2error(j,j,:,1);
+        plotC = blasData.averageL2error(j,j,:,3);
         plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
         hold on
         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
+        plot(cpuDimsPowers, squeeze(plotC(1:blasDimLen)), 'g+-');
         title(['Average L2 error - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
         hold off
         
         subplot(numel(tols), 3, (j-1)*3+3);
[…]

### 2010-11-07 results/random_data/bench_PALM.m
new data with new GPU and BLAS data

@@ -8,7 +8,7 @@ maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];
 lambda = 10;
-ratio = 1:3;
+ratio = 1:5;
 
 h = figure;
 

### 2010-11-07 results/random_data/bench_PALM.m
new plots

@@ -44,7 +44,9 @@ for i=ratio
         hold on
         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
         plot(cpuDimsPowers, squeeze(plotC(1:blasDimLen)), 'g+-');
-        title(['Average runtime - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        title(['Tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        ylabel('Average runtime (s)');
+        xlabel('Height of A');
         hold off
         
         subplot(numel(tols), 3, (j-1)*3+2);
@@ -55,7 +57,9 @@ for i=ratio
         hold on
         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
         plot(cpuDimsPowers, squeeze(plotC(1:blasDimLen)), 'g+-');
-        title(['Average L2 error - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        title(['Tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        ylabel('Average L2 error');
+        xlabel('Height of A');
         hold off
         
         subplot(numel(tols), 3, (j-1)*3+3);
@@ -66,7 +70,9 @@ for i=ratio
         hold on
         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
         plot(cpuDimsPowers, squeeze(plotC(1:blasDimLen)), 'g+-');
-        title(['Average iterations - tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        title(['Tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        ylabel('Average iterations');
+        xlabel('Height of A');
         hold off
     end
     

### 2010-11-07 results/random_data/bench_PALM.m
updated plots with legend

@@ -49,6 +49,10 @@ for i=ratio
         xlabel('Height of A');
         hold off
         
+        if j == 1
+            legend('GPU', 'CPU', 'BLAS');
+        end
+        
         subplot(numel(tols), 3, (j-1)*3+2);
         plotA = gpuData.averageL2error(j,j,:,2);
         plotB = cpuData.averageL2error(j,j,:,1);

### 2010-11-08 results/random_data/bench_PALM.m
removing old data for recognition. random_data - plots for runtime are... matrix dimension vs runtime

@@ -23,30 +23,43 @@ for i=ratio
     
     cpuData = load(cpuName);
     
-    cpuName = [name 'BLAS.mat'];
+    blasName = [name 'BLAS.mat'];
     
-    blasData = load(cpuName);
+    blasData = load(blasName);
     
-    gpuDimsPowers = gpuData.dimsPowers(gpuData.dimsPowers <= 6000);
-    gpuDimLen     = numel(gpuDimsPowers);
-    cpuDimsPowers = cpuData.dimsPowers(cpuData.dimsPowers <= 6000);
-    cpuDimLen     = numel(cpuDimsPowers);
+    gpuDimsPowers  = gpuData.dimsPowers(gpuData.dimsPowers <= 6000);
+    gpuDimLen      = numel(gpuDimsPowers);
+    cpuDimsPowers  = cpuData.dimsPowers(cpuData.dimsPowers <= 6000);
+    cpuDimLen      = numel(cpuDimsPowers);
     blasDimsPowers = blasData.dimsPowers(blasData.dimsPowers <= 6000);
     blasDimLen     = numel(blasDimsPowers);
     
+    gpuDimsPowers  = gpuDimsPowers * (10 + i) / 10 * 32 / (2^10);
+    cpuDimsPowers  = cpuDimsPowers * (10 + i) / 10 * 32 / (2^10);
+    blasDimsPowers = blasDimsPowers * (10 + i) / 10 * 32 / (2^10);
+        
     for j=1:numel(tols)
         
         subplot(numel(tols), 3, (j-1)*3+1);
+        gpuOutIter  = gpuData.averageIteration(j,j,:,2);
+        gpuInIter   = gpuData.averageTotalInnerIteration(j,j,:,2);
+        cpuOutIter  = cpuData.averageIteration(j,j,:,2);
+        cpuInIter   = cpuData.averageTotalInnerIteration(j,j,:,2);
+        blasOutIter = blasData.averageIteration(j,j,:,2);
+        blasInIter  = blasData.averageTotalInnerIteration(j,j,:,2);
+        
+        
+        
         plotA = gpuData.averageRunTime(j,j,:,2,2);
         plotB = cpuData.averageRunTime(j,j,:,1,2);
         plotC = blasData.averageRunTime(j,j,:,1,3);
         plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
         hold on
[…]

### 2010-11-08 results/recognition/bench_rec.m

@@ -6,58 +6,78 @@ tol_ints = [1e-3 1e-4 1e-5 1e-6];
 
 h = figure;
 
-name = [type '_rec_'];
+sizes = {'small', 'big'};
 
-gpuName = [name 'GPU.mat'];
-
-gpuData = load(gpuName);
-
-cpuName = [name 'CPU.mat'];
-
-cpuData = load(cpuName);
-
-gpuNumUsers = gpuData.numUsers(gpuData.numUsers < 50);
-gpuNUlen    = numel(gpuNumUsers);
-cpuNumUsers = cpuData.numUsers(cpuData.numUsers < 50);
-cpuNUlen    = numel(cpuNumUsers);
-
-for j=1:numel(tols)
+for size=sizes
     
-    plotNum = 1;
+    size = size{1};
     
-    for k = 1:numel(tol_ints);
+    name = [type '_rec_'];
+    
+    gpuName = [name 'GPU_' size '.mat'];
+    
+    gpuData = load(gpuName);
+    
+    cpuName = [name 'CPU_' size '.mat'];
+    
+    cpuData = load(cpuName);
+    
+    blasName = [name 'BLAS_' size '.mat'];
+    
+    blasData = load(blasName);
+    
+    gpuNumUsers = gpuData.numUsers(gpuData.numUsers < 50);
+    gpuNUlen    = numel(gpuNumUsers);
+    cpuNumUsers = cpuData.numUsers(cpuData.numUsers < 50);
+    cpuNUlen    = numel(cpuNumUsers);
+    blasNumUsers = blasData.numUsers(blasData.numUsers < 50);
+    blasNUlen    = numel(blasNumUsers);
+    
+    for j=1:numel(tols)
         
-        subplot(numel(tol_ints), 2, plotNum);
-        plotA = gpuData.averageRunTime(j,k,:,2,2);
-        plotB = cpuData.averageRunTime(j,k,:,1,2);
-        plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
-        hold on
-        plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
-        title(['Average runtime - outer tolerance ' num2str(tols(j)) '- inner tolerance ' num2str(tol_ints(k))]);
-        xlabel('Number of Users');
-        hold off
+        plotNum = 1;
         
-        subplot(numel(tol_ints), 2, plotNum+1);
[…]

### 2010-11-09 paper/main.tex
Removing cruft...
changed two subsection titles
started section 2 and 5

@@
 
-	%Major steps:
-	%while:
-		%mu *= rho
-		%lambda += mu(y - Ax - e)
-		%while:
-			%e <- shrink(y-Ax+lambda/mu, 1/mu)
-			%x <- Apinvt'*(y-e+lamba/mu)
+The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
+\begin{equation}
+F(x) = f(x) + \lambda g(x).
+\end{equation}
+
+Because the optimal lagrange multiplier values are not known, ALM iterates between estimating the solution and the lagrange multipier values.
+
+In order to solve the recognition stage, we consider the $l_1$ problem:
+\begin{equation}
+\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
+\end{equation}
+which can be formulated as
+\begin{equation}
+L_u(x,y) = \|x\|_1 + <\y, \b - A\x - \e> + \frac{\mu}2 \| \b-A\x-\e \|_2^2
+\end{equation}
+
+where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.
+
 We summarize the entire ALM
@@
 
+In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
+\begin{equation}
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
+\end{equation}
+which is solved in a similar fashion.
+
 We need to define the following soft-thresholding operator for a
@@
 
-\section{Efficient Implementation on the CPU}
+\section{CPU Implementation}
 % With a reshuffling of the order of operations and a change of the termination
@@
 
-\section{Efficient Implementation on the CPU}
+\section{GPU Implementation}
 \subsection{GPU Hardware Parallelism}
@@
 
+In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU Cuda implementations of the Primal ALM algorithm on synthetic data to find the point at which the GPU version runs faster than CPU versions.
+To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations at 5000 and run the algorithm to convergence.
+CPU code was run on an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480
+We ran the A matrix with different ratios between m:n from 10:1:5 for m:2000:25:8000 with an x vector of 10 percent sparsity.
+Tolerance levels were empirically seen to allow best performance with the inner loop tolerance set at one order of magnitude lower than the outer loop tolerance.
+
+For figures ...
+
+As the dimensions of A increase, so does the significance of the size of AtA.  Until the A and AtA matrix reach x Mb, the CPU version runs faster than the GPU.
+
 Optimization:
 \begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
+\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
 \end{equation}

### 2010-11-10 results/random_data/bench_PALM.m
new plots

@@ -34,10 +34,10 @@ for i=ratio
     blasDimsPowers = blasData.dimsPowers(blasData.dimsPowers <= 6000);
     blasDimLen     = numel(blasDimsPowers);
     
-    gpuDimsPowers  = gpuDimsPowers * (10 + i) / 10 * 32 / (2^10);
-    cpuDimsPowers  = cpuDimsPowers * (10 + i) / 10 * 32 / (2^10);
-    blasDimsPowers = blasDimsPowers * (10 + i) / 10 * 32 / (2^10);
-        
+    gpuDimsPowers  = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    cpuDimsPowers  = (cpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    blasDimsPowers = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    
     for j=1:numel(tols)
         
         subplot(numel(tols), 3, (j-1)*3+1);
@@ -48,8 +48,6 @@ for i=ratio
         blasOutIter = blasData.averageIteration(j,j,:,2);
         blasInIter  = blasData.averageTotalInnerIteration(j,j,:,2);
         
-        
-        
         plotA = gpuData.averageRunTime(j,j,:,2,2);
         plotB = cpuData.averageRunTime(j,j,:,1,2);
         plotC = blasData.averageRunTime(j,j,:,1,3);
@@ -59,26 +57,39 @@ for i=ratio
         plot(blasDimsPowers, squeeze(plotC(1:blasDimLen)), 'g+-');
         title(['Tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
         ylabel('Average speed (kb/s)');
-        xlabel('A Matrix (in kb)');
+        xlabel('A & G Matrix (in Mb)');
         hold off
         
         if j == 1
             legend('GPU', 'CPU', 'BLAS');
         end
-        
+%         
+%         subplot(numel(tols), 3, (j-1)*3+2);
+%         plotA = gpuData.averageL2error(j,j,:,2);
+%         plotB = cpuData.averageL2error(j,j,:,1);
+%         plotC = blasData.averageL2error(j,j,:,3);
+%         plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
+%         hold on
+%         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
[…]

### 2010-11-11 results/recognition/bench_rec.m
new plots for recognition

@@ -2,7 +2,7 @@ function rtn = bench_rec()
 
 type = 'PALM';
 tols = [1e-3 1e-4 1e-5];
-tol_ints = [1e-3 1e-4 1e-5 1e-6];
+tol_ints = [1e-4 1e-5 1e-6];
 
 h = figure;
 
@@ -26,48 +26,46 @@ for size=sizes
     
     blasData = load(blasName);
     
-    gpuNumUsers = gpuData.numUsers(gpuData.numUsers < 50);
+    gpuNumUsers = gpuData.numUsers(gpuData.numUsers < 240);
     gpuNUlen    = numel(gpuNumUsers);
-    cpuNumUsers = cpuData.numUsers(cpuData.numUsers < 50);
+    cpuNumUsers = cpuData.numUsers(cpuData.numUsers < 240);
     cpuNUlen    = numel(cpuNumUsers);
-    blasNumUsers = blasData.numUsers(blasData.numUsers < 50);
+    blasNumUsers = blasData.numUsers(blasData.numUsers < 240);
     blasNUlen    = numel(blasNumUsers);
     
     for j=1:numel(tols)
+        k = j;
         
         plotNum = 1;
         
-        for k = 1:numel(tol_ints);
-            
-            subplot(numel(tol_ints), 2, plotNum);
-            plotA = gpuData.averageRunTime(j,k,:,2,2);
-            plotB = cpuData.averageRunTime(j,k,:,1,2);
-            plotC = blasData.averageRunTime(j,k,:,3,2);
-            plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
-            hold on
-            plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
-            plot(blasNumUsers, squeeze(plotC(1:blasNUlen)), 'g+-');
-            title(['Outer tolerance: ' num2str(tols(j)) '   inner tolerance: ' num2str(tol_ints(k))]);
-            ylabel('Average runtime (s)');
-            xlabel('Number of Users');
-            hold off
-            
-            subplot(numel(tol_ints), 2, plotNum+1);
-            plotA = gpuData.averageIteration(j,k,:,2);
-            plotB = cpuData.averageIteration(j,k,:,1);
-            plotC = blasData.averageIteration(j,k,:,3);
-            plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
[…]

### 2011-01-14 paper/main.tex
Added example of adding a figure to latex document
updated section 4 (GPU implementation)

@@
 \begin{equation}
-L_u(x,y) = \|x\|_1 + <\y, \b - A\x - \e> + \frac{\mu}2 \| \b-A\x-\e \|_2^2
+L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
 \end{equation}
@@
 
+The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.  
+GPUs are comprised of several streaming multiprocessors, which contain 32 floating-point units.
+The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called warps.
+These warps are then grouped into higher levels called thread blocks and grids.
+The improvement in speed comes from the memory structure of the GPU.
+The GPU contains an inverted memory hierarchy so that the L2 cache stays close to the cores, providing the increase in performance.
+It essentially turns the GPU from being memory-limited to computational-limited.
+GPU memory also differs from CPU memory in that it requires data transfers over PCI-Express.
+The cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
+The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
+Each streaming multiprocessor is analogous to a single CPU except that it can run multiple threads simultaneously...
+
+%You should give a brief (1-2 paragraphs) overview of the GPU architecture -- most people have heard of GPUs and what they can do by now, so it isn't worth spending a whole lot of space in the paper describing it.
+
+%You should define the levels of the Cuda thread and memory hierarchies: i.e.
+warps, thread blocks, grids and register file, shared memory, global memory.
+Also mention that the GPU memories are all distinct from the CPU memories, and data transfers go over PCI-Express -- but the cost of those transfers are amortized by doing lots of work (i.e.
+solving an l1 minimization after transferring the matrix once).
+It is worth spending a little space describing how the GPU is a multi-core processor (just like the CPUs), but with much more sophisticated vector processing hardware.
+That is: on a Fermi there are 15 cores, and each one has two 16-wide vector floating-point units.
+
+%The GPU-algorithm is the same as the CPU algorithm: primary- and dual- ALM methods.
+The implementations differ in that you have to parallelize the GPU code at an additional level, and you can't use the GPU's caches to communicate between cores (only within a core).
+On the CPU, you only need to parallelize for the 4 or 8 cores -- and perhaps add SSE instructions when possible (usually this is inside of the BLAS library) -- any communication between cores happens implicitly via cache coherence.
+For the GPU, you need to parallelize for the 15 cores, but also for the 32-way SIMD parallelism of Warps.
+Any intra-core communication happens via the cache (aka scratchpad, or __shared__ memory) and within-block barriers (i.e.
+__syncthreads()), and any inter-core communication happens via the GPU's DRAM and the implicit barriers between grid launches.
+
 cuda cores, SIMD width, scheduling restrictions, bandwidth
@@
 Plot Solver Bandwidth in GB/s vs. Problem size in MB
+
+The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
+The algorithm is the same as the CPU algorithm except that we had to parallelize the GPU at an additional level.
+The CPU portion of the code acts as a conductor, making calls to GPU functions in C, while the GPU portion does the computation.
 
@@
 
-For figures ...
+%An example of image inclusion for Victor:
+\begin{figure*}
+\centering
+\includegraphics[width=3.5in]{figures/PALM_benchmark_ratio_1.png}
+\caption{This is where the caption goes.}
+\label{fig:uniqueidentifierforthisimage}
+\end{figure*}
 

### 2011-01-14 paper/main.tex
minor additions to section 5

@@
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
-The algorithm is the same as the CPU algorithm except that we had to parallelize the GPU at an additional level.
-The CPU portion of the code acts as a conductor, making calls to GPU functions in C, while the GPU portion does the computation.
+The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an additional level.
+The CPU portion of the code serves as the conductor, making calls to GPU functions and directing the flow of the code, while the GPU portion does all of the computation.
 
@@
 
-In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU Cuda implementations of the Primal ALM algorithm on synthetic data to find the point at which the GPU version runs faster than CPU versions.
+In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU Cuda implementations of the Primal ALM algorithm on synthetic data to find at which point the GPU implementation runs faster than CPU implementation.
 To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations at 5000 and run the algorithm to convergence.
@@
 Tolerance levels were empirically seen to allow best performance with the inner loop tolerance set at one order of magnitude lower than the outer loop tolerance.
+
+To generate a meaningful comparison between different hardware
+architectures, we compare on a per-board basis:  For CPU implementations, the benchmark makes use of
+as all of the cores in as many cpu's are present.  For GPU implementations, the benchmark makes the best use
+of the entire GPU chip (most GPU boards have a single GPU chip).
 
@@
 with other benchmarks on face data.
-
-To generate a meaningful comparison between different hardware
-architectures, we compare on a per-board basis:
-
-For CPU implementations, the benchmark makes use of
-as all of the cores in as many cpu's are present.
-
-For GPU implementations, the benchmark makes the best use
-of the entire GPU chip (most GPU boards have a single GPU chip).
 

### 2011-01-20 paper/main.tex
made some revisions to the tex

@@
 
-Show Pseudocode for one variant
-\subsection{Computational Complexity}
+% Write about the general algorithm and then say they have to be adapted for the two problems.
+% Refer to book chapter
+
+Shown above is the general algorithm, which has been adapted for the two problems (alignment and recognition).  Refer to book...
+
+%Show Pseudocode for one variant
+%\subsection{Computational Complexity}
 The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$), 
 in terms of both computation cost and memory-bandwidth.
-Therefore, assuming the iterative solver always takes a constant number of iterations,
-the computational complexity of the overall algorithm is linear in the problem size.
+As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n < m.
+
+% Complexity of entire algorithm
 
@@
 The improvement in speed comes from the memory structure of the GPU.
-The GPU contains an inverted memory hierarchy so that the L2 cache stays close to the cores, providing the increase in performance.
+The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance.
 It essentially turns the GPU from being memory-limited to computational-limited.
@@
 The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
-Each streaming multiprocessor is analogous to a single CPU except that it can run multiple threads simultaneously...
+Each streaming multiprocessor is analogous to a single CPU except that it can run multiple threads simultaneously.
+All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.
 
@@
 Also mention that the GPU memories are all distinct from the CPU memories, and data transfers go over PCI-Express -- but the cost of those transfers are amortized by doing lots of work (i.e.
-solving an l1 minimization after transferring the matrix once).
+solving an $\ell_1$ minimization after transferring the matrix once).
 It is worth spending a little space describing how the GPU is a multi-core processor (just like the CPUs), but with much more sophisticated vector processing hardware.
@@
 cuda cores, SIMD width, scheduling restrictions, bandwidth
-\subsection{Solving a Single L1 minimization}
-This section is about Victor's CUBLAS implementations.
-Plot Solver Bandwidth in GB/s vs. Problem size in MB
+\subsection{Solving a Single $\ell_1$ minimization}
 
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
+CUBLAS are the BLAS libraries consisting of all single precision BLAS calls which are provided by Nvidia and are optimized for GPU performance.
+In addition, we coded kernel (GPU) methods for vector-vector operations to replace operations that would have taken multiple BLAS calls.
 The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an additional level.
-The CPU portion of the code serves as the conductor, making calls to GPU functions and directing the flow of the code, while the GPU portion does all of the computation.
+The CPU portion of the code serves as the conductor, making calls to GPU functions and directing the flow of the code while the GPU portion does all of the actual computation.
 
-\subsection{Solving multiple L1 minimizations in parallel}
-This section is about Drew's kernels for solving multiple L1 minimizations in parllel
+\subsection{Solving multiple $\ell_1$ minimizations in parallel}
+This section is about Drew's kernels for solving multiple $\ell_1$ minimizations in parllel
 
-How to efficiently perform an L1 minimization on a single core, quote performance on a large batch.
+How to efficiently perform an $\ell_1$ minimization on a single core, quote performance on a large batch.
 
@@
 
-\section{Benchmarks on Synthetic Data}
-\subsection{PALM L1 minimization benchmark}
+\section{Benchmarks}
 
-In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU Cuda implementations of the Primal ALM algorithm on synthetic data to find at which point the GPU implementation runs faster than CPU implementation.
+In this section, we present __ sets of experiment to benchmark the performance of the PALM algorithm on CPU and GPU platforms.
+To generate a meaningful comparison between different hardware architectures, we compare on a per-board basis:  For CPU implementations, the benchmark makes use of as all of the cores in as many cpu's are present.
+For GPU implementations, the benchmark makes the best use of the entire GPU chip (most GPU boards have a single GPU chip).
+Both implementations utilized single precision floating point datatypes and were verified to output results differing by no more than 10^-4.
+
+To construct the benchmark for synthetic data, we first created an mxn matrix A with entries \+- 1 each with a 50\% chance.
+Then we created a sparse vector x0 with k\% nonzero random entries.
+We also created a sparse vector yk with the same number of non-zero entries as x.
+Let y0 = A*x0 + yk.
+We timed the time taken for the ALM algorithm using y0, A, and other inputs.
+The other inputs included, maximum number of inner/outer loop iterations, and inner/outer loop tolerance level.
+% give a discusison of how to construct the benchmark:
+% one paragraph for protocol on how to construct synthetic data
+% second P for protocol on how to construct the face data
+
+\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}
+
+In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA implementations of the Primal ALM algorithm on synthetic data to find at which point the GPU implementation runs faster than CPU implementation.
 To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations at 5000 and run the algorithm to convergence.
-CPU code was run on an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480
+CPU code was run on an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.
+GPU code was run on a GTX480
 We ran the A matrix with different ratios between m:n from 10:1:5 for m:2000:25:8000 with an x vector of 10 percent sparsity.
@@
 
-To generate a meaningful comparison between different hardware
-architectures, we compare on a per-board basis:  For CPU implementations, the benchmark makes use of
-as all of the cores in as many cpu's are present.  For GPU implementations, the benchmark makes the best use
-of the entire GPU chip (most GPU boards have a single GPU chip).
+Results were plotted as total size of A and A'A vs Gb/s and size of A vs time elapsed to run a single $\ell_1$ solver.
 
@@
 
+% change picture to only include CPU/GPU not MATLAB
+
 As the dimensions of A increase, so does the significance of the size of AtA.  Until the A and AtA matrix reach x Mb, the CPU version runs faster than the GPU.
-
-Optimization:
-\begin{equation}
-\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
-\end{equation}
-
-$A$ is $m \times n$
-$m = 5120$
-$n = 32$
-
-One of these minimizations is performed per 
-alignment iteration in our face recognition pipeline.
-
-All of the following are implementations of primal ALM.
-Primal ALM contains outer and inner iterations.
-To simplify the benchmark, the number of outer iterations
-is clamped to 50, and the number of inner iterations is
-clamped to 50 iterations per outer loopsretrsitrsi.  Running with
-a convergence criterion will typically result in fewer
-iterations, depending on the data, so these numbers are
-for comparison across hardware, and may look worse than
-with other benchmarks on face data.
-
-Since different platforms may perform most efficiently when
-executing multiple L1 minimizations simultaneously,
-we quote the average time required to solve a single L1
-minimization problem instance.
-
-schur:  Mac Pro, dual socket, quad core, hyperthreading intel nehalem @ 2.66 GHz
-
-Nvidia GTX480: This is the current state of the art consumer card.  
-It is clocked slightly faster than the pro version of the card, so it 
-should have slightly faster single precision float performance. (double precision is crippled, though)
-GTX480 has 15 cores, each of which has 2x32-wide floating point pipelines.
-All Fermi GPU's have 768KB L2 cache.
-
-Zihan's OpenCV implementation: *** 190 ms per L1 minimization ***
-amortized over 16 instances,
-four of which run simultaneously.  Tuning note:  the cpus
-can support up to 16 logical threads, but 4 instances seem
-sufficient to saturate the memory architecture.  Using more
-than 5 instances results in lower performance, perhaps due
-to less efficient usage of the cpu caches.
-
-Zihan's MATLAB implementation: *** 60 ms per L1 minimization ***
-This is amortized over 100 sequential instances.  All 16 logical
-cores were used, i.e. showed 100% usage in htop.  Pararllelism
-is a result of MATLAB's usage of Intel's MKL, a parallel blas
-implementation. I believe performance is very high to
-efficient cache usage:
-
-The A matrix takes up 5120*32*8 bytes = 1.25 MB
-The two processors have 16MB L3 Cache in total.
-The two processors have 2MB L2 Cache in total.
-
-Drew's pure CUDA, single kernel implementation: *** 36 ms per L1 minimization ***
-This implementation performs L1 minimization on 60 instances of L1.
-They are all run as a single cuda kernel.  
-Each instance of L1 is mapped onto a cuda block. (tuned to 64 threads / block)
-This means that the GPU hardware will schedule the multiple instances to run
-on the (15 for GTX480) cuda cores as concurrently (15? 30?) as possible.  
-For this to work, all of the data has to be ready at the same time.
-
-Drew's pure CUDA, streams implementation: *** 70 ms per L1 minimization ***
-This implementation performs L1 minimization on 60 instances of L1.
-Each instance of L1 is mapped onto a cuda block. (tuned to 128 threads / block)
-Each instance is run in a separate single-block CUDA kernel.
-The 60 kernels are added to 15 cuda streams.
-The GPU (hardware? driver?) then runs the kernels, up to 15 concurrently.
-
-Victor's CUDA/cublas mixed implementation: *** 302 ms per L1 minimization ***
-This implementation makes use of hand-written kernels for vector-vector operations,
-and cublas for matrix-vector operations.  There are two implementations under development:
-1. Perform a single L1 minimization using the entire GPU.
-2. Use cuda streams to perform multiple L1 minimizations concurrently on the GPU.
-Benchmarked on Drew's GTX480
-
-Drew's OpenMP implementation: *** 200 ms per L1 minimization***
-Benchmarked on schur(Dual Socket Quad Core Nehalem), all compiler optimizations turned on.
-This divides the matrix operations into 8 chunks, each of which is handled by a thread
-on a different CPU core.  Fine-grained parallelism is achieved through SSE2 vector
-floating point operations.  This implementation should ultimately beat the MATLAB/MKL
-implementation, especially for small problem sizes, since the primary advantage
-over MKL is a reduction in the number of times image-sized temporary variables
-are written and read from memory.
-
-Drew's Single-Threaded MKL implementation: *** 226 ms per L1 minimization ***
-Benchmarked on schur(Dual Socket Quad Core Nehalem), all compiler optimizations turned on.
-Uses MKL for single-threaded matrix-vector multiplies.  
-Saves ~almost as many writes of temporary vector variables as the previous OpenMP implementation.
-Shrinkage operator improved; at least with g++ there is now no branching or function call overhead.
-This implementation uses a single CPU core...
-... so if linear speedup can be achieved, could possibly get *** 28 ms per minimization ***
-
-%\begin{algorithm}[thb]
-%\caption{\bf (Deformable Sparse Recovery and Classification for
-%Face Recognition)} \label{alg:deformable-src}
-%\begin{algorithmic}[1]
-%\STATE {\bf Input:} Frontal training images $A_i \in \Re^{m\times n_i}, i=1,2,\ldots,K$ for $K$ subjects,  a test image
-%$\bb\in\Re^m$ and a deformation group $T$. 
-%\STATE
-%{\bf for} each subject $i$, 
-%\STATE \hspace{3mm} $\tau^{(0)}
-%\leftarrow I$. 
-%\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
-%\STATE \hspace{6mm}
-%$\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ
-%\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
-%\tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
-%%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \bb + J \Delta \tau = A_k \x + \e \end{array}\right.$
-%\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
-%\subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
-%\STATE
-%\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-%\tau$; 
-%\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
-%the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
-%residuals $\|\e\|_1$. \STATE Compute an average transformation
-%$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
-%\STATE Update $\bb \leftarrow \bb \circ \bar{\tau}$ and $\tau_i
-%\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
-%k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
-%\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
-%\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
-%$\ell^1$-minimization problem:
-%$$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \bb = A \x + \e.$$
-%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-%\STATE {\bf Output:} $\mbox{identity}(\bb) = \arg\min_i r_i(\bb)$.
-%\end{algorithmic}
-%\end{algorithm}
-
-%We summarize the entire ALM
-%algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
-%largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-%in \cite{YangJ2009-pp} and set $\mu = 2m / \|\bb\|_1$.
-%\begin{algorithm}[h]
-%\caption{\bf (Augmented Lagrange Multiplier Method for Global
-%Recognition)}
-%\begin{algorithmic}[1]
-%\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
-%$\x_1 = \mathbf{0}$, $\e_1 = \bb$, $\blamda_1 =
-%\mathbf{0}$.
-%\WHILE{not converged ($k = 1,2,\ldots$)}
-%\STATE $\e_{k+1} = \textup{shrink}\left(\bb - A\x_k +
-%\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
-%\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
-%\WHILE{not converged ($l = 1,2,\ldots$)}
-%\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
-%\frac{1}{\gamma}A^T\left(\bb - A\z_l - \e_{k+1} +
-%\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
-%\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-%\sqrt{1+4t_l^2}\right)$;
-%\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
-%\ENDWHILE
-%\STATE $\x_{k+1} \leftarrow \w_{l}$;
-%\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
-%\ENDWHILE \STATE
-%{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-%\end{algorithmic}
-%\label{alg:alm}
-%\end{algorithm}
 

### 2011-01-26 paper/main.tex
Fixed latex markup errors that were preventing build.

@@
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
-CUBLAS are the BLAS libraries consisting of all single precision BLAS calls which are provided by Nvidia and are optimized for GPU performance.
-In addition, we coded kernel (GPU) methods for vector-vector operations to replace operations that would have taken multiple BLAS calls.
-The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an additional level.
-The CPU portion of the code serves as the conductor, making calls to GPU functions and directing the flow of the code while the GPU portion does all of the actual computation.
+CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
+which are provided by Nvidia and are optimized for GPU performance. In
+addition, we coded kernel (GPU) methods for vector-vector operations to replace
+operations that would have taken multiple BLAS calls. The algorithm is similar
+to the CPU algorithm except that we had to parallelize the code at an
+additional level. The CPU portion of the code serves as the conductor, making
+calls to GPU functions and directing the flow of the code while the GPU portion
+does all of the actual computation.
 
@@
 
-In this section, we present __ sets of experiment to benchmark the performance of the PALM algorithm on CPU and GPU platforms.
-To generate a meaningful comparison between different hardware architectures, we compare on a per-board basis:  For CPU implementations, the benchmark makes use of as all of the cores in as many cpu's are present.
-For GPU implementations, the benchmark makes the best use of the entire GPU chip (most GPU boards have a single GPU chip).
-Both implementations utilized single precision floating point datatypes and were verified to output results differing by no more than 10^-4.
+In this section, we present INSERTNUMBERHERE sets of experiment to benchmark the performance
+of the PALM algorithm on CPU and GPU platforms. To generate a meaningful
+comparison between different hardware architectures, we compare on a per-board
+basis:  For CPU implementations, the benchmark makes use of as all of the cores
+in as many cpu's are present.  For GPU implementations, the benchmark makes the
+best use of the entire GPU chip (most GPU boards have a single GPU chip). Both
+implementations utilized single precision floating point datatypes and were
+verified to output results differing by no more than $10^{-4}$.
 
-To construct the benchmark for synthetic data, we first created an mxn matrix A with entries \+- 1 each with a 50\% chance.
-Then we created a sparse vector x0 with k\% nonzero random entries.
-We also created a sparse vector yk with the same number of non-zero entries as x.
-Let y0 = A*x0 + yk.
-We timed the time taken for the ALM algorithm using y0, A, and other inputs.
-The other inputs included, maximum number of inner/outer loop iterations, and inner/outer loop tolerance level.
-% give a discusison of how to construct the benchmark:
-% one paragraph for protocol on how to construct synthetic data
-% second P for protocol on how to construct the face data
+To construct the benchmark for synthetic data, we first created an $m \times n$
+matrix $A$ with entries $\pm 1$ each with a $50\%$ chance.  Then we created a
+sparse vector $\x_0$ with $k\%$ nonzero random entries.  We also created a
+sparse vector $y_k$ with the same number of non-zero entries as $\x$.  Let
+$\y_0 = A \x_0 + \y_k$.  We timed the time taken for the ALM algorithm using
+$\y_0$, $A$, and other inputs.  The other inputs included, maximum number of
+inner/outer loop iterations, and inner/outer loop tolerance level.
+% give a discusison of how to construct the benchmark: one paragraph for
+% protocol on how to construct synthetic data second P for protocol on how to
+% construct the face data
 
@@
 
-In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA implementations of the Primal ALM algorithm on synthetic data to find at which point the GPU implementation runs faster than CPU implementation.
-To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations at 5000 and run the algorithm to convergence.
-CPU code was run on an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.
-GPU code was run on a GTX480
-We ran the A matrix with different ratios between m:n from 10:1:5 for m:2000:25:8000 with an x vector of 10 percent sparsity.
-Tolerance levels were empirically seen to allow best performance with the inner loop tolerance set at one order of magnitude lower than the outer loop tolerance.
+In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA
+implementations of the Primal ALM algorithm on synthetic data to find at which
+point the GPU implementation runs faster than CPU implementation.  To simply
+the benchmark, we cap the number of inner iterations at 50 and outer iterations
+at 5000 and run the algorithm to convergence.  CPU code was run on an Intel
+Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480 We
+ran the A matrix with different ratios between m:n from 10:1:5 for
+m:2000:25:8000 with an x vector of 10 percent sparsity.  Tolerance levels were
+empirically seen to allow best performance with the inner loop tolerance set at
+one order of magnitude lower than the outer loop tolerance.
 
-Results were plotted as total size of A and A'A vs Gb/s and size of A vs time elapsed to run a single $\ell_1$ solver.
+Results were plotted as total size of A and A'A vs Gb/s and size of A vs time
+elapsed to run a single $\ell_1$ solver.
 

### 2011-02-02 paper/main.tex
Fleshed out computational complexity discussion.

@@
 it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
-that arithmetic logic.
+than arithmetic logic.
 
@@
 \section{Characteristics of the ALM Algorithm}
-Describe the two variants of the ALM algorithm
+%Describe the two variants of the ALM algorithm
 
-The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
+%The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
+%\begin{equation}
+%F(x) = f(x) + \lambda g(x).
+%\end{equation}
+
+%Because the optimal lagrange multiplier values are not known, ALM iterates between estimating the solution and the lagrange multipier values.
+
+%In order to solve the recognition stage, we consider the $l_1$ problem:
+%\begin{equation}
+%\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
+%\end{equation}
+%which can be formulated as
+%\begin{equation}
+%L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
+%\end{equation}
+
+%where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.
+
+We define the following soft-thresholding operator for a
+scalar $x$ and a scalar $\alpha \geq 0$:
 \begin{equation}
-F(x) = f(x) + \lambda g(x).
+\textup{shrink}(x,\alpha) = \textup{sign}(x)\cdot \max \{|x| - \alpha, 0\},
 \end{equation}
+In practice, this operator will be applied elementwise to a vector $\x$ with a single $\lambda$,
+and will be notated as $\textup{shrink}(\x,\alpha)$.
 
-Because the optimal lagrange multiplier values are not known, ALM iterates between estimating the solution and the lagrange multipier values.
-
-In order to solve the recognition stage, we consider the $l_1$ problem:
-\begin{equation}
-\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
-\end{equation}
-which can be formulated as
-\begin{equation}
-L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
-\end{equation}
-
-where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.
-
-We summarize the entire ALM
-algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
+We summarize the ALM algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
 largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
@@
 
-In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
-\begin{equation} 
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
-\end{equation}
-which is solved in a similar fashion.
+\subsection{Computational Complexity}
+Any discussion of the speed of an algorithm should discuss the comutational complexity of the algorithm.  For the ALM
+algorithms considered in this paper, this is rather simple.
+The Matrix-vector computions are $O(mn)$, and the vector-vector operations are $O(m)$, 
+in terms of both computation cost and memory-bandwidth.  Since A is very tall, 
+the computation of $A^\dagger$, the pseudoinverse of $A$, is best computed via $A^\dagger = (A^TA)^-1 A^T$.
+The complexity of computing $G = A^T A$ is $O(n*n*m)$.  $G$ is sufficiently small that it can be inverted via
+Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation of $A$ finishes with a matrix
+multiplication, with cost $O(n^2 m)$.
+Fortunately, $A^\dagger$ only needs to be computed once per instance of the $\ell_1$ minimization problem.
 
-We need to define the following soft-thresholding operator for a
-scalar $x$ and a scalar $\alpha \geq 0$:
-\begin{equation}
-\textup{shrink}(x,\alpha) = \textup{sign}(x)\cdot \max \{|x| - \alpha, 0\},
-\end{equation}
-In practice, this operator will be applied elementwise to a vector $\x$ with a single $\lambda$,
-and will be notated as $\textup{shrink}(\x,\alpha)$.
-
-% Write about the general algorithm and then say they have to be adapted for the two problems.
-% Refer to book chapter
-
-Shown above is the general algorithm, which has been adapted for the two problems (alignment and recognition).  Refer to book...
-
-%Show Pseudocode for one variant
-%\subsection{Computational Complexity}
-The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$), 
-in terms of both computation cost and memory-bandwidth.
-As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n < m.
-
-% Complexity of entire algorithm
+Therefore, for problems with large $n$, we expect the computation of the pseudoinverse to dominate.  However,
+for the problem sizes needed in face recognition problems, we find that the computation time is dominated
+by the repeated computation of matrix-vector and matrix-matrix multiplications in the inner loop.
 
@@
 
-\subsection{Solving multiple $\ell_1$ minimizations in parallel}
-This section is about Drew's kernels for solving multiple $\ell_1$ minimizations in parllel
+\section{Benchmarks on a Single Large $\ell_1$ Problem}
 
-How to efficiently perform an $\ell_1$ minimization on a single core, quote performance on a large batch.
+In this section, we present $2$ sets of experiment to benchmark the performance
+of the PALM algorithm on CPU and GPU platforms. 
+The first set of experiments will compare implementations that solve a single
+large instance of an $ell_1$ minimization problem.  
+The second set of experiments will compare implementations that must solve
+many instances of $\ell_1$ minimization.
 
-How to solve multiple instances in smaller batches without losing concurrency, using CUDA streams.
+To generate a meaningful comparison between different hardware architectures,
+in all cases we compare the performance of hardware architectures on a
+per-board basis:  For CPU implementations, the benchmark makes use of all of
+the cores in as many cpu's are present.  For GPU implementations, the benchmark
+makes the best use of the entire GPU chip (most GPU boards have a single GPU
+chip).
 
-...Mention effect of compiler verion on streams implementation
-
-\section{Benchmarks}
-
-In this section, we present INSERTNUMBERHERE sets of experiment to benchmark the performance
-of the PALM algorithm on CPU and GPU platforms. To generate a meaningful
-comparison between different hardware architectures, we compare on a per-board
-basis:  For CPU implementations, the benchmark makes use of as all of the cores
-in as many cpu's are present.  For GPU implementations, the benchmark makes the
-best use of the entire GPU chip (most GPU boards have a single GPU chip). Both
-implementations utilized single precision floating point datatypes and were
-verified to output results differing by no more than $10^{-4}$.
+Unless otherwise specified, all implementations utilize single precision
+floating point datatypes.
+% and were verified to output results differing by no
+%more than $10^{-4}$.
 
@@
 
+\section{Benchmarks on Solving Many $\ell_1$ Problems Concurrently}
+
+%This section is about Drew's kernels for solving multiple $\ell_1$ minimizations in parllel
+
+%How to efficiently perform an $\ell_1$ minimization on a single core, quote performance on a large batch.
+
+%How to solve multiple instances in smaller batches without losing concurrency, using CUDA streams.
+
+%...Mention effect of compiler verion on streams implementation
+
 \section{Conclusion}

### 2011-02-02 paper/main.tex
Added discussion of CPU Parallelism

@@
 
+% With a reshuffling of the order of operations and a change of the termination
+% condition, the inner loop takes the form of two matrix-vector multiplications
+% and a sequence of vector-vector operations.
+
 \subsection{Computational Complexity}
@@
 \section{CPU Implementation}
-% With a reshuffling of the order of operations and a change of the termination
-% condition, the inner loop takes the form of two matrix-vector multiplications
-% and a sequence of vector-vector operations.
 \subsection{CPU Hardware Parallelism}
-High-level discussion of CPU parallelism, i.e.
-thread-level, vector-level, cache size, cache locality
-\subsection{Tuning the GEMV operation}
-When matrix-vector operations are called sequentially on the same data,
-the operation can be blocked so that each thread operates on the same block of data for each operation.
+The baseline architecture for our experiments is 
+a typical high end workstation with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.
+Each processor has its own memory interface, and is directly connected to half of the
+RAM installed in the machine.  The amount of RAM installed exceeds the amount used
+by the algorithms, and is not an important performance consideration.
 
-For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
+Each core has a 32 KB L1 data cache and a 256 KB L2 cache. With Hyperthreading enabled, each core can
+support two threads simultaneously.  For floating point math, each core also has a vector unit capable of
+performing the same arithmetic operation on 4 single precision floating point values simultaneously.
+\footnote{The core also has scalar floating point units, but some compilers, notably gcc, generate
+code that uses the vector units exclusively for floating point math.}
 
-For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
+There are thus two important levels of parallelism that need to be exploited to use
+a modern cpu efficiently: core-level parallelism and vector-level parallelism.
+For the implementations discussed here, we will primarily use the OpenMP programming model 
+to express core-level parallelism.  Depending on the section of the code, we will achieve vector-level parallelism 
+implicitly through either the auto-vectorization features of Intel's icc compiler, 
+or through the use of a highly optimized BLAS library (Intel's MKL).
 
-For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
+Each processor has 8MB of L3 cache that is shared by all 4 cores.
+Overall, this means the algorithm has approximately 16MB of on-chip cache available.  
+For the low resolution images used in sparse representation based face recognition,
+this is enough cache to hold over 800 image-sized variables, meaning that for small data sets
+it is possible to run completely using on-chip cache.
 
-\subsection{Tuning the Soft-Thresholding Operator}
-...Make sure there are no branches
+%\subsection{Tuning the GEMV operation}
+%When matrix-vector operations are called sequentially on the same data,
+%the operation can be blocked so that each thread operates on the same block of data for each operation.
+%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
+%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
+%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
 
-...Make sure there are no function calls
-
-...Make sure the loop gets automatically vectorized
-
-...Block the computation into work for different cpu cores
+%\subsection{Tuning the Soft-Thresholding Operator}
+%...Make sure there are no branches
+%...Make sure there are no function calls
+%...Make sure the loop gets automatically vectorized
+%...Block the computation into work for different cpu cores
 

### 2011-02-02 paper/main.tex
Reflowed some sections, Re-organized a bit.

@@
 
-\subsection{Computational Complexity}
-Any discussion of the speed of an algorithm should discuss the comutational complexity of the algorithm.  For the ALM
-algorithms considered in this paper, this is rather simple.
-The Matrix-vector computions are $O(mn)$, and the vector-vector operations are $O(m)$, 
-in terms of both computation cost and memory-bandwidth.  Since A is very tall, 
-the computation of $A^\dagger$, the pseudoinverse of $A$, is best computed via $A^\dagger = (A^TA)^-1 A^T$.
-The complexity of computing $G = A^T A$ is $O(n*n*m)$.  $G$ is sufficiently small that it can be inverted via
-Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation of $A$ finishes with a matrix
-multiplication, with cost $O(n^2 m)$.
-Fortunately, $A^\dagger$ only needs to be computed once per instance of the $\ell_1$ minimization problem.
+\subsection{Computational Complexity} Any discussion of the speed of an
+algorithm should discuss the comutational complexity of the algorithm.  For the
+ALM algorithms considered in this paper, this is rather simple.  The
+Matrix-vector computions are $O(mn)$, and the vector-vector operations are
+$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
+very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
+computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
+A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
+Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
+of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
+Fortunately, $A^\dagger$ only needs to be computed once per instance of the
+$\ell_1$ minimization problem.
 
-Therefore, for problems with large $n$, we expect the computation of the pseudoinverse to dominate.  However,
-for the problem sizes needed in face recognition problems, we find that the computation time is dominated
-by the repeated computation of matrix-vector and matrix-matrix multiplications in the inner loop.
+Therefore, for problems with large $n$, we expect the computation of the
+pseudoinverse to dominate.  However, for the problem sizes needed in face
+recognition problems, we find that the computation time is dominated by the
+repeated computation of matrix-vector and matrix-matrix multiplications in the
+inner loop.
 
-\section{CPU Implementation}
+\section{Hardware Parallelism}
+
 \subsection{CPU Hardware Parallelism} 
-The baseline architecture for our experiments is 
-a typical high end workstation with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.
-Each processor has its own memory interface, and is directly connected to half of the
-RAM installed in the machine.  The amount of RAM installed exceeds the amount used
-by the algorithms, and is not an important performance consideration.
+The baseline architecture for our experiments is a typical workstation,
+with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each
+processor has its own memory interface, and is directly connected to half of
+the RAM installed in the machine.  The amount of RAM installed exceeds the
+amount used by the algorithms, and is not an important performance
+consideration.
 
-Each core has a 32 KB L1 data cache and a 256 KB L2 cache. With Hyperthreading enabled, each core can
-support two threads simultaneously.  For floating point math, each core also has a vector unit capable of
-performing the same arithmetic operation on 4 single precision floating point values simultaneously.
-\footnote{The core also has scalar floating point units, but some compilers, notably gcc, generate
-code that uses the vector units exclusively for floating point math.}
+Each core has a 32 KB L1 data cache and a 256 KB L2 cache. With Hyperthreading
+enabled, each core can support two threads simultaneously.  For floating point
+math, each core also has a vector unit capable of performing the same
+arithmetic operation on four single precision floating point values
+simultaneously.  \footnote{The core also has scalar floating point units, but
+some compilers, notably gcc, generate code that uses the vector units
+exclusively for floating point math.}
 
-There are thus two important levels of parallelism that need to be exploited to use
-a modern cpu efficiently: core-level parallelism and vector-level parallelism.
-For the implementations discussed here, we will primarily use the OpenMP programming model 
-to express core-level parallelism.  Depending on the section of the code, we will achieve vector-level parallelism 
-implicitly through either the auto-vectorization features of Intel's icc compiler, 
-or through the use of a highly optimized BLAS library (Intel's MKL).
+There are thus two important levels of parallelism that need to be exploited to
+use a modern cpu efficiently: core-level parallelism and vector-level
+parallelism.  For the implementations discussed here, we will primarily use the
+OpenMP programming model to express core-level parallelism.  Depending on the
+section of the code, we will achieve vector-level parallelism implicitly
+through either the auto-vectorization features of Intel's icc compiler, or
+through the use of a highly optimized BLAS library (Intel's MKL).
 
-Each processor has 8MB of L3 cache that is shared by all 4 cores.
-Overall, this means the algorithm has approximately 16MB of on-chip cache available.  
-For the low resolution images used in sparse representation based face recognition,
-this is enough cache to hold over 800 image-sized variables, meaning that for small data sets
-it is possible to run completely using on-chip cache.
+Each processor has 8MB of L3 cache that is shared by all four cores.  Overall,
+this means the algorithm has approximately 16MB of on-chip cache available.
+For the low resolution images used in sparse representation based face
+recognition, this is enough cache to hold over 800 image-sized variables,
+meaning that for small data sets it is possible to run completely using on-chip
+cache.
 
+\subsection{GPU Hardware Parallelism}
+
+The ALM algorithm requires many matrix-vector operations, this gives manycore
+processors such as GPUs a computational advantage for large problem sizes.
+GPUs are comprised of several streaming multiprocessors, which contain 32
+floating-point units.  The GPU executes programs as SIMT (Single Instruction
+Multiple Threads) in groups of 32 threads called warps.  These warps are then
+grouped into higher levels called thread blocks and grids.  The improvement in
+speed comes from the memory structure of the GPU.   The GPU contains an
+inverted memory hierarchy so that the L1 cache stays close to the cores,
+providing the increase in performance.  It essentially turns the GPU from being
+memory-limited to computational-limited.  GPU memory also differs from CPU
+memory in that it requires data transfers over PCI-Express.  The cost of these
+transfers are amortized by doing large amounts of computation at once,
+requiring code to be run completely on the GPU.  The GPU can be analogous to a
+multi-core CPU processor, except with more sophisticated vector processing
+hardware.  Each streaming multiprocessor is analogous to a single CPU except
+that it can run multiple threads simultaneously.  All of these differences
+allow GPUs to perform intensive calculations such as matrix-vector operations
+at much faster rates than CPUs.
+
+%You should define the levels of the Cuda thread and memory hierarchies: i.e.
+%warps, thread blocks, grids and register file, shared memory, global memory.
+%Also mention that the GPU memories are all distinct from the CPU memories, and
+%data transfers go over PCI-Express -- but the cost of those transfers are
+%amortized by doing lots of work (i.e. solving an $\ell_1$ minimization after
+%transferring the matrix once). It is worth spending a little space describing
+%how the GPU is a multi-core processor (just like the CPUs), but with much more
+%sophisticated vector processing hardware. That is: on a Fermi there are 15
+%cores, and each one has two 16-wide vector floating-point units.
+
+%The GPU-algorithm is the same as the CPU algorithm: primary- and dual- ALM
+%methods. The implementations differ in that you have to parallelize the GPU
+%code at an additional level, and you can't use the GPU's caches to communicate
+%between cores (only within a core). On the CPU, you only need to parallelize
+%for the 4 or 8 cores -- and perhaps add SSE instructions when possible
+%(usually this is inside of the BLAS library) -- any communication between
+%cores happens implicitly via cache coherence. For the GPU, you need to
+%parallelize for the 15 cores, but also for the 32-way SIMD parallelism of
+%Warps. Any intra-core communication happens via the cache (aka scratchpad, or
+%__shared__ memory) and within-block barriers (i.e. __syncthreads()), and any
+%inter-core communication happens via the GPU's DRAM and the implicit barriers
+%between grid launches.
+
+% In the followingn 2 sections, just describe implementations; single number
+% benchmarks and microbenchmarks are ok, but save big graphs for experiments
+% section.
+
+% This section is Victor's 
+\section{Single Instance Implementations}
+
+\subsection{Single Instance CPU Implementations}
+
+\subsection{Single Instance GPU Implementations}
+
+The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
+CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
+which are provided by Nvidia and are optimized for GPU performance. In
+addition, we coded kernel (GPU) methods for vector-vector operations to replace
+operations that would have taken multiple BLAS calls. The algorithm is similar
+to the CPU algorithm except that we had to parallelize the code at an
+additional level. The CPU portion of the code serves as the conductor, making
+calls to GPU functions and directing the flow of the code while the GPU portion
+does all of the actual computation.
+
+% This section is Drew's
+\section{Multiple Instance Implementations}
+
+\subsection{Multiple Instance CPU Implementations}
 %\subsection{Tuning the GEMV operation}
@@
 %...Block the computation into work for different cpu cores
+\subsection{Multiple Instance GPU Implementations}
 
-\section{GPU Implementation}
-\subsection{GPU Hardware Parallelism}
-High-level description of GPU parallelism, i.e.
+% Describingn the sequence of GPU implementations Drew has tried...
 
-The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.  
-GPUs are comprised of several streaming multiprocessors, which contain 32 floating-point units.
-The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called warps.
-These warps are then grouped into higher levels called thread blocks and grids.
-The improvement in speed comes from the memory structure of the GPU.
-The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance.
-It essentially turns the GPU from being memory-limited to computational-limited.
-GPU memory also differs from CPU memory in that it requires data transfers over PCI-Express.
-The cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
-The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
-Each streaming multiprocessor is analogous to a single CPU except that it can run multiple threads simultaneously.
-All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.
-
-%You should give a brief (1-2 paragraphs) overview of the GPU architecture -- most people have heard of GPUs and what they can do by now, so it isn't worth spending a whole lot of space in the paper describing it.
-
-%You should define the levels of the Cuda thread and memory hierarchies: i.e.
-warps, thread blocks, grids and register file, shared memory, global memory.
-Also mention that the GPU memories are all distinct from the CPU memories, and data transfers go over PCI-Express -- but the cost of those transfers are amortized by doing lots of work (i.e.
-solving an $\ell_1$ minimization after transferring the matrix once).
-It is worth spending a little space describing how the GPU is a multi-core processor (just like the CPUs), but with much more sophisticated vector processing hardware.
-That is: on a Fermi there are 15 cores, and each one has two 16-wide vector floating-point units.
-
-%The GPU-algorithm is the same as the CPU algorithm: primary- and dual- ALM methods.
-The implementations differ in that you have to parallelize the GPU code at an additional level, and you can't use the GPU's caches to communicate between cores (only within a core).
-On the CPU, you only need to parallelize for the 4 or 8 cores -- and perhaps add SSE instructions when possible (usually this is inside of the BLAS library) -- any communication between cores happens implicitly via cache coherence.
-For the GPU, you need to parallelize for the 15 cores, but also for the 32-way SIMD parallelism of Warps.
-Any intra-core communication happens via the cache (aka scratchpad, or __shared__ memory) and within-block barriers (i.e.
-__syncthreads()), and any inter-core communication happens via the GPU's DRAM and the implicit barriers between grid launches.
-
-cuda cores, SIMD width, scheduling restrictions, bandwidth
-\subsection{Solving a Single $\ell_1$ minimization}
-
-The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
-CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
-which are provided by Nvidia and are optimized for GPU performance. In
-addition, we coded kernel (GPU) methods for vector-vector operations to replace
-operations that would have taken multiple BLAS calls. The algorithm is similar
-to the CPU algorithm except that we had to parallelize the code at an
-additional level. The CPU portion of the code serves as the conductor, making
-calls to GPU functions and directing the flow of the code while the GPU portion
-does all of the actual computation.
-
-\section{Benchmarks on a Single Large $\ell_1$ Problem}
-
-In this section, we present $2$ sets of experiment to benchmark the performance
+\section{Benchmarks}
+In this section, we present two sets of experiment to benchmark the performance
 of the PALM algorithm on CPU and GPU platforms. 
@@
 makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).
+chip).  Unless otherwise specified, all implementations utilize single
+precision floating point datatypes.
 
-Unless otherwise specified, all implementations utilize single precision
-floating point datatypes.
-% and were verified to output results differing by no
-%more than $10^{-4}$.
+\subsection{Single Large $\ell_1$ Problem, Synthetic Data}
+In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA
+implementations of the Primal ALM algorithm on synthetic data to find at which
+point the GPU implementation runs faster than CPU implementation.
 
@@
 
-\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}
-
-In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA
-implementations of the Primal ALM algorithm on synthetic data to find at which
-point the GPU implementation runs faster than CPU implementation.  To simply
-the benchmark, we cap the number of inner iterations at 50 and outer iterations
-at 5000 and run the algorithm to convergence.  CPU code was run on an Intel
-Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480 We
-ran the A matrix with different ratios between m:n from 10:1:5 for
+To simply the benchmark, we cap the number of inner iterations at 50 and outer
+iterations at 5000 and run the algorithm to convergence.  CPU code was run on
+an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a
+GTX480 We ran the A matrix with different ratios between m:n from 10:1:5 for
 m:2000:25:8000 with an x vector of 10 percent sparsity.  Tolerance levels were
@@
 
-%An example of image inclusion for Victor:
 \begin{figure*}
@@
 
-% change picture to only include CPU/GPU not MATLAB
+As the dimensions of $A$ increase, so does the significance of the size of
+$A^TA$.  Until the $A$ and $A^TA$ reach ???? Mb (SHOULD BE MB?), the CPU version runs faster
+than the GPU.
 
-As the dimensions of A increase, so does the significance of the size of AtA.  Until the A and AtA matrix reach x Mb, the CPU version runs faster than the GPU.
+\subsection{Multiple Small $\ell_1$ Problems, Synthetic Data}
 
-\section{Benchmarks on Solving Many $\ell_1$ Problems Concurrently}
-
-%This section is about Drew's kernels for solving multiple $\ell_1$ minimizations in parllel
-
-%How to efficiently perform an $\ell_1$ minimization on a single core, quote performance on a large batch.
-
-%How to solve multiple instances in smaller batches without losing concurrency, using CUDA streams.
-
-%...Mention effect of compiler verion on streams implementation
+\subsection{Alignment Stage, Real Face Data}
 
 \section{Conclusion}
-...How lessons from this paper extend (or don't extend) to other similar problems in computer vision
-
-...Reiterate how the appropriate hardware choice depends on the problem size.
 

### 2011-02-02 paper/main.tex
minor edits to section 2

--- 2011-02-02 paper/main.tex
@@
 it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
-than arithmetic logic.
+that arithmetic logic.
 
@@
 
-%The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
-%\begin{equation}
-%F(x) = f(x) + \lambda g(x).
-%\end{equation}
+In this section, we describe the general version ALM algorithm and then describe how it can be adapted to solve our problem.
 
-%Because the optimal lagrange multiplier values are not known, ALM iterates between estimating the solution and the lagrange multipier values.
+The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
+\begin{equation}
+F(x) = f(x) + \lambda g(x).
+\end{equation}
+which can be formulated as an augmented Langrange function given by
+\begin{equation}
+L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
+\end{equation}
 
-%In order to solve the recognition stage, we consider the $l_1$ problem:
-%\begin{equation}
-%\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
-%\end{equation}
-%which can be formulated as
-%\begin{equation}
-%L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
-%\end{equation}
+where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.
 
-%where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.
+For an optimal $y*$ which satisfies the second-order sufficiency conditions, for a sufficiently large $mu$, it can be shown that
+\begin{equation}
+x* = \arg min L_u(x,y*)
+\end{equation}
 
-We define the following soft-thresholding operator for a
+However, because $y*$ is generally unknown, ALM methods simultaneously estimate the optimal solution and Lagrange multipliers iteratively in the following procedure.
+
+We also need to define the following soft-thresholding operator for a
 scalar $x$ and a scalar $\alpha \geq 0$:
@@
 
-We summarize the ALM algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
+We summarize the entire ALM
+algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
 largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
@@
 
-\subsection{Computational Complexity}
-Any discussion of the speed of an algorithm should discuss the comutational complexity of the algorithm.  For the ALM
-algorithms considered in this paper, this is rather simple.
-The Matrix-vector computions are $O(mn)$, and the vector-vector operations are $O(m)$, 
-in terms of both computation cost and memory-bandwidth.  Since A is very tall, 
-the computation of $A^\dagger$, the pseudoinverse of $A$, is best computed via $A^\dagger = (A^TA)^-1 A^T$.
-The complexity of computing $G = A^T A$ is $O(n*n*m)$.  $G$ is sufficiently small that it can be inverted via
-Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation of $A$ finishes with a matrix
-multiplication, with cost $O(n^2 m)$.
-Fortunately, $A^\dagger$ only needs to be computed once per instance of the $\ell_1$ minimization problem.
+%Show Pseudocode for one variant
+%\subsection{Computational Complexity}
+The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$), 
+in terms of both computation cost and memory-bandwidth.
+As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n $<$ m.
 
-Therefore, for problems with large $n$, we expect the computation of the pseudoinverse to dominate.  However,
-for the problem sizes needed in face recognition problems, we find that the computation time is dominated
-by the repeated computation of matrix-vector and matrix-matrix multiplications in the inner loop.
+% Complexity of entire algorithm
+
+\subsection{Adapations}
+
+In this section, we describe how to adapt the general ALM algorithm for the alignment and recognition problem.  Refer to book...
+
+In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
+\begin{equation} 
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
+\end{equation}
+which is solved in a similar fashion.
 
@@
 The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.  
-GPUs are comprised of several streaming multiprocessors, which contain 32 floating-point units.
+GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi), which each contain 32 floating-point units.
+Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.
 The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called warps.
 These warps are then grouped into higher levels called thread blocks and grids.
-The improvement in speed comes from the memory structure of the GPU.
-The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance.
+%The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance.
 It essentially turns the GPU from being memory-limited to computational-limited.
-GPU memory also differs from CPU memory in that it requires data transfers over PCI-Express.
+GPU memory also differs from CPU memory in that it requires data transfers over PCI-Express, which have limited bandwidth.
 The cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
 The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
-Each streaming multiprocessor is analogous to a single CPU except that it can run multiple threads simultaneously.
+Each thread block is a virtualized multiprocessor, which each thread as a virtualized scalar processor.
 All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.
+
+%Each SMP is analogous to a single CPU except that it can run multiple threads simultaneously.
 
@@
 
-cuda cores, SIMD width, scheduling restrictions, bandwidth
+%cuda cores, SIMD width, scheduling restrictions, bandwidth
 \subsection{Solving a Single $\ell_1$ minimization}
@@
 addition, we coded kernel (GPU) methods for vector-vector operations to replace
-operations that would have taken multiple BLAS calls. The algorithm is similar
+operations that would have taken multiple BLAS calls.
+Because data transfer from the RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.
+The algorithm is similar
 to the CPU algorithm except that we had to parallelize the code at an
 additional level. The CPU portion of the code serves as the conductor, making
-calls to GPU functions and directing the flow of the code while the GPU portion
-does all of the actual computation.
+calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.
 
-\section{Benchmarks on a Single Large $\ell_1$ Problem}
+\subsection{Solving multiple $\ell_1$ minimizations in parallel}
+This section is about Drew's kernels for solving multiple $\ell_1$ minimizations in parllel
 
-In this section, we present $2$ sets of experiment to benchmark the performance
-of the PALM algorithm on CPU and GPU platforms. 
-The first set of experiments will compare implementations that solve a single
-large instance of an $ell_1$ minimization problem.  
-The second set of experiments will compare implementations that must solve
-many instances of $\ell_1$ minimization.
+How to efficiently perform an $\ell_1$ minimization on a single core, quote performance on a large batch.
 
-To generate a meaningful comparison between different hardware architectures,
-in all cases we compare the performance of hardware architectures on a
-per-board basis:  For CPU implementations, the benchmark makes use of all of
-the cores in as many cpu's are present.  For GPU implementations, the benchmark
-makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).
+How to solve multiple instances in smaller batches without losing concurrency, using CUDA streams.
 
-Unless otherwise specified, all implementations utilize single precision
-floating point datatypes.
-% and were verified to output results differing by no
-%more than $10^{-4}$.
+...Mention effect of compiler verion on streams implementation
+
+\section{Benchmarks}
+
+In this section, we present INSERTNUMBERHERE sets of experiment to benchmark the performance
+of the PALM algorithm on CPU and GPU platforms. To generate a meaningful
+comparison between different hardware architectures, we compare on a per-board
+basis:  For CPU implementations, the benchmark makes use of as all of the cores
+in as many cpu's are present.  For GPU implementations, the benchmark makes the
+best use of the entire GPU chip (most GPU boards have a single GPU chip). Both
+implementations utilized single precision floating point datatypes and were
+verified to output results differing by no more than $10^{-4}$.
 
@@
 matrix $A$ with entries $\pm 1$ each with a $50\%$ chance.  Then we created a
-sparse vector $\x_0$ with $k\%$ nonzero random entries.  We also created a
+sparse vector $\x_0$ with $10\%$ nonzero random entries.  We also created a
 sparse vector $y_k$ with the same number of non-zero entries as $\x$.  Let
@@
 
-In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA
+In this section, we benchmark the performance of BLAS MKL and GPU CUDA
 implementations of the Primal ALM algorithm on synthetic data to find at which
@@
 
-\section{Benchmarks on Solving Many $\ell_1$ Problems Concurrently}
-
-%This section is about Drew's kernels for solving multiple $\ell_1$ minimizations in parllel
-
-%How to efficiently perform an $\ell_1$ minimization on a single core, quote performance on a large batch.
-
-%How to solve multiple instances in smaller batches without losing concurrency, using CUDA streams.
-
-%...Mention effect of compiler verion on streams implementation
-
 \section{Conclusion}

### 2011-02-02 paper/main.tex
fixed merger (i think...)

--- 2011-02-02 paper/main.tex
@@
 
-We summarize the ALM algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
-largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
+We summarize the entire ALM
+algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
 in \cite{YangJ2009-pp} and set $\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
@@
 
+%Show Pseudocode for one variant
+%\subsection{Computational Complexity}
+The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$), 
+in terms of both computation cost and memory-bandwidth.
+As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n $<$ m.
+
+% Complexity of entire algorithm
+
+\subsection{Adapations}
+
+In this section, we describe how to adapt the general ALM algorithm for the alignment and recognition problem.  Refer to book...
+
+In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
+\begin{equation} 
+\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
+\end{equation}
+which is solved in a similar fashion.
+
+\section{CPU Implementation}
 % With a reshuffling of the order of operations and a change of the termination
@@
 \subsection{GPU Hardware Parallelism}
+High-level description of GPU parallelism, i.e.
+
+The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.  
+GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi), which each contain 32 floating-point units.
+Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.
+The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called warps.
+These warps are then grouped into higher levels called thread blocks and grids.
+%The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance.
+It essentially turns the GPU from being memory-limited to computational-limited.
+GPU memory also differs from CPU memory in that it requires data transfers over PCI-Express, which have limited bandwidth.
+The cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
+The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
+Each thread block is a virtualized multiprocessor, which each thread as a virtualized scalar processor.
+All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.
+
+%Each SMP is analogous to a single CPU except that it can run multiple threads simultaneously.
+
+%You should give a brief (1-2 paragraphs) overview of the GPU architecture -- most people have heard of GPUs and what they can do by now, so it isn't worth spending a whole lot of space in the paper describing it.
+
+%You should define the levels of the Cuda thread and memory hierarchies: i.e.
+warps, thread blocks, grids and register file, shared memory, global memory.
+Also mention that the GPU memories are all distinct from the CPU memories, and data transfers go over PCI-Express -- but the cost of those transfers are amortized by doing lots of work (i.e.
+solving an $\ell_1$ minimization after transferring the matrix once).
+It is worth spending a little space describing how the GPU is a multi-core processor (just like the CPUs), but with much more sophisticated vector processing hardware.
+That is: on a Fermi there are 15 cores, and each one has two 16-wide vector floating-point units.
+
+%The GPU-algorithm is the same as the CPU algorithm: primary- and dual- ALM methods.
+The implementations differ in that you have to parallelize the GPU code at an additional level, and you can't use the GPU's caches to communicate between cores (only within a core).
+On the CPU, you only need to parallelize for the 4 or 8 cores -- and perhaps add SSE instructions when possible (usually this is inside of the BLAS library) -- any communication between cores happens implicitly via cache coherence.
+For the GPU, you need to parallelize for the 15 cores, but also for the 32-way SIMD parallelism of Warps.
+Any intra-core communication happens via the cache (aka scratchpad, or __shared__ memory) and within-block barriers (i.e.
+__syncthreads()), and any inter-core communication happens via the GPU's DRAM and the implicit barriers between grid launches.
+
+%cuda cores, SIMD width, scheduling restrictions, bandwidth
+\subsection{Solving a Single $\ell_1$ minimization}
 
@@
 \subsection{Single Instance GPU Implementations}
-
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
@@
 addition, we coded kernel (GPU) methods for vector-vector operations to replace
-operations that would have taken multiple BLAS calls. The algorithm is similar
+operations that would have taken multiple BLAS calls.
+Because data transfer from the RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.
+The algorithm is similar
 to the CPU algorithm except that we had to parallelize the code at an
 additional level. The CPU portion of the code serves as the conductor, making
-calls to GPU functions and directing the flow of the code while the GPU portion
-does all of the actual computation.
+calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.
 
@@
 
-To simply the benchmark, we cap the number of inner iterations at 50 and outer
-iterations at 5000 and run the algorithm to convergence.  CPU code was run on
-an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a
-GTX480 We ran the A matrix with different ratios between m:n from 10:1:5 for
-m:2000:25:8000 with an x vector of 10 percent sparsity.  Tolerance levels were
+\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}
+
+In this section, we benchmark the performance of BLAS MKL and GPU CUDA
+implementations of the Primal ALM algorithm on synthetic data to find at which
+point the GPU implementation runs faster than CPU implementation.  To simply
+the benchmark, we cap the number of inner iterations at 50 and outer iterations
+at 5000 and run the algorithm to convergence.  CPU code was run on an Intel
+Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480 We
+ran the A matrix with different ratios between m:n from 10:1:5 for m:2000:25:8000 with an x vector of 10 percent sparsity.  Tolerance levels were
 empirically seen to allow best performance with the inner loop tolerance set at

### 2011-02-09 paper/main.tex
removed a repeated section (GPU implementation) and revised it.

@@
 
-\section{Characteristics of the ALM Algorithm}
+\section{ALM Algorithm}
 %Describe the two variants of the ALM algorithm
@@
 \begin{algorithm}[h]
-\caption{\bf (Augmented Lagrange Multiplier Method Used in Alignment Inner Loop)}
+\caption{\bf (Augmented Lagrange Multiplier Method FIX)}
 \begin{algorithmic}[1]
@@
 \WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
 \WHILE{not converged ($l = 1,2,\ldots$)}
@@
 %\subsection{Computational Complexity}
-The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$), 
-in terms of both computation cost and memory-bandwidth.
+%The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$),  in terms of both computation cost and memory-bandwidth.
 As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n $<$ m.
@@
 
+% put this computational complexity in section 2?
 \subsection{Computational Complexity} Any discussion of the speed of an
@@
 \subsection{GPU Hardware Parallelism}
-High-level description of GPU parallelism, i.e.
+%High-level description of GPU parallelism, i.e.
 
 The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi), which each contain 32 floating-point units.
+GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi) which each contain 32 floating-point units.
 Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.
@@
 These warps are then grouped into higher levels called thread blocks and grids.
-%The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance.
-It essentially turns the GPU from being memory-limited to computational-limited.
-GPU memory also differs from CPU memory in that it requires data transfers over PCI-Express, which have limited bandwidth.
-The cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
+The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance and a computational advantage over CPUs.
+Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and 768kb of L2 cache.
+This increase in performance doesn't come for free.
+Because the GPU has it's own memory system, any data the GPU uses must be transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
+However, the cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
+
 The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
@@
 
-%Each SMP is analogous to a single CPU except that it can run multiple threads simultaneously.
+%\subsection{Solving a Single $\ell_1$ minimization}
 
-%You should give a brief (1-2 paragraphs) overview of the GPU architecture -- most people have heard of GPUs and what they can do by now, so it isn't worth spending a whole lot of space in the paper describing it.
-
-%You should define the levels of the Cuda thread and memory hierarchies: i.e.
-warps, thread blocks, grids and register file, shared memory, global memory.
-Also mention that the GPU memories are all distinct from the CPU memories, and data transfers go over PCI-Express -- but the cost of those transfers are amortized by doing lots of work (i.e.
-solving an $\ell_1$ minimization after transferring the matrix once).
-It is worth spending a little space describing how the GPU is a multi-core processor (just like the CPUs), but with much more sophisticated vector processing hardware.
-That is: on a Fermi there are 15 cores, and each one has two 16-wide vector floating-point units.
-
-%The GPU-algorithm is the same as the CPU algorithm: primary- and dual- ALM methods.
-The implementations differ in that you have to parallelize the GPU code at an additional level, and you can't use the GPU's caches to communicate between cores (only within a core).
-On the CPU, you only need to parallelize for the 4 or 8 cores -- and perhaps add SSE instructions when possible (usually this is inside of the BLAS library) -- any communication between cores happens implicitly via cache coherence.
-For the GPU, you need to parallelize for the 15 cores, but also for the 32-way SIMD parallelism of Warps.
-Any intra-core communication happens via the cache (aka scratchpad, or __shared__ memory) and within-block barriers (i.e.
-__syncthreads()), and any inter-core communication happens via the GPU's DRAM and the implicit barriers between grid launches.
-
-%cuda cores, SIMD width, scheduling restrictions, bandwidth
-\subsection{Solving a Single $\ell_1$ minimization}
-
-The ALM algorithm requires many matrix-vector operations, this gives manycore
-processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors, which contain 32
-floating-point units.  The GPU executes programs as SIMT (Single Instruction
-Multiple Threads) in groups of 32 threads called warps.  These warps are then
-grouped into higher levels called thread blocks and grids.  The improvement in
-speed comes from the memory structure of the GPU.   The GPU contains an
-inverted memory hierarchy so that the L1 cache stays close to the cores,
-providing the increase in performance.  It essentially turns the GPU from being
-memory-limited to computational-limited.  GPU memory also differs from CPU
-memory in that it requires data transfers over PCI-Express.  The cost of these
-transfers are amortized by doing large amounts of computation at once,
-requiring code to be run completely on the GPU.  The GPU can be analogous to a
-multi-core CPU processor, except with more sophisticated vector processing
-hardware.  Each streaming multiprocessor is analogous to a single CPU except
-that it can run multiple threads simultaneously.  All of these differences
-allow GPUs to perform intensive calculations such as matrix-vector operations
-at much faster rates than CPUs.
-
-%You should define the levels of the Cuda thread and memory hierarchies: i.e.
-%warps, thread blocks, grids and register file, shared memory, global memory.
-%Also mention that the GPU memories are all distinct from the CPU memories, and
-%data transfers go over PCI-Express -- but the cost of those transfers are
-%amortized by doing lots of work (i.e. solving an $\ell_1$ minimization after
-%transferring the matrix once). It is worth spending a little space describing
-%how the GPU is a multi-core processor (just like the CPUs), but with much more
-%sophisticated vector processing hardware. That is: on a Fermi there are 15
-%cores, and each one has two 16-wide vector floating-point units.
-
-%The GPU-algorithm is the same as the CPU algorithm: primary- and dual- ALM
-%methods. The implementations differ in that you have to parallelize the GPU
-%code at an additional level, and you can't use the GPU's caches to communicate
-%between cores (only within a core). On the CPU, you only need to parallelize
-%for the 4 or 8 cores -- and perhaps add SSE instructions when possible
-%(usually this is inside of the BLAS library) -- any communication between
-%cores happens implicitly via cache coherence. For the GPU, you need to
-%parallelize for the 15 cores, but also for the 32-way SIMD parallelism of
-%Warps. Any intra-core communication happens via the cache (aka scratchpad, or
-%__shared__ memory) and within-block barriers (i.e. __syncthreads()), and any
-%inter-core communication happens via the GPU's DRAM and the implicit barriers
-%between grid launches.
+%The ALM algorithm requires many matrix-vector operations, this gives manycore
+%processors such as GPUs a computational advantage for large problem sizes.
+%GPUs are comprised of several streaming multiprocessors, which contain 32
+%floating-point units.  The GPU executes programs as SIMT (Single Instruction
+%Multiple Threads) in groups of 32 threads called warps.  These warps are then
+%grouped into higher levels called thread blocks and grids.  The improvement in
+%speed comes from the memory structure of the GPU.   The GPU contains an
+%inverted memory hierarchy so that the L1 cache stays close to the cores,
+%providing the increase in performance.  It essentially turns the GPU from being
+%memory-limited to computational-limited.  GPU memory also differs from CPU
+%memory in that it requires data transfers over PCI-Express.  The cost of these
+%transfers are amortized by doing large amounts of computation at once,
+%requiring code to be run completely on the GPU.  The GPU can be analogous to a
+%multi-core CPU processor, except with more sophisticated vector processing
+%hardware.  Each streaming multiprocessor is analogous to a single CPU except
+%that it can run multiple threads simultaneously.  All of these differences
+%allow GPUs to perform intensive calculations such as matrix-vector operations
+%at much faster rates than CPUs.
 
@@
 The first set of experiments will compare implementations that solve a single
-large instance of an $ell_1$ minimization problem.  
+large instance of an $ell_1$ minimization problem on synthetic data to find at which
+point the GPU implementation runs faster than CPU implementation.  
 The second set of experiments will compare implementations that must solve
@@
 \subsection{Single Large $\ell_1$ Problem, Synthetic Data}
-In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU CUDA
-implementations of the Primal ALM algorithm on synthetic data to find at which
-point the GPU implementation runs faster than CPU implementation.
+%\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}
 
@@
 inner/outer loop iterations, and inner/outer loop tolerance level.
-% give a discusison of how to construct the benchmark: one paragraph for
-% protocol on how to construct synthetic data second P for protocol on how to
-% construct the face data
 
-\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}
-
-In this section, we benchmark the performance of BLAS MKL and GPU CUDA
-implementations of the Primal ALM algorithm on synthetic data to find at which
-point the GPU implementation runs faster than CPU implementation.  To simply
-the benchmark, we cap the number of inner iterations at 50 and outer iterations
+To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations
 at 5000 and run the algorithm to convergence.  CPU code was run on an Intel

### 2011-02-09 paper/main.tex
Added automated conversion from .dia to proper .pdf in Makefile
Added alignment runtime plot
testing merge.  started on lit review

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=3.5in]{figures/pipeline.pdf}
+\caption{The Full Recognition Pipeline}
+\label{fig:pipeline}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.5in]{figures/alignment_l1.pdf}
+\caption{Alignment Stage L1 Solver}
+\label{fig:alignment_l1}
+\end{figure}
+
 \subsection{Literature review}
@@
 
+Past work in parallelizing $\ell_1$-minimization has been limited.
+Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and Borghi at UCLA developed a proximal based algorithm using Moreau-Yosida regularization.
+
+In a review paper by Allen Yang, they list 5 fast methods that solve $\ell_1$-minimization including gradient projection, homotopy, iterative shrinkage-thresholding, proximal gradient, and augmented lagrangian methods.
+According to the review paper, they have determined that Homotopy and the Augmented Lagrangian Multiplier methods achieve the best recognition rates while maintaining a low computational cost.
+A closer look at these two algorithms will show that
+
 \subsection{Contributions of the paper}
 
-\section{ALM Algorithm}
+\section{Augmented Lagrangian Multiplier Algorithm}
 %Describe the two variants of the ALM algorithm

### 2011-02-09 paper/main.tex
testing merge
revisions to lit review

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
+\caption{GPU Alignment Stage Time vs. Size of Database}
+\label{fig:alignment_runtime}
+\end{figure}
+
 \subsection{Literature review}
@@
 In a review paper by Allen Yang, they list 5 fast methods that solve $\ell_1$-minimization including gradient projection, homotopy, iterative shrinkage-thresholding, proximal gradient, and augmented lagrangian methods.
-According to the review paper, they have determined that Homotopy and the Augmented Lagrangian Multiplier methods achieve the best recognition rates while maintaining a low computational cost.
-A closer look at these two algorithms will show that
+According to the review paper, they have determined that Homotopy and the Augmented Lagrangian Multiplier methods (ALM) achieve the best recognition rates while maintaining a low computational cost.
+A closer look at these two algorithms will show that while the Homotopy method will produce the exact solution, the algorithm structure and functions do not lend themselves parallelization well.
+On the other hand, the ALM algorithm computation mainly consists of matrix-vector operations, which lends itself to parallelization.
 

### 2011-02-09 paper/main.tex
Moved figure down to where it goes.
Restructured paper

@@
 
-\begin{figure}
-\centering
-\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
-\caption{GPU Alignment Stage Time vs. Size of Database}
-\label{fig:alignment_runtime}
-\end{figure}
-
 \subsection{Literature review}
@@
 \section{Augmented Lagrangian Multiplier Algorithm}
-%Describe the two variants of the ALM algorithm
 
@@
 
-%Show Pseudocode for one variant
-%\subsection{Computational Complexity}
-%The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$),  in terms of both computation cost and memory-bandwidth.
-As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n $<$ m.
+%In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
+%\begin{equation} 
+%\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
+%\end{equation}
+%which is solved in a similar fashion.
 
-% Complexity of entire algorithm
-
-\subsection{Adapations}
-
-In this section, we describe how to adapt the general ALM algorithm for the alignment and recognition problem.  Refer to book...
-
-In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
-\begin{equation} 
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
-\end{equation}
-which is solved in a similar fashion.
-
-\section{CPU Implementation}
+% A note on alignment l1...
 % With a reshuffling of the order of operations and a change of the termination
@@
 
-% put this computational complexity in section 2?
-\subsection{Computational Complexity} Any discussion of the speed of an
+\subsection{Alignment Stage L1 Computational Complexity} Any discussion of the speed of an
 algorithm should discuss the comutational complexity of the algorithm.  For the
@@
 \section{Hardware Parallelism} 
+In this section we give a brief overview of the hardware architectures used
+in this paper.
+
+We feel it is most meaningful to compare hardware architectures on a per-board basis,
+and in a typical hardware configuration.  GPU's are most commonly shipped with a single
+GPU chip per PCI-express card.  
+%For multicore CPU architectures, a typical workstation
+%will consist of
+
+For CPU implementations, the benchmarks makes use of all of
+the cores in as many cpu's are present.  For GPU implementations, the benchmark
+makes the best use of the entire GPU chip (most GPU boards have a single GPU
+chip).  Unless otherwise specified, all implementations utilize single
+precision floating point datatypes.
 
@@
 
-%\subsection{Solving a Single $\ell_1$ minimization}
+% This section is Drew's
+\section{Iterative Alignment Stage Implementation}
+
+\subsection{Iterative Alignment on the CPU}
+%\subsection{Tuning the GEMV operation}
+%When matrix-vector operations are called sequentially on the same data,
+%the operation can be blocked so that each thread operates on the same block of data for each operation.
+%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
+%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
+%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
+
+%\subsection{Tuning the Soft-Thresholding Operator}
+%...Make sure there are no branches
+%...Make sure there are no function calls
+%...Make sure the loop gets automatically vectorized
+%...Block the computation into work for different cpu cores
+\subsection{Iterative Alignment on the GPU}
+% Describingn the sequence of GPU implementations Drew has tried...
+
+\subsection{Benchmark on Real Face Data}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
+\caption{GPU Alignment Stage Time vs. Size of Database}
+\label{fig:alignment_runtime}
+\end{figure}
+
+\section{Recognition Stage Implementation}
+% This section is Victor's
 
@@
 
-% This section is Victor's 
-\section{Single Instance Implementations}
-
-\subsection{Single Instance CPU Implementations}
-
-\subsection{Single Instance GPU Implementations}
+\subsection{ GPU Implementation}
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
@@
 
-% This section is Drew's
-\section{Multiple Instance Implementations}
-
-\subsection{Multiple Instance CPU Implementations}
-%\subsection{Tuning the GEMV operation}
-%When matrix-vector operations are called sequentially on the same data,
-%the operation can be blocked so that each thread operates on the same block of data for each operation.
-%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
-%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
-%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
-
-%\subsection{Tuning the Soft-Thresholding Operator}
-%...Make sure there are no branches
-%...Make sure there are no function calls
-%...Make sure the loop gets automatically vectorized
-%...Block the computation into work for different cpu cores
-\subsection{Multiple Instance GPU Implementations}
-
-% Describingn the sequence of GPU implementations Drew has tried...
-
-\section{Benchmarks}
-In this section, we present two sets of experiment to benchmark the performance
-of the PALM algorithm on CPU and GPU platforms. 
-The first set of experiments will compare implementations that solve a single
-large instance of an $ell_1$ minimization problem on synthetic data to find at which
-point the GPU implementation runs faster than CPU implementation.  
-The second set of experiments will compare implementations that must solve
-many instances of $\ell_1$ minimization.
-
-To generate a meaningful comparison between different hardware architectures,
-in all cases we compare the performance of hardware architectures on a
-per-board basis:  For CPU implementations, the benchmark makes use of all of
-the cores in as many cpu's are present.  For GPU implementations, the benchmark
-makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Unless otherwise specified, all implementations utilize single
-precision floating point datatypes.
-
-\subsection{Single Large $\ell_1$ Problem, Synthetic Data}
-%\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}
+\subsection{Benchmark on Synthetic Data}
 
@@
 
-\subsection{Multiple Small $\ell_1$ Problems, Synthetic Data}
-
-\subsection{Alignment Stage, Real Face Data}
+\section{Full Pipeline Benchmarks}
 

### 2011-02-10 paper/main.tex
lit review

--- 2011-02-09 paper/main.tex
@@
 \subsection{Literature review}
-[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
+%[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
+%[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.
+%[Zero in and justify why we choose ALM for our implementation.]
 
-[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.]
-[Allen]
+Standard solvers for $\ell_1$-minimization include orthogonal matching pursuit, basis pursuit and LASSO which cast the $\ell_1$-minimization as either a linear program or quadratic program.
 
-[Zero in and justify why we choose ALM for our implementation.]
+These methods can be accelerated by more contemporary methods such as gradient projection, homotopy, iterative shrinkage-thresholding, proximal gradient, and augmented lagrangian methods.
+According to analysis of these five algorithms applied to face recognition, Yang et al have determined that Homotopy and the Augmented Lagrangian Multiplier methods (ALM) achieve the best recognition rates while maintaining a low computational cost.
+A closer look at these two algorithms will show that while the Homotopy method will produce the exact solution, the algorithm structure and functions do not lend themselves parallelization well.
+On the other hand, while the ALM finds an approximate solution, its computation mainly consists of matrix-vector operations, which lends itself to parallelization.
 
 Past work in parallelizing $\ell_1$-minimization has been limited.
-Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and Borghi at UCLA developed a proximal based algorithm using Moreau-Yosida regularization.
-
-In a review paper by Allen Yang, they list 5 fast methods that solve $\ell_1$-minimization including gradient projection, homotopy, iterative shrinkage-thresholding, proximal gradient, and augmented lagrangian methods.
-According to the review paper, they have determined that Homotopy and the Augmented Lagrangian Multiplier methods (ALM) achieve the best recognition rates while maintaining a low computational cost.
-A closer look at these two algorithms will show that while the Homotopy method will produce the exact solution, the algorithm structure and functions do not lend themselves parallelization well.
-On the other hand, the ALM algorithm computation mainly consists of matrix-vector operations, which lends itself to parallelization.
+Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and Borghi at UCLA developed a proximal point algorithm using Moreau-Yosida regularization, which is special form of the $\ell_1$-minimization based on proximal gradient.
 

### 2011-02-11 paper/main.tex
minor change

@@
 Past work in parallelizing $\ell_1$-minimization has been limited.
-Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and Borghi at UCLA developed a proximal point algorithm using Moreau-Yosida regularization, which is special form of the $\ell_1$-minimization based on proximal gradient.
+In 2009, Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and in 2009 Borghi at UCLA developed a proximal point algorithm using Moreau-Yosida regularization, which is special form of the $\ell_1$-minimization based on proximal gradient.
 
@@
 \end{equation}
-which can be formulated as an augmented Langrange function given by
+which can be formulated as an augmented Lagrange function given by
 \begin{equation}
@@
 % and a sequence of vector-vector operations.
+
+In order to adapt the general algorithm to face alignment and recognition, we formulated the 
+$\ell_1$-minimization problem as ___ as described in [book].
 

### 2011-02-14 paper/main.tex
Added proof of concept for getting latex into diagrams.

@@
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+
+% Needed for getting formulas in dia diagrams.
+\usepackage{pstricks}
+\usepackage{tikz}
 
@@
 \centering
-\includegraphics[width=3.5in]{figures/pipeline.pdf}
+%\includegraphics[width=3.5in]{figures/pipeline.pdf}
+\input{figures/pipeline.tex}
 \caption{The Full Recognition Pipeline}
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=3.5in]{figures/alignment_l1.pdf}
-\caption{Alignment Stage L1 Solver}
-\label{fig:alignment_l1}
-\end{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=3.5in]{figures/alignment_l1.pdf}
+%\caption{Alignment Stage L1 Solver}
+%\label{fig:alignment_l1}
+%\end{figure}
 
@@
 In order to adapt the general algorithm to face alignment and recognition, we formulated the 
-$\ell_1$-minimization problem as ___ as described in [book].
+$\ell_1$-minimization problem as described in \cite{}.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
-\caption{GPU Alignment Stage Time vs. Size of Database}
-\label{fig:alignment_runtime}
-\end{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
+%\caption{GPU Alignment Stage Time vs. Size of Database}
+%\label{fig:alignment_runtime}
+%\end{figure}
 
@@
 \centering
-\includegraphics[width=3.5in]{figures/PALM_benchmark_ratio_1.png}
+\includegraphics[width=3.5in]{figures/PALM_benchmark_ratio_1}
 \caption{This is where the caption goes.}

### 2011-02-14 paper/main.tex
Diagram getting included properly.
Alignment and font still need work.
One solution might be to typeset
a full-size version of each diagram
with appropriately scaled font, convert
to ps, and include that.
Re-wrote part of introduction, Simplified Diagram

@@
 case for many problems in computer vision.
-The techniques presented in this paper target these applications, as well
-as others where the dictionary is dense and unstructured. $A$ must be loaded every time
-it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
-than arithmetic logic.
+The techniques presented in this paper target these applications, 
+but are relevant to other non-vision applications where the dictionary is dense and unstructured. 
+%$A$ must be loaded every time
+%it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
+%than arithmetic logic.
 
-% Narrow focus for applications to Recognition Application
 The primary motivating application for this paper is face recognition for access control applications.
@@
 
-For this idea to work, all of the images must be aligned. An iterative alignment
-algorithm for this purpose was described as part of the \cite{Wagner2009-CVPR}, which was
-the first paper to expand \cite{Wright2009-PAMI} into a functioning face recognition system.
-This iterative alignment routine is based on repeatedly solving the following optimization problem:
-\begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_k \x + \e.
-\end{equation}
-Where $A_k$ contains the training images for user $k$ as well as several vectors consisting of the
-Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
-both the current representation coefficients, as well as an update to the transformation parameters.
+For this idea to work, all of the images must be aligned. An iterative
+alignment algorithm for this purpose was described described in
+\cite{Wagner2009-CVPR}.  The algorithm essentially extends Lucas-Kanade
+iterative alignment to use the $\ell_1$ norm as a robust error function, and to
+simultaneously estimate the illumination in terms of the gallery illuminations.
+In combination with a custom training image acquisition system, this resulted
+in a complete face recognition system that achieved good recognition rates even
+on data captured in realistic situations.  Unfortunately, the computational
+cost of this algorithm was still too high for production use.  This speed was
+somewhat improved in \cite{Wagner2010-PAMI}, which showed improved performance
+by using Augmented Lagrange Multiplier (ALM) based $\ell_1$ minimizations
+rather than interior point methods.  This brought alignment time down to about
+half a second per training user, this is still far too slow to be useful in
+access control scenariors where the gallery contains hundreds of users.
+The primary contribution of this paper is to demonstrate how massively
+parallel hardware can dramatically improve recognition time in this application,
+finally making it practical for real-world use.
+
+We now provide an overview of the recognition pipeline proposed 
+in \cite{Wagner2010-PAMI}.  The full recognition pipeline can be divided into 5 major stages.
+
+(... I will describe the stages in more detail with equations...)
+
+These stages are summarized in the diagram in Figure \ref{fig:pipeline}.
+
+%This iterative alignment routine is based on repeatedly solving the following optimization problem:
+%\begin{equation}
+%\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_i \x + \e.
+%\end{equation}
+%Where $A_i$ contains the training images for user $i$ as well as several vectors consisting of the
+%Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
+%both the current representation coefficients, as well as an update to the transformation parameters.
 
@@
 \centering
-%\includegraphics[width=3.5in]{figures/pipeline.pdf}
-\input{figures/pipeline.tex}
+{\tiny \input{figures/pipeline_simplified.tex}}
 \caption{The Full Recognition Pipeline}
@@
 \end{figure}
+
+\begin{algorithm}[thb]
+\caption{\bf (Deformable Sparse Recovery and Classification for
+Face Recognition)} \label{alg:deformable-src}
+\begin{algorithmic}[1]
+\begin{small}
+\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
+$\bb\in\Re^m$ and a deformation group $T$. 
+\STATE
+{\bf for} each subject $i$, 
+\STATE \hspace{3mm} $\tau^{(0)}
+\leftarrow I$. 
+\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
+\STATE \hspace{6mm}
+$\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ
+\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
+\tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
+%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \bb + J \Delta \tau = A_k \x + \e \end{array}\right.$
+\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
+\subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
+\STATE
+\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
+\tau$; 
+\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
+the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
+residuals $\|\e\|_1$. \STATE Compute an average transformation
+$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
+\STATE Update $\bb \leftarrow \bb \circ \bar{\tau}$ and $\tau_i
+\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
+k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
+\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
+\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
+$\ell^1$-minimization problem: \hspace{2mm}
+$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \bb = A \x + \e.$
+\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
+\STATE {\bf Output:} $\mbox{identity}(\bb) = \arg\min_i r_i(\y)$.
+\end{small}
+\end{algorithmic}
+\end{algorithm}
 

### 2011-02-17 paper/main.tex
Added Allen's extensive changes.
old version moved to main_bak.tex
for temporary convenience, to be removed soon.

--- 2011-02-14 paper/main.tex
@@
 
-\usepackage{../authorkit/cvpr}
+\usepackage{iccv}
 \usepackage{times}
@@
 
+\def\arr{{\mathcal{A}}}
+\newcommand{\Null}{{\mbox{Null}}}
+\renewcommand{\mathbf}{\boldsymbol}
+\renewcommand{\Re}{{\mathbb{R}}}
+\def\tt{{\boldsymbol{t}}}
+\def\ee{{\boldsymbol{e}}}
+\def\aa{{\boldsymbol{a}}}
+\def\bb{{\boldsymbol{b}}}
+\def\dd{{\boldsymbol{d}}}
+\def\ss{{\boldsymbol{s}}}
+\def\cc{{\boldsymbol{c}}}
+\def\pp{{\boldsymbol{p}}}
+\def\chat{{\boldsymbol{\hat{\c}}}}
+\def\nn{{\boldsymbol{n}}}
+\def\rr{{\boldsymbol{n}}}
+\def\m{{\boldsymbol{m}}}
+\def\X{{\boldsymbol{X}}}
+\def\L{{\boldsymbol{L}}}
+\def\l{{\boldsymbol{l}}}
+\def\Y{{\boldsymbol{Y}}}
+\def\V{{\boldsymbol{V}}}
+\def\uu{{\boldsymbol{u}}}
+\def\vv{{\boldsymbol{v}}}
+\def\yy{{\boldsymbol{y}}}
+\def\xx{{\boldsymbol{x}}}
+\def\ww{{\boldsymbol{w}}}
+\def\zz{{\boldsymbol{z}}}
+\def\uu{{\boldsymbol{u}}}
+\def\F{{\boldsymbol{F}}}
+\def\veronese{{\nu}}
+\def\Qveronese{{\mu}}
+\def\H{{\mathcal{H}}}
+\def\Q{{\mathcal{Q}}}
+\def\E{{\mathcal{E}}}
+\def\L{{\mathcal{L}}}
+\def\M{{\mathcal{M}}}
+\def\A{{\mathcal{A}}}
+\def\B{{\mathcal{B}}}
+\def\F{{\mathcal{F}}}
+\def\field{\mathbb{F}}
+\def\QED{~\rule[-1pt]{5pt}{5pt}\par\medskip}
+
 % Needed for getting formulas in dia diagrams.
@@
 
-\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
+\def\iccvPaperID{****} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 % Pages are numbered in submission mode, and unnumbered in camera-ready
-\ifcvprfinal\pagestyle{empty}\fi
+\ificcvfinal\pagestyle{empty}\fi
 \begin{document}
@@
 %\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
-\title{Efficient Parallelization of $\ell_1$-Minimization for Face Recognition
-\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
+\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
+%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}
+}
 
@@
 
-%Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization.
-$\ell_1$-minimization is one of the most powerful tools to be added to the engineer's
-toolbox in years.  It has strong theoretical motivation for both for use in recovering sparse
-solutions to systems of linear equations, and has been shown to perform well even in the
-presence of a combination of sparse and dense error.  Since solutions of linear systems
-arise in practically every branch of engineering and there are very few application which
-do {\em not} benefit from a robust error function, the applications are far to numerous to
-itemize here.  
-For many applications of $\ell_1$-minimization solving a system $b=Ax$, $A$ often has a
-special structure that can be leveraged to compute portions of $A$ on the fly
-as they are needed, speeding up perfomance of the minimization dramatically.
-Unfortunately, there are many import applications where this does not hold true.
-This paper targets applications where $\bb$ and $A$ contain
-image data, and thus must be loaded every time they are used, as is the 
-case for many problems in computer vision.
-The techniques presented in this paper target these applications, as well
-as others where the dictionary is dense and unstructured. $A$ must be loaded every time
-it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
-than arithmetic logic.
+$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent years due to important applications
+in compressive sensing \cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.
+The problem refers to finding the minimum $\ell$-norm solution to an underdetermined linear system $\bb=A\xx$:
+\begin{equation}
+\min \|\xx\|_1\quad \mbox{ subj. to }\quad \bb = A\xx.
+\label{eq:l1min}
+\end{equation}
+It has now been well understood that, under certain conditions \cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is
+also the \emph{sparsest} solution to the system \eqref{eq:l1min}.
 
-% Narrow focus for applications to Recognition Application
-The primary motivating application for this paper is face recognition for access control applications.
-Recently, advances in automatic face recognition have been made by re-casting
-it as a sparse representation problem.  This core of this technique consists of
-re-casting recognition as a sparse representation problem. This is done by  
-stacking the test image into a vector $\bb$, stacking the training images for
-all of the users into the columns of a matrix $A$, and solving the following
-minimization problem:
+Among its many applications, $\ell_1$-min has been recently used to reformulate image-based face recognition as a sparse representation problem \cite{WrightJ2009-PAMI}.
+If we stack the training images of $K$ subject classes $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$ into the columns of a matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, given a new query image in the same vector form $\bb\in\Re^d$, \emph{sparsity-based classification} (SBC) solves the following minimization problem:
 \begin{equation}
-\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
+\min_{\xx, \ee} \| \xx \|_1 + \|\ee\|_1 \quad \subj \quad \bb = A \xx + \ee.
+\label{eq:l1min_denoise}
 \end{equation}
-This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
-Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
-correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
-that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
-The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.
+If the sparsest solutions for $\x$ and $\e$ are recovered, $\ee$ provides a means to compensate corrupted pixels that main occlude some part of the query image, and the dominant nonzero coefficients in $\xx$ reveals the membership of $\bb$ based on the training image labels associated with $A$.
 
-For this idea to work, all of the images must be aligned. An iterative alignment
-algorithm for this purpose was described as part of the \cite{Wagner2009-CVPR}, which was
-the first paper to expand \cite{Wright2009-PAMI} into a functioning face recognition system.
-This iterative alignment routine is based on repeatedly solving the following optimization problem:
+In this paper, we study parallelization of $\ell$-min on many-core CPU and GPU architectures.
+Firstly, although $\ell_1$-min \eqref{eq:l1min} is a convex program, conventional algorithms such as interior-point methods \cite{ChenS2001-SIAM,TibshiraniR1996} are known to suffer from poor scalability for large-scale applications such as face recognition.
+Recently, a number of accelerated algorithms have been proposed, which explicitly take advantage of the special structure of $\ell_1$-min problems \cite{LorisI2009,YangA2010-ICIP}.
+We investigate parallelization of a state-of-the-art $\ell_1$-min solution based on the classical framework of \emph{augmented Lagrange multiplier} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
+
+Secondly, while solving \eqref{eq:l1min_denoise} can achieve exceptional recognition accuracy on public databases, its success relies on the assumption that the query image is well aligned with the training images.
+In \cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align a query image to one subject class.
+The face alignment algorithm in turn solves the following minimization problem:
 \begin{equation}
-\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_k \x + \e.
+\hat{\tau}_i = \arg\min_{\xx, \ee, \tau_i} \|\ee\|_1\quad \mbox{subj. to}\quad \bb\circ\tau_i = A_i\xx + \ee,
+\label{eq:l1min_alignment}
 \end{equation}
-Where $A_k$ contains the training images for user $k$ as well as several vectors consisting of the
-Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
-both the current representation coefficients, as well as an update to the transformation parameters.
+where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations (e.g., affine or homography) acting on the image domain, which aligns $\bb$ with the training images from the $i$-th class $A_i$.
+The algorithm essentially extends Lucas-Kanade iterative alignment \cite{LucasB1981} to using the $\ell_1$ norm as a robust error function.
+Although the original solution \cite{WagnerA2009-CVPR} was implemented based on standard interior-point methods, its speed has been improved by ALM in a more recent paper \cite{WagnerA2011-PAMI}.
+In this paper, in addition to solving the generic $\ell_1$-min objectives \eqref{eq:l1min} or \eqref{eq:l1min_denoise}, we also discuss how to properly customize the implementation to accelerate the face alignment step \eqref{eq:l1min_alignment} on many-core CPUs and GPUs.
 
+Finally, we present extensive benchmark to compare the performance of the generic ALM $\ell_1$-min algorithm and the corresponding face recognition routines on massively parallel CPUs and GPUs. 
+To this end, the face recognition system has been fully implemented on a state-of-the-art workstation, as shown in Figure \ref{fig:pipeline}.
+Since any algorithm parallelization is highly dependent on the specific computer hardware, the architecture of our test system will be explained in detail in Section \ref{sec:parallelism}.
+We show that a highly accurate and robust face recognition system can be constructed to recognize face images from a moderate database of several hundred subjects.
+Note that with the possible continual increase of the number of subject classes, it is possible to modularize and dispatch different stages of the face recognition routine to a network of workstations to form an even larger scale networked system.
+Nevertheless, this scenario is not the focus of this paper.
+The source code of our algorithms, which is based on BLAS and Nvidia CUDA libraries, will be made available on our website to aid peer evaluation.
 \begin{figure}
 \centering
-%\includegraphics[width=3.5in]{figures/pipeline.pdf}
-\input{figures/pipeline.tex}
-\caption{The Full Recognition Pipeline}
+{\tiny \input{figures/pipeline_simplified.tex}}
+\caption{The face recognition pipeline implemented on massively parallel hardware.}
 \label{fig:pipeline}
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=3.5in]{figures/alignment_l1.pdf}
-%\caption{Alignment Stage L1 Solver}
-%\label{fig:alignment_l1}
-%\end{figure}
+\subsection{Literature Review}
+In this section, we briefly review the literature of $\ell_1$-min and existing works on its parallelization.
+Traditionally, $\ell_1$-min (or the so-called basis pursuit (BP)) has been formulated as a linear program \cite{ChenS2001-SIAM}.
+Several variations of the solution are also well known in optimization, including a noisy approximation via quadratic programming called the LASSO \cite{TibshiraniR1996} and truncated Newton interior-point method (TNIPM) \cite{KimS2007}.
 
-\subsection{Literature review}
-%[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
-%[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.
-%[Zero in and justify why we choose ALM for our implementation.]
+One of the drawbacks of most interior-point methods for $\ell_1$-min (including BP, LASSO, and TNIPM) is that they require the solution sequence to follow an interior path, which is often computationally expensive.
+To mitigate these issues, an approach called \emph{Homotopy} has been studied to accelerate the speed of $\ell_1$-min.
+Homotopy methods for recovering sparse signals were first studied in the context of LASSO \cite{OsborneM2000}, which inspired a solution to the \emph{forward stagewise linear regression} problem called LARS \cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms for BP in \cite{MalioutovD2005,DonohoD2006}.
 
-Standard solvers for $\ell_1$-minimization include orthogonal matching pursuit, basis pursuit and LASSO which cast the $\ell_1$-minimization as either a linear program or quadratic program.
+Although Homotopy can be shown to exactly estimate BP when the solution is sufficiently sparse \cite{DonohoD2006}, the algorithm still involves expensive matrix operations such as matrix-matrix multiplication and solving least-squares problems with varying $A$ matrices.
+Hence, it is not efficient to implement Homotopy on parallel hardware.
+In Section \ref{sec:ALM}, we contend that ALM is a better choice among Homotopy and other existing $\ell_1$-min solvers to parallelize on many-core CPUs and GPUs.
+The algorithm belongs to a category of approximate $\ell_1$-min solutions called \emph{iterative shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}.
+The (sequential) implementation of IST algorithms mainly involves elementary operations such as vector algebra and matrix-vector multiplication.
+Therefore, when the dimension of the problem becomes high, IST-type algorithms are particularly suitable for systems in parallel-computing environments.
+In \cite{YangA2010-ICIP}, the authors further showed that, compared to other IST-type algorithms, ALM is able to achieve good estimation accuracy comparable to BP and Homotopy, while at the same time significantly improve the speed compared to these exact $\ell_1$-min solutions.
+Due to these facts, in this paper, we choose ALM as the core algorithm to implement the complete parallel face recognition pipeline.
 
-These methods can be accelerated by more contemporary methods such as gradient projection, homotopy, iterative shrinkage-thresholding, proximal gradient, and augmented lagrangian methods.
-According to analysis of these five algorithms applied to face recognition, Yang et al have determined that Homotopy and the Augmented Lagrangian Multiplier methods (ALM) achieve the best recognition rates while maintaining a low computational cost.
-A closer look at these two algorithms will show that while the Homotopy method will produce the exact solution, the algorithm structure and functions do not lend themselves parallelization well.
-On the other hand, while the ALM finds an approximate solution, its computation mainly consists of matrix-vector operations, which lends itself to parallelization.
+In terms of the past works in parallel $\ell_1$-min, the literature has been limited, to the best of our knowledge.
+In \cite{BorghiA2010}, Borghi et al.
+developed a special proximal gradient $\ell_1$-min algorithm based on Moreau-Yosida regularization.
+In \cite{MurphyM2010}, Murphy et al.
+presented parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia programmable GPUs.
 
-Past work in parallelizing $\ell_1$-minimization has been limited.
-In 2009, Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and in 2009 Borghi at UCLA developed a proximal point algorithm using Moreau-Yosida regularization, which is special form of the $\ell_1$-minimization based on proximal gradient.
+%\subsection{Contributions of the Paper}
 
-\subsection{Contributions of the paper}
+\section{Augmented Lagrange Multiplier Method}
+\label{sec:ALM}
 
-\section{Augmented Lagrangian Multiplier Algorithm}
+For the completeness of the paper, in this section, we briefly describe the ALM algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze its complexity.
+Parallelization of ALM for face alignment and recognition will be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition}, respectively.
 
-In this section, we describe the general version ALM algorithm and then describe how it can be adapted to solve our problem.
+Lagrange multiplier methods have been frequently used in convex programming to eliminate equality constraints via adding a significant penalty term to the cost function for infeasible points.
+ALM methods differ from other penalty-based approaches by simultaneously estimating the optimal solution and Lagrange multipliers in an iterative fashion.
+For $\ell_1$-min, the augmented Lagrange function is defined as:
+\begin{equation}
+L_\mu(\xx,\yy) = \|\xx\|_1 + \yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2,
+\end{equation}
+where $\mu > 0$ is a constant that penalizes infeasibility and $\yy$ is a vector of lagrange multipliers.
 
-The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
+In Lagrange Multiplier Theory, if there exists a Lagrangian $y^*$ that satisfies the second-order sufficiency conditions for optimality \cite{BertsekasD2003}, then for a sufficiently large $mu$, the optimal $\ell_1$-min solution also minimizes
 \begin{equation}
-F(x) = f(x) + \lambda g(x).
-\end{equation}
-which can be formulated as an augmented Lagrange function given by
-\begin{equation}
-L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
+\xx^* = \arg \min L_\mu(\xx,\yy^*).
+\label{eq:optimal-ALM}
 \end{equation}
 
-where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.
+In practice, the optimal values for the triplet $(\xx^*, \yy^*, \uu)$ are all unknown.
+Furthermore, it has been observed that if solving \eqref{eq:optimal-ALM} with a large initial value of $\mu$ tends to lead to slower convergence speed \cite{WrightS2008,YangA2010-ICIP}.
+In \cite{BertsekasD2003,YangJ2009}, an alternating procedure has been proposed to iteratively update $\xx$ and $\yy$:
+\begin{equation}
+\left \{
+\begin{array}{lll}
+\xx_{k+1} & = & \arg\min_{\xx} \, L_{\mu_k} (\xx,\yy_k)\\
+\yy_{k+1} & = & \yy_k + \mu_k (\bb - A\xx_{k+1}) \\
+\end{array}
+\right . ,
+\label{eq:alm}
+\end{equation}
+where $\left\{\mu_{k}\right\}$ is a monotonically increasing positive sequence. The iteration terminates when the estimates $(\xx, \yy)$ converges.
 
-For an optimal $y*$ which satisfies the second-order sufficiency conditions, for a sufficiently large $mu$, it can be shown that
-\begin{equation}
-x* = \arg min L_u(x,y*)
-\end{equation}
+Note that in the iterative procedure outlined in \eqref{eq:alm}, the second step only involves vector algebra and matrix-vector multiplication.
+Therefore, the procedure is computationally efficient only if it is easier to minimize the augmented Lagrange function compared to solving the original problem \eqref{eq:l1min} directly.
+In fact, the first step that minimizes $L_{\mu_k} (\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009}, whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
+Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm.
+ \begin{algorithm}[h]
+\caption{Augmented Lagrange Multiplier (ALM)}
+{\bf INPUT:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$, $\tau\leftarrow$ largest eigenvalue of $A^TA$, and a constant $\rho>1$.
+\begin{algorithmic}[1]
+\WHILE{not converged ($k = 1,2,\ldots$)} \STATE $t_1 \leftarrow 1$, $\zz_1
+\leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ \WHILE{not
+converged ($l = 1,2,\ldots$)} \STATE $\uu_{l+1}  \leftarrow
+\mathrm{soft}\left(\zz_l - \frac{1}{\tau}A^T\left(A\zz_l - \bb
+- \frac{1}{\mu_k}\yy_k\right), \frac{1}{\mu_k\tau}\right)$
+\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
+\sqrt{1+4t_l^2}\right)$ \STATE $\zz_{l+1} \leftarrow \uu_{l+1}
++ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ \ENDWHILE \STATE$
+\xx_{k+1} \leftarrow \uu_{l+1}$ \STATE $\yy_{k+1} \leftarrow
+\yy_k + \mu_k (\bb - A\xx_{k+1})$ \STATE $\mu_{k+1} \leftarrow
+\rho\cdot\mu_k$ \ENDWHILE 
+\end{algorithmic}
 
-However, because $y*$ is generally unknown, ALM methods simultaneously estimate the optimal solution and Lagrange multipliers iteratively in the following procedure.
-
-We also need to define the following soft-thresholding operator for a
-scalar $x$ and a scalar $\alpha \geq 0$:
-\begin{equation}
-\textup{shrink}(x,\alpha) = \textup{sign}(x)\cdot \max \{|x| - \alpha, 0\},
-\end{equation}
-In practice, this operator will be applied elementwise to a vector $\x$ with a single $\lambda$,
-and will be notated as $\textup{shrink}(\x,\alpha)$.
-
-We summarize the entire ALM
-algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
-in \cite{YangJ2009-pp} and set $\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
-\begin{algorithm}[h]
-\caption{\bf (Augmented Lagrange Multiplier Method FIX)}
-\begin{algorithmic}[1]
-\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
-\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
-\ENDWHILE
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
-\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{algorithmic}
+{\bf OUTPUT:} $\xx^* \leftarrow \xx_k$.
 \label{alg:alm}
@@
 
-%In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
-%\begin{equation} 
-%\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
-%\end{equation}
-%which is solved in a similar fashion.
-
-% A note on alignment l1...
-% With a reshuffling of the order of operations and a change of the termination
-% condition, the inner loop takes the form of two matrix-vector multiplications
-% and a sequence of vector-vector operations.
-
-In order to adapt the general algorithm to face alignment and recognition, we formulated the 
-$\ell_1$-minimization problem as described in \cite{}.
-
-\subsection{Alignment Stage L1 Computational Complexity} Any discussion of the speed of an
-algorithm should discuss the comutational complexity of the algorithm.  For the
-ALM algorithms considered in this paper, this is rather simple.  The
-Matrix-vector computions are $O(mn)$, and the vector-vector operations are
-$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
-very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
-computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
-A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
-Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
-of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
-Fortunately, $A^\dagger$ only needs to be computed once per instance of the
-$\ell_1$ minimization problem.
-
-Therefore, for problems with large $n$, we expect the computation of the
-pseudoinverse to dominate.  However, for the problem sizes needed in face
-recognition problems, we find that the computation time is dominated by the
-repeated computation of matrix-vector and matrix-matrix multiplications in the
-inner loop.
-
 \section{Hardware Parallelism}
-In this section we give a brief overview of the hardware architectures used
-in this paper.
-
-We feel it is most meaningful to compare hardware architectures on a per-board basis,
-and in a typical hardware configuration.  GPU's are most commonly shipped with a single
-GPU chip per PCI-express card.  
-%For multicore CPU architectures, a typical workstation
-%will consist of
-
-For CPU implementations, the benchmarks makes use of all of
-the cores in as many cpu's are present.  For GPU implementations, the benchmark
-makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Unless otherwise specified, all implementations utilize single
+\label{sec:parallelism}
+In this section we discuss the hardware architectures used in this paper. Aside from the $\ell_1$-min algorithm, our choice
+of the architectures plays a significant role in many practical decisions we make in Section \ref{sec:alignment} and Section \ref{sec:recognition}.
+For CPU implementation, the system would make use of all the cores in as many CPUs are present on one motherboard.  For GPU implementation, the system would
+make the best use of a single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with a single GPU chip per PCI-express card).
+Unless otherwise specified, all implementations utilize single
 precision floating point datatypes.
@@
 \subsection{CPU Hardware Parallelism} 
-The baseline architecture for our experiments is a typical workstation,
-with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each
+The baseline architecture for our experiments is a Linux workstation with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each
 processor has its own memory interface, and is directly connected to half of
@@
 
-Each core has a 32 KB L1 data cache and a 256 KB L2 cache. With Hyperthreading
+Each core has a 32\,KB L1 data cache and a 256\,KB L2 cache. With Hyperthreading
 enabled, each core can support two threads simultaneously.  For floating point
-math, each core also has a vector unit capable of performing the same
+instructions, each core also has a vector unit capable of performing the same
 arithmetic operation on four single precision floating point values
@@
 some compilers, notably gcc, generate code that uses the vector units
-exclusively for floating point math.}
+exclusively for floating point instructions.}
 
-There are thus two important levels of parallelism that need to be exploited to
-use a modern cpu efficiently: core-level parallelism and vector-level
+There are thus two important levels of parallelism that need to be exploited to efficiently
+use a modern CPU: core-level parallelism and vector-level
 parallelism.  For the implementations discussed here, we will primarily use the
-OpenMP programming model to express core-level parallelism.  Depending on the
-section of the code, we will achieve vector-level parallelism implicitly
+OpenMP programming model to express core-level parallelism {\bf AY: What is OpenMP? citation needed}.  Depending on the
+section of the code{\bf AY: What is a section of the code?}, we will achieve vector-level parallelism implicitly
 through either the auto-vectorization features of Intel's icc compiler, or
-through the use of a highly optimized BLAS library (Intel's MKL).
+through the use of a highly optimized BLAS library (Intel's MKL). {\bf AY: citations needed, what is icc, what is the relation between BLAS and MKL? Are they equivalent? not clear in the context.}
 
-Each processor has 8MB of L3 cache that is shared by all four cores.  Overall,
+Each processor further has 8MB of L3 cache that is shared by all four cores.  Overall,
 this means the algorithm has approximately 16MB of on-chip cache available.
 For the low resolution images used in sparse representation based face
-recognition, this is enough cache to hold over 800 image-sized variables,
+recognition, this is enough cache to hold over 800 image-sized variables, {\bf AY: how low resolution? which sparse representation problem? ``image-sized variables'' means simply ``image vectors''?}
 meaning that for small data sets it is possible to run completely using on-chip
@@
 
-The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi) which each contain 32 floating-point units.
+The ALM algorithm requires many matrix-vector operations, this gives many-core processors such as GPUs a computational advantage for large problem sizes.
+GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi) which each contains 32 floating-point units.
 Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.
-The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called warps.
-These warps are then grouped into higher levels called thread blocks and grids.
+The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called \emph{warps}.
+These warps are then grouped into higher levels called \emph{thread blocks} and \emph{grids}.
 The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance and a computational advantage over CPUs.
 Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and 768kb of L2 cache.
-This increase in performance doesn't come for free.
-Because the GPU has it's own memory system, any data the GPU uses must be transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
+However, this increase in performance does not come for free.
+As the GPU has its own memory system, any data the GPU uses must be transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
 However, the cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
 
-The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
+In summary, a programmable GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
 Each thread block is a virtualized multiprocessor, which each thread as a virtualized scalar processor.
@@
 % This section is Drew's
-\section{Iterative Alignment Stage Implementation}
+\section{Parallelization of the Face Alignment Stage}
+\label{sec:alignment}
+{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
 
-\subsection{Iterative Alignment on the CPU}
+\subsection{CPU Implementation}
+
+{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
 %\subsection{Tuning the GEMV operation}
@@
 %...Block the computation into work for different cpu cores
-\subsection{Iterative Alignment on the GPU}
-% Describingn the sequence of GPU implementations Drew has tried...
+\subsection{GPU Implementation}
+{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
 
-\subsection{Benchmark on Real Face Data}
+\section{Parallelization of the Face Recognition Stage}
+\label{sec:recognition}
+{\bf AY: Add an algorithm here describe the recognition block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
-%\caption{GPU Alignment Stage Time vs. Size of Database}
-%\label{fig:alignment_runtime}
-%\end{figure}
+\subsection{CPU Implementation}
+{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
 
-\section{Recognition Stage Implementation}
-% This section is Victor's
-
-%The ALM algorithm requires many matrix-vector operations, this gives manycore
-%processors such as GPUs a computational advantage for large problem sizes.
-%GPUs are comprised of several streaming multiprocessors, which contain 32
-%floating-point units.  The GPU executes programs as SIMT (Single Instruction
-%Multiple Threads) in groups of 32 threads called warps.  These warps are then
-%grouped into higher levels called thread blocks and grids.  The improvement in
-%speed comes from the memory structure of the GPU.   The GPU contains an
-%inverted memory hierarchy so that the L1 cache stays close to the cores,
-%providing the increase in performance.  It essentially turns the GPU from being
-%memory-limited to computational-limited.  GPU memory also differs from CPU
-%memory in that it requires data transfers over PCI-Express.  The cost of these
-%transfers are amortized by doing large amounts of computation at once,
-%requiring code to be run completely on the GPU.  The GPU can be analogous to a
-%multi-core CPU processor, except with more sophisticated vector processing
-%hardware.  Each streaming multiprocessor is analogous to a single CPU except
-%that it can run multiple threads simultaneously.  All of these differences
-%allow GPUs to perform intensive calculations such as matrix-vector operations
-%at much faster rates than CPUs.
-
-% In the followingn 2 sections, just describe implementations; single number
-% benchmarks and microbenchmarks are ok, but save big graphs for experiments
-% section.
-
-\subsection{ GPU Implementation}
+\subsection{GPU Implementation}
+{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
@@
 
-\subsection{Benchmark on Synthetic Data}
+\section{Experiment}
+{\bf AY: Describe the benchmark protocol here. Both for simulation and real face data.}
 
-To construct the benchmark for synthetic data, we first created an $m \times n$
-matrix $A$ with entries $\pm 1$ each with a $50\%$ chance.  Then we created a
-sparse vector $\x_0$ with $10\%$ nonzero random entries.  We also created a
-sparse vector $y_k$ with the same number of non-zero entries as $\x$.  Let
-$\y_0 = A \x_0 + \y_k$.  We timed the time taken for the ALM algorithm using
-$\y_0$, $A$, and other inputs.  The other inputs included, maximum number of
-inner/outer loop iterations, and inner/outer loop tolerance level.
+\subsection{Simulation}
 
-To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations
-at 5000 and run the algorithm to convergence.  CPU code was run on an Intel
-Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480 We
-ran the A matrix with different ratios between m:n from 10:1:5 for m:2000:25:8000 with an x vector of 10 percent sparsity.  Tolerance levels were
-empirically seen to allow best performance with the inner loop tolerance set at
-one order of magnitude lower than the outer loop tolerance.
-
-Results were plotted as total size of A and A'A vs Gb/s and size of A vs time
-elapsed to run a single $\ell_1$ solver.
-
-\begin{figure*}
-\centering
-\includegraphics[width=3.5in]{figures/PALM_benchmark_ratio_1}
-\caption{This is where the caption goes.}
-\label{fig:uniqueidentifierforthisimage}
-\end{figure*}
-
-As the dimensions of $A$ increase, so does the significance of the size of
-$A^TA$.  Until the $A$ and $A^TA$ reach ???? Mb (SHOULD BE MB?), the CPU version runs faster
-than the GPU.
-
-\section{Full Pipeline Benchmarks}
+\subsection{Face Recognition Benchmark}
 
@@
 {\small
-\bibliographystyle{../authorkit/ieee}
+\bibliographystyle{ieee}
 \bibliography{faces}

### 2011-02-17 paper/main.tex
Moved Allen's additional defs to mydefs.sty
Reflowed long lines

@@
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
-
-\def\arr{{\mathcal{A}}}
-\newcommand{\Null}{{\mbox{Null}}}
-\renewcommand{\mathbf}{\boldsymbol}
-\renewcommand{\Re}{{\mathbb{R}}}
-\def\tt{{\boldsymbol{t}}}
-\def\ee{{\boldsymbol{e}}}
-\def\aa{{\boldsymbol{a}}}
-\def\bb{{\boldsymbol{b}}}
-\def\dd{{\boldsymbol{d}}}
-\def\ss{{\boldsymbol{s}}}
-\def\cc{{\boldsymbol{c}}}
-\def\pp{{\boldsymbol{p}}}
-\def\chat{{\boldsymbol{\hat{\c}}}}
-\def\nn{{\boldsymbol{n}}}
-\def\rr{{\boldsymbol{n}}}
-\def\m{{\boldsymbol{m}}}
-\def\X{{\boldsymbol{X}}}
-\def\L{{\boldsymbol{L}}}
-\def\l{{\boldsymbol{l}}}
-\def\Y{{\boldsymbol{Y}}}
-\def\V{{\boldsymbol{V}}}
-\def\uu{{\boldsymbol{u}}}
-\def\vv{{\boldsymbol{v}}}
-\def\yy{{\boldsymbol{y}}}
-\def\xx{{\boldsymbol{x}}}
-\def\ww{{\boldsymbol{w}}}
-\def\zz{{\boldsymbol{z}}}
-\def\uu{{\boldsymbol{u}}}
-\def\F{{\boldsymbol{F}}}
-\def\veronese{{\nu}}
-\def\Qveronese{{\mu}}
-\def\H{{\mathcal{H}}}
-\def\Q{{\mathcal{Q}}}
-\def\E{{\mathcal{E}}}
-\def\L{{\mathcal{L}}}
-\def\M{{\mathcal{M}}}
-\def\A{{\mathcal{A}}}
-\def\B{{\mathcal{B}}}
-\def\F{{\mathcal{F}}}
-\def\field{\mathbb{F}}
-\def\QED{~\rule[-1pt]{5pt}{5pt}\par\medskip}
 
@@
 \section{Introduction} 
-
-$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent years due to important applications
-in compressive sensing \cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.
-The problem refers to finding the minimum $\ell$-norm solution to an underdetermined linear system $\bb=A\xx$:
+$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
+years due to important applications in compressive sensing
+\cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  The
+problem refers to finding the minimum $\ell$-norm solution to an
+underdetermined linear system $\bb=A\xx$:
 \begin{equation}
@@
 \end{equation}
-It has now been well understood that, under certain conditions \cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is
-also the \emph{sparsest} solution to the system \eqref{eq:l1min}.
+It has now been well understood that, under certain conditions
+\cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is also
+the \emph{sparsest} solution to the system \eqref{eq:l1min}.
 
-Among its many applications, $\ell_1$-min has been recently used to reformulate image-based face recognition as a sparse representation problem \cite{WrightJ2009-PAMI}.
-If we stack the training images of $K$ subject classes $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$ into the columns of a matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, given a new query image in the same vector form $\bb\in\Re^d$, \emph{sparsity-based classification} (SBC) solves the following minimization problem:
+Among its many applications, $\ell_1$-min has been recently used to reformulate
+image-based face recognition as a sparse representation problem
+\cite{WrightJ2009-PAMI}.  If we stack the training images of $K$ subject
+classes $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$ into the
+columns of a matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, given a new
+query image in the same vector form $\bb\in\Re^d$, \emph{sparsity-based
+classification} (SBC) solves the following minimization problem:
 \begin{equation}
@@
 \end{equation}
-If the sparsest solutions for $\x$ and $\e$ are recovered, $\ee$ provides a means to compensate corrupted pixels that main occlude some part of the query image, and the dominant nonzero coefficients in $\xx$ reveals the membership of $\bb$ based on the training image labels associated with $A$.
+If the sparsest solutions for $\x$ and $\e$ are recovered, $\ee$ provides a
+means to compensate corrupted pixels that main occlude some part of the query
+image, and the dominant nonzero coefficients in $\xx$ reveals the membership of
+$\bb$ based on the training image labels associated with $A$.
 
-In this paper, we study parallelization of $\ell$-min on many-core CPU and GPU architectures.
-Firstly, although $\ell_1$-min \eqref{eq:l1min} is a convex program, conventional algorithms such as interior-point methods \cite{ChenS2001-SIAM,TibshiraniR1996} are known to suffer from poor scalability for large-scale applications such as face recognition.
-Recently, a number of accelerated algorithms have been proposed, which explicitly take advantage of the special structure of $\ell_1$-min problems \cite{LorisI2009,YangA2010-ICIP}.
-We investigate parallelization of a state-of-the-art $\ell_1$-min solution based on the classical framework of \emph{augmented Lagrange multiplier} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
+In this paper, we study parallelization of $\ell$-min on many-core CPU and GPU
+architectures. Firstly, although $\ell_1$-min \eqref{eq:l1min} is a convex
+program, conventional algorithms such as interior-point methods
+\cite{ChenS2001-SIAM,TibshiraniR1996} are known to suffer from poor scalability
+for large-scale applications such as face recognition. Recently, a number of
+accelerated algorithms have been proposed, which explicitly take advantage of
+the special structure of $\ell_1$-min problems
+\cite{LorisI2009,YangA2010-ICIP}. We investigate parallelization of a
+state-of-the-art $\ell_1$-min solution based on the classical framework of
+\emph{augmented Lagrange multiplier} (ALM)
+\cite{BertsekasD2003,YangA2010-ICIP}.
 
-Secondly, while solving \eqref{eq:l1min_denoise} can achieve exceptional recognition accuracy on public databases, its success relies on the assumption that the query image is well aligned with the training images.
-In \cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align a query image to one subject class.
-The face alignment algorithm in turn solves the following minimization problem:
+Secondly, while solving \eqref{eq:l1min_denoise} can achieve exceptional
+recognition accuracy on public databases, its success relies on the assumption
+that the query image is well aligned with the training images. In
+\cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align
+a query image to one subject class. The face alignment algorithm in turn solves
+the following minimization problem:
 \begin{equation}
@@
 \end{equation}
-where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations (e.g., affine or homography) acting on the image domain, which aligns $\bb$ with the training images from the $i$-th class $A_i$.
-The algorithm essentially extends Lucas-Kanade iterative alignment \cite{LucasB1981} to using the $\ell_1$ norm as a robust error function.
-Although the original solution \cite{WagnerA2009-CVPR} was implemented based on standard interior-point methods, its speed has been improved by ALM in a more recent paper \cite{WagnerA2011-PAMI}.
-In this paper, in addition to solving the generic $\ell_1$-min objectives \eqref{eq:l1min} or \eqref{eq:l1min_denoise}, we also discuss how to properly customize the implementation to accelerate the face alignment step \eqref{eq:l1min_alignment} on many-core CPUs and GPUs.
+where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations
+(e.g., affine or homography) acting on the image domain, which aligns $\bb$
+with the training images from the $i$-th class $A_i$. The algorithm essentially
+extends Lucas-Kanade iterative alignment \cite{LucasB1981} to using the
+$\ell_1$ norm as a robust error function.  Although the original solution
+\cite{WagnerA2009-CVPR} was implemented based on standard interior-point
+methods, its speed has been improved by ALM in a more recent paper
+\cite{WagnerA2011-PAMI}. In this paper, in addition to solving the generic
+$\ell_1$-min objectives \eqref{eq:l1min} or \eqref{eq:l1min_denoise}, we also
+discuss how to properly customize the implementation to accelerate the face
+alignment step \eqref{eq:l1min_alignment} on many-core CPUs and GPUs.
 
-Finally, we present extensive benchmark to compare the performance of the generic ALM $\ell_1$-min algorithm and the corresponding face recognition routines on massively parallel CPUs and GPUs. 
-To this end, the face recognition system has been fully implemented on a state-of-the-art workstation, as shown in Figure \ref{fig:pipeline}.
-Since any algorithm parallelization is highly dependent on the specific computer hardware, the architecture of our test system will be explained in detail in Section \ref{sec:parallelism}.
-We show that a highly accurate and robust face recognition system can be constructed to recognize face images from a moderate database of several hundred subjects.
-Note that with the possible continual increase of the number of subject classes, it is possible to modularize and dispatch different stages of the face recognition routine to a network of workstations to form an even larger scale networked system.
-Nevertheless, this scenario is not the focus of this paper.
-The source code of our algorithms, which is based on BLAS and Nvidia CUDA libraries, will be made available on our website to aid peer evaluation.
+Finally, we present extensive benchmark to compare the performance of the
+generic ALM $\ell_1$-min algorithm and the corresponding face recognition
+routines on massively parallel CPUs and GPUs.  To this end, the face
+recognition system has been fully implemented on a state-of-the-art
+workstation, as shown in Figure \ref{fig:pipeline}. Since any algorithm
+parallelization is highly dependent on the specific computer hardware, the
+architecture of our test system will be explained in detail in Section
+\ref{sec:parallelism}.  We show that a highly accurate and robust face
+recognition system can be constructed to recognize face images from a moderate
+database of several hundred subjects. Note that with the possible continual
+increase of the number of subject classes, it is possible to modularize and
+dispatch different stages of the face recognition routine to a network of
+workstations to form an even larger scale networked system. Nevertheless, this
+scenario is not the focus of this paper. The source code of our algorithms,
+which is based on BLAS and Nvidia CUDA libraries, will be made available on our
+website to aid peer evaluation.
 \begin{figure}
@@
 \subsection{Literature Review} 
-In this section, we briefly review the literature of $\ell_1$-min and existing works on its parallelization.
-Traditionally, $\ell_1$-min (or the so-called basis pursuit (BP)) has been formulated as a linear program \cite{ChenS2001-SIAM}.
-Several variations of the solution are also well known in optimization, including a noisy approximation via quadratic programming called the LASSO \cite{TibshiraniR1996} and truncated Newton interior-point method (TNIPM) \cite{KimS2007}.
+In this section, we briefly review the literature of $\ell_1$-min and existing
+works on its parallelization.  Traditionally, $\ell_1$-min (or the so-called
+basis pursuit (BP)) has been formulated as a linear program
+\cite{ChenS2001-SIAM}. Several variations of the solution are also well known
+in optimization, including a noisy approximation via quadratic programming
+called the LASSO \cite{TibshiraniR1996} and truncated Newton interior-point
+method (TNIPM) \cite{KimS2007}.
 
-One of the drawbacks of most interior-point methods for $\ell_1$-min (including BP, LASSO, and TNIPM) is that they require the solution sequence to follow an interior path, which is often computationally expensive.
-To mitigate these issues, an approach called \emph{Homotopy} has been studied to accelerate the speed of $\ell_1$-min.
-Homotopy methods for recovering sparse signals were first studied in the context of LASSO \cite{OsborneM2000}, which inspired a solution to the \emph{forward stagewise linear regression} problem called LARS \cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms for BP in \cite{MalioutovD2005,DonohoD2006}.
+One of the drawbacks of most interior-point methods for $\ell_1$-min (including
+BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
+interior path, which is often computationally expensive. To mitigate these
+issues, an approach called \emph{Homotopy} has been studied to accelerate the
+speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were
+first studied in the context of LASSO \cite{OsborneM2000}, which inspired a
+solution to the \emph{forward stagewise linear regression} problem called LARS
+\cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms
+for BP in \cite{MalioutovD2005,DonohoD2006}.
 
-Although Homotopy can be shown to exactly estimate BP when the solution is sufficiently sparse \cite{DonohoD2006}, the algorithm still involves expensive matrix operations such as matrix-matrix multiplication and solving least-squares problems with varying $A$ matrices.
-Hence, it is not efficient to implement Homotopy on parallel hardware.
-In Section \ref{sec:ALM}, we contend that ALM is a better choice among Homotopy and other existing $\ell_1$-min solvers to parallelize on many-core CPUs and GPUs.
-The algorithm belongs to a category of approximate $\ell_1$-min solutions called \emph{iterative shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}.
-The (sequential) implementation of IST algorithms mainly involves elementary operations such as vector algebra and matrix-vector multiplication.
-Therefore, when the dimension of the problem becomes high, IST-type algorithms are particularly suitable for systems in parallel-computing environments.
-In \cite{YangA2010-ICIP}, the authors further showed that, compared to other IST-type algorithms, ALM is able to achieve good estimation accuracy comparable to BP and Homotopy, while at the same time significantly improve the speed compared to these exact $\ell_1$-min solutions.
-Due to these facts, in this paper, we choose ALM as the core algorithm to implement the complete parallel face recognition pipeline.
+Although Homotopy can be shown to exactly estimate BP when the solution is
+sufficiently sparse \cite{DonohoD2006}, the algorithm still involves expensive
+matrix operations such as matrix-matrix multiplication and solving
+least-squares problems with varying $A$ matrices. Hence, it is not efficient to
+implement Homotopy on parallel hardware. In Section \ref{sec:ALM}, we contend
+that ALM is a better choice among Homotopy and other existing $\ell_1$-min
+solvers to parallelize on many-core CPUs and GPUs. The algorithm belongs to a
+category of approximate $\ell_1$-min solutions called \emph{iterative
+shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}. The
+(sequential) implementation of IST algorithms mainly involves elementary
+operations such as vector algebra and matrix-vector multiplication. Therefore,
+when the dimension of the problem becomes high, IST-type algorithms are
+particularly suitable for systems in parallel-computing environments. In
+\cite{YangA2010-ICIP}, the authors further showed that, compared to other
+IST-type algorithms, ALM is able to achieve good estimation accuracy comparable
+to BP and Homotopy, while at the same time significantly improve the speed
+compared to these exact $\ell_1$-min solutions. Due to these facts, in this
+paper, we choose ALM as the core algorithm to implement the complete parallel
+face recognition pipeline.
 
-In terms of the past works in parallel $\ell_1$-min, the literature has been limited, to the best of our knowledge.
-In \cite{BorghiA2010}, Borghi et al.
-developed a special proximal gradient $\ell_1$-min algorithm based on Moreau-Yosida regularization.
-In \cite{MurphyM2010}, Murphy et al.
-presented parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia programmable GPUs.
-
-%\subsection{Contributions of the Paper}
+In terms of the past works in parallel $\ell_1$-min, the literature has been
+limited, to the best of our knowledge. In \cite{BorghiA2010}, Borghi et al.
+developed a special proximal gradient $\ell_1$-min algorithm based on
+Moreau-Yosida regularization. In \cite{MurphyM2010}, Murphy et al. presented
+parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia
+programmable GPUs.
 
@@
 
-For the completeness of the paper, in this section, we briefly describe the ALM algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze its complexity.
-Parallelization of ALM for face alignment and recognition will be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition}, respectively.
+For the completeness of the paper, in this section, we briefly describe the ALM
+algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze
+its complexity. Parallelization of ALM for face alignment and recognition will
+be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition},
+respectively.
 
-Lagrange multiplier methods have been frequently used in convex programming to eliminate equality constraints via adding a significant penalty term to the cost function for infeasible points.
-ALM methods differ from other penalty-based approaches by simultaneously estimating the optimal solution and Lagrange multipliers in an iterative fashion.
-For $\ell_1$-min, the augmented Lagrange function is defined as:
-\begin{equation}
-L_\mu(\xx,\yy) = \|\xx\|_1 + \yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2,
-\end{equation}
-where $\mu > 0$ is a constant that penalizes infeasibility and $\yy$ is a vector of lagrange multipliers.
+Lagrange multiplier methods have been frequently used in convex programming to
+eliminate equality constraints via adding a significant penalty term to the
+cost function for infeasible points. ALM methods differ from other
+penalty-based approaches by simultaneously estimating the optimal solution and
+Lagrange multipliers in an iterative fashion.  For $\ell_1$-min, the augmented
+Lagrange function is defined as: \begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
+\yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2, \end{equation}
+where $\mu > 0$ is a constant that penalizes infeasibility and $\yy$ is a
+vector of lagrange multipliers.
 
-In Lagrange Multiplier Theory, if there exists a Lagrangian $y^*$ that satisfies the second-order sufficiency conditions for optimality \cite{BertsekasD2003}, then for a sufficiently large $mu$, the optimal $\ell_1$-min solution also minimizes
+In Lagrange Multiplier Theory, if there exists a Lagrangian $y^*$ that
+satisfies the second-order sufficiency conditions for optimality
+\cite{BertsekasD2003}, then for a sufficiently large $mu$, the optimal
+$\ell_1$-min solution also minimizes
 \begin{equation}
@@
 
-In practice, the optimal values for the triplet $(\xx^*, \yy^*, \uu)$ are all unknown.
-Furthermore, it has been observed that if solving \eqref{eq:optimal-ALM} with a large initial value of $\mu$ tends to lead to slower convergence speed \cite{WrightS2008,YangA2010-ICIP}.
-In \cite{BertsekasD2003,YangJ2009}, an alternating procedure has been proposed to iteratively update $\xx$ and $\yy$:
+In practice, the optimal values for the triplet $(\xx^*, \yy^*, \uu)$ are all
+unknown. Furthermore, it has been observed that if solving
+\eqref{eq:optimal-ALM} with a large initial value of $\mu$ tends to lead to
+slower convergence speed \cite{WrightS2008,YangA2010-ICIP}. In
+\cite{BertsekasD2003,YangJ2009}, an alternating procedure has been proposed to
+iteratively update $\xx$ and $\yy$:
 \begin{equation}
@@
 \end{equation}
-where $\left\{\mu_{k}\right\}$ is a monotonically increasing positive sequence. The iteration terminates when the estimates $(\xx, \yy)$ converges.
+where $\left\{\mu_{k}\right\}$ is a monotonically increasing positive sequence.
+The iteration terminates when the estimates $(\xx, \yy)$ converges.
 
-Note that in the iterative procedure outlined in \eqref{eq:alm}, the second step only involves vector algebra and matrix-vector multiplication.
-Therefore, the procedure is computationally efficient only if it is easier to minimize the augmented Lagrange function compared to solving the original problem \eqref{eq:l1min} directly.
-In fact, the first step that minimizes $L_{\mu_k} (\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009}, whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
+Note that in the iterative procedure outlined in \eqref{eq:alm}, the second
+step only involves vector algebra and matrix-vector multiplication. Therefore,
+the procedure is computationally efficient only if it is easier to minimize the
+augmented Lagrange function compared to solving the original problem
+\eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k}
+(\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise
+iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009},
+whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
 Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm.
+
  \begin{algorithm}[h]
@@
 
-\section{Hardware Parallelism}
-\label{sec:parallelism}
-In this section we discuss the hardware architectures used in this paper. Aside from the $\ell_1$-min algorithm, our choice
-of the architectures plays a significant role in many practical decisions we make in Section \ref{sec:alignment} and Section \ref{sec:recognition}.
-For CPU implementation, the system would make use of all the cores in as many CPUs are present on one motherboard.  For GPU implementation, the system would
-make the best use of a single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with a single GPU chip per PCI-express card).
-Unless otherwise specified, all implementations utilize single
-precision floating point datatypes.
+\section{Hardware Parallelism} \label{sec:parallelism} 
+In this section we discuss the hardware architectures used in this paper. Aside
+from the $\ell_1$-min algorithm, our choice of the architectures plays a
+significant role in many practical decisions we make in Section
+\ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
+the system would make use of all the cores in as many CPUs are present on one
+motherboard.  For GPU implementation, the system would make the best use of a
+single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with
+a single GPU chip per PCI-express card).  Unless otherwise specified, all
+implementations utilize single precision floating point datatypes.
 
 \subsection{CPU Hardware Parallelism} 
-The baseline architecture for our experiments is a Linux workstation with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each
-processor has its own memory interface, and is directly connected to half of
-the RAM installed in the machine.  The amount of RAM installed exceeds the
-amount used by the algorithms, and is not an important performance
-consideration.
+The baseline architecture for our experiments is a Linux workstation with two
+quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
+has its own memory interface, and is directly connected to half of the RAM
+installed in the machine.  The amount of RAM installed exceeds the amount used
+by the algorithms, and is not an important performance consideration.
 
@@
 
-There are thus two important levels of parallelism that need to be exploited to efficiently
-use a modern CPU: core-level parallelism and vector-level
+There are thus two important levels of parallelism that need to be exploited to
+efficiently use a modern CPU: core-level parallelism and vector-level
 parallelism.  For the implementations discussed here, we will primarily use the
-OpenMP programming model to express core-level parallelism {\bf AY: What is OpenMP? citation needed}.  Depending on the
-section of the code{\bf AY: What is a section of the code?}, we will achieve vector-level parallelism implicitly
+OpenMP programming model to express core-level parallelism {\bf AY: What is
+OpenMP? citation needed}.  Depending on the section of the code{\bf AY: What is
+a section of the code?}, we will achieve vector-level parallelism implicitly
 through either the auto-vectorization features of Intel's icc compiler, or
-through the use of a highly optimized BLAS library (Intel's MKL). {\bf AY: citations needed, what is icc, what is the relation between BLAS and MKL? Are they equivalent? not clear in the context.}
+through the use of a highly optimized BLAS library (Intel's MKL). {\bf AY:
+citations needed, what is icc, what is the relation between BLAS and MKL? Are
+they equivalent? not clear in the context.}
 
-Each processor further has 8MB of L3 cache that is shared by all four cores.  Overall,
-this means the algorithm has approximately 16MB of on-chip cache available.
-For the low resolution images used in sparse representation based face
-recognition, this is enough cache to hold over 800 image-sized variables, {\bf AY: how low resolution? which sparse representation problem? ``image-sized variables'' means simply ``image vectors''?}
-meaning that for small data sets it is possible to run completely using on-chip
-cache.
+Each processor further has 8MB of L3 cache that is shared by all four cores.
+Overall, this means the algorithm has approximately 16MB of on-chip cache
+available.  For the low resolution images used in sparse representation based
+face recognition, this is enough cache to hold over 800 image-sized variables,
+{\bf AY: how low resolution? which sparse representation problem? ``image-sized
+variables'' means simply ``image vectors''?} meaning that for small data sets
+it is possible to run completely using on-chip cache.
 
 \subsection{GPU Hardware Parallelism}
-%High-level description of GPU parallelism, i.e.
 
-The ALM algorithm requires many matrix-vector operations, this gives many-core processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi) which each contains 32 floating-point units.
-Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.
-The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called \emph{warps}.
-These warps are then grouped into higher levels called \emph{thread blocks} and \emph{grids}.
-The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance and a computational advantage over CPUs.
-Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and 768kb of L2 cache.
-However, this increase in performance does not come for free.
-As the GPU has its own memory system, any data the GPU uses must be transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
-However, the cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.
+The ALM algorithm requires many matrix-vector operations, this gives many-core
+processors such as GPUs a computational advantage for large problem sizes.
+GPUs are comprised of several streaming multiprocessors (SMP) (16 for the
+Nvidia Fermi) which each contains 32 floating-point units.  Each SMP contains
+its own L1-cache and shared memory and all SMPs share a common L2-cache.  The
+GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups
+of 32 threads called \emph{warps}.  These warps are then grouped into higher
+levels called \emph{thread blocks} and \emph{grids}. The GPU contains an
+inverted memory hierarchy so that the L1 cache stays close to the cores,
+providing the increase in performance and a computational advantage over CPUs.
+Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and
+768kb of L2 cache. However, this increase in performance does not come for
+free. As the GPU has its own memory system, any data the GPU uses must be
+transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
+However, the cost of these transfers are amortized by doing large amounts of
+computation at once, requiring code to be run completely on the GPU.
 
-In summary, a programmable GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.
-Each thread block is a virtualized multiprocessor, which each thread as a virtualized scalar processor.
-All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.
+In summary, a programmable GPU can be analogous to a multi-core CPU processor,
+except with more sophisticated vector processing hardware.  Each thread block
+is a virtualized multiprocessor, which each thread as a virtualized scalar
+processor.  All of these differences allow GPUs to perform intensive
+calculations such as matrix-vector operations at much faster rates than CPUs.
 
-% This section is Drew's
 \section{Parallelization of the Face Alignment Stage}

### 2011-02-18 paper/main.tex
Added paper ID
revised section 3 GPU implementation started section 6

@@
 \documentclass[10pt,twocolumn,letterpaper]{article}
+% ICCV Rules:
+% 8 page limit (6 free + 2 paid)
+% 10MB pdf file size limit
+% 30MB supplementary materials file size limit (PDF or ZIP)
 
@@
 
-\def\iccvPaperID{****} % *** Enter the CVPR Paper ID here
+\def\iccvPaperID{257} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 %\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
-\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
+%\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
+\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition
 %\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}
@@
 scenario is not the focus of this paper. The source code of our algorithms,
-which is based on BLAS and Nvidia CUDA libraries, will be made available on our
+which is based on CUBLAS and Nvidia CUDA libraries, will be made available on our
 website to aid peer evaluation.
@@
 processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors (SMP) (16 for the
-Nvidia Fermi) which each contains 32 floating-point units.  Each SMP contains
+GPUs are comprised of several streaming multiprocessors (SMP) (15 for our
+NVIDIA Fermi) which each contains 32 floating-point units.  Each SMP contains
 its own L1-cache and shared memory and all SMPs share a common L2-cache.  The
@@
 of 32 threads called \emph{warps}.  These warps are then grouped into higher
-levels called \emph{thread blocks} and \emph{grids}. The GPU contains an
-inverted memory hierarchy so that the L1 cache stays close to the cores,
-providing the increase in performance and a computational advantage over CPUs.
-Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and
+levels called \emph{thread blocks} and \emph{grids}.
+
+Fermi's inverted memory hierarchy (aggregate l1 capacity is greater than l2 capacity) is designed to support massively parallel workloads in which there is more task-private data than shared data, providing an increase in performance and a computational advantage over CPUs.
+For our system, each SMP has 64 kb of L1 cache with a total of 960 kb of L1 cache/shared memory and
 768kb of L2 cache. However, this increase in performance does not come for
@@
 However, the cost of these transfers are amortized by doing large amounts of
-computation at once, requiring code to be run completely on the GPU.
+computation at once, requiring code to be run completely on the GPU.  In 2011, NVIDIA is developing a new ARM-based CPU system, dubbed 
+"Project Denver" to fix performance issues such as the limited bandwidth due to PCI-E \cite{Dally-Web}.
 
@@
 processor.  All of these differences allow GPUs to perform intensive
-calculations such as matrix-vector operations at much faster rates than CPUs.
+calculations such as large scale matrix-vector operations at much faster rates than CPUs.
 
@@
 {\bf AY: Add an algorithm here describe the recognition block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
+{\bf VS: Why did we modify the algorithm for the recognition stage? }
+
+\begin{algorithm}[t]
+\caption{\bf (Augmented Lagrange Multiplier Method for
+Recognition)}
+\begin{algorithmic}[1]
+\begin{small}
+\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
+\mathbf{0}$.
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
+\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
+\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
+\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
+\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
+\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
+\sqrt{1+4t_l^2}\right)$;
+\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
+\ENDWHILE
+\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
+\ENDWHILE \STATE
+{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+\end{small}
+\end{algorithmic}
+\label{alg:alm_rec} 
+\end{algorithm}
 
@@
 {\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
+The CPU ALM algorithm was written in C and compiled with Intel's MKL libraries, which provide
+automatic parallelization.
 
 \subsection{GPU Implementation}
-{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
+{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a GPU architecture.}
 The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
@@
 
+In this section, we present four experiments to benchmark the performance of our algorithm on CPU and GPU platforms. 
+To generate a meaningful comparison between different hardware architectures,
+in all cases we compare the performance of hardware architectures on a
+per-board basis:  For CPU implementations, the benchmark makes use of all of
+the cores in as many cpu's are present.  For GPU implementations, the benchmark
+makes the best use of the entire GPU chip (most GPU boards have a single GPU
+chip).  Unless otherwise specified, all implementations utilize single
+precision floating point datatypes.
+
+The first experiment is designed to compare the two implementations of PALM
+that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm}).  
+The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
+randomly generated from the standard normal and normalized with unit norm.  The ratio of $d/n$ varied from
+$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
+The ground truth signals, $x_0$ and $y_k$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
+The measurement vector was generated with the equation $\y_0 = A \x_0 + \y_k$.    We set the number of inner iterations at 50 and outer iterations at 5000.
+We also varied the different inner and tolerance levels from 1e-3 to 1e-6.
+We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.
+
+The x set of experiments will compare the CPU and GPU versions of the entire pipeline, 
+which includes the alignment and recognition stage.
+
+We pl
+
 \subsection{Simulation}
+
+We plot the resulted of the simulation benchmark below.  The x-axis represents the size of the 
+$A$ and $A'A$ matrix and the y-axis represents the average amount of time it to complete
+a single problem instance.
+
+The observations of this experiment are summarized below:
+1.  With similar accuracy rates, the GPU implementation runs faster than the CPU when
 

### 2011-02-18 paper/main.tex
fixed algorithm alm_rec and updated some numbers

@@
 processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors (SMP) (15 for our
+GPUs are comprised of several streaming multiprocessors (SMP) (14 for our
 NVIDIA Fermi) which each contains 32 floating-point units.  Each SMP contains
@@
 Fermi's inverted memory hierarchy (aggregate l1 capacity is greater than l2 capacity) is designed to support massively parallel workloads in which there is more task-private data than shared data, providing an increase in performance and a computational advantage over CPUs.
-For our system, each SMP has 64 kb of L1 cache with a total of 960 kb of L1 cache/shared memory and
+For our system, each SMP has 64 kb of L1 cache with a total of 896 kb of L1 cache/shared memory and
 768kb of L2 cache. However, this increase in performance does not come for
@@
 
+Below is the psuedocode solving the minimization problem posed in \eqref{eq:l1min_denoise}.
+
 \begin{algorithm}[t]
@@
 \begin{small}
-\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
+\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\e_1 = \bb$, $\y_1 =
 \mathbf{0}$.
 \WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
-\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
+\STATE $\e_{k+1} = \textup{shrink}\left(\bb - A\x_k +
+\frac{1}{\mu}\y_k, \frac{1}{\mu}\right)$;
 \STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
@@
 \STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
+\frac{1}{\gamma}A^T\left(\bb - A\z_l - \e_{k+1} +
+\frac{1}{\mu}\y_k\right), \frac{1}{\mu\gamma}\right)$;
 \STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
@@
 \ENDWHILE
-\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
+\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\y_{k+1} \leftarrow \y_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
 \ENDWHILE \STATE
@@
 addition, we coded kernel (GPU) methods for vector-vector operations to replace
-operations that would have taken multiple BLAS calls.
-Because data transfer from the RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.
-The algorithm is similar
-to the CPU algorithm except that we had to parallelize the code at an
+operations that would have taken multiple BLAS calls. Because data transfer from the 
+RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.  
+The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an
 additional level. The CPU portion of the code serves as the conductor, making
@@
 The first experiment is designed to compare the two implementations of PALM
-that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm}).  
+that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm_rec}).  
 The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
-randomly generated from the standard normal and normalized with unit norm.  The ratio of $d/n$ varied from
+randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
 $10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
@@
 
-We pl
-
 \subsection{Simulation}
@@
 We plot the resulted of the simulation benchmark below.  The x-axis represents the size of the 
-$A$ and $A'A$ matrix and the y-axis represents the average amount of time it to complete
+$A$ and $A^TA$ matrix and the y-axis represents the average amount of time it to complete
 a single problem instance.

### 2011-02-19 paper/main.tex
Applied already discussed changes to the intro.

--- 2011-02-18 paper/main.tex
@@
 years due to important applications in compressive sensing
-\cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  The
-problem refers to finding the minimum $\ell$-norm solution to an
+\cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  
+$\ell_1$-min refers to finding the minimum $\ell$-norm solution to an
 underdetermined linear system $\bb=A\xx$:
@@
 \end{equation}
-It has now been well understood that, under certain conditions
+It is now well established that, under certain conditions
 \cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is also
@@
 If the sparsest solutions for $\x$ and $\e$ are recovered, $\ee$ provides a
-means to compensate corrupted pixels that main occlude some part of the query
-image, and the dominant nonzero coefficients in $\xx$ reveals the membership of
-$\bb$ based on the training image labels associated with $A$.
+means to compensate for pixels that are corrupted due to occlusion of some part of the query
+image, and the dominant nonzero coefficients of $\xx$ reveal the membership of
+$\bb$ based on the training image labels associated with $A$.  Since $A$ ideally
+contains many images per user taken under different illuminations, 
+$A\x$ also acts as a linear illumination model for the test image $\bb$.
 
 In this paper, we study parallelization of $\ell$-min on many-core CPU and GPU
-architectures. Firstly, although $\ell_1$-min \eqref{eq:l1min} is a convex
+architectures. Although $\ell_1$-min \eqref{eq:l1min} is a convex
 program, conventional algorithms such as interior-point methods
@@
 for large-scale applications such as face recognition. Recently, a number of
-accelerated algorithms have been proposed, which explicitly take advantage of
+accelerated algorithms have been proposed that explicitly take advantage of
 the special structure of $\ell_1$-min problems
@@
 state-of-the-art $\ell_1$-min solution based on the classical framework of
-\emph{augmented Lagrange multiplier} (ALM)
-\cite{BertsekasD2003,YangA2010-ICIP}.
+\emph{augmented Lagrange multiplier} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
 
-Secondly, while solving \eqref{eq:l1min_denoise} can achieve exceptional
+While solving \eqref{eq:l1min_denoise} can achieve exceptional
 recognition accuracy on public databases, its success relies on the assumption
@@
 (e.g., affine or homography) acting on the image domain, which aligns $\bb$
-with the training images from the $i$-th class $A_i$. The algorithm essentially
-extends Lucas-Kanade iterative alignment \cite{LucasB1981} to using the
-$\ell_1$ norm as a robust error function.  Although the original solution
-\cite{WagnerA2009-CVPR} was implemented based on standard interior-point
-methods, its speed has been improved by ALM in a more recent paper
-\cite{WagnerA2011-PAMI}. In this paper, in addition to solving the generic
-$\ell_1$-min objectives \eqref{eq:l1min} or \eqref{eq:l1min_denoise}, we also
-discuss how to properly customize the implementation to accelerate the face
-alignment step \eqref{eq:l1min_alignment} on many-core CPUs and GPUs.
+with the training images from the $i$-th class $A_i$. The algorithm 
+extends Lucas-Kanade iterative alignment \cite{LucasB1981} to both to use the
+$\ell_1$ norm as a robust error function, and also simultaneously 
+to estimate the illumination model $A\x$.  
+The original solution \cite{WagnerA2009-CVPR} was implemented based 
+on standard interior-point
+methods, and its speed has been since improved by ALM in a more recent paper
+\cite{WagnerA2011-PAMI}. However, primarily due the high per-user cost
+of the alignment step, the recognition system was still not fast enough
+for real-time recognition against datasets of hundreds of users.
+In this paper, in addition to solving the generic
+$\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
+discuss how to properly accelerate the iterative face alignment step
+\eqref{eq:l1min_alignment} on multi-core CPUs and GPUs.
 
@@
 recognition system has been fully implemented on a state-of-the-art
-workstation, as shown in Figure \ref{fig:pipeline}. Since any algorithm
-parallelization is highly dependent on the specific computer hardware, the
-architecture of our test system will be explained in detail in Section
-\ref{sec:parallelism}.  We show that a highly accurate and robust face
-recognition system can be constructed to recognize face images from a moderate
-database of several hundred subjects. Note that with the possible continual
-increase of the number of subject classes, it is possible to modularize and
-dispatch different stages of the face recognition routine to a network of
-workstations to form an even larger scale networked system. Nevertheless, this
-scenario is not the focus of this paper. The source code of our algorithms,
-which is based on CUBLAS and Nvidia CUDA libraries, will be made available on our
-website to aid peer evaluation.
+workstation, as shown in Figure \ref{fig:pipeline}. 
+Since the parallelization of any algorithm
+is highly dependent on the target hardware, we will give a brief overview of
+the architecture of our test system in detail in Section
+\ref{sec:parallelism}.  We will show that a highly accurate and robust face
+recognition system can be constructed to recognize face images from a 
+database of several hundred subjects quickly enough to be practical for
+access control applications. 
+Note that for applications targeting thousands of users or more, it would likely be
+necessary to distribute the work over a network of machines; this paper
+focuses on the (shared memory) parallelism available in a single machine.
+The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture
+will be made available for evaluation by other researchers.
 \begin{figure}
@@
 {\tiny \input{figures/pipeline_simplified.tex}}
-\caption{The face recognition pipeline implemented on massively parallel hardware.}
+\caption{Overview of the face recognition pipeline.}
 \label{fig:pipeline}
@@
 BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
-interior path, which is often computationally expensive. To mitigate these
+interior path{\bf DW: Following the interior path is what made them faster than
+traditional linear programming techniques that follow the edge of the feasible set;
+does Homotopy actually go outside the feasible set?  Just want to check if this is
+an accurate representation...}, which is often computationally expensive. To mitigate these
 issues, an approach called \emph{Homotopy} has been studied to accelerate the
@@
 Although Homotopy can be shown to exactly estimate BP when the solution is
-sufficiently sparse \cite{DonohoD2006}, the algorithm still involves expensive
-matrix operations such as matrix-matrix multiplication and solving
-least-squares problems with varying $A$ matrices. Hence, it is not efficient to
-implement Homotopy on parallel hardware. In Section \ref{sec:ALM}, we contend
+sufficiently sparse \cite{DonohoD2006}, the algorithm still involves 
+computationally expensive operations such as matrix-matrix multiplication and 
+the solution of least-squares problems with varying $A$ matrices. 
+In Section \ref{sec:ALM}, we contend
 that ALM is a better choice among Homotopy and other existing $\ell_1$-min
-solvers to parallelize on many-core CPUs and GPUs. The algorithm belongs to a
+solvers for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
 category of approximate $\ell_1$-min solutions called \emph{iterative
-shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}. The
-(sequential) implementation of IST algorithms mainly involves elementary
-operations such as vector algebra and matrix-vector multiplication. Therefore,
+shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}.
+IST algorithms mainly involves elementary operations such as vector 
+algebra and matrix-vector multiplication. Therefore,
 when the dimension of the problem becomes high, IST-type algorithms are

### 2011-02-19 paper/main.tex
revisions to section 6. added a citation to bib

--- 2011-02-18 paper/main.tex
@@
 whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
-Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm.
+Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm \footnote{While we
+present the ALM algorithm in the primal domain, there also
+exists an implementation ALM in the dual domain.}.
 
@@
 
-Below is the psuedocode solving the minimization problem posed in \eqref{eq:l1min_denoise}.
+Below is the algorithm for the face recognition stage in Figure \ref{fig:pipeline}, which
+solves the $\ell_1$ minimization problem posed in \eqref{eq:l1min_denoise}.
 
 \begin{algorithm}[t]
-\caption{\bf (Augmented Lagrange Multiplier Method for
-Recognition)}
+\caption{\bf (Face Recognition via ALM)}
 \begin{algorithmic}[1]
@@
 operations that would have taken multiple BLAS calls. Because data transfer from the 
-RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.  
+RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector once.  
 The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an
@@
 chip).  Unless otherwise specified, all implementations utilize single
-precision floating point datatypes.
+precision floating point datatypes.  For all of the experiments, we set the number 
+of inner loop iterations at 50 and outer loop iterations at 5000.
 
 The first experiment is designed to compare the two implementations of PALM
-that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm_rec}).  
+that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm_rec}) by
+comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
+vector $\xx_0$.  
 The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
@@
 $10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
-The ground truth signals, $x_0$ and $y_k$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
-The measurement vector was generated with the equation $\y_0 = A \x_0 + \y_k$.    We set the number of inner iterations at 50 and outer iterations at 5000.
+The ground truth signals, $\xx_0$ and $\yy_k$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
+Because the ground truth signal $\xx_0$ and $yy_k$ are known, the algorithm terminates
+when $norm(\bx-\bx_0) < tolerance$.
+The measurement vector was generated with the equation $\y_0 = A \x_0 + \y_k$.   
 We also varied the different inner and tolerance levels from 1e-3 to 1e-6.
@@
 
-The x set of experiments will compare the CPU and GPU versions of the entire pipeline, 
-which includes the alignment and recognition stage.
+%The x set of experiments will compare the CPU and GPU versions of the entire pipeline, 
+%which includes the alignment and recognition stage.
+
+The second experiment is to compare the amount of time to solve the face alignment
+stage for each implementation.  Because the alignment stage is not affected by the resolution of the image,
+we fix the resolution at 60x80.
+{\bf VS:  this accuracy number depends on the number of subjects in the A matrix right?  So we either 
+need to fix the number of subjects to compare against or vary it...}
+
+The third experiment is to compare accuracy rate of the face recognition algorithm at different
+image resolutions.
+
+The fourth experiment is to compare the speed of the recognition stage w.r.t the number of 
+subjects in the matrix $A$.
+
+Our final experiment is to compare the speeds of CPU and GPU implementations entire pipeline in Figure \ref{fig:pipeline}.
 

### 2011-02-19 paper/main.tex
Wrote section describing iterative alignment
computation, including a couple new algorithms.
Also added back in my discussion of computational
complexity.

--- 2011-02-19 paper/main.tex
@@
 \begin{equation}
-\hat{\tau}_i = \arg\min_{\xx, \ee, \tau_i} \|\ee\|_1\quad \mbox{subj. to}\quad \bb\circ\tau_i = A_i\xx + \ee,
+\hat{\tau}_i = \arg\min_{\xx, \ee, \tau_i} \|\ee\|_1\quad \mbox{subj. to}\quad \tilde\bb\circ\tau_i = A_i\xx + \ee,
 \label{eq:l1min_alignment}
@@
 where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations
-(e.g., affine or homography) acting on the image domain, which aligns $\bb$
+(e.g., affine or homography) acting on the image domain, which aligns $\tilde\bb$
 with the training images from the $i$-th class $A_i$. The algorithm 
@@
 \label{sec:alignment}
-{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
+%{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
+This section provides a brief explanation of the iterative alignment stage
+of the recognition pipeline.  
+The inputs of the algorithm are the test image $f(x,y)$ stacked into an $M$-
+dimensional vector $\f$, and the aligned and stacked training images $A_i$ for
+each training user $i$. 
+The algorithm repeatedly solves a linearized version of the minimization problem
+in \eqref{eq:l1min_alignment}.  This involves two computational steps:
+the computation of the linearization, and the solution of the minimization problem.
+
+The computation of the linearization is complicated by the need to avoid trivial 
+solutions where the algorithm
+progressively zooms in on a black pixel of the test image.
+This is achieved by minimizing the $\ell_1$ error between the illumination model $Ax$ and the cropped,
+resampled, and normalized test image $\tilde\bb$, rather than the un-normalized version as in 
+conventional Lucas-Kanade iterative alignment.
+This involves a bit of calculus:
+\begin{equation}
+\frac{\partial}{\partial \tau_p} \frac{\bb}{||\bb||} 
+= \frac{1}{||\bb||} \frac{\partial \bb}{\partial \tau_p} - 
+\frac{\bb}{||\bb||^3} \bb^T \frac{\partial \bb}{\partial \tau_p}
+\end{equation}
+If the warping $\tau$ is constrained to be a similarity transform,
+it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
+reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
+and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
+once per test image. {bf DW: I could go into the rest of the math for this, but I'm worried about space}
+
+The iterative alignment stage is summarized as Algorithm \ref{alg:iterative_alignment}.
+\begin{algorithm}[thb]
+\caption{\bf (Iterative Alignment Stage)} \label{alg:iterative_alignment}
+\begin{algorithmic}[1]
+\begin{small}
+\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
+$\f\in\Re^M$ 
+\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
+\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
+\STATE {\bf for} each gallery subject $i$, 
+\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
+\STATE \hspace{6mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
+\STATE \hspace{6mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
+\STATE \hspace{6mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
+\STATE \hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
+\tau$; 
+\end{small}
+\end{algorithmic}
+\end{algorithm}
+
+The solution of the minimization problem in Algorithm
+\ref{alg:iterative_alignment} Step 8 can be solved via a
+variant of ALM, which is summarized as Algorithm \ref{alg:alignment_alm}.
+$\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of
+parameter $\mu$, we take the same strategy as in \cite{YangJ2009-pp} and set
+$\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
+
+\begin{algorithm}[h]
+\caption{\bf (Augmented Lagrange Multiplier Method For Alignment)}
+\begin{algorithmic}[1]
+\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
+$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
+\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
+\ENDWHILE
+\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
+\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
+\ENDWHILE \STATE
+{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
+\end{algorithmic}
+\label{alg:alignment_alm}
+\end{algorithm}
+
+\subsection{Computational Complexity} Any discussion of the speed of an
+algorithm should discuss the comutational complexity of the algorithm.  For the
+alignment stage ALM algorithm \ref{alg:alignment_alm}, this is rather simple.  The
+Matrix-vector computions are $O(mn)$, and the vector-vector operations are
+$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
+very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
+computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
+A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
+Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
+of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
+Fortunately, $A^\dagger$ only needs to be computed once per instance of the
+$\ell_1$ minimization problem.
+
+Therefore, for problems with large $n$, we expect the computation of the
+pseudoinverse to dominate.  However, for the problem sizes needed in face
+recognition problems, we find that the computation time is dominated by the
+repeated computation of matrix-vector and matrix-matrix multiplications in the
+inner loop.
+
+%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
 

### 2011-02-22 results/random_data/bench_PALM.m
adding benchmarks for algorithm 1: the generic PALM benchmark. averageIterations and averageL2error for BLAS and GPU are equal!  which is awesome!

@@ -1,6 +1,6 @@
 function rtn = bench_PALM()
 
-type = 'PALM';
+type = 'PALM1';
 n  = 10; % denominator for ratio
 ks = 0.1;
 tols = [1e-4 1e-5];
@@ -19,49 +19,45 @@ for i=ratio
     
     gpuData = load(gpuName);
     
-    cpuName = [name 'CPU.mat'];
-    
-    cpuData = load(cpuName);
-    
     blasName = [name 'BLAS.mat'];
     
     blasData = load(blasName);
     
-    gpuDimsPowers  = gpuData.dimsPowers(gpuData.dimsPowers <= 6000);
-    gpuDimLen      = numel(gpuDimsPowers);
-    cpuDimsPowers  = cpuData.dimsPowers(cpuData.dimsPowers <= 6000);
-    cpuDimLen      = numel(cpuDimsPowers);
-    blasDimsPowers = blasData.dimsPowers(blasData.dimsPowers <= 6000);
-    blasDimLen     = numel(blasDimsPowers);
+    gpuDimsPowers  = gpuData.dimsPowers;
+    gpuDimLen      = numel(gpuData.dimsPowers <= 6000);
+    blasDimsPowers = blasData.dimsPowers;
+    blasDimLen     = numel(blasData.dimsPowers <= 6000);
     
-    gpuDimsPowers  = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
-    cpuDimsPowers  = (cpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
-    blasDimsPowers = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
-    
+    gpuASize       = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    blasASize      = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+        
     for j=1:numel(tols)
         
         subplot(numel(tols), 3, (j-1)*3+1);
-        gpuOutIter  = gpuData.averageIteration(j,j,:,2);
-        gpuInIter   = gpuData.averageTotalInnerIteration(j,j,:,2);
-        cpuOutIter  = cpuData.averageIteration(j,j,:,2);
-        cpuInIter   = cpuData.averageTotalInnerIteration(j,j,:,2);
-        blasOutIter = blasData.averageIteration(j,j,:,2);
[…]

### 2011-02-22 paper/main.tex
revised the synthetic experiment description.  now we benchmark the generic ALM algorithm for the 1st benchmark

@@
 The first experiment is designed to compare the two implementations of PALM
-that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm_rec}) by
+that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm}) by
 comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
@@
 $10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
-The ground truth signals, $\xx_0$ and $\yy_k$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
-Because the ground truth signal $\xx_0$ and $yy_k$ are known, the algorithm terminates
+The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
+Because the ground truth signal $\xx_0$ is known, the algorithm terminates
 when $norm(\bx-\bx_0) < tolerance$.
-The measurement vector was generated with the equation $\y_0 = A \x_0 + \y_k$.   
+The measurement vector was generated with the equation $\y_0 = A \x_0$.   
 We also varied the different inner and tolerance levels from 1e-3 to 1e-6.

### 2011-02-25 results/random_data/bench_PALM_paper.m
removed old images. adding benchmarks that compare  speed, l2error to ground truth, number of iterations of the gpu and cpu code

@@ -1,0 +1,109 @@
+function rtn = bench_PALM_paper()
+
+type = 'PALM1';
+n  = 10; % denominator for ratio
+ks = 0.1;
+maxIter = 5000;
+iterations = 10;
+dimsPowers = [500:100:900 1000:1000:10000];
+lambda = 10;
+ratio = 1:5;
+%tols = gpuData.tols;
+%tol_ints = gpuData.tol_ints;
+tols = [1e-2 1e-3];
+tol_ints = [1e-2 1e-3 1e-4];
+
+h = figure;
+
+for i=ratio
+    name = [type '_' num2str(n) '_' num2str(i) '_'];
+    
+    gpuName = [name 'GPU.mat'];
+    
+    gpuData = load(gpuName);
+    
+    blasName = [name 'BLAS.mat'];
+    
+    blasData = load(blasName);
+
+    gpuDimsPowers  = gpuData.dimsPowers;
+    gpuDimLen      = numel(gpuData.dimsPowers <= 6000);
+    blasDimsPowers = blasData.dimsPowers;
+    blasDimLen     = numel(blasData.dimsPowers <= 6000);
+    
+    gpuASize       = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    blasASize      = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    
+    for j=1:numel(tols)
+        for k=1:numel(tol_ints)
+            
+            subplot(numel(tol_ints), 3, (k-1)*3+1);
+            gpuOutIter  = squeeze(gpuData.averageIteration(j,k,:,2))';
+            gpuInIter   = squeeze(gpuData.averageTotalInnerIteration(j,k,:,2))';
+            blasOutIter = squeeze(blasData.averageIteration(j,k,:,3))';
+            blasInIter  = squeeze(blasData.averageTotalInnerIteration(j,k,:,3))';
+            
+            % Calculations for totalSize (aka the number of times we use each
+            % matrix * size of the matrix
+            % A is mxn.  G is nxn.  (3*m*n) * totalOutIterations + (n^2)
+            % * total Inner iterations.  Is that correct?
+            gpuTotalSize    = gpuDimsPowers.^2 .* ((3 * (i/10) .* gpuOutIter)  + (i/10)^2 .* gpuInIter) * 32 / (2^30);
[…]

### 2011-02-25 results/random_data/bench_PALM.m
adding image for the paper.

@@ -1,14 +1,17 @@
-function rtn = bench_PALM()
+function rtn = bench_PALM_paper()
 
 type = 'PALM1';
 n  = 10; % denominator for ratio
 ks = 0.1;
-tols = [1e-4 1e-5];
 maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];
 lambda = 10;
 ratio = 1:5;
+%tols = gpuData.tols;
+%tol_ints = gpuData.tol_ints;
+tols = [1e-2 1e-3];
+tol_ints = [1e-2 1e-3 1e-4];
 
 h = figure;
 
@@ -30,80 +33,91 @@ for i=ratio
     
     gpuASize       = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
     blasASize      = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
-        
+    
     for j=1:numel(tols)
-        
-        subplot(numel(tols), 3, (j-1)*3+1);
-        gpuOutIter  = squeeze(gpuData.averageIteration(j,j,:,2))';
-        gpuInIter   = squeeze(gpuData.averageTotalInnerIteration(j,j,:,2))';
-        blasOutIter = squeeze(blasData.averageIteration(j,j,:,3))';
-        blasInIter  = squeeze(blasData.averageTotalInnerIteration(j,j,:,3))';
-        
-        % Calculations for totalSize (aka the number of times we use each
-        % matrix * size of the matrix
-        % A is mxn.  G is nxn.  (3*m*n) * totalOutIterations + (n^2)
-        % * total Inner iterations.  Is that correct?
-        gpuTotalSize    = gpuDimsPowers.^2 .* ((3 * (i/10) .* gpuOutIter)  + (i/10)^2 .* gpuInIter) * 32 / (2^30);
-        blasTotalSize   = blasDimsPowers.^2 .* ((3 * (i/10) .* blasOutIter)  + (i/10)^2 .* blasInIter) * 32 / (2^30);
-
-        plotA = squeeze(gpuData.averageRunTime(j,j,:,2,2))';
-        plotC = squeeze(blasData.averageRunTime(j,j,:,3,2))';
-        plot(gpuASize(1:gpuDimLen), gpuTotalSize(1:gpuDimLen)./plotA(1:gpuDimLen), 'bx-');
-        hold on
[…]

### 2011-02-25 results/random_data/bench_PALM_paper.m

@@ -15,7 +15,7 @@ tol_ints = [1e-2 1e-3 1e-4];
 
 h = figure;
 
-for i=ratio
+for i=2:3
     name = [type '_' num2str(n) '_' num2str(i) '_'];
     
     gpuName = [name 'GPU.mat'];
@@ -25,85 +25,34 @@ for i=ratio
     blasName = [name 'BLAS.mat'];
     
     blasData = load(blasName);
-
+    
     gpuDimsPowers  = gpuData.dimsPowers;
     gpuDimLen      = numel(gpuData.dimsPowers <= 6000);
     blasDimsPowers = blasData.dimsPowers;
     blasDimLen     = numel(blasData.dimsPowers <= 6000);
     
-    gpuASize       = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
-    blasASize      = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    subplot(1, 2, i-1);
+    plotA = squeeze(gpuData.averageRunTime(2,1,:,2,2))';
+    plotC = squeeze(blasData.averageRunTime(2,1,:,3,2))';
+    plot(gpuDimsPowers(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
+    hold on
+    plot(blasDimsPowers(1:blasDimLen), plotC(1:blasDimLen), 'g+-');
+    title(['Tolerance 1e-3, Inner Tolerance 1e-2']);
+    ylabel('Elapsed time (s)');
+    xlabel(['A matrix dimension (m x 0.' num2str(i) 'm)']);
+    hold off
+    legend('gpu', 'cpu');
     
-    for j=1:numel(tols)
-        for k=1:numel(tol_ints)
-            
-            subplot(numel(tol_ints), 3, (k-1)*3+1);
-            gpuOutIter  = squeeze(gpuData.averageIteration(j,k,:,2))';
-            gpuInIter   = squeeze(gpuData.averageTotalInnerIteration(j,k,:,2))';
-            blasOutIter = squeeze(blasData.averageIteration(j,k,:,3))';
-            blasInIter  = squeeze(blasData.averageTotalInnerIteration(j,k,:,3))';
-            
-            % Calculations for totalSize (aka the number of times we use each
-            % matrix * size of the matrix
-            % A is mxn.  G is nxn.  (3*m*n) * totalOutIterations + (n^2)
-            % * total Inner iterations.  Is that correct?
[…]

### 2011-02-26 paper/figures/user_alignment_rank_histogram.m
added another plot which plots percentage that user is in training set vs number of users in training set

@@ -1,0 +1,24 @@
+load('sorted_user_ranks.dat');
+
+h = figure;
+[a, x] = hist(sorted_user_ranks, 100);
+a = a / sum(a);
+bar(x,a);
+
+b = [a(1)];
+for i=2:numel(a)
+    b(i) = a(i) + b(i-1);
+end
+plot(x,b);
+
+xlabel('Number of users with lower residual that test user');
+ylabel('Percentage that user lies in training set');
+title('Percentage that usre lies in training set based on alignment residuals')
+
+set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
+set (gcf,'Units','Inches');                     %   Using units of inches
+pos=get(gcf,'position');                        %   Save position values
+set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
+set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
+
+plot2svg(['user_alignment_rank_plot.svg'], h, 'png');

### 2011-02-26 paper/figures/user_alignment_rank_histogram.m
new plots for reduced training set plot

@@ -9,11 +9,11 @@ b = [a(1)];
 for i=2:numel(a)
     b(i) = a(i) + b(i-1);
 end
-plot(x,b);
+b = b;
+plot(x+1,b);
 
-xlabel('Number of users with lower residual that test user');
-ylabel('Percentage that user lies in training set');
-title('Percentage that usre lies in training set based on alignment residuals')
+xlabel('Number of users in reduced training set');
+ylabel('Probability that user lies in reduced training set');
 
 set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
 set (gcf,'Units','Inches');                     %   Using units of inches

### 2011-02-27 paper/main.tex
adding some experiment writeups

@@
 The measurement vector was generated with the equation $\y_0 = A \x_0$.   
-We also varied the different inner and tolerance levels from 1e-3 to 1e-6.
+We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
 We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.
@@
 we fix the resolution at 60x80.
-{\bf VS:  this accuracy number depends on the number of subjects in the A matrix right?  So we either 
-need to fix the number of subjects to compare against or vary it...}
 
-The third experiment is to compare accuracy rate of the face recognition algorithm at different
-image resolutions.
+Since we use a reduced training set for the recognition stage, our third experiment was to 
+find the relationship between the probability that the true subject is in the training set vs the
+ number of users in the reduced training set and the speed of the recognition stage w.r.t.
+ the number of subjects in the training set.
 
-The fourth experiment is to compare the speed of the recognition stage w.r.t the number of 
-subjects in the matrix $A$.
+The fourth experiment is to compare accuracy rate of the face recognition algorithm at different
+image resolutions.  We fixed the number of users to be $enter here$ which was shown in figure  $enter here$  have 95\% probability
+to have the true user in the training set, inner loop tolerance to 1e-2 and outer loop tolerance to 1e-3.
 
@@
 
-We plot the resulted of the simulation benchmark below.  The x-axis represents the size of the 
-$A$ and $A^TA$ matrix and the y-axis represents the average amount of time it to complete
-a single problem instance.
+We plot the most significant results of the simulation benchmark below.  The x-axis represents the size of the 
+$A$ matrix and the y-axis represents the average amount of time it to complete a single problem instance.
 
 The observations of this experiment are summarized below:
-1.  With similar accuracy rates, the GPU implementation runs faster than the CPU when
+1.  With similar accuracy rates, the GPU implementation runs faster than the CPU at the crossover points 
+noted on the plots.
+2.  From the results, it was seen that tolerances...
+
+\subsection{Face Alignment Benchmark}
+
+\subsection{Face Recognition Benchmarks}
 

### 2011-03-01 paper/main.tex
Fixed compile bug; hyperref needs to be uspackaged before algorithm

@@
 \usepackage{iccv}
+\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
 \usepackage{times}
@@
 
-% Packages added by us
 \usepackage{amsfonts}
 \usepackage{mydefs}
+\usepackage{algorithm}
 \usepackage{algorithmic}
-\usepackage{algorithm}
-
-\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
 
@@
 BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
-interior path{\bf DW: Following the interior path is what made them faster than
-traditional linear programming techniques that follow the edge of the feasible set;
-does Homotopy actually go outside the feasible set?  Just want to check if this is
-an accurate representation...}, which is often computationally expensive. To mitigate these
+interior path
+%{\bf DW: Following the interior path is what made them faster than
+%traditional linear programming techniques that follow the edge of the feasible set;
+%does Homotopy actually go outside the feasible set?  Just want to check if this is
+%an accurate representation...}
+, which is often computationally expensive. To mitigate these
 issues, an approach called \emph{Homotopy} has been studied to accelerate the
@@
 whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
-Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm \footnote{While we
+Algorithm \eqref{alg:alm} summarizes the ALM $\ell_1$-min algorithm \footnote{While we
 present the ALM algorithm in the primal domain, there also
@@
 a single GPU chip per PCI-express card).  Unless otherwise specified, all
-implementations utilize single precision floating point datatypes.
+implementations utilize single precision floating point datatypes.  Since on-chip
+memory (cache) is often orders of magnitude faster than off-chip memory, the sizes
+and speeds of on-chip memory (cache) have big performance implications.  We give
+a brief overview of the caches that are available in our target architectures,
+and defer discussion of their performance implications to sections
+\ref{sec:alignment_implementation_cpu} and
+\ref{sec:alignment_implementation_gpu}.
 
@@
 quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
-has its own memory interface, and is directly connected to half of the RAM
+has its own memory interface, and is directly interfaced to half of the RAM
 installed in the machine.  The amount of RAM installed exceeds the amount used
@@
 
-Each core has a 32\,KB L1 data cache and a 256\,KB L2 cache. With Hyperthreading
+%2MB = 2**21 bytes
+%64*64*4 = 2**14 bytes = 16384 bytes per image
+%... so can fit 2**7  = 128 images.
+Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache.  
+Between the 8 cores, this is a total of 2\,MB L2 cache.
+For the low resolution images used in sparse representation based
+face recognition (on the order of $64\times64$ pixel images), this corresponds to
+128 images.
+
+With Hyperthreading
 enabled, each core can support two threads simultaneously.  For floating point
@@
 arithmetic operation on four single precision floating point values
-simultaneously.  \footnote{The core also has scalar floating point units, but
-some compilers, notably gcc, generate code that uses the vector units
-exclusively for floating point instructions.}
+simultaneously.
+
+%16MB = 2**24 bytes = 16777216 bytes
+%64*64*4 = 2**14 bytes = 16384 bytes per image
+%... so can fit 2**10  = 1024 images.
+Each processor further has 8MB of L3 cache that is shared by all four cores.
+Overall, this means the algorithm has approximately 16MB of on-chip cache
+available.
+
+it is possible
+to simultaneously hold up to 1024 image vectors in on-chip cache at a time.
 
@@
 efficiently use a modern CPU: core-level parallelism and vector-level
-parallelism.  For the implementations discussed here, we will primarily use the
-OpenMP programming model to express core-level parallelism {\bf AY: What is
-OpenMP? citation needed}.  Depending on the section of the code{\bf AY: What is
-a section of the code?}, we will achieve vector-level parallelism implicitly
-through either the auto-vectorization features of Intel's icc compiler, or
-through the use of a highly optimized BLAS library (Intel's MKL). {\bf AY:
-citations needed, what is icc, what is the relation between BLAS and MKL? Are
-they equivalent? not clear in the context.}
+parallelism. CPU tools for exploiting these two levels of parallelism are
+relatively mature.  Heavily optimized Basic Linear Algebra Subprograms (BLAS) are
+commercially available that make good use of both levels of CPU parallelism.
+For operations that do not fit well with the BLAS API, we 
+achieve both vector-level and core-level parallelism through a combination of
+the OpenMP API \cite{dagum2002openmp} and the automatic vectorization 
+facilities of Intel's C++ compiler (ICC)\cite{dulong1999overview}.
+% DW: I'm really not sure these citations are even really relevant;
+% our target audience will just google icc openmp and get more up-to-date
+% info anyway.
+\subsection{GPU Hardware Parallelism}
+Compared to CPU architectures, GPU architectures dedicate less of their transistors
+to cache, and more of them for arithmetic logic. Our target GPU architecture
+is the Nvidia Fermi architecture.  CUDA based GPUs are comprised of several 
+streaming multiprocessors (SMP) (currently 14 to 16 for a high end GPU).
 
-Each processor further has 8MB of L3 cache that is shared by all four cores.
-Overall, this means the algorithm has approximately 16MB of on-chip cache
-available.  For the low resolution images used in sparse representation based
-face recognition, this is enough cache to hold over 800 image-sized variables,
-{\bf AY: how low resolution? which sparse representation problem? ``image-sized
-variables'' means simply ``image vectors''?} meaning that for small data sets
-it is possible to run completely using on-chip cache.
+The GPU executes programs as SIMT (Single Instruction Multiple Threads) in
+groups of 32 threads called \emph{warps}.  These warps are then grouped into
+higher levels called \emph{thread blocks} and \emph{grids}.  It is impossible
+to communicate between SMPs during gpu program execution, computational tasks
+are typically broken up into multiple gpu programs (kernels) that are executed
+serially.
 
-\subsection{GPU Hardware Parallelism}
+Each SMP contains its own L1-cache, which is divided between hardware managed
+and software managed ("shared") memory.  Additionally, all SMPs share a common
+L2-cache.  For our system, each SMP has 64 kb of L1 cache with a total of 896
+kb of L1 cache/shared memory and 768kb of L2 cache. The relatively small
+amount of cache on the GPU is ballanced by a significantly higher bandwidth between the
+processor chip and off-chip memory (DRAM) compared to the CPU.
 
-The ALM algorithm requires many matrix-vector operations, this gives many-core
-processors such as GPUs a computational advantage for large problem sizes.
-GPUs are comprised of several streaming multiprocessors (SMP) (14 for our
-NVIDIA Fermi) which each contains 32 floating-point units.  Each SMP contains
-its own L1-cache and shared memory and all SMPs share a common L2-cache.  The
-GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups
-of 32 threads called \emph{warps}.  These warps are then grouped into higher
-levels called \emph{thread blocks} and \emph{grids}.
+The GPU has its own memory system, any data the GPU uses must first be
+transferred from CPU DRAM to GPU DRAM over PCI-Express.  For our application,
+this transfer overhead is amortized over a large amount of computation and is
+not a major concern.
 
-Fermi's inverted memory hierarchy (aggregate l1 capacity is greater than l2 capacity) is designed to support massively parallel workloads in which there is more task-private data than shared data, providing an increase in performance and a computational advantage over CPUs.
-For our system, each SMP has 64 kb of L1 cache with a total of 896 kb of L1 cache/shared memory and
-768kb of L2 cache. However, this increase in performance does not come for
-free. As the GPU has its own memory system, any data the GPU uses must be
-transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
-However, the cost of these transfers are amortized by doing large amounts of
-computation at once, requiring code to be run completely on the GPU.  In 2011, NVIDIA is developing a new ARM-based CPU system, dubbed 
-"Project Denver" to fix performance issues such as the limited bandwidth due to PCI-E \cite{Dally-Web}.
-
-In summary, a programmable GPU can be analogous to a multi-core CPU processor,
-except with more sophisticated vector processing hardware.  Each thread block
-is a virtualized multiprocessor, which each thread as a virtualized scalar
-processor.  All of these differences allow GPUs to perform intensive
-calculations such as large scale matrix-vector operations at much faster rates than CPUs.
+Compared to a typical CPU, a GPU has much higher peak computational bandwidth
+as well as data bandwith to DRAM.  However, since our application is primarily
+of matrix-vector operations and image resampling operations, it has a very low
+arithmetic intensity (i.e. we perform a very small number of arithmetic
+operations for every number that has to be loaded), and thus performance is
+typically constrained by memory (and cache) bandwidth.  At large problem sizes
+the CPU is limited by its relatively low bandwidth to DRAM compared to the gpu,
+while at small problem sizes the GPU is limited by the relatively high latency
+of its access to DRAM.
 
@@
 and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
-once per test image. {bf DW: I could go into the rest of the math for this, but I'm worried about space}
-
-The iterative alignment stage is summarized as Algorithm \ref{alg:iterative_alignment}.
+once per test image.  The iterative alignment stage is summarized as Algorithm \eqref{alg:iterative_alignment}.
 \begin{algorithm}[thb]
@@
 $\f\in\Re^M$ 
-\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
+%\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
 \STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
@@
 The solution of the minimization problem in Algorithm
-\ref{alg:iterative_alignment} Step 8 can be solved via a
-variant of ALM, which is summarized as Algorithm \ref{alg:alignment_alm}.
+\eqref{alg:iterative_alignment} Step 7 can be solved via a
+variant of ALM, which is summarized as Algorithm \eqref{alg:alignment_alm}.
 $\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of
@@
 \end{algorithm}
+%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
 
@@
 algorithm should discuss the comutational complexity of the algorithm.  For the
-alignment stage ALM algorithm \ref{alg:alignment_alm}, this is rather simple.  The
-Matrix-vector computions are $O(mn)$, and the vector-vector operations are
+alignment stage ALM algorithm \eqref{alg:alignment_alm}, this is rather simple.
+The Matrix-vector computions are $O(mn)$, and the vector-vector operations are
 $O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
@@
 recognition problems, we find that the computation time is dominated by the
-repeated computation of matrix-vector and matrix-matrix multiplications in the
-inner loop.
+bandwidth required by repeated computation of matrix-vector and matrix-matrix
+multiplications in the inner loop.
 
-%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
+I summary, the alignment stage essentially two levels of available parallelism:
+at the higher level there are multiple alignent problems that are solved
+independently.  At a lower level, the first-order linear algebraic operations
+exhibit parallelism within image operations, i.e. at a pixel level.
 
 \subsection{CPU Implementation} 
+\label{sec:alignment_implementation_cpu}
+
+This section discusses how the iterative
+alignment problem can be implemented on CPU hardware.  The implementations
+presented in \cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} were written
+using the OpenCV library.
+
+OpenCV wraps either the operating system's built-in BLAS library, or Math
+Kernel Library (MKL), Intel's BLAS implementation, and thus achieves vector
+(i.e. pixel)-level parallelism implicitly.  The authors of
+\cite{WagnerA2011-PAMI} used OpenMP to schedule instances of the alignment
+problem to the CPU cores in a 1:1 fashion.  With eight alignment problems for
+ten training users being solved concurrently, and with 20-40 images per
+training user, all of the required data fits into L3 cache.  This provided a
+significant, though sub-linear improvement over a single-core implementation.
+
+In the proposed CPU implementation, we instead solve a single alignment
+instance at a time, using a multi-core BLAS library.  Recall that the L2 cache
+can old up to 128 image vectors' worth of data.  Since we're only aligning data
+for 1 gallery user at a time the working set of the problem fits into L2 cache,
+resulting in significantly higher image alignment speed than previous cpu
+implementations.
 
@@
 \subsection{GPU Implementation}
+\label{sec:alignment_implementation_gpu}
 {\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
@@
 The first experiment is designed to compare the two implementations of PALM
-that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm}) by
+that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
 comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse

### 2011-03-01 paper/main.tex
Starting alignment GPU section
Working on gpu implementation discussion

@@
 \usepackage{amssymb}
-
 \usepackage{amsfonts}
-\usepackage{mydefs}
 \usepackage{algorithm}
@@
 \usepackage{tikz}
+
+\usepackage{mydefs}
 
@@
 
-This section discusses how the iterative
-alignment problem can be implemented on CPU hardware.  The implementations
-presented in \cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} were written
-using the OpenCV library.
+This section discusses how the iterative alignment problem can be implemented
+on CPU hardware.  The implementations presented in \cite{WagnerA2009-CVPR} and
+\cite{WagnerA2011-PAMI} were written using the OpenCV library.
 
@@
 
-In the proposed CPU implementation, we instead solve a single alignment
-instance at a time, using a multi-core BLAS library.  Recall that the L2 cache
-can old up to 128 image vectors' worth of data.  Since we're only aligning data
-for 1 gallery user at a time the working set of the problem fits into L2 cache,
-resulting in significantly higher image alignment speed than previous cpu
-implementations.
+During the course of development, many implementation strategies were tried.
+Listed in order of increasing performance based on a synthetic benchmark of
+the $\ell_1$ minimization step, they included:
+\begin{enumerate}
+\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
+\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
+\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
+\end{enumerate}
 
-{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
+In the proposed CPU implementation of the iterative alignment stage, we
+therefore perform parallelization in a manner similar to the matlab code:
+solving a single problem instance at a time, relying on the threaded version of
+MKL BLAS for both multi-core and vector parallelism.  Why is this
+implementation faster?  Recall that the L2 cache can old up to 128 image
+vectors' worth of data.  Since we're only aligning data for 1 gallery user at a
+time the working set of the problem fits into L2 cache, resulting in
+significantly higher image alignment speed than previous cpu implementations.
+
 %\subsection{Tuning the GEMV operation}
@@
 %...Block the computation into work for different cpu cores
+
 \subsection{GPU Implementation}
 \label{sec:alignment_implementation_gpu}
-{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
+
+This section discusses how to implement algorithm
+\eqref{alg:iterative_alignment} on the GPU.  Like for the CPU case, many GPU
+implementation strategies were explored during the course of our development,
+and are only mentioned briefly.  Listed in order of increasing performance
+based on a synthetic benchmark of the $\ell_1$ minimization step, they
+included:
+\begin{enumerate}
+\item Solving a single problem instance at a time using Cublas\footnote{Nvidia provided BLAS implementation} (302 ms)
+\item Pure CUDA C implementation solving one problem per SMP using cuda streams\footnote{Streams is a CUDA feature that allows multiple
+kernels to be run concurrently} (70 ms)
+\item Pure CUDA C implementation solving one problem per SMP using one large kernel call (36 ms)
+\end{enumerate}
+
+We attribute the low performance of Cublas in this benchmark to the relatively
+low problem size. Cublas is intended to take advantage of the resources of the
+entire gpu to solve a single large problem, and our problem size is too small
+to hide the memory latency.  Effects of this will be seen again in the
+benchmarks in section \ref{sec:simulation}.
+
+Based on the strength of the results for our single-kernel implementation,
 
@@
 \subsection{Simulation}
+\label{sec:simulation}
 
@@
 \subsection{Face Alignment Benchmark}
+\label{sec:alignment_benchmark}
 
 \subsection{Face Recognition Benchmarks}
+\label{sec:recognition_benchmark}
 
 \subsection{Face Recognition Benchmark}
+\label{sec:pipeline_benchmark}
 

### 2011-03-01 paper/main.tex
adding a simulation plot which plots matrix size vs speed for tol,tol_int = 1e-1, 1e-5.  Figure doesn't import...
Working on GPU implementation section

@@
 
+% CPU CACHE DISCUSSION
 %2MB = 2**21 bytes
@@
 For the low resolution images used in sparse representation based
-face recognition (on the order of $64\times64$ pixel images), this corresponds to
-128 images.
-
-With Hyperthreading
-enabled, each core can support two threads simultaneously.  For floating point
-instructions, each core also has a vector unit capable of performing the same
-arithmetic operation on four single precision floating point values
-simultaneously.
-
+face recognition (on the order of $64\times64$ pixel images), the L2 cache
+can in principle hold up to 128 image vectors.
 %16MB = 2**24 bytes = 16777216 bytes
@@
 Overall, this means the algorithm has approximately 16MB of on-chip cache
-available.
+available.  This corresponds to 1024 image vectors.
 
-it is possible
-to simultaneously hold up to 1024 image vectors in on-chip cache at a time.
+With Hyperthreading enabled, each core can support two threads simultaneously.
+For floating point instructions, each core also has a vector processing unit
+(SSE) capable of performing the same arithmetic operation on four single
+precision floating point values simultaneously.
 
@@
 
+% GPU CACHE DISCUSSION
+% for L2 cache
+% 768KB = 786432 bytes
+% 64*64*4 = 2**14 bytes = 16384 bytes per image
+% 48 images
 Each SMP contains its own L1-cache, which is divided between hardware managed
 and software managed ("shared") memory.  Additionally, all SMPs share a common
-L2-cache.  For our system, each SMP has 64 kb of L1 cache with a total of 896
-kb of L1 cache/shared memory and 768kb of L2 cache. The relatively small
-amount of cache on the GPU is ballanced by a significantly higher bandwidth between the
-processor chip and off-chip memory (DRAM) compared to the CPU.
+L2-cache.  For our system, each SMP has 64 KB of L1 cache with a total of 896
+KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount
+of cache (1/23 as much as CPU L3) on the GPU is ballanced by a significantly
+higher bandwidth between the processor chip and off-chip memory (DRAM) compared
+to the CPU.
 
@@
 
-Based on the strength of the results for our single-kernel implementation,
+Based on the strength of the results for our single-kernel implementation, we chose to follow
+a single kernel strategy for the alignment stage.
+
+% DREW Idea... in hindsight, maybe we should still try many kernel launches, even if it
+% means individual instances can't terminate early.
+
+% Should really verify somehow that we're maxing out GPU bandwidth
 

### 2011-03-01 results/random_data/bench_PALM_chart.m
matlab script for new plot!

@@ -1,0 +1,134 @@
+%function rtn = bench_PALM_paper()
+
+type = 'PALM1';
+n  = 10; % denominator for ratio
+ks = 0.1;
+maxIter = 5000;
+iterations = 10;
+dimsPowers = [500:100:900 1000:1000:10000];
+lambda = 10;
+ratio = 1:5;
+%tols = gpuData.tols;
+%tol_ints = gpuData.tol_ints;
+
+ratios = [];
+tols   = [];
+tol_ints = [];
+
+xsizes = [];
+x1sizes = [];
+xn1sizes = [];
+
+xtimes = [];
+x1sizes = [];
+xn1sizes = [];
+
+l2diff = [];
+blasl2err = [];
+gpul2err = [];
+
+for i=2:4
+    name = [type '_' num2str(n) '_' num2str(i) '_'];
+    
+    gpuName = [name 'GPU.mat'];
+    gpuData = load(gpuName);
+    
+    blasName = [name 'BLAS.mat'];
+    blasData = load(blasName);
+    
+    ratio = i;
+    for j=1:numel(gpuData.tols)
+        for k=1:numel(gpuData.tol_ints)
+            ratios  = [ratios i];
+            tols    = [tols gpuData.tols(j)];
+            tol_ints  = [tol_ints gpuData.tol_ints(k)];
+            gpul2err  = [ gpul2err gpuData.averageL2error(j,k,:,2)];
+            blasl2err = [ blasl2err blasData.averageL2error(j,k,:,3)];
+            l2diff    = [l2diff norm(squeeze(gpuData.averageL2error(j,k,:,2))-squeeze(blasData.averageL2error(j,k,:,3)))];
+            %             plot(squeeze(gpuData.averageRunTime(j,k,:,2,2)));
+            %             hold on;
+            %             plot(squeeze(blasData.averageRunTime(j,k,:,3,2)), 'r');
+            %             hold off;
+            xpoint  = find(squeeze(gpuData.averageRunTime(j,k,:,2,2)) > squeeze(blasData.averageRunTime(j,k,:,3,2))==0,1);
+            if isempty(xpoint);
+                xpoint = -1;
+                xtime = -1;
+            else
+                xtime = squeeze(gpuData.averageRunTime(j,k,xpoint,2,2));
+                xpoint = gpuData.dimsPowers(xpoint);
+            end
+            xsizes = [xsizes xpoint];
+            xtimes = [xtimes xtime];
+            
[…]

### 2011-03-01 results/random_data/bench_PALM_chart.m
revised the new plots that plot size vs tolerance.

@@ -7,28 +7,19 @@ maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];
 lambda = 10;
-ratio = 1:5;
 %tols = gpuData.tols;
 %tol_ints = gpuData.tol_ints;
 
-ratios = [];
-tols   = [];
-tol_ints = [];
+% ratio, tol, tol_int, l2diff, blasl2err, gpul2err
+data = [];
 
-xsizes = [];
-x1sizes = [];
-xn1sizes = [];
+% xn1size, xn1time, xsize, xtime, x1size, x1time
+level_sets = [];
+level_times = [];
+i = 0;
 
-xtimes = [];
-x1sizes = [];
-xn1sizes = [];
-
-l2diff = [];
-blasl2err = [];
-gpul2err = [];
-
-for i=2:4
-    name = [type '_' num2str(n) '_' num2str(i) '_'];
+for ratio=2:4
+    name = [type '_' num2str(n) '_' num2str(ratio) '_'];
     
     gpuName = [name 'GPU.mat'];
     gpuData = load(gpuName);
@@ -36,99 +27,87 @@ for i=2:4
     blasName = [name 'BLAS.mat'];
     blasData = load(blasName);
     
-    ratio = i;
     for j=1:numel(gpuData.tols)
         for k=1:numel(gpuData.tol_ints)
-            ratios  = [ratios i];
-            tols    = [tols gpuData.tols(j)];
-            tol_ints  = [tol_ints gpuData.tol_ints(k)];
-            gpul2err  = [ gpul2err gpuData.averageL2error(j,k,:,2)];
-            blasl2err = [ blasl2err blasData.averageL2error(j,k,:,3)];
-            l2diff    = [l2diff norm(squeeze(gpuData.averageL2error(j,k,:,2))-squeeze(blasData.averageL2error(j,k,:,3)))];
+            i = i + 1;
+            data(i,1)  = ratio;
+            data(i,2)  = gpuData.tols(j);
+            data(i,3)  = gpuData.tol_ints(k);
+            data(i,6)  = mean(squeeze(gpuData.averageL2error(j,k,:,2)));
+            data(i,5)  = mean(squeeze(blasData.averageL2error(j,k,:,3)));
+            data(i,4)  = mean(squeeze(gpuData.averageL2error(j,k,:,2))-squeeze(blasData.averageL2error(j,k,:,3)));
             %             plot(squeeze(gpuData.averageRunTime(j,k,:,2,2)));
[…]

### 2011-03-01 paper/main.tex
Working on GPU implementation section

@@
 
-Based on the strength of the results for our single-kernel implementation, we chose to follow
-a single kernel strategy for the alignment stage.
+Our proposed GPU implementation for the iterative alignment stage consists of a
+single kernel that performs all of the test image alignment problems.  Each of
+instance of the alignment problem is assigned its own thread block, and the GPU
+hardware schedules as many of these thread blocks to run concurrently as the 
+hardware will allow.
 
-% DREW Idea... in hindsight, maybe we should still try many kernel launches, even if it
+% DREW Idea... in hindsight, maybe we should still try many kernel launches. 
+% instances can still exit early, and can still update a global variable signaling
+% that they're done
+
+%even if it
 % means individual instances can't terminate early.
@@
 
+\begin{figure}[sim_plot]
+\begin{center}
+      \includegraphics{../results/random_data/time_vs_matrix_size_constant_tol.svg}
+\end{center}
+\caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
+\end{figure}
+
 The observations of this experiment are summarized below:

### 2011-03-05 paper/main.tex
Finished drafting GPU parallelism section

@@
 higher levels called \emph{thread blocks} and \emph{grids}.  It is impossible
-to communicate between SMPs during gpu program execution, computational tasks
+to communicate between SMPs during gpu program execution, so computational tasks
 are typically broken up into multiple gpu programs (kernels) that are executed
@@
 
+One possible implementation of the iterative alignment stage would be to
+perform the optimization for all instances simultaneously, but to perform each
+linear algebraic operation in a separate gpu kernel call.  The primary
+advantage of this implementation strategy is that every kernel would be
+performing a very simple operation requiring fewer registers per thread.  The
+registers in a GPU are shared between all executing threads thus, barring other
+hardware limitations, the fewer registers each thread requires, the more
+threads can be scheduled to run simultaneously.  The primary downside of this
+approach is that it forces synchronization between all problem instances.  It will
+also have increased kernel launch overhead and potentially less efficient use of
+the GPU's L2 cache.
+% DREW Idea... in hindsight, maybe we should still try many kernel launches. 
+% instances can still exit early, and can still update a global variable signaling
+% that they're done
+
 Our proposed GPU implementation for the iterative alignment stage consists of a
@@
 
-% DREW Idea... in hindsight, maybe we should still try many kernel launches. 
-% instances can still exit early, and can still update a global variable signaling
-% that they're done
-
-%even if it
-% means individual instances can't terminate early.
+The full resolution test image, along with its numerical derivatives in both
+image directions, are stored in GPU texture memory. This  allows the
+recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
+versions of the test image and its derivatives, to exploit the GPU's special
+purpose hardware for bilinear interpolation.
 

### 2011-03-08 results/random_data/bench_PALM_chart.m
updated size vs tolerance GPU vs CPU crossover plots

@@ -41,7 +41,10 @@ for ratio=2:4
             %             plot(squeeze(blasData.averageRunTime(j,k,:,3,2)), 'r');
             %             hold off;
             
-            for m=-1:1
+            count = 0;
+            
+            for m=-1:0.5:1
+                count = count + 1;
                 xpoint  = find(squeeze(gpuData.averageRunTime(j,k,:,2,2)) + m/2 > squeeze(blasData.averageRunTime(j,k,:,3,2))==0,1);
                 if isempty(xpoint);
                     xpoint = inf;
@@ -50,8 +53,8 @@ for ratio=2:4
                     xtime = squeeze(gpuData.averageRunTime(j,k,xpoint,2,2));
                     xpoint = gpuData.dimsPowers(xpoint);
                 end
-                level_sets(i,m+2) = xpoint;
-                level_times(i,m+2) = xtime;
+                level_sets(i,count) = xpoint;
+                level_times(i,count) = xtime;
             end
         end
     end
@@ -89,15 +92,15 @@ for ratio=unique(ratios)'
         end
         
         plotX = [plotX; xpoint];
-        plotY = [plotY; unique_tols(j) unique_tols(j) unique_tols(j)];
+        plotY = [plotY; unique_tols(j) unique_tols(j) unique_tols(j) unique_tols(j) unique_tols(j)];
     end
         
     semilogy(plotX, plotY, '-o');
     set(gca, 'ydir', 'reverse');
     xlabel('A matrix');
-    xlabel(['A matrix (m x ' num2str(i/10) 'm)']);
+    xlabel(['A matrix (m x ' num2str(ratio/10) 'm)']);
     xlim([min(min(plotX))-500 max(max(plotX))+500]);
-    legend('GPU slower by 1 second', 'GPU = CPU speed', 'GPU faster by 1 second');
+    legend('GPU slower by 1 second','GPU slower by 1/2 second', 'GPU = CPU speed', 'GPU faster by 1/2 second', 'GPU faster by 1 second');
     %axis([]);
     
     %hold off;

### 2011-03-08 paper/main.tex
Working on writing for the implementation and experiments sections

@@
 \ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
-the system would make use of all the cores in as many CPUs are present on one
-motherboard.  For GPU implementation, the system would make the best use of a
+the system makes use of all the cores in as many CPUs are present on one
+motherboard.  For GPU implementation, the system makes the best use of a
 single many-core Nvidia GPU chip on a PCI card (GPUs are commonly shipped with
@@
 
-\subsection{CPU Implementation}
-{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
-The CPU ALM algorithm was written in C and compiled with Intel's MKL libraries, which provide
-automatic parallelization.
+\subsection{CPU Implementation} Our CPU implementation of Algorithm
+\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
+nicely onto the BLAS API (such as SGEMV, SAXPY {\bf? Victor, fill in here}), we
+call Intel's MKL BLAS library, which internally takes advantage of the SSE
+units on all available CPU cores. 
+{\bf TODO: Our current implementation links against the MKL
+library provided internally with matlab; we are in the process of updating the code to link directly against a newer version
+of MKL}
 
-\subsection{GPU Implementation}
-{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a GPU architecture.}
-The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
-CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
-which are provided by Nvidia and are optimized for GPU performance. In
-addition, we coded kernel (GPU) methods for vector-vector operations to replace
-operations that would have taken multiple BLAS calls. Because data transfer from the 
-RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector once.  
-The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an
-additional level. The CPU portion of the code serves as the conductor, making
-calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.
+For operations that do not map nicely onto the BLAS API, such as the shrinkage
+operator, {\bf Argue that these parts are optimized, and/or demonstrate that
+performance of these parts is insignificant}.
 
-\section{Experiment}
-{\bf AY: Describe the benchmark protocol here. Both for simulation and real face data.}
+\subsection{GPU Implementation} Our GPU implementation of Algorithm
+\ref{alg:alm_rec} is similar to the CPU implementation described in the
+previous section, in that only one problem is solved at a time, and each linear
+algebraic operation is parallelized separately.
 
-In this section, we present four experiments to benchmark the performance of our algorithm on CPU and GPU platforms. 
+For most of the operations, calls to MKL CPU BLAS library map directly onto
+similar calls in Nvidia's GPU BLAS library (CUBLAS).  The CUBLAS library is
+designed to utilize all of the SM's on the GPU.
+
+Operations that do not map well onto the BLAS API (mostly combined
+vector-vector operations) were implemented as gpu kernels (in C for CUDA).
+Even for operations that could have been implemented via multiple BLAS calls,
+this was found to have better performance due to reducede bandwidth and kernel
+call overhead {\bf Victor, verify or correct this}.
+
+In order to avoid expensive data transfer across the bus connecting the GPU
+card and the CPU motherboard (the PCI express bus), all of the data is
+transferred to GPU DRAM once, and all non-trivial tasks are performed on the
+GPU on data stored in GPU DRAM.  The sole exception to the above implementation
+strategy is that the pseudoinverse of A is still computed on the CPU via the
+LAPACK implementation that is part of Intel's MKL. This operation need only be
+performed (and results uploaded to GPU DRAM) once before GPU computation
+starts. {\bf we should prove justification for A pseudoinverse being cheap,
+either here or perhaps later in a pie chart}
+
+\section{Experiments}
+
+In this section, we benchmark the performance of our algorithm on CPU and GPU platforms. 
+In order to show how our $\ell_1$ minimization algorithms scale with problem size,
+we will begin with benchmarks on synthetic data.  We will then progress to demonstrating 
+the speed of our implementations as applied to the face recognition problem.
+
 To generate a meaningful comparison between different hardware architectures,
@@
 chip).  Unless otherwise specified, all implementations utilize single
-precision floating point datatypes.  For all of the experiments, we set the number 
-of inner loop iterations at 50 and outer loop iterations at 5000.
+precision floating point datatypes.
 
@@
 \begin{center}
-      \includegraphics{../results/random_data/time_vs_matrix_size_constant_tol.svg}
+      \includegraphics{../results/random_data/time_vs_matrix_size_constant_tol}
 \end{center}

### 2011-03-08 paper/figures/copy_data_from_code_repo.sh
Working on adding alignment stage runtime plot

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+EXTERNAL_DATA_LOCATION=~/cvpr2011/pipeline/singleblock/
+
+# Be careful... no trailing spaces!!
+EXTERNAL_DATA="alignment_runtime_graph_reference_cpu.dat
+alignment_runtime_graph_reference_gpu.dat"
+
+echo "${EXTERNAL_DATA}" | parallel cp -u ${EXTERNAL_DATA_LOCATION}{} ./{}

### 2011-03-08 paper/main.tex
Pretty much rewrote experiments section

@@
 
-In this section, we benchmark the performance of our algorithm on CPU and GPU platforms. 
-In order to show how our $\ell_1$ minimization algorithms scale with problem size,
-we will begin with benchmarks on synthetic data.  We will then progress to demonstrating 
-the speed of our implementations as applied to the face recognition problem.
+In this section, we benchmark the performance of our algorithm on CPU and GPU
+platforms.  In order to show how our $\ell_1$ minimization algorithms scale
+with problem size, in Section \ref{sec:simulation} we will begin with
+benchmarks on synthetic data.  We will then progress in Sections
+\ref{sec:alignment_benchmark} and \ref{sec:recognition_benchmark} to
+demonstrating the speed of our implementations as applied to the alignment and
+recognition stages of the face recognition problem.
 
@@
 
-The first experiment is designed to compare the two implementations of PALM
-that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
+\subsection{Simulations on Random Data}
+\label{sec:simulation}
+
+The first experiment compares the performance of our two proposed implementations of 
+the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
 comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
@@
 The measurement vector was generated with the equation $\y_0 = A \x_0$.   
-We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
-We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.
+%We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
+We averaged the time taken for the ALM algorithm to solve 10 instances of a problem.
 
-%The x set of experiments will compare the CPU and GPU versions of the entire pipeline, 
-%which includes the alignment and recognition stage.
-
-The second experiment is to compare the amount of time to solve the face alignment
-stage for each implementation.  Because the alignment stage is not affected by the resolution of the image,
-we fix the resolution at 60x80.
-
-Since we use a reduced training set for the recognition stage, our third experiment was to 
-find the relationship between the probability that the true subject is in the training set vs the
- number of users in the reduced training set and the speed of the recognition stage w.r.t.
- the number of subjects in the training set.
-
-The fourth experiment is to compare accuracy rate of the face recognition algorithm at different
-image resolutions.  We fixed the number of users to be $enter here$ which was shown in figure  $enter here$  have 95\% probability
-to have the true user in the training set, inner loop tolerance to 1e-2 and outer loop tolerance to 1e-3.
-
-Our final experiment is to compare the speeds of CPU and GPU implementations entire pipeline in Figure \ref{fig:pipeline}.
-
-\subsection{Simulation}
-\label{sec:simulation}
-
-We plot the most significant results of the simulation benchmark below.  The x-axis represents the size of the 
-$A$ matrix and the y-axis represents the average amount of time it to complete a single problem instance.
-
-\begin{figure}[sim_plot]
+The results of this benchmark can be found in Figure \ref{fig:random_data}.
+The x-axis represents the size of the $A$ matrix and the y-axis represents the
+average amount of time it to complete a single problem instance.  A can be seen
+in Figure \ref{fig:random_data}, the GPU implementation tends to be faster than
+the CPU implementation at solving a single large problem, whereas the the CPU
+implementation is faster at solving a single small problem.  
+\begin{figure}
+\label{fig:random_data}
 \begin{center}
@@
 
-The observations of this experiment are summarized below:
-1.  With similar accuracy rates, the GPU implementation runs faster than the CPU at the crossover points 
-noted on the plots.
-2.  From the results, it was seen that tolerances...
+\subsection{Face Alignment Stage Benchmark} \label{sec:alignment_benchmark}
+This section presents benchmarks of our CPU and GPU implementations of the
+alignment stage of the face recognition pipeline described in Algorithm
+\ref{alg:iterative_alignment}.  The iterative alignment seeks a similarity
+transformation between the coordinate frame of the "camera" resolution test
+image and the coordinate frame of the "canonical" resolution images which the
+images are compared.  We use a $64$ pixel wide square window in the canonical
+frame.  The training images are aligned by manually clicking the two outer eye
+corners in the camera frame, and then warping the images via a similarity
+transformation such that the eye corners lie at a pair of canonical eye
+locations in the canonical frame.  The canonical eye corners are chosen such
+that for well aligned images, the canonical window contains exclusively pixels
+from the subject's face.
 
-\subsection{Face Alignment Benchmark}
-\label{sec:alignment_benchmark}
+For experiments on face data we use subsets of the CMU Multi-PIE Face Database.
+For gallery images we use frontal images from session 1, which contains 20
+images per user taken under different illuminations. For test images we use
+images from session 2.
 
-\subsection{Face Recognition Benchmarks}
+In Figure \ref{fig:alignment_stage_runtime} we report the time it takes
+to align a single test image against each of the users in the gallery.
+We vary the size of the gallery to show how implementation scales in the
+number of gallery users.
+\begin{figure}
+\centering
+\includegraphics[width=3.5in]{figures/alignment_stage_runtime.pdf}
+\caption{Alignment Stage Time vs. Size of Training Database}
+\label{fig:alignment_stage_runtime}
+\end{figure}
+
+\subsection{Face Recognition Stage Benchmarks}
 \label{sec:recognition_benchmark}
+Since we use a reduced training set for the recognition stage, our third
+experiment was to find the relationship between the probability that the true
+subject is in the training set vs the number of users in the reduced training
+set and the speed of the recognition stage w.r.t.  the number of subjects in
+the training set.
 
@@
 \label{sec:pipeline_benchmark}
+The fourth experiment compares the recognition rate of the face recognition algorithm at different
+image resolutions.  We fixed the number of users to be $enter here$ which was shown in figure  $enter here$  have 95\% probability
+to have the true user in the training set, inner loop tolerance to 1e-2 and outer loop tolerance to 1e-3.
+
+Our final experiment is to compare the speeds of CPU and GPU implementations entire pipeline in Figure \ref{fig:pipeline}.
 

### 2011-03-08 paper/main.tex
Added 2 of Victor's plots to the paper

@@
 \begin{figure}
-\label{fig:random_data}
 \begin{center}
-      \includegraphics{../results/random_data/time_vs_matrix_size_constant_tol}
+\includegraphics[width=3.5in]{../results/random_data/time_vs_matrix_size_constant_tol}
 \end{center}
 \caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
+\label{fig:random_data}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{../results/random_data/size_vs_speed_crossover_ratio_2}
+\end{center}
+\caption{\bf CAPTION NEEDED}
+\label{fig:random_data2}
 \end{figure}
@@
 number of gallery users.
-\begin{figure}
-\centering
-\includegraphics[width=3.5in]{figures/alignment_stage_runtime.pdf}
-\caption{Alignment Stage Time vs. Size of Training Database}
-\label{fig:alignment_stage_runtime}
-\end{figure}
+%\begin{figure}
+%\centering
+%\include{figures/alignment_stage_runtime.tex}
+%%\includegraphics[width=3.5in]{figures/alignment_stage_runtime.pdf}
+%\caption{Alignment Stage Runtime vs. Size of Training Database}
+%\label{fig:alignment_stage_runtime}
+%\end{figure}
 

### 2011-03-09 paper/main.tex
Got Alignment stage runtime plot into the paper
Added alignment stage rank plot

@@
 number of gallery users.
-%\begin{figure}
-%\centering
-%\include{figures/alignment_stage_runtime.tex}
-%%\includegraphics[width=3.5in]{figures/alignment_stage_runtime.pdf}
-%\caption{Alignment Stage Runtime vs. Size of Training Database}
-%\label{fig:alignment_stage_runtime}
-%\end{figure}
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_runtime_graph}
+\caption{Alignment Stage Runtime vs. Size of Training Database}
+\label{fig:alignment_stage_runtime}
+\end{figure}
+
+Previous publications used a hard threshold on the number of subjects kept for the recognition
+stage as a compromise between recognition speed and potential recognition rate.
+In order to better motivate 
+a choice for a threshold of the number of users kept for the recognition stage,
+we varied the number of users kept for the recognition stage and estimated the probability
+of the test user being among the set of training users kept for the recognition stage.
+As can be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly diminishing
+returns as the number of kept users is increased beyond ten.  
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/user_alignment_rank_plot}
+\caption{Probability of the test user being kept for the recognition stage vs. Number of traning
+users kept for the recognition stage}
+\label{fig:user_alignment_rank_plot}
+\end{figure}
 

### 2011-03-09 paper/main.tex
Removed redundant paragraph.

@@
 \label{sec:recognition_benchmark}
-Since we use a reduced training set for the recognition stage, our third
-experiment was to find the relationship between the probability that the true
-subject is in the training set vs the number of users in the reduced training
-set and the speed of the recognition stage w.r.t.  the number of subjects in
-the training set.
 
-\subsection{Face Recognition Benchmark}
+\subsection{Face Recognition Pipeline Benchmarks}
 \label{sec:pipeline_benchmark}

### 2011-03-09 results/gpu_bench/bench_GPU.m
adding matlab files to create plots and updated some plots adding some papers for the lit-review into repo adding gpu and gpu_cache benchmarks (not really needed)

@@ -1,0 +1,99 @@
+function rtn = bench_PALM()
+
+type = 'benchGPU';
+n  = 10; % denominator for ratio
+ks = 0.1;
+tols = [1e-4];
+maxIter = 5000;
+iterations = 10;
+dimsPowers = [500:100:900 1000:1000:10000];
+lambda = 10;
+ratio = 1:5;
+
+h = figure;
+
+for i=ratio
+    name = [type '_' num2str(n) '_' num2str(i) '_'];
+    
+    gpuName = [name 'GPU.mat'];
+    
+    gpuData = load(gpuName);
+        
+    gpuDimsPowers  = gpuData.dimsPowers;
+    gpuDimLen      = numel(gpuData.dimsPowers <= 6000);
+
+    gpuASize       = (gpuDimsPowers.^2 * i / 10 * 32) / (2^20);
+        
+    for j=1:numel(tols)
+        
+        subplot(numel(tols), 3, (j-1)*3+1);
+        gpuOutIter  = squeeze(gpuData.averageIteration(j,j,:,2))'*5;
+        
+        % Calculations for totalSize (aka the number of times we use each
+        % matrix * size of the matrix
+        % A is mxn.  G is nxn.  (3*m*n) * totalOutIterations + (n^2)
+        % * total Inner iterations.  Is that correct?
+        gpuTotalSize    = gpuASize .* gpuOutIter / (2^10);
+
+        plotA = squeeze(gpuData.averageRunTime(j,j,:,2,2))'; % multiply by 5 because I had A1 to A5 all run "maxIter" times
+        plot(gpuASize(1:gpuDimLen), gpuTotalSize(1:gpuDimLen)./plotA(1:gpuDimLen), 'bx-');
+        title(['ratio 10 to ' num2str(i)]);
+        ylabel('Average speed (Gbits/s)');
+        xlabel('A (in Mbits)');
+        hold off
+        
+        if j == 1
+            legend('GPU');
+        end
+%         
+%         subplot(numel(tols), 3, (j-1)*3+2);
+%         plotA = gpuData.averageL2error(j,j,:,2);
+%         plotB = cpuData.averageL2error(j,j,:,1);
+%         plotC = blasData.averageL2error(j,j,:,3);
+%         plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
+%         hold on
+%         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
[…]

### 2011-03-09 results/gpu_cache_bench/bench_PALM.m

@@ -1,0 +1,99 @@
+function rtn = bench_PALM()
+
+type = 'benchGPU';
+n  = 10; % denominator for ratio
+ks = 0.1;
+tols = [1e-4];
+maxIter = 5000;
+iterations = 10;
+dimsPowers = [500:100:900 1000:1000:10000];
+lambda = 10;
+ratio = 1:5;
+
+h = figure;
+
+for i=ratio
+    name = [type '_' num2str(n) '_' num2str(i) '_'];
+    
+    gpuName = [name 'GPU.mat'];
+    
+    gpuData = load(gpuName);
+        
+    gpuDimsPowers  = gpuData.dimsPowers;
+    gpuDimLen      = numel(gpuData.dimsPowers <= 6000);
+
+    gpuASize       = (gpuDimsPowers.^2 * i / 10 * 32) / (2^20);
+        
+    for j=1:numel(tols)
+        
+        subplot(numel(tols), 3, (j-1)*3+1);
+        gpuOutIter  = squeeze(gpuData.averageIteration(j,j,:,2))';
+        
+        % Calculations for totalSize (aka the number of times we use each
+        % matrix * size of the matrix
+        % A is mxn.  G is nxn.  (3*m*n) * totalOutIterations + (n^2)
+        % * total Inner iterations.  Is that correct?
+        gpuTotalSize    = gpuASize .* gpuOutIter / (2^10);
+
+        plotA = squeeze(gpuData.averageRunTime(j,j,:,2,2))'; % multiply by 5 because I had A1 to A5 all run "maxIter" times
+        plot(gpuASize(1:gpuDimLen), gpuTotalSize(1:gpuDimLen)./plotA(1:gpuDimLen), 'bx-');
+        title(['Tolerance ' num2str(tols(j)) ' ratio ' num2str(i)]);
+        ylabel('Average speed (Gbits/s)');
+        xlabel('A & G Matrix (in Mbits)');
+        hold off
+        
+        if j == 1
+            legend('GPU');
+        end
+%         
+%         subplot(numel(tols), 3, (j-1)*3+2);
+%         plotA = gpuData.averageL2error(j,j,:,2);
+%         plotB = cpuData.averageL2error(j,j,:,1);
+%         plotC = blasData.averageL2error(j,j,:,3);
+%         plot(gpuDimsPowers, squeeze(plotA(1:gpuDimLen)), 'bx-');
+%         hold on
+%         plot(cpuDimsPowers, squeeze(plotB(1:cpuDimLen)), 'r+-');
[…]

### 2011-03-09 results/random_data/bench_PALM.m

@@ -1,4 +1,4 @@
-function rtn = bench_PALM_paper()
+function rtn = bench_PALM()
 
 type = 'PALM1';
 n  = 10; % denominator for ratio
@@ -7,11 +7,11 @@ maxIter = 5000;
 iterations = 10;
 dimsPowers = [500:100:900 1000:1000:10000];
 lambda = 10;
-ratio = 1:5;
+ratio = 2:4;
 %tols = gpuData.tols;
 %tol_ints = gpuData.tol_ints;
-tols = [1e-2 1e-3];
-tol_ints = [1e-2 1e-3 1e-4];
+tols = [1e-1 1e-2];
+tol_ints = [1e-1 1e-2 1e-3 1e-4 1e-5];
 
 h = figure;
 
@@ -31,8 +31,8 @@ for i=ratio
     blasDimsPowers = blasData.dimsPowers;
     blasDimLen     = numel(blasData.dimsPowers <= 6000);
     
-    gpuASize       = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
-    blasASize      = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 32 / (2^20);
+    gpuASize       = (gpuDimsPowers.^2 * i / 10 * (1 + i / 10)) * 4 / (2^20);
+    blasASize      = (blasDimsPowers.^2 * i / 10 * (1 + i / 10)) * 4 / (2^20);
     
     for j=1:numel(tols)
         for k=1:numel(tol_ints)
@@ -47,8 +47,8 @@ for i=ratio
             % matrix * size of the matrix
             % A is mxn.  G is nxn.  (3*m*n) * totalOutIterations + (n^2)
             % * total Inner iterations.  Is that correct?
-            gpuTotalSize    = gpuDimsPowers.^2 .* ((3 * (i/10) .* gpuOutIter)  + (i/10)^2 .* gpuInIter) * 32 / (2^30);
-            blasTotalSize   = blasDimsPowers.^2 .* ((3 * (i/10) .* blasOutIter)  + (i/10)^2 .* blasInIter) * 32 / (2^30);
+            gpuTotalSize    = gpuDimsPowers.^2 .* ((3 * (i/10) .* gpuOutIter)  + (i/10)^2 .* gpuInIter) * 4 / (2^30);
[…]

### 2011-03-09 results/random_data/bench_PALM_paper.m

@@ -1,4 +1,4 @@
-function rtn = bench_PALM_paper()
+%function rtn = bench_PALM_paper()
 
 type = 'PALM1';
 n  = 10; % denominator for ratio
@@ -10,49 +10,60 @@ lambda = 10;
 ratio = 1:5;
 %tols = gpuData.tols;
 %tol_ints = gpuData.tol_ints;
-tols = [1e-2 1e-3];
-tol_ints = [1e-2 1e-3 1e-4];
+tols = [1e-1 1e-2 1e-3 1e-4];
+tol_ints = [1e-1 1e-2 1e-3 1e-4 1e-5];
 
 h = figure;
 
-for i=2:3
-    name = [type '_' num2str(n) '_' num2str(i) '_'];
-    
-    gpuName = [name 'GPU.mat'];
-    
-    gpuData = load(gpuName);
-    
-    blasName = [name 'BLAS.mat'];
-    
-    blasData = load(blasName);
-    
-    gpuDimsPowers  = gpuData.dimsPowers;
-    gpuDimLen      = numel(gpuData.dimsPowers <= 6000);
-    blasDimsPowers = blasData.dimsPowers;
-    blasDimLen     = numel(blasData.dimsPowers <= 6000);
-    
-    subplot(1, 2, i-1);
-    plotA = squeeze(gpuData.averageRunTime(2,1,:,2,2))';
-    plotC = squeeze(blasData.averageRunTime(2,1,:,3,2))';
-    plot(gpuDimsPowers(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
-    hold on
-    plot(blasDimsPowers(1:blasDimLen), plotC(1:blasDimLen), 'g+-');
-    title(['Tolerance 1e-3, Inner Tolerance 1e-2']);
-    ylabel('Elapsed time (s)');
-    xlabel(['A matrix dimension (m x 0.' num2str(i) 'm)']);
-    hold off
-    legend('gpu', 'cpu');
+for ratio_i=2:4
     
     
-end
-set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
-set (gcf,'Units','Inches');                     %   Using units of inches
-pos=get(gcf,'position');                        %   Save position values
-set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
-set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
+    for i=2:4
[…]

### 2011-03-09 results/random_data/bench_PALM_paper1.m

@@ -37,9 +37,10 @@ for i=2:4
     subplot(1, 3, i-1);
     plotA = squeeze(gpuData.averageRunTime(1,5,:,2,2))';
     plotC = squeeze(blasData.averageRunTime(1,5,:,3,2))';
-    plot(gpuASize(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
+    %plot(gpuASize(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
+    plot(gpuDimsPowers(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
     hold on
-    plot(blasASize(1:blasDimLen), plotC(1:blasDimLen), 'g+-');
+    plot(blasDimsPowers(1:blasDimLen), plotC(1:blasDimLen), 'g+-');
     %title(['Tolerance ' num2str(tols(1)) ', Inner Tolerance ' num2str(tol_ints(5))]);
     ylabel('Elapsed time (s)');
     %axis([0 100 0 30])

### 2011-03-09 results/recognition/bench_rec_test.m

@@ -1,0 +1,55 @@
+function rtn = bench_rec()
+
+type = 'PALM';
+tols = [1e-3 1e-4 1e-5];
+
+h = figure;
+
+name = [type '_rec_'];
+
+gpuName = [name 'GPU.mat'];
+
+gpuData = load(gpuName);
+
+cpuName = [name 'CPU.mat'];
+
+cpuData = load(cpuName);
+
+gpuNumUsers = gpuData.numUsers(gpuData.numUsers < 50);
+gpuNUlen    = numel(gpuNumUsers);
+cpuNumUsers = cpuData.numUsers(cpuData.numUsers < 50);
+cpuNUlen    = numel(cpuNumUsers);
+
+for j=1:numel(tols)
+    
+    subplot(numel(tols), 2, (j-1)*2+1);
+    plotA = gpuData.averageRunTime(j,:,2,2);
+    plotB = cpuData.averageRunTime(j,:,1,2);
+    plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
+    hold on
+    plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
+    title(['Average runtime - tolerance ' num2str(tols(j))]);
+    xlabel('Number of Users');
+    hold off
+    
+    subplot(numel(tols), 2, (j-1)*2+2);
+    plotA = gpuData.averageIteration(j,:,2);
+    plotB = cpuData.averageIteration(j,:,1);
+    plot(gpuNumUsers, squeeze(plotA(1:gpuNUlen)), 'bx-');
+    hold on
+    plot(cpuNumUsers, squeeze(plotB(1:cpuNUlen)), 'r+-');
+    title(['Average iterations - tolerance ' num2str(tols(j))]);
+    xlabel('Number of Users');
+    hold off
+end
+
+set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
+set (gcf,'Units','Inches');                     %   Using units of inches
+pos=get(gcf,'position');                        %   Save position values
+set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
+set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
+
+plot2svg([type '_benchmark.svg'], h, 'png');
+%saveas(h, [type '_benchmark_ratio_' num2str(i)], 'jpg');
+
+end

### 2011-03-10 main.tex
Reorganizing paper repo

--- 2011-03-09 paper/main.tex
@@
 \begin{center}
-\includegraphics[width=3.5in]{../results/random_data/time_vs_matrix_size_constant_tol}
+\includegraphics[width=3.5in]{results/random_data/time_vs_matrix_size_constant_tol}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=3.5in]{../results/random_data/size_vs_speed_crossover_ratio_2}
+\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_2}
 \end{center}

### 2011-03-10 pami_paper_for_reference.tex
Reorganizing paper repo

%% bare_jrnl.tex
%% V1.1
%% 2002/08/13
%% by Michael Shell
%% mshell@ece.gatech.edu
%%
%% NOTE: This text file uses MS Windows line feed conventions. When (human)
%% reading this file on other platforms, you may have to use a text
%% editor that can handle lines terminated by the MS Windows line feed
%% characters (0x0D 0x0A).
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.6 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.ieee.org
%% and/or
%% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
%%
%% This code is offered as-is - no warranty - user assumes all risk.
%% Free to use, distribute and modify.

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% Testflow can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/testflow

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print. Authors are encouraged to use U.S. letter paper when
% submitting to IEEE. Use the testflow package mentioned above to verify
% correct handling of both paper sizes by the author's LaTeX system.
%
% Also note that the "draftcls", not "draft", option should be used if
% it is desired that the figures are to be displayed in draft mode.

% *************************************************************************
% You can now use this template both for submitting to peer review and,
% if your paper is accepted, for sending in final publication materials.
% To change from peer review format (single column, double-spaced) to
% final publication format (double column, single-spaced), just move the
% comment-line sign (%) from one \documentclass line to the other.
% The first is for peer review format(single column), the second is for final publication(double column).

\documentclass[12pt,journal,draftcls,letterpaper,onecolumn]{IEEEtran}
%\documentclass[9.5pt,journal,final,finalsubmission,twocolumn]{IEEEtran}
% *************************************************************************

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{pdfsync}
\usepackage{helvet}
\usepackage{url}
\usepackage{color}

% Include other packages here, before hyperref.
\usepackage{mydefs}
\usepackage{algorithm}
\usepackage{algorithmic}

\newcommand{\subj}{\textup{subj to}}

\newcommand{\jw}[1]{{\bf \textcolor{blue}{John: #1}}}
%\newcommand{\aw}[1]{\bf \color{greem} Drew: #1}
%\newcommand{\ym}[1]{\bf \color{red}{ Yi: #1}}

\newcommand{\zz}[1]{{\bf \textcolor{red}{Zihan: #1}}}

\begin{document}
%
% paper title
\title{Towards a Practical Face Recognition System: \\ Robust Alignment and
Illumination by Sparse Representation}
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
\author{Andrew~Wagner,~\IEEEmembership{Student~Member,~IEEE,}
        John~Wright,~\IEEEmembership{Member,~IEEE,}
        Arvind~Ganesh,~\IEEEmembership{Student~Member,~IEEE,}
        Zihan~Zhou,~\IEEEmembership{Student~Member,~IEEE,}
        Hossein~Mobahi,
        and~Yi~Ma,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
\thanks{A. Wagner, A. Ganesh, Z. Zhou, and Y. Ma are with the Dept. of Electrical and Computer Engineering at the University of Illinois at Urbana-Champaign. H. Mobahi is with the Computer Science Dept. at the University of Illinois at Urbana-Champaign.  J. Wright and Y. Ma are with Microsoft Research Asia.  Corresponding author: Andrew Wagner, awagner@illinois.edu, 1308 W. Main st. Urbana, IL 61801, (312) 343-1380.}}% <-this % stops a space

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~1, No.~8,~August~2002}{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.

% If you want to put a publisher's ID mark on the page
% (can leave text blank if you just want to see how the
% text height on the first page will be reduced by IEEE)
%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}

% use only for invited papers
%\specialpapernotice{(Invited Paper)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%% ABSTRACT
\IEEEcompsoctitleabstractindextext{
\vspace{-15mm}
\begin{abstract}
Many classic and contemporary face recognition algorithms work well on public data sets, but degrade sharply when they are  used in a real recognition system. This is mostly due to the difficulty of simultaneously handling variations in illumination, image misalignment, and occlusion in the test image. We consider a scenario where the training images are well controlled, and test images are only loosely controlled. We propose a conceptually simple face recognition system that achieves a high degree of robustness and stability to illumination variation, image misalignment, and partial occlusion. The system uses tools from sparse representation to align a test face image to a set of frontal training images. The region of attraction of our alignment algorithm is computed empirically for public face datasets such as Multi-PIE. We demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under uncontrolled illumination. In order to evaluate how our algorithms work under practical testing conditions, we have implemented a complete face recognition system, including a projector-based training acquisition system. Our system can efficiently and effectively recognize faces under a variety of realistic conditions, using only frontal images under the proposed illuminations as training.\vspace{0mm}
\end{abstract}

% make the title area

\begin{IEEEkeywords}
Face Recognition, Face Alignment, Illumination Variation, Occlusion and Corruption,
Sparse Representation,  Error Correction, Validation and Outlier Rejection.
\end{IEEEkeywords}}

\maketitle

\IEEEdisplaynotcompsoctitleabstractindextext

\IEEEpeerreviewmaketitle

%%%%%%%%% BODY TEXT
\section{Introduction}
%Automatic face recognition remains one of the most active areas in computer vision.

\PARstart{F}{ace} recognition applications to date have fallen into roughly two
categories.  Face recognition has recently seen a lot of success in a family of
less-demanding applications such as online image search and family photo album
organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto).
At the other end of the tractability spectrum there are the terrorist watchlist
and mass surveillance applications that have for the most part dominated the
field of face recognition research.  However, there are many face recognition
that fall roughly between these extremes, where very high recognition
performance is desired, but the users in the gallery are still allies of the
system rather than adversaries.  These applications include access control for
secure facilities (e.g., prisons, office buildings), computer systems,
automobiles, or automatic teller machines, where controlled gallery images can
be obtained in advance.  These applications are very interesting due to their
potential sociological impact.  Since the gallery subjects are allies, rather
than opponents, of the recognition system, this creates the possibility of
carefully controlling the acquisition of the training data. While the same can
be said for other biometrics such as fingerprints and iris recognition, face
recognition has the potential of working with test data that is much less
controlled, allowing the access control system to be made less intrusive to the
users of the system. To some extent, the goal of this paper is to show how a
reliable face recognition system can be built for this restricted,
but still important, scenario.

Very few recognition systems specifically target applications where many well-
controlled training images are available.  Of these, the classical holistic
subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
are well known for their speed and simplicity, as well as for their natural
extension to linear illumination models.  However, their performance has been
shown to be extremely brittle not only to alignment variation, but to even
minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
is slightly open. One promising recent direction, set forth in
\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
sparse representation of the test image in terms of the training set as a
whole, up to some sparse error due to occlusion. A \emph{sparse
representation-based classification} (SRC) method is then proposed for
recognition. The main idea is that the sparse nonzero coefficients should
concentrate on the training samples with the same class label as the test
sample. SRC has demonstrated striking recognition performance despite severe
occlusion or corruption by solving a simple convex program.  
%For this reason, the final recognition stage of our proposed system will
%consist of SRC, and the alignment stage that precedes it will be based on
%similar techniques.

\begin{figure} \centering \begin{tabular}{cc}
\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
\hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
\hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
\hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
\end{tabular} \caption{{\bf Effects of registration and illumination on
Recognition}. In this example we identify the girl among 20 subjects, by
computing the sparse representation of her input face with respect to the
entire training set. The absolute sum of the coefficients associated with each
subject is plotted on the right. We also show the faces reconstructed with each
subject's training images weighted by the associated sparse coefficients. The
red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The
input face is from Viola and Jones' face detector (the black box) and all 38
illuminations specified in Section \ref{sec:illumination} are used in the
training.  {\bf Middle:} The input face is well-aligned (the white box) with
the training by our algorithm specified in Section \ref{sec:registration} but
only 24 frontal illuminations are used in the training for recognition (see
Section \ref{sec:illumination}). {\bf Bottom:} The input face is well aligned and 
a sufficient set (all 38) of
illuminations are used in the training. Both are necessary for correct recognition
using SRC.} \label{fig:promo}
\vspace{-.5in}
\end{figure}
%\subsection{Compound Effect of Registration and Illumination}

Unfortunately, while the work \cite{Wright2009-PAMI} achieves impressive results, it does not deal with misalignment
between the test and training images, and it requires a rich set of illuminations
in the gallery images for good performance.  
%Furthermore, the experiments were carried out on public datasets taken under
%controlled laboratory conditions such as Extended Yale B
%\cite{Georghiades2001-PAMI}, which do not contain realistic test images.
%two critical aspects of real world face recognition must be addressed:  significant variations in both the {\em image domain} and in
%the {\em image value}.
We illustrate the compounded effect of both alignment and illumination 
with an example in Figure \ref{fig:promo}.
The task is to identify the girl among 20 subjects. If the test
face image, say obtained from an off-the-shelf face detector,
has even a small amount of registration error against the
training images (caused by mild pose, scale, or misalignment),
the sparse representation obtained using the method of \cite{Wright2009-PAMI} is no longer informative, 
even if sufficient illuminations are present in the training, as shown in Figure
\ref{fig:promo}(top). Additionally, in order to span
the illuminations of a typical indoor (or outdoor) environment,
illuminations from behind the subject are needed in the
training set. Otherwise, even for perfectly aligned test images,
the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
informative, as shown by the example in Figure
\ref{fig:promo}(middle). Clearly, both good alignment, as well
as sufficient training images are needed to ensure success of the sparsity-based 
recognition method proposed by \cite{Wright2009-PAMI}.  
%Unfortunately, most public face databases lack images with a
%significant component of rear (more than 90 degrees from
%frontal) illumination, either for training or testing.
In this paper, we examine how to handle alignment and illumination simultaneously
in the sparse representation framework, bringing the method proposed in \cite{Wright2009-PAMI}
closer to practical use.

\subsection{Related Work}

We briefly review existing techniques for recognition, image registration, and
handling of illumination variation.  Our system is based purely on 2D
techniques.  This fact immediately distinguishes our approach from systems
that either require a 3D data as an input, or attempt to estimate a 3D model
from 2D input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
techniques can achieve better robustness to pose variation given
a sufficiently accurate 3D model, for access control applications where only 
moderate pose variation is present, the proposed method will be more
than sufficient.   Note that 2D images of faces
under varying illuminations already contain 3D shape-related information,\footnote{In principle,
one can recover the 3D shape of the face from multiple illuminations using photometric stereo.} 
and this information can be leveraged by 2D algorithms for alignment and recognition 
even if shape is not reconstructed explicitly. 
%In addition, the proposed sparsity-based robust
%alignment and recognition techniques are not limited to 2D data and can be easily extended
%to 3D data if necessary.

% Alignment
In holistic recognition algorithms, correspondence between points in the test
image and in the training must be achieved.  A long line of research exists on
using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related
Active Shape Models \cite{cootes1992active} to register images against a
relatively high-dimensional model of plausible face appearances, often
leveraging face-specific contours.  While these model-based techniques have
advantages in dealing with variations in expression and pose, they may add
unnecessary complexity to applications where subjects normally present a neutral face
or only have moderate expression. We prefer to focus on deformations with far fewer 
degrees of freedom, i.e. similarity transformations, and to use the training images themselves 
as the appearance model. Iterative registration in this spirit dates at least back to the 
Lucas-Kanade algorithm \cite{lucas1981iterative}.

Whereas much of the early work on image registration is aimed at the
problem of registering nearly identical images, say by minimizing a sum of
squared distances or maximizing normalized correlation, here we must confront several physical factors
simultaneously: misalignment, illumination variations, and corrupted pixels.
As we discuss further below, illumination variation can be dealt by
expressing the test image as a linear combination of an appropriate set of
training images. Similar representations have been exploited in
illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
For robustness to gross errors, the $\ell^1$-norm of the residual is a more
appropriate objective function than the classical $\ell^2$-norm. Its use here
is loosely motivated by theoretical results due to Cand\`{e}s and Tao
\cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
lead us to pose the registration problem as the search for a set of
transformations and illumination coefficients that minimize the $\ell^1$-norm
of the representation error. We solve this problem using a generalized
Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm
minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of
these problems can also be solved efficiently using recently developed
first-order techniques for $\ell^1$-minimization, which are reviewed in
\cite{YangA2010-pp}.

% Illumination
Researchers have tried various techniques to deal with illumination variation.
In almost all recognition algorithms where only a single gallery image is
available per individual, illumination effects are regarded as a nuisance that
must be removed before the algorithm can continue.  This is typically done by
making statistical assumptions about how illumination affects the image, and
using those assumptions to extract a new representation that is claimed to be
illumination invariant.  Recent examples include \cite{chen2006total} and
\cite{zhou2007appearance}.  However, despite these efforts, truly
illumination-invariant features are difficult to obtain from a single input
image.  We argue that if one has the luxury of designing the acquisition system 
and the application demands a high recognition rate,
it is then unwise to limit the gallery to a
single image per person.  We therefore take the strategy of sampling many
gallery images of each individual under varying illuminations.  These images
are used as the basis for either a convex cone model
\cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model
\cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
projector based light stage.  While similar systems have been used for
other applications, to our knowledge, we
are the first to use projectors to indirectly illuminate a subject's face for
the purpose of face recognition.

\subsection{Contributions} In this paper, we show how 
registration and illumination
can be simultaneously addressed within a robust sparse representation
framework. We show that face registration, a challenging
nonlinear problem, can be solved by a series of linear programs
that iteratively minimize the sparsity of the registration
error. This leads to an efficient and effective alignment
algorithm for face images that works for a large range of
variation in translation, rotation, and scale, even when
the face is only partially visible due to eyeglasses, closed
eyes and open mouth, sensor saturation, etc.  We also propose a
sufficient set of training illuminations
that is capable of linearly representing typical indoor and outdoor
lighting, along with a practical hardware system for capturing
them.

% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.  While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.

% We use tools from sparse representation
% to align a test face image with a set of frontal training
% images in the presence of registration error, illumination variation, and partial occlusion. The region
% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions. We show
% that our system can efficiently and effectively recognize faces under
% a variety of realistic conditions, using only
% frontal images under the proposed illuminations as training.

%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.  We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.  Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.  We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.  We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.

We then demonstrate the effectiveness of the proposed new
methods with a complete face recognition system that is {\em
simple, stable, and scalable}. The proposed system performs
robust automatic recognition of subjects from loosely
controlled probe images taken both indoors and outdoors, 
using a gallery of 
frontal views of the subjects' faces under the proposed
illuminations. An off-the-shelf face
detector\footnote{We use the OpenCV
implementation of the Viola and Jones' face detector
\cite{Viola2004-IJCV}.} is used to detect faces in the test images.

We conduct extensive experiments on the proposed system with
both public databases and a face database that is collected by
our own acquisition system. Our experimental results on
large-scale public face databases show that our algorithm
indeed achieves very good performance on these databases,
exceeding or competing with the state-of-the-art algorithms. 
Additionally, our experimental results on our own database
clearly demonstrate that our system not only works well with
images taken under controlled laboratory conditions, but is
capable of handling practical indoor and outdoor illuminations as well.

\noindent{\em Organization of this paper:} In Section \ref{sec:registration},
we derive our robust registration and recognition algorithm within the sparse
representation framework. We elaborate on algorithmic implementation issues,
conduct region of attraction experiments with respect to both 2D in-plane
deformation and 3D pose variation, and discuss its relationship to existing
work. Section \ref{sec:illumination} is dedicated to our training acquisition
system. Using this system, we investigate empirically how many training
illuminations are required to handle practical illumination variations, and
suggest a sufficient set of 38 training illuminations. Extensive experiments on
a large-scale public database and on our own database are conducted in Section
\ref{sec:multipie} and Section \ref{sec:own-data}, respectively, to verify the
proposed system. Section \ref{sec:conclusion} concludes our work with
discussion of promising future directions.

\section{Robust Alignment}\label{sec:registration}
As demonstrated in Figure \ref{fig:promo}(top), the main
limitation of the {\em sparse representation and
classification} (SRC) algorithm of \cite{Wright2009-PAMI} is
the assumption of pixel-accurate alignment between the test
image and the training set. This leads to brittleness under
pose and misalignment, making it inappropriate for deployment
outside a laboratory setting. In this section, we show how this
weakness can be rectified while still preserving the conceptual
simplicity and good recognition performance of SRC.

SRC assumes access to a database of multiple registered
training images per subject, taken under varying illuminations.
The images of subject $i$, stacked as vectors, form a matrix
$A_i \in \Re^{m \times n_i}$. Taken together, all of the images
form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ]
\in \Re^{m \times n}$. As argued in \cite{Wright2009-PAMI}, a
well-aligned test image $\y_0$ can be represented as a sparse
linear combination $A \x_0$ of all of the images in the
database,\footnote{We assume the training illuminations are sufficient. We will address how to ensure illumination
sufficiency in the next section.} plus a sparse error $\e_0$
due to corrupted pixels. The sparse representation can be recovered by
minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
vector, denoted by $\|\cdot\|_1$, is the sum of absolute values of its entries.} of
$\x$ and $\e$:
\begin{equation}
\min_{\x,\e} \, \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
\label{eqn:robust-l1}
\end{equation}
Now suppose that $\y_0$ is subject to some pose or
misalignment, so that instead of observing $\y_0$, we observe
the warped image $\y = \y_0 \circ \tau^{-1}$, for some
transformation $\tau \in T$ where $T$ is a finite-dimensional
group of transformations acting on the image domain.  The
transformed image $\y$ no longer has a sparse representation of
the form $\y = A \x_0 + \e_0$, and naively applying the
algorithm of \cite{Wright2009-PAMI} is no longer appropriate,
as seen in Figure \ref{fig:promo}(top).

\subsection{Batch and Individual Alignment} If the
true deformation $\tau^{-1}$ can be found, then we can apply
its inverse $\tau$ to the test image and it again becomes
possible to find a sparse representation of the resulting
image, as $\y \circ \tau = A \x_0 + \e_0$.\footnote{In the terminology of \cite{baker2004lucas}, this formulation is ``Forward Additive''.}
  This sparsity
provides a strong cue for finding the correct deformation
$\tau$: conceptually, one would like to seek a transformation
$\tau$ that allows the sparsest representation, via
\begin{equation} \label{eqn:L1-L1-conceptual}
\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e.
\end{equation}
For fixed $\tau$, this problem is jointly convex in $\x$ and
$\e$. However, as a simultaneous optimization over the
coefficients $\x$, error representation $\e$, and
transformation $\tau$, it is a difficult, nonconvex
optimization problem. One source of difficulty is the presence
of multiple faces in the matrix $A$:
\eqref{eqn:L1-L1-conceptual} has many local minima that
correspond to aligning $\y$ to different subjects. In this
sense, the misaligned recognition problem differs from the
well-aligned version studied in \cite{Wright2009-PAMI}. For the
well-aligned case, it is possible to directly solve for a
global representation, with no concern for local minima. With
possible misalignment, it is more appropriate to seek the best
alignment of the test face with each subject $i$:
\begin{equation} \label{eqn:per-subject-L1}
\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.
\end{equation}
We no longer penalize $\| \x \|_1$, since $A_i$ consists of
only images of subject $i$ and so $\x$ is no longer expected to
be sparse.

\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector. We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
\begin{equation}
\y\circ \tau + J \Delta \tau = A_i \x + \e.
\end{equation}
Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
Jacobian of $\y \circ \tau$ with respect to the transformation
parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
above equation is underdetermined if we allow the registration
error $\e$ to be arbitrary. At the correct alignment we
expect the test image to differ from $A_i \x$ only
for the minority of the pixels corrupted by occlusions. Thus,
we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
\begin{equation}
\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
\label{eqn:L1-align}
\end{equation}
This is different from the popular choice that
minimizes the $\ell^2$-norm of the registration error:
\begin{equation}
\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
\label{eqn:L2-align}
\end{equation}
which is also equivalent to finding the deformation step
$\Delta  \tau$ by solving the least-square problem:
$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
\|_2$. Empirically, we find that if there is only small noise
between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
$\eqref{eqn:L2-align}$ have similar performance.  However, if
there are occlusions in $\y_0$, sequential
$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
better than sequential $\ell^2$-minimization
\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
example.

The scheme \eqref{eqn:L1-align} can be viewed as a generalized Gauss-Newton method for minimizing the composition of a nonsmooth objective function (the $\ell^1$-norm) with a differentiable mapping from transformation parameters to transformed images. Such algorithms date at least back to the 1970's  \cite{Cromme1978-NM,Jittorntrum1980-NM}, and continue to attract attention today \cite{Lewis2008-TR}. While space precludes a detailed discussion of their properties, we should mention that the scheme \eqref{eqn:L1-align} is known to converge quadratically in the neighborhood of any local optimum of the $\ell^1$-norm. In practice, this means that $\approx$ 10 to 15 iterations suffice to reach the desired solution. We refer the interested reader to \cite{Jittorntrum1980-NM,Osborne1990-JAMSSB} and the references therein.

\begin{figure}
\centering
{\small
\begin{tabular}{cccc}
\includegraphics[height=1.2in]{figures_cvpr/L1_cropped} &
\includegraphics[height=1.2in]{figures_cvpr/y_warp_L1} &
\includegraphics[height=1.2in]{figures_cvpr/y_hat_L1} &
\includegraphics[height=1.2in]{figures_cvpr/e_L1} \\
\includegraphics[height=1.2in]{figures_cvpr/L2_cropped} &
\includegraphics[height=1.2in]{figures_cvpr/y_warp_L2} &
\includegraphics[height=1.2in]{figures_cvpr/y_hat_L2} &
\includegraphics[height=1.2in]{figures_cvpr/e_L2} \\
(a) & (b) & (c) & (d)
\end{tabular}
\vspace{-.1in}}
\caption{{\bf Comparing alignment of a subject wearing sunglasses by
$\ell^1$ and $\ell^2$ minimization.}
{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
given by the face detector, {\em Red (solid):} Alignment result shown on the same
face; (b) warped testing image using the estimated transformation $\y_0$;
(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
\vspace{-.4in}
\end{figure}

In addition to normalizing the training images (which is done
once), it is important to normalize the warped testing image
$\y \circ \tau$ as the algorithm runs.  Without normalization,
the algorithm may fall into a degenerate global minimum
corresponding to zooming in on a dark region of the test
image.  Normalization is done by replacing the linearization of
$\y \circ \tau$ with a linearization of the normalized version
$\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.
The proposed alignment algorithm can be easily extended to work
in a {\em multiscale} fashion, with benefits both in
convergence behavior and computational cost.  The alignment
algorithm is simply run to completion on progressively less
downsampled versions of the training and testing images, using
the result of one level to initialize the next.

\subsection{Robust Recognition by Sparse Representation} Once
the best transformation $\tau_i$ has been computed for each
subject $i$, the training sets $A_i$ can be aligned to $\y$,
and a global sparse representation problem of the form
\eqref{eqn:robust-l1} can be solved to obtain a discriminative
representation in terms of the entire training set. Moreover,
the per-subject alignment residuals $\| \e \|_1$ can be used to
prune unpromising candidates from the global optimization,
leaving a much smaller and more efficiently solvable problem.
The complete optimization procedure is summarized as Algorithm
\ref{alg:deformable-src}. The parameter $S$ in our algorithm is the number of subjects
considered together to provide a sparse representation for the
test image. If $S = 1$, the algorithm reduces to classification
by registration error; but considering the test image might be
an invalid subject, we typically choose $S = 10$. Since valid
images have a sparse representation in terms of this larger
set, we can reject invalid test images using the {\em sparsity
concentration index} proposed in \cite{Wright2009-PAMI}.
The function $\delta_i(\x)$ in Algorithm \ref{alg:deformable-src}
selects coefficients from the vector $\x$ corresponding to subject $i$.

Another important free parameter in Algorithm
\ref{alg:deformable-src} is the class of deformations $T$. In
our experiments, we typically use 2D similarity
transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{Here, SE stands for Special
Euclidean, i.e., 2D rigid transformations.  The $\Re_+$ accounts for the scale.}, for removing alignment error incurred by face
detector, or 2D projective transformations, $T =
\mathbb{GL}(3)\footnote{Here, GL stands for General Linear.
This class of transformations is able to represent distortion
in a perspective image of a planar object.}$, for handling some
pose variation.

In Algorithm \ref{alg:deformable-src}, we also implement a simple heuristic which improves the
performance of our system, based on the observation that the face detector output may be poorly centered on
the face, and may contain a significant amount of the background.  Therefore, before the recognition stage,
instead of aligning the training sets to the
original $\y$ directly obtained from the face detector, we
compute an average transformation $\bar{\tau}$ from $\tau_{k_1},
\tau_{k_2}, \ldots, \tau_{k_S}$ of the top $S$ classes, which
is believed to be better centered, and update $\y$ according to
$\bar{\tau}$. For the 2D similarity transformations, which are
used in our system when initialized by the face detector, a
transformation $\tau$ can be parameterized as $\tau = (\tau^1,
\tau^2, \tau^3, \tau^4)$, where $\tau^1$ and $\tau^2$ represent
the translations in $x$- and $y$-axis, $\tau^3$ represents the
rotation angle and $\tau^4$ represents the scale. Then the
average transformation is simply obtained by taking the
component-wise mean:
\begin{displaymath}
\bar{\tau}^i = (\tau_{k_1}^i + \tau_{k_2}^i + \cdots +
\tau_{k_S}^i) / S, i = 1,2,3,4.
\end{displaymath}
Finally, the training sets are aligned to the new $\y$.

\begin{algorithm}[thb]
\caption{\bf (Deformable Sparse Recovery and Classification for
Face Recognition)} \label{alg:deformable-src}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\b\in\Re^m$ and a deformation group $T$. 
\STATE
{\bf for} each subject $i$, 
\STATE \hspace{3mm} $\tau^{(0)}
\leftarrow I$. 
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
\STATE \hspace{6mm}
$\tilde \b(\tau) \leftarrow \frac{\b \circ \tau}{\|\b \circ
\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
\tau} \tilde\b(\tau)  \bigr|_{\tau^{(j)}} $;
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \b + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
\subj \; \tilde \b + J \Delta \tau = A_i \x + \e.$ 
\STATE
\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$; 
\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
residuals $\|\e\|_1$. \STATE Compute an average transformation
$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
\STATE Update $\b \leftarrow \b \circ \bar{\tau}$ and $\tau_i
\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
$\ell^1$-minimization problem: \hspace{2mm}
$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \b = A \x + \e.$
\STATE Compute residuals $r_i(\b) = \| {\b} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\b) = \arg\min_i r_i(\b)$.
\end{small}
\end{algorithmic}
\end{algorithm}

The transformation $\tau$ defines a mapping between the
coordinates of pixels in the large original image and a smaller
(un)warped image. The pixels of the small image are stacked into a vector. To prevent aliasing
artifacts in the downsampled image, one should apply a
smoothing filter to the original image. For a simple implementation, a rectangular window with
regular sampling can used, but in general, the small image need
not be regularly sampled in pixel coordinates.  For example,
the sample locations could be arbitrarily selected from within a
``face shaped'' area. We will discuss how choosing
different windows can affect the performance of our algorithm
in Section \ref{sec:multipie}. 
%The strength of the
%filter depends on how much the warping downsamples the image.
%Although the scale component of the warping may vary with each
%step of the iterative alignment, it is safe in practice to
%choose a fixed conservative smoothing filter (erring slightly
%on the side of more smoothing). Whenever a transformation
%operator is not shown for a vector in our notation (i.e. for
%the columns of $A$ or $J$), it means that a suitable
%transformation has already been applied to a larger image
%(otherwise, the dimensions of the arrays would not agree).

\subsection{System Implementation}
In this section, we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
%The most classical
%solution to these $\ell^1$-minimization problems is the so
%called \emph{primal-dual interior-point} method
%\cite{BoydS2004}. Conceptually, it operates in an iterative
%fashion, and a linear system is solved at each iteration.
%Therefore, its complexity per iteration is at least $O(n^3)$,
%where $n$ is the total number of training images for the face
%recognition problem. Meanwhile, it is also observed in
%\cite{YangA2010-pp} that its computational cost with respect to
%the dimension of each sample, $m$, grows much fast than other
%alternative methods. Hence for real-world applications, as the
%image resolution and number of classes grow, this method will
%soon become too slow to be useful.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
%We refer the interested readers to
%\cite{YangA2010-pp} for a comprehensive survey in the
%literature and a detailed study of the performance of five
%representative approaches on solving the two $\ell^1$-minimization
%problems in Step 6 and 14 of
%Algorithm~\ref{alg:deformable-src}.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.
%method, is always
%among the most accurate and fastest algorithms and achieves
%better performance in both effectiveness and efficiency
%compared to other first-order methods.
%Most importantly, as first-order methods, they require only
%lightweight vector operations and matrix-vector
%multiplications, contrast to many other methods, such as
%primal-dual interior-point method and the homotopy method
%\cite{DonohoD2006-pp}, which all involve expensive operations
%such as solving a linear system. For the rest of this section,
%we briefly describe the ALM method.

We use Step 14 as an example to illustrate the ALM
method, since solving Step 6 is very similar. Recall that in
Step 14 the problem we are interested in is:
\begin{equation}
\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \y =
A \x + \e.
\end{equation}
Its corresponding augmented Lagrangian function is
\begin{equation}
L_\mu (\x,\e,\blamda) = \|\x\|_1 + \|\e\|_1 + \langle \blamda, \y-A\x - \e \rangle + \frac{\mu}{2} \|\y - A\x - \e\|_2^2,
\end{equation}
where $\blamda$ is the Lagrange multiplier and $\mu > 0$ is a
penalty parameter. The ALM method seeks a saddlepoint of $L_\mu
(\x,\e,\blamda)$ by alternating between optimizing with respect
to the primal variables $\x, \e$ and updating the dual variable
$\blamda$, with the other fixed, as follows:
\begin{equation}
\left \{
\begin{array}{lll}
(\x_{k+1},\e_{k+1})  =  \arg\min_{(\x,\e)} \, L_{\mu} (\x,\e,\blamda_k),\\
\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
\end{array}
\right .
\label{eq:alm}
\end{equation}
Although updating $\blamda$ is trivial,
minimizing $L_{\mu} (\x,\e,\blamda_k)$ with respect to both
$\x$ and $\e$ could still be costly. To further reduce the
complexity of the problem, we adopt an approach used in
\cite{YangJ2009-pp}, called \emph{alternating direction
method of multipliers} (ADM) \cite{Glowinski1975-TR}, which alternates between minimizing $L_{\mu} (\x,\e,\blamda_k)$
over $\x$ (with $\e$ fixed) and over $\e$ (with $\x$ fixed). After solving these two subproblems, the Lagrange multiplier $\blamda$ is updated, yielding an iteration of the form:
\begin{equation}
\left \{
\begin{array}{lll}
\e_{k+1}  =  \arg\min_{\e} \, L_{\mu} (\x_k,\e,\blamda_k),\\
\x_{k+1}  =  \arg\min_{\x} \, L_{\mu} (\x,\e_{k+1},\blamda_k),\\
\blamda_{k+1}  =  \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1}). \\
\end{array}
\right .
\label{eq:adm}
\end{equation}
As the objective function is convex and alternation is between two
terms, this procedure is guaranteed to converge to a global optimum (see \cite{YangJ2009-pp} and references therein).

In order to discuss the solution to the above subproblems, we
need to define the following soft-thresholding operator for a
vector $\x$ and a scalar $\alpha \geq 0$:
\begin{equation}
\textup{shrink}(\x,\alpha) = \textup{sign}(\x)\cdot \max \{|\x| - \alpha, 0\},
\end{equation}
where all the operations are performed component-wise. It is
easy to show that the subproblem with respect to $\e$ has a
closed-form solution given by the soft-thresholding operator:
\begin{equation}
\e_{k+1} = \textup{shrink}(\y - A\x_k + \mu^{-1}\blamda_k, \mu^{-1}).
\end{equation}
To solve the subproblem associated with $\x$, we
apply a first-order $\ell^1$-minimization method,
called \emph{fast iterative shrinkage-threshold algorithm}
(FISTA) \cite{BeckA2009}. The main idea of FISTA is to
iteratively minimize a quadratic approximation $Q(\x, \z)$ to
$L_{\mu} (\x,\e_{k+1},\blamda_k)$ around a point $\z$, which is
carefully chosen in order to achieve a good convergence
rate. We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu = 2m / \|\y\|_1$.
\begin{algorithm}[t]
\caption{\bf (Augmented Lagrange Multiplier Method for Global
Recognition)}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{small}
\end{algorithmic}
\label{alg:alm} 
\end{algorithm}

%
%Note
%that, however, Algorithm~\ref{alg:alm} differs from
%\cite{YangJ2009-pp} mainly in the way that the sub-problem
%associated with $\x$ is solved. In \cite{YangJ2009-pp}, only an
%approximation to the solution is obtained by just computing one
%iteration of the Conjugate Gradient algorithm. As pointed out
%by \cite{YangA2010-pp}, although this further reduces the
%complexity of the algorithm, it converges very slowly when the
%condition number of the matrix $A^TA$ is large, which is indeed
%the case for the face recognition problem.

We have selected this algorithm because it strikes the best
balance between speed, accuracy, and scalability for our problem out of
many algorithms that we have tested. We refer the interested reader to
\cite{YangA2010-pp} for a more detailed discussion of competing
approaches.  On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory, 
running on our database containing images size $80\times 60$ 
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm~\ref{alg:deformable-src} takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is 
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!

\subsection{Simulations and Experiments on Region of
Attraction} We will now present three experimental results
demonstrating the effectiveness of the individual alignment
procedure outlined in the previous section. They show the sufficiency of the
region of attraction, verify effectiveness of the multiscale extension,
and show stability to small pose variations.  We delay large-scale recognition experiments
to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
have discussed the issue of illumination in the next section.

\noindent {1) \em 2D Deformation.}  We first verify the
    effectiveness of our alignment algorithm with images
    from the CMU Multi-PIE Database \cite{Gross2008-FGR}.
    We select all the subjects in Session 1, use 7
    illuminations per person from Session 1 for training,
    and test on one new illumination from Session
    2.\footnote{The training are illuminations $\{0, 1, 7,
    13, 14, 16, 18\}$ of \cite{Gross2008-FGR}, and the
    testing is the illumination 10. } We manually select
    eye corners in both training and testing as the ground
    truth for registration. We downsample the images to
    $80\times 60$ pixels\footnote{Unless otherwise stated,
    this will be the default resolution at which we prepare
    all our training and testing datasets.} and the distance between the two outer
    eye corners is normalized to be 50 pixels for each
    person. We introduce artificial deformation to the
    testing image with a combination of translation,
    rotation and scaling. We further use the alignment
    error $\|\e\|_1$ as an indicator of success. Let $r_0$
    be the alignment error obtained by aligning a test
    image to the training images without any artificial
    perturbation. When the test image is artificially
    perturbed and aligned, resulting in an alignment error
    $r$, we consider the alignment successful if $|r - r_0 | \leq
    0.01r_0$. Figure \ref{fig:attraction} shows the
    percentage of successful registrations for all subjects
    for each artificial deformation. The results suggest
    that our algorithm works extremely well with
    translation up to 20\% of the eye distance (or 10
    pixels) in all directions and up to $30^\circ$ in-plane
    rotation. We have also tested our alignment algorithm
    with scale variation and it can handle up to 15\%
    change in scale.

We have gathered the statistics of the Viola and Jones'
 face detector on the Multi-PIE dataset. For 4,600 frontal
 images of 230 subjects under 20 different illuminations,
 using manual registration as the ground truth, the average
 misalignment error of the detected faces is about 6 pixels
 and the average variation in scale is 8\%. This falls
 safely inside the region of attraction for our alignment
 algorithm.
\begin{figure}
\centering
{\small
\begin{tabular}{ccc}
\includegraphics[height=1.4in]{figures_pami/x_y_roa.png} &
\includegraphics[height=1.4in]{figures_pami/y_theta_roa.png} &
\includegraphics[height=1.4in]{figures_pami/y_s_roa.png}\\
(a)&(b)&(c)
\end{tabular}
\begin{tabular}{cccc}
\includegraphics[height=1.1in]{figures_pami/x_tr.png} &
\includegraphics[height=1.1in]{figures_pami/y_tr.png} &
\includegraphics[height=1.1in]{figures_pami/theta.png} &
\includegraphics[height=1.1in]{figures_pami/scale.png}\\
(d)&(e)&(f)& (g)
\end{tabular}
\vspace{-.2in}
}
\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.  The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees. {\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions. (b) Simultaneous translation in $y$ direction and in-plane rotation. (c) Simultaneous translation in $y$ direction and scale variation. {\bf Bottom row:} (d) Translation in $x$ direction only. (e) Translation in $y$ direction only. (f) In-plane rotation only. (g) Scale variation only.}
\label{fig:attraction}
\vspace{-.4in}
\end{figure}

\noindent{2) \em Multiscale Implementation.}
        %The
    %gain here is a reduction in the computational cost. We
    %briefly explain how this works here. In our system, a
    %truncated Gaussian kernel is used for blurring. In
    %order to ensure accurate operation, we change the
    %window size of the Gaussian kernel in proportion to its
    %standard deviation. Consequently, achieving a large
    %amount of blur requires a large convolution
    %mask\footnote{It is possible to apply a low-bandwidth
    %Gaussian kernel multiple times to achieve the same
    %effect of a single high-bandwidth kernel. However, the
    %time required will be at least as great as performing
    %the single large blur.} and hence is computationally
    %expensive.
    %To avoid using a large window, we observe that, by
    %down-sampling the image, a smaller blur kernel can
    %achieve similar blurring effect. However, working at a
    %lower resolution degrades the alignment precision.
    %Therefore, we use a multi-scale framework to achieve
    %computational efficiency and high precision at the same
    %time. It starts from the smallest scale with largest
%blur. The strong blur can handle large misalignment, but
%the low resolution results in poor precision. We then use
%the estimated alignment to initialize a similar process
%that works at a higher resolution and with smaller blur.
%This process repeats and thus the solution is computed
%progressively.
Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost. Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
%To see the relationship between number of scales and the
%standard deviation $\sigma$ of the Gaussian kernel applied
%between scales,
We use the same 7 illuminations in
Session 1 as training, and all 20 illuminations in the same
session as testing. We introduce artificial deformation in
both $x$ and $y$ directions up to 16 pixels in the
$80\times 60$ frame, with a step size of 4 pixels, i.e.,
$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
\{-16,-12,\ldots,12,16\}$. We consider an alignment
successful if the estimated coordinates of the eye-corners
are within 1 pixel from the ground truth in the original
image.  In Figure \ref{fig:multiscale}, we report the
alignment success rate, averaged over the artificially
perturbed initial deformations, as a function of the
standard deviation of the Gaussian kernel $\sigma$, for
three choices of the number of scales. As one can see,
using multiscale indeed improves the performance, and when
3 scales are used, a smaller convolution kernel can achieve
a similar performance compared to a much larger kernel when
only 2 scales are used.
\begin{figure}
\centering
\includegraphics[height=1.8in]{figures_pami/multiscale.png}
\vspace{-.3in}
\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations. A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
\label{fig:multiscale}
\vspace{-.5in}
\end{figure}

\noindent{3) \em 3D Pose Variation.} As densely sampled pose and
    illumination face images are not available in any of
    the public databases, including Multi-PIE, we have
    collected our own dataset using our own system (to be
    introduced in the next section). We use frontal face
    images of a subject under the 38 illuminations proposed
    in the next section as training. For testing, we
    collect images of the subject under a typical indoor
    lighting condition at pose ranging from $-90^\circ$ to
    $+90^\circ$ with step size 5.625$^\circ$, a total of 33
    poses. We use Viola and Jones' face detector to
    initialize our alignment algorithm.
%whenever it succeeds. For extreme poses where the detector fails (usually for poses $> 45^\circ$), we use manual initialization by putting a reasonable square window around the face.
Figure \ref{fig:pose-alignment} shows that our algorithm works reasonably well with poses up
to $\pm 45^\circ$.
Note that this level of out-of-plane
 pose variation is beyond what we intend to handle with our formulation.
%While we don't intend to perform 2D recognition with this much pose,
%the alignment is at least stable.
\begin{figure}
\centering
{\small
\begin{tabular}{ccccc}
\includegraphics[height=1in]{figures_cvpr/5} &
\includegraphics[height=1in]{figures_cvpr/7} &
\includegraphics[height=1in]{figures_cvpr/09} &
\includegraphics[height=1in]{figures_cvpr/11} &
\includegraphics[height=1in]{figures_cvpr/13} \vspace{-.1in} \\ 
(a) & (b) & (c) & (d) & (e)\\
\includegraphics[height=1in]{figures_cvpr/15} &
\includegraphics[height=1in]{figures_cvpr/17} &
\includegraphics[height=1in]{figures_cvpr/19} &
\includegraphics[height=1in]{figures_cvpr/21} &
\includegraphics[height=1in]{figures_cvpr/3} \vspace{-.1in} \\ 
(f) & (g) & (h) & (i) & (j) \vspace{-.1in}
\end{tabular}
}
\caption{{\bf 2D Alignment of test images with different poses to frontal training images.} {\bf (a) to (i):}  plausible alignment for pose from $-45^{\circ}$
to $+45^{\circ}$. {\bf (j):} a case when the algorithm fails for an extreme pose ($>45^{\circ}$).
}\label{fig:pose-alignment} \vspace{-.2in}
\end{figure}

\subsection{Comparison with Related Work}
Our modification to SRC roots solidly in the tradition of
adding deformation-robustness to face recognition algorithms
\cite{Cootes2001-PAMI,Gross2006-PAMI,Wiskott1997-PAMI}.
However, the only previous work to investigate face alignment
in the context of sparse signal representation and SRC is the
work of \cite{Huang2008-CVPR}. They consider the case where the
training images themselves are misaligned and allow one
deformation per training image. They linearize the training
rather than the test, which is computationally more costly as
it effectively triples the size of the training set. In
addition, as they align the test image to all subjects
simultaneously, it potentially is more prone to local minima
when the number of subjects increases, as we will see in the
following experimental comparisons.
\begin{enumerate}
\item {\em Extended Yale B.} In this experiment, we have
    used the same experimental settings as in
    \cite{Huang2008-CVPR}. 20 subjects are selected and
    each has 32 frontal images (selected at random) as
    training and another 32 for testing. An artificial
    translation of 10 pixels (in both $x$ and $y$
    directions) is introduced to the test image. For our
    algorithm we downsample all the images to $88\times
    80$ for memory reasons, whereas the work of
    \cite{Huang2008-CVPR} uses random projections.
Note that the use of cropped images in this experiment introduces image boundary effects.
    Our
    algorithm achieves the recognition rate 93.7\%,
    compared to 89.1\% recognition rate reported in
    \cite{Huang2008-CVPR}.
\item {\em CMU Multi-PIE.} In this experiment, we choose
    all subjects from the CMU Multi-PIE database, 7
    training images from Session 1 and 1 test image from
    Session 2 per person. The setting is exactly the same
    as the previous experiment on 2D deformation. We again
    work with downsampled images of size $80\times 60$ pixels. An
    artificial translation of 5 pixels (in both $x$ and $y$
    directions) was induced in the test image. The
    algorithm of \cite{Huang2008-CVPR} achieves a
    recognition rate of 67.5\%,\footnote{That algorithm has
    two free parameters - $l$ and $d$, which govern the tradeoff between
	accuracy and run-time. For this experiment
    we chose $l = 1$ and $d = 514$.} while ours achieves 92.2\%.
    \end{enumerate}

\section{Handling Practical Illumination Variation}\label{sec:illumination}
In the above section, we have made the assumption that the test image, although taken under some arbitrary illumination, can be linearly represented by a finite number of training illuminations.  Under what conditions is this a reasonable assumption to make?  What can we say from first principles about how the training images should be chosen?

\subsection{The Illumination Model}

% % META: Most of this is about understanding the Basi Result.  For the Paper, don't try to derive the result; just state it.
% % Derivation of linearity using strong assumptions
% First, let us assume that
% the illumination if the subject's face is distant.  This model
% will be a good approximation as long as distance to the nearest
% light source is much larger than the person's face.  If we
% further assume that the object is convex so that there is no
% self-shadowing, the illumination incident on a surface patch
% will depend only on the surface normal, and not on the location on the surface.
% Furthermore, if the object is Lambertian, the image intensity will not depend on the angle between patch surface normal and the sensor, or on the depth of the patch.  Since each pixel in the image corresponds to
% some patch on the object, the vector of image intensities is a
% linear function of the radiance of the corresponding portions
% of the object.
% % Motivation of low dimensionality from spherical harmonics, low-pass interpretation
% Consider the vector space of all illuminations
% of the object.  These are just positive functions (or more
% generally, distributions) defined on the sphere, where each
% point on the sphere corresponds to a direction of incoming
% light.  Consider the vector space formed by the light exiting
% the object.  This too is a positive function defined on the
% sphere, where now each point on the sphere corresponds to the
% normal of a patch on the object (or any other patch with the
% same normal).  Under the Lambertian assumption, the radiance of
% a patch depends on the cosine of the incident angle of incoming
% light, and is integrated over the half-sphere of illumination
% that the patch can "see". Thus, the surface of the object acts
% on the sphere in a manner very similar to a half-cosine
% approximation to a low-pass filter for a one-dimensional
% signal.  Thus the energy leaving the object is
% disproportionately concentrated in the subspace corresponding
% to low spatial frequencies.  Since the image is a linear
% function of this, the space of images of an object will also
% tend to fall on (the positive portion of) a subspace.   In
% \cite{Basri2003-PAMI}, Basri showed using spherical harmonic
% basis functions that nine basis illuminations (corresponding to
% the lowest frequency spherical harmonics) result in training
% images that do a good job of linearly interpolating all other
% training images.  It should be noted, however, that these basis
% illuminations are not strictly positive, and thus neither are
% the training images they generate (when rendered on a computer
% that can handle negative illumination).  \footnote{Note that
% while the spherical harmonic basis functions have regions where
% they are negative, this does no necessary preclude this
% theoretical result from being put to use in practice.  If the
% geometry of the training illumination system were well
% calibrated, it would be possible to generate approximations of
% the positive and (rectified) negative components of the basis
% illuminations separately, taking the difference of the
% resulting two images, and storing it using a signed datatype.
% Combinations of these images may not be positive, but they may
% still be good enough to be useful.}

The strongest theoretical results so far regarding the relationship
between illumination and the resulting sets of images is due to Basri and Jacobs \cite{Basri2003-PAMI}.
The main result of this paper is that for convex Lambertian objects, distant illuminations, and fixed pose,
%Under those assumptions, the incident and reflected light are represented using distributions on a sphere and can thus be studied in a spherical harmonic basis.
%The Lambertian nature of the object acts as a low-pass filter between the incident and reflected light distributions,
%and as a result, the set of images of the object end up lying very
%close to the subspace corresponding to low frequency spherical harmonics.\footnote{This is highly condensed; please refer to \cite{Basri2003-PAMI} for a  complete description of this result.}
all images of the object can be well approximated by linear combinations of
nine (properly chosen) basis images.  The basis images have mixed sign, and
their illuminations consist of the lowest frequency spherical harmonics.
While this is a very important result for understanding the image
formation process, the direct application of this result in most practical
systems is misguided for several reasons.
% Non-physicality
%Neither the basis illuminations, nor the basis images used
%in the Basri analysis are physical.  Like Fourier bases, they have negative components,
%and are thus not, at least directly, physically realizable.
% This alone is not
% a showstopper, however: one could, for example, decompose the basis images into
% their positive and negative parts and use them separately.  If the images
% are being rendered from a 3D model, ``negative'' light can be permitted.
%There are ways to engineer around this difficulty\footnote{such as allowing negative light while rendering from a 3D model, decomposing mixed-sign illuminations and images into positive negative parts, and enforcing positivity only in the output image}, but the photometric and geometric assumptions behind this model limit its applicability even more severely.
%Studio photographers and movie-makers
%go to great lengths to control their effects through makeup and lighting.
% Not only does specularity allow some of the high angular frequency components of the
% incident illumination to affect the outgoing illumination, it even prevents
% the relationship between incoming and outgoing light from even being expressable as a convolution.
% Indeed, in the extreme case of an image of a mirror-surface sphere, the image is basically a warped
% version of the sphere's incident illumination, and there is no reason to believe the set of
% images is any lower dimensional than the set of illuminations.
Specularities, self-shadowing, and inter-reflections all dramatically affect the appearance of face images,
and they all do so in a way that violates the modeling assumptions of the Basri analysis.
%Specifically, self-shadowing and inter-reflection break the simple relationship between the outgoing light
%and the resulting images, and specularity even invalidates the representation of outgoing light that is used.
% itself,
%since it is no longer independent of viewing angle.
%%If the surface has a specular component, the relationship between incoming and outgoing light
%is no longer expressible as a convolution, and if there is self-shadowing or inter-reflection the illumination
%of each patch will depend not only on its location on the surface, but on the illumination of other pixels in the image.

Fortunately, even with these effects, for most materials the relationship between
illumination and image is still linear,\footnote{Materials that break
this assumption include fluorescent materials and the photochromic (``Transition'') lenses
in some eyeglasses.  Most materials emit light in proportion to their
incident light.} provided the sensor has a linear response curve.\footnote{Proper handling of gamma encoding is an important consideration for
practitioners.  Most cameras apply a non-linear and often undocumented response
curve to captured images.  A slight degradation of performance will occur if
gamma compressed images are treated as if they were linear.  We recommend the use
of cameras with well documented response curves that can be inverted when the 
image file is loaded.}
For a more in-depth study 
of the relationship between illumination and images, we refer the reader to
\cite{belhumeur1998set}.
While the relationship between illuminations and images is linear,
only positive weights are allowed; the space of all images of an object with
fixed pose and varying illumination is a convex cone lying in the positive
orthant. The question becomes, how many images does it take to do a good job
of representing images sampled from this cone?

It has been observed in various empirical studies that
one can get away with using a small number of frontal
illuminations to linearly represent a wide range of new frontal
illuminations, when they are all taken under the same laboratory conditions
\cite{Georghiades2001-PAMI}. This is the case for many public
face datasets, including AR, ORL, PIE, and Multi-PIE.
Unfortunately, we have found that in practice, a training
database consisting purely of frontal illuminations is not
sufficient to linearly represent images of a faces taken
under typical indoor or outdoor conditions (see the experiment
conducted in Section \ref{sec:own-data}). As illustrated by the
example in Figure \ref{fig:promo}, an insufficient number of
training illuminations can result in recognition failure. To
ensure our algorithm works in practice, we need to find a set
of training illuminations that are indeed {\em sufficient} to
linearly represent a wide variety of practical indoor and
outdoor illuminations.

\subsection{Projector-based Illumination System}

We have designed a system that can acquire frontal images of a subject while
simultaneously illuminating the subject from all directions above horizontal. A sketch of the
system is shown in Figure \ref{fig:system}: The illumination
system consists of four projectors that display various bright
patterns onto the three white walls in the corner of a dark
room.  The light reflects off of the walls and illuminates the
user's head indirectly.  After taking the frontal illuminations
we rotate the chair by 180 degrees and take pictures from the
opposite direction.  Having two cameras speeds the process
since only the chair needs to be moved in between frontal and
rear illuminations. Our projector-based system has several
advantages over flash-based illumination systems for face recognition:
\begin{itemize}
\item The illuminations can be modified in software, rather than hardware.
\item It is easy to capture many different illuminations quickly.
\item Good coverage and distant illumination can be achieved simultaneously.
\item There is no need to mount anything on the walls or construct a large dome.
\item The system can be assembled from off-the-shelf hardware.
\end{itemize}
\begin{figure}
\centerline{\hspace{-0.1in}\includegraphics[height=2.3in]{figures_cvpr/camera_rig.pdf}\vspace{-.2in}}
\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
\vspace{-.3in}
\label{fig:system}
\end{figure}
With our projector system, our choice of illuminations is
constrained only by the need to achieve a good
SNR\footnote{Since illuminations with more pixels illuminated
will have a better SNR (provided they don't saturate), there is
an engineering tradeoff between the SNR and the number of
training images.}, avoid saturation, and achieve
a reasonably short acquisition time.  Two simplifying
assumptions that we make are that every pixel is either turned
fully on or off in every illumination, and that the illuminated
regions do not overlap.

Assuming that each pixel is fully on or off enables us to guarantee
that each illumination image has the same overall intensity, merely
by guaranteeing that we illuminate the same number of pixels in each image.\footnote{Since DLP projectors may have dramatically different response
curves depending on the mode they are in, it is not advisable to simply normalize each illumination image by its mean.}
Since our algorithm depends only on  the
linearity between the illuminations and the images, and not on the
relative intensities of the illuminations, the designer has the freedom to choose the overall intensity of the illuminations
to prevent saturation or low SNR, in a sort of offline exposure control.

%Assuming that no pixels are shared by multiple illuminations is important;
%for example, any other non-negative basis illuminations formed from positive linear
%combinations of our basis illuminations would span a cone of lower
%volume, and thus the resulting training images might not be
%able to represent as wide a variety of testing images without using
%negative coefficients in $\x$.
%\footnote{This is true if $\x>0$ is enforced;
%the effect of negative coefficients in $\x$ is still under study.}

Assuming that the sequentially illuminated regions do not overlap results in a
set of training images that span a larger cone than a similar number of
overlapping regions.  This results in training images that require fewer
negative coefficients in $\x$ to represent test images under natural
illuminations.  The effect of negative coefficients in $\x$ appears to depend
partly on how the test images are taken and is still under study.

{\em Relationship to existing work:} Most light stages used for face recognition have
been constructed for the purpose of creating public data sets to study
illumination invariance \cite{Georghiades2001-PAMI, Gross2008-FGR}.  Many other
light stages have been used for computer graphics purposes
\cite{debevec2000acquiring, jones2005performance}.
The light source can be
moved around manually \cite{masselus2002free}, but this may result in poor
consistency of illuminations between users.  Structured light applications use projectors to
directly illuminate the face (or other object) \cite{zhang2002rapid} for 3D
reconstruction, but this is very disturbing to the user. 
Y.\ Schechner \cite{schechner2007multiplexing}
studies techniques for multiplexing illumination that can dramatically reduce the noise 
of the demultiplexed images for certain classes of objects and cameras.  
While these techniques have not been incorporated into the current
system, they fit elegantly into our framework and will likely be used
in future implementations.  We stress that use of this multiplexing technique
is independent from the choice of original (directional) illuminations.

\subsection{Choice of Illumination Patterns}

We ran two experiments to guide our choice of illuminations for
our large-scale experiments:
\begin{enumerate}
\item {\em Coverage Experiment.} In the first experiment we
    attempt to determine what coverage of the sphere is
    required to achieve good interpolation for test images.
    The subject was illuminated by 100 (50 front, 50 back)
    illuminations arranged in concentric rings centered at
    the front camera.  Subsets of the training images were
    chosen, starting at the front camera and adding a ring
    at a time.  Each time a ring was added to the training
    illumination set, the average $\ell^1$ registration
    error (residual) for a set of test images taken under
    sunlight was computed and plotted in Figure
    \ref{fig:illumination-patterns}(a).  The more rings of
    training illuminations are added, the lower the
    representation error becomes, with diminishing returns.
%\end{itemize}
%\begin{itemize}
\item {\em Granularity Experiment.} In the second
    experiment we attempt to determine how finely divided
    the illumination sphere should be.  At the first
    granularity level, the projectors  illuminate the
    covered area uniformly.  At each subsequent granularity
    level each illuminated cell is divided in two along its
    longer side but intensity doubled.  For each
    granularity level the average $\ell^1$ registration
    error is computed as in the coverage experiment and
    shown in Figure \ref{fig:illumination-sufficiency}(b).
    Again, diminishing returns are observed as more
    illuminations are added.
\end{enumerate}
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[height=1.5in]{figures_cvpr/coverage_experiment_asplode.png} &
\includegraphics[height=1.5in]{figures_cvpr/final_cvpr_illuminations_asplode.png}  \\
(a) Coverage Experiment & (b) Chosen Illumination Patterns 
\end{tabular}\vspace{-2mm}
\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.  For rear illuminations the sequence is reversed.  In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations. The four rectangular regions correspond to the four projectors.  }
\label{fig:illumination-patterns}\label{fig:illumination-sufficiency}
\vspace{-.2in}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[height=1.5in]{figures_cvpr/illum_results/coverage_sunset.pdf} &
\includegraphics[height=1.5in]{figures_cvpr/illum_results/granularity_sunset.pdf} \\
(a) Coverage & (b) Granularity
\end{tabular}\vspace{-2mm}
\caption{{\bf Study of sufficient illuminations.} The average $\ell^1$ registration residual versus different illumination training sets. }
\label{fig:illumination-sufficiency}
\vspace{-.5in}
\end{figure}

 In the plot for the
coverage experiment, Figure \ref{fig:illumination-sufficiency}(a),
 we clearly see two plateau regions: one is after 4 rings
and one is after 10 rings. The first four rings represent the
typical frontal illuminations, which are present in most public
face datasets; however, we see that the residual stabilizes
after 10 rings which include some illuminations from the back
of the subject. This suggests that although the frontal
illuminations account for most of the illumination on the face,
some illuminations from the back are needed in the training set to
represent images with illumination coming from all directions.
In the plot for the granularity experiment, Figure
\ref{fig:illumination-sufficiency}(b), we observe that the
residual reaches a plateau after four divisions, corresponding
to a total of 32 illuminations. Based on the results from both
experiments, we decide to partition the area covered by the
first 10 rings into a total of 38 cells, whose layout is
explained in Figure \ref{fig:illumination-patterns}(b). For
our large-scale experiments, we have collected those
illuminations for all our subjects.\footnote{It is possible
that with further experimentation a reduced set of illuminations
can be found that performs as well or better.}

See below for the 38 training images of one subject:
\begin{figure}[h]
\centering
\includegraphics[width=6.3in]{figures_cvpr/training.png}
\vspace{-.3in}
\end{figure}

%\subsection{Should Positivity in the Representation Coefficients
%be Enforced?} One critical issue in linear illumination models
%is whether to enforce nonnegativity in the coefficients $\x$,
%i.e. whether to model illumination using a cone or a subspace.
%Some authors, i.e. \cite{Basri2003-PAMI}, have chosen to allow
%negative components in the coefficient vector $\x$ when
%representing one image as a sum of other images, and only
%enforce that the linear combination $A\x$ be positive
%everywhere.  Unfortunately, for non-convex objects such as
%faces, enforcing non-negativity of the representation is not
%sufficient to guarantee that the resulting image of the object
%is physical.  For example, consider a Lambertian planar object
%with a cylindrical hole drilled in it that is aligned with the
%optical axis of an orthographic camera.  Consider two images:
%one under illumination from a distant point source aligned with
%the hole, and one off of the hole axis.  Positively weighting
%the on-axis image and negatively weighting the on-axis image
%can result in an image where the bottom of the hole is more
%strongly illuminated than the surrounding plane.  The image is
%positive, but it is clearly non-physical.  It is worth
%emphasizing that this phenomenon is a result of some points on
%an object seeing a restricted subset of the illumination sphere
%(a violation of convexity),  and not a result of the
%directionality of the light chosen in the thought experiment.
%While the number of coefficients that actually end up negative
%after optimization is usually very small, we have observed
%pathological cases where the shadows next to the user's nose in
%the reconstructed image invert in value.

%Instead of using a smaller set of training images, allowing the
%coefficients to go negative to try to represent more test
%illuminations, and risking over-fitting, we decided to enforce
%non-negativity of the coefficients.  Nonnegative combinations
%of images are guaranteed to correspond to physically plausible
%images.  all physical illuminations unless the training images
%actually span the boundary of the illumination cone. Because we
%have a flexible acquisition system, we can directly generate a
%set of illuminations that span most of the illumination cone,
%without resorting to negative coefficients and risking
%overfitting.  Thus, in Algorithm 1, we have enforced that the
%coefficients $\x$ be non-negative.

%One of the main factors that complicates face recognition, and computer vision in general, is that two images of the same person's face can be very different, even if the pose is carefully controlled and all points on the person's face are visible in both images.  The most safe, reliable, and general assumption that can be made about the imaging process is that it is linear with respect to the intensity of different light sources.  The set of images of a given object is therefore a cone in the non-negative orthant of the pixel basis.  All recognition algorithms that rely on vision for training rely on images sampled from this cone.

\section{Tests on Public Databases}\label{sec:multipie}
In this section and the next section, we conduct comprehensive
experiments on large-scale face databases to verify the performance
of our algorithm and system. We
first test on the largest public face database available that
is suitable for testing our algorithm, the CMU Multi-PIE.
One shortcoming of the CMU Multi-PIE database for our purposes
is that there is no separate set of test images taken under natural illuminations;
we are left to choose which sets of images to use for testing and training.
To challenge our algorithm, we choose only a small set of illuminations for the training set,
yet we include all illuminations in the testing set. In the following section, we will test our
algorithm on a face dataset that is collected by our own
system. The goal for that experiment will be to show that with a sufficient set of
training illuminations for each subject, our algorithm indeed
works stably and robustly with practical illumination,
misalignment, pose, and occlusion, as already indicated by our
experiment shown in Figure \ref{fig:promo}(bottom).

CMU Multi-PIE provides the most extensive test set among public
datasets. This database contains images of 337 subjects across
simultaneous variation in pose, expression, and illumination.
Of these 337 subjects, we use all of the 249 subjects present
in Session 1 as the training set. The remaining 88 subjects are
treated as ``impostors'', or invalid images. For each of the
249 training subjects, we include frontal images of 7 frontal
illuminations,\footnote{They are illuminations
$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
directional illumination, we subtract the ambient-illuminated
image 0.} taken with neutral expression. As suggested by the
work of \cite{Georghiades2001-PAMI}, these extreme frontal
illuminations would be sufficient to linearly represent other
frontal illuminations, as will also be corroborated by the next
experiment on our own dataset. For the test set, we use all 20
illuminations from Sessions 2-4, which were recorded at
distinct times over a period of several months. The dataset is
challenging due to the large number of subjects, and due to
natural variation in subject appearance over time. Table
\ref{tab:MultiPIE-recognition} shows the result of our
algorithm on each of the 3 testing sessions. Our algorithm
achieves recognition rates above $90\%$ for all three sessions,
with input {\em directly} obtained from the Viola and Jones'
face detector; there was no manual intervention for the test images.
To demonstrate that the sparse representation based recognition step is indeed beneficial
even when there are no impostors, we include results for recognition
based only on the alignment error residuals(i.e. S=1), shown in row 6.

\begin{table}
\caption{Recognition rates on large-scale Multi-PIE database.}
\vspace{-.2in}
\centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & Session 2 & Session 3 & Session 4  \\
\hline
\hline
LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
\hline
NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
\hline
NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
\hline
LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
\hline
\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
\hline
{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
\hline
{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
\hline
{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
\hline
\end{tabular}
\label{tab:MultiPIE-recognition} }
\end{table}
\begin{figure}
\centering
{\small
\begin{tabular}{cccccc}
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{-.1in} \\
(a) & (b) & (c) & (d) & (e) & (f) \vspace{-.1in}
\end{tabular}
}
\caption{{\bf Representative failures from Multi-PIE}. {\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2. Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
\label{fig:failed-examples}
\vspace{-.5in}
\end{figure}

\subsection{Improving the Performance by Choosing Different Sampling Windows}
Our algorithm's errors are mostly caused by a few subjects who
significantly change their appearances between sessions (such
as hair, facial hair, and eyeglasses). Some representative
examples are shown in Figure \ref{fig:failed-examples}. For those subjects, alignment and recognition fail on
almost all test illuminations.

Meanwhile, this observation also suggests that we might be able
to improve the performance of our method by carefully choosing
a face region which is less affected by the above factors for
recognition. In particular, since the forehead region is likely
to be affected by the change of hair style, we try replacing
the previous $80 \times 60$ canonical frame with a new
window that better excludes the forehead. We adjust the
resolution of the window to keep $m$ approximately constant. In addition,
we cut off two lower corners of the $80 \times 60$ canonical frame, motivated by
the observation that in many cases the corners
actually contain background. An example of the new window
is shown in Figure~\ref{fig:new-mask}.

\begin{figure}
\centering
{\small
\begin{tabular}{cc}
\includegraphics[height=1.5in]{figures_pami/example.png} &
\includegraphics[height=1.5in]{figures_pami/example_new.png} \vspace{-.1in} \\
(a) & (b) \vspace{-.1in}
\end{tabular}
}
\caption{{\bf Choosing different sampling windows.} (a) The default window. (b) The newly proposed window.}
\label{fig:new-mask}
\vspace{-.3in}
\end{figure}

Table \ref{tab:MultiPIE-recognition} shows that the
recognition rates on Multi-PIE indeed increase with this new
window. In addition, Figure \ref{fig:failed-examples}(a), (b),
and (c) illustrate three representative subjects for which the
recognition rates of our algorithm are significantly boosted
with the new window. However, we should mention that the best choice of the window
is problem-specific and there is
not a simple guideline to follow. For example, although the new
window performs better on Multi-PIE, the same window does not
help at all on our own database, which will be introduced in
the next section. This is because most of the training and
testing images in our database are taken on the same day so the
variation in hair style is very small. Hence, excluding the
forehead part may actually result in loss of useful
discriminative information.

\subsection{Comparison to Existing Work}
We compare our result to baseline linear-projection-based
algorithms, such as Nearest Neighbor (NN), Nearest Subspace
(NS) \cite{Lee2005-PAMI}, and Linear Discriminant Analysis
(LDA) \cite{Belhumeur1997-PAMI}.\footnote{We do not list
results on PCA \cite{Turk1991-CVPR} as its performance is
always below that of Nearest Subspace.} Since these algorithms
assume pixel-accurate alignment, they are not expected to work
well if the test is not well aligned with the training. In
Table~\ref{tab:MultiPIE-recognition}, we report the results of
those algorithms with two types of input: 1.\ the output of the
Viola and Jones' detector, indicated by a subscript ``$d$'';
2.\ the input face is aligned to the training with manually
selected outer eye corners, indicated by a subscript ``$m$''.
Notice that even with careful manual registration, these
baseline algorithms perform significantly worse than our
algorithm, which uses input directly from the face detector.
The performance drop of the LDA algorithm on Multi-PIE reported here
seems to agree with that reported already in
\cite{Gross2008-FGR}.

We also compare our result to the very recent work
\cite{Yang2010-CVPR}. Notice that in \cite{Yang2010-CVPR}, the
initial registration is again obtained from manually selected
outer eye corners. Then, a supervised hierarchical sparse
coding model based on local image descriptors is trained, which
enjoys certain translation invariant properties. With the same
training and testing sets, \cite{Yang2010-CVPR} is able to
handle any remaining misalignment and achieves state-of-the-art
performance on the CMU Multi-PIE database.
Table~\ref{tab:MultiPIE-recognition} shows that our algorithm
achieves similar or better performance on different sessions of
Multi-PIE.

Finally, we compare our result to Local Binary Patterns (LBP)
\cite{Ahonen2006-PAMI}, a local appearance descriptor which is
able to capture fine details of facial appearance and texture.
Due to its robustness to variations in illumination, facial
expression, aging and other changes, LBP has achieved the
state-of-the-art face recognition performance in the scenario
when only one sample per person is used for training
\cite{Tan06facerecognition}. In this paper, we follow the same
steps as in \cite{Ahonen2006-PAMI} to construct an LBP
descriptor for each training and testing sample. The $80\times
60$ face region is first divided into a regular $10\times 10$
grid of cells, each of size $8\times 6$ pixels. Within each
cell, the histogram of 59 uniform binary patterns is then
computed, where the patterns are generated by thresholding 8
neighboring pixels in a circle of radius 2 using the central
pixel value. Finally, the local histograms are concatenated to
produce the global descriptor vector. As suggested in
\cite{Ahonen2006-PAMI}, the recognition is performed using a
nearest neighbor classifier with Chi square distance as the
distance measure and we report the recognition rates with the
same two types of input as before. As shown in
Table~\ref{tab:MultiPIE-recognition}, LBP achieves competitive
recognition rates given manually aligned training and testing
samples, which demonstrates its robustness to illumination
changes. However, its performance is very sensitive to
misalignment.

Although our algorithm is not designed for
recognition when there is only a single gallery image per user,
we compare its
performance with LBP within this setting for completeness.
For this experiment, we use the FERET dataset
\cite{phillips1998feret}, which contains five standard
partitions: `fa' is the gallery containing 1196 frontal images
of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
sets of probe images. The testing sets differ from the training
in facial expression (`fb'), illumination (`fc'), aging (`dup1'
) and long aging (`dup2'). In fact, except for `fb', we notice
significant changes of illumination in all the other three test
sets. For the training, we again crop and normalize the face
region from each original image to an $80\times 60$ window
using manually marked eye coordinates \cite{Deng2010-PR}. In
Table~\ref{tab:FERET-recognition}, we report the performance of
our algorithm on the four test sets, with input directly
obtained from the Viola and Jones' detector, and the
performance of LBP with input obtained from both face detector
and manual alignment using the eye coordinates.

\begin{table}
\caption{Recognition rates on single gallery image FERET
dataset} \vspace{-.1in} \centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & fb & fc & dup1 & dup2 \\
\hline
\hline
LBP$_d$ (LBP$_m$) & 54.8 (96.6)\% & 10.3 (58.8)\% & 29.8 (71.6)\% & 19.8 (61.5)\% \\
\hline
{Algorithm 1} & 95.2\% & 28.4\% & 46.1\% & 20.3\% \\
\hline
\end{tabular}
\label{tab:FERET-recognition} } \vspace{-.4in}
\end{table}

As expected, our algorithm does not perform well except for
`fb', in which the illumination is similar to the training and 
the mere variation in facial expression is handled well by the sparse error 
model. For the other three test sets, our algorithm fails because the illumination 
changes and other variations seriously violate the assumptions of our method. 
On the other hand, while LBP achieves the best recognition rates given
manually aligned training and testing samples, its performance
degrades drastically when the input is obtained directly from
the face detector. It is also worth noting that similar poor
performance of LBP, as well as other descriptors, has been
observed on the Labeled Face in the Wild (LFW) database, where
the training is uncontrolled and limited and the input is
directly obtained from the face detector \cite{Wolf2008-ECCV}.

All of these experimental results confirm that
both illumination and alignment need to be simultaneously handled
well in order to achieve accurate face recognition, even when there is
no obvious occlusion or corruption in the test.

\subsection{Subject Validation}

We test the algorithms' ability to reject invalid images of the
88 subjects not appearing in the training database. As
mentioned before, the \emph{sparsity concentration index} (SCI)
is used as the outlier rejection rule. Given the sparse
representation $\x$ of a test image with respect to $K$
training classes, the SCI measures how concentrated the
coefficients are on a single class in the dataset and is
defined as in \cite{Wright2009-PAMI}:
\begin{displaymath}
\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
\|\x\|_1 - 1}{K - 1} \in [0,1] .
\end{displaymath}
It is easy to see that if $\textup{SCI}(\x) = 1$, the test
image is represented using images from one single subject
class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
evenly over all classes. Thus, we can choose a threshold $t \in
[0,1]$ for the proposed method and accept a test image as valid
if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
invalid. We compare this classifier to classifiers based on
thresholding the error residuals of NN, NS, LDA, and LBP.

Figure \ref{fig:roc-multipie} plots the receiver operating
characteristic (ROC) curves, which are generated by sweeping
the threshold $t$ through the entire range of possible values
for each algorithm.\footnote{Rejecting invalid images not in
the entire database is much more difficult than deciding if two
face images are the same subject. Figure 8 should not be
confused with typical ROC curves for face similarity, e.g.,
\cite{PhillipsP2007}.} We can see that SCI based approach
significantly outperforms the other three algorithms. Similar
contrasts between our algorithm and baseline algorithms were
also observed for SRC in \cite{Wright2009-PAMI}, though on much
smaller datasets.

\begin{figure}
\centerline{
\includegraphics[width=3in]{figures_pami/pami_roc2.png}
}\vspace{-.2in}
\caption{{\bf ROC curves} for subject validation on Multi-PIE database.}\label{fig:roc-multipie}
\vspace{-.35in}
\end{figure}

\subsection{Recognition with Synthetic Random Block Occlusion}

We further test the robustness of our $\ell^1$-norm based
algorithm to synthetic occlusion. We simulate various levels of
occlusion from 10\% to 50\% by replacing a randomly located
block of the face image with an image of a baboon, as shown in
Figure~\ref{fig:multipie-occ-rec}. In this experiment, to avoid
any other factors that may contribute to extra occlusion of the
face (such as the change of hair style), we choose illumination
10 from Session 1\footnote{This is the same session as the
training set.} as testing. The rest of the experimental setting
remains unchanged. The table in
Figure~\ref{fig:multipie-occ-rec} shows that our algorithm is
indeed capable of handling a moderate amount of occlusion. For
example, at 20\% occlusion, our algorithm still achieves 94.9\%
recognition rate.\vspace{-.1in}

\begin{figure}[t]
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ10.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ20.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ30.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ40.png}
\includegraphics[scale=0.75,clip=true]{figures_pami/multipie_occ/occ50.png}
\vspace{-.5in}
\caption{{\bf Recognition under varying level of
random block occlusion.} The above row shows examples of occluded test images with occlusion level from 10\% to 50\%. Our method maintains high recognition rates up to 30\% occlusion:} \vspace{1mm}
{\small
\begin{tabular}{|c|c|c|c|c|c| }
\hline
Percent occluded & 10\% & 20\% & 30\% & 40\% & 50\%  \\
\hline
\hline
Recognition rate & 99.6\% & 94.9\% & 79.6\% & 46.5\% & 19.8\% \\
\hline
\end{tabular}
}
\label{fig:multipie-occ-rec}
\vspace{-.5in}
\end{figure}

\subsection{Recognition with Pose and Expression} We now run tests of
our algorithm on a subset of the images from Multi-PIE with pose and expression variation in the test set, although we do not model these variations explicitly.
Using the same training set as above, we test our algorithm on
images in Session 2 with $15^\circ$ pose, for all 20
illuminations. As expected, the recognition rate drops to 78.0\%. We also test our
algorithm on images in Session 3 with smile, again for all 20
illuminations. The recognition rate is 64.8\%. Of course, it is reasonable to expect that
the performance of our method will be significantly improved if pose and expression data
are available in the training.

\section{Tests on Our Own Database}\label{sec:own-data} Using the training acquisition
system we described in Section \ref{sec:illumination}, and shown in Figure
\ref{fig:system}, we have collected the frontal view of 109
subjects {\em without eyeglasses} under 38 illuminations shown
in Figure \ref{fig:illumination-patterns}. For testing our
algorithm, we have also taken 935 images of these subjects with
a different camera under a variety of practical conditions.

\subsection{Necessity of Rear Illuminations} To see how
training illuminations affect the performance of our algorithm
in practice, we now compare how well a few frontal
illuminations can linearly represent: 1. other frontal illuminations
taken under the same laboratory conditions, and 2. typical
indoor and outdoor illuminations. To this end, we use the face
database acquired by our system and use 7 illuminations per
subject as training. The illuminations are chosen to be similar
to the 7 illuminations used in the previous experiment on
Multi-PIE.\footnote{We use the illuminations $\{6, 9, 12,
13, 18, 21, 22\}$ shown in Figure
\ref{fig:illumination-patterns}(b) to mimic the illuminations
 $\{0, 1, 6, 7, 13, 14, 18\}$ in Multi-PIE.} We then test
our algorithm on the remaining $24 - 7 = 17$ frontal
illuminations for all the subjects. The recognition rate is
$99.8\%$, nearly perfect. We also test our algorithm on 310
indoor images and 168 outdoor images of these subjects taken
under a variety of lighting conditions (category 1 and 2
specified below), similar to the one shown in Figure
\ref{fig:promo}, and the recognition rates for indoor and
outdoor images drop down to $94.2\%$ and $89.2\%$,
respectively. This is a strong indication that 
frontal illuminations taken under laboratory conditions
are insufficient for representing test images under typical indoor and
outdoor illuminations.\vspace{-0.2in}

\begin{figure}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1318.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1521.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1673.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1732.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_1941.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_indoor/DSC_3766.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1574.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1622.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_1641.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3522.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3707.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/normal_outdoor/DSC_3772.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1397.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1532.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1556.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1585.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_1688.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/glasses/DSC_4035.JPG} 
\vspace{-.2in}
 \caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
 \vspace{-.1in}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_1565.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3656.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_3827.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4090.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4106.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses/DSC_4126.JPG} \\
\vspace{2mm}
\centering
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_1611.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3528.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3744.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_3995.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4030.JPG}
\includegraphics[scale=0.75,clip=true]{figures_pami/uiuc_example/sunglasses_failed/DSC_4095.JPG}
\vspace{-.2in}
 \caption{{\bf Representative examples of category C4}. Top row: successful examples with our method using overlapping blocks. Bottom row: failures with our method using overlapping blocks.}\label{fig:examples4}
\vspace{-.5in}
\end{figure}

\subsection{Large-Scale Test with Sufficient Training
Illuminations} Now we use all 109 subjects and 38 illuminations
in the training and test on 935 images taken under a variety of
practical illuminations and conditions. We have manually partitioned the test images into four main
categories:
\begin{description}
\item[C1:] 310 \emph{indoor} images of 72 subjects without
    eyeglasses, frontal view
    (Fig.~\ref{fig:examples1-3}, row 1).
\item[C2:] 168 \emph{outdoor} images of 48 subjects without
    eyeglasses, frontal view
    (Fig.~\ref{fig:examples1-3}, row 2).
\item[C3:] 211 images of 32 subjects with \emph{eyeglasses}
    (Fig.~\ref{fig:examples1-3}, row 3).
\item[C4:] 246 images of 56 subjects with \emph{sunglasses}
    (Fig.~\ref{fig:examples4}).
\end{description}
We apply Viola and Jones' face detector on these images and
directly use the detected faces as the input to our algorithm.
Table~\ref{tab:UIUC-recognition} reports the performance of our
algorithm on each category.
\begin{table}
\vspace{-.1in}	
\centering \caption{Recognition rates on our own database.}
\vspace{-.2in}
\begin{tabular}{|c|c|c|c|c| }
\hline
Test Category & C1 & C2 & C3 & C4  \\
\hline
\hline
Recognition Rate & 98.4\% & 95.8\% & 95.1\% & 40.9\% \\
\hline
\end{tabular}
\label{tab:UIUC-recognition}
\vspace{-.4in}
\end{table}
Since our focus is on face
recognition, the errors do not include failures of the face
detector on some of the more challenging images.
As one can see, our algorithm achieves higher than 95\%
recognition rates on categories 1-3. Furthermore, using the
full set of 38 illuminations indeed improves the performance of
our system under practical illumination conditions compared to
only using a small subset of 7 illuminations. However, the
performance dramatically drops when the faces are occluded by
various types of sunglasses, which could cover up to 40\% of
the entire face. Given the previous experimental results on
synthetic random block occlusions, and given that the
illuminations are more challenging, the result is not
surprising. In the next subsection, we will show how additional
assumptions can be used to improve the recognition performance.

\subsection{Improving the Performance with Occlusion using Overlapping Blocks}
A traditional approach to improve the performance of face
recognition under severe occlusion is to use subregions instead
the entire face as a whole. This idea has been explored in many
earlier works; see \cite{Pentland1994-CVPR, Wright2009-PAMI}
for examples. Since in most real world cases the occlusion is contiguous, it is reasonable to argue that a minority of the
subregions are likely to be affected by the occlusion. In this
paper, we adopt the same idea and partition the face into four
overlapping blocks to better handle sunglasses. This
scheme is illustrated in Figure~\ref{fig:occ-block}. Notice
that in this example three out of the four blocks are partially
or almost completely occluded. In our experiment, each block is
of size $90\times 48$ and covers about two-fifths of the entire
face. The testing and training sets are partitioned in the same
way. We then independently apply
Algorithm~\ref{alg:deformable-src} and compute a sparse
representation after registration for each block independently
with respect to the training set. The recognition
results for individual blocks are then aggregated by voting.

\begin{figure}
\centerline{
\includegraphics[width=4in]{figures_pami/occ_block.png}\vspace{-0.2in}}
\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses. (b) The four overlapping blocks. (c) The sparse representation is calculated after alignment for each block independently. The red lines correspond to his true identity. (d) The true identity is successfully recovered by voting based on the SCI scores.}
\label{fig:occ-block}
\vspace{-.5in}
\end{figure}

In this experiment, we found that the using the \emph{sparsity
concentration index} (SCI) scores for voting achieves higher
recognition rate than the residual measure used in Algorithm~\ref{alg:deformable-src}, on
category 4 (sunglasses) of our database. The recognition rate
is increased to 78.3\%, compared to 40.9\% obtained without
this partition scheme. This is another evidence of the superior
ability of SCI on subject validation, since a heavily occluded
block can be regarded as an outlier for recognition and should
be rejected while voting.

However, we should point out that a major problem with this
approach is that occlusion cannot always be expected to fall within
 any fixed partition of the face image. Therefore, the
proposed scheme should only be viewed as an example which shows
that the performance under occlusion can be boosted by
leveraging local information of a face as well as global information. We 
leave the investigation of more general models (e.g., MRF \cite{ZhouZ2009}) for face
recognition with both misalignment and occlusion as an
interesting future work.

\section{Conclusion}\label{sec:conclusion}
Using a well-though-out combination of existing ideas
(iterative image alignment, l1-error function, SRC, using projectors for
illumination), we have proposed a system for recognizing human faces
from images taken under practical conditions that is conceptually simple, well
motivated, and competitive with state-of-the-art recognition systems for access
control scenarios.

The system achieves extremely stable performance under
a wide range of variations in illumination, misalignment, and even under small amounts of
pose and occlusion. We achieve very good recognition performance on
large-scale tests with public datasets as well as our practical face
images, while using only frontal 2D images in the gallery and no
explicit 3D face model.
Our system could potentially be extended to better handle large pose
and expression, either by incorporating training images with different poses or
expressions or by explicitly modeling and compensating the associated deformations
in the alignment stage.

Another important direction for future
investigation is to extend the alignment algorithm to better
tackle contiguous occlusion. We have demonstrated that misalignment can be naturally handled within the
sparse representation framework. More complicated models for
spatial continuity, such as Markov random fields, have also
been successfully integrated into the computation of a sparse
representation of well-aligned test images
\cite{Cevher2008-NIPS, ZhouZ2009}. A unified approach
for face alignment and recognition in the presence of
contiguous occlusion remains an open problem.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%
% Use this command to get the appendices' numbers in "A", "B" instead of the
% default capitalized Roman numerals ("I", "II", etc.).
% However, the capital letter form may result in awkward subsection numbers
% (such as "A-A"). Capitalized Roman numerals are the default.
%\useRomanappendicesfalse
%
%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgement
\section*{Acknowledgment}
% optional entry into table of contents (if used)
%\addcontentsline{toc}{section}{Acknowledgment}
This work was supported by grants NSF IIS 08-49292, NSF ECCS
07-01676, and ONR N00014-09-1-0230. JW thanks Allen Yang of UC Berkeley EECS and Robert Fossum of UIUC
Mathematics for discussions related to this work, and 
acknowledges support from a Microsoft Fellowship and
the Lemelson-Illinois Student Prize. %} %\vspace{-0.2in}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section
% NOTE: BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/

% can use a bibliography generated by BibTeX as a .bbl file
% standard IEEE bibliography style from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/IEEEtran/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,faces}
\bibliography{faces}

% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% biography section
%
% If you had an eps/pdf photo file (graphicx package needed)
% the extra braces prevent the LaTeX parser from getting confused
% when it sees the complicated \includegraphics command within an
% optional argument. You can create your own macro to make things
% simpler here.
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell.eps}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%
%\begin{biography}{Andrew Wagner}
%Biography text here.
%\end{biography}
%
%\begin{biography}{John Wright}
%Biography text here.
%\end{biography}
%
%\begin{biography}{Arvind Ganesh}
%Biography text here.
%\end{biography}
%
%\begin{biography}{Zihan Zhou}
%Biography text here.
%\end{biography}
%
%\begin{biography}{Yi Ma}
%Biography text here.
%\end{biography}
%

% if you will not have a photo
%\begin{biographynophoto}{John Doe}
%Biography text here.
%\end{biographynophoto}

% insert where needed to balance the two columns on the last page
%\newpage

%\begin{biographynophoto}{Jane Doe}
%Biography text here.
%$\end{biographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2011-03-15 main.tex
In the middle of working on changing terminology

@@
 at the higher level there are multiple alignent problems that are solved
-independently.  At a lower level, the first-order linear algebraic operations
-exhibit parallelism within image operations, i.e. at a pixel level.
+independently, or {\bf problem-level} parallelism.  At a lower level, the first-order linear algebraic operations
+exhibit parallelism within image operations, i.e. at a pixel level.  We term
+this {\bf pixel-level} parallelism.
 
@@
 
+One of the main contributions of this paper is a more efficent mapping of the
+two levels of parallelism available in iterative face alignment to the two
+levels of concurrency available in atypical multicore CPU architecture.
+
+We argue that, in constrast with previous implementations, which mapped problem
+level parallelism onto core-level concurrency, it is more efficient to
+completely disregard problem-level parallelism, and instead map pixel-level
+parallelism onto both core-level and vector-level cpu concurrency.  In other words,
+the implementation we advocate performs alignment of the test image against the gallery
+images of a single user at a time.  Each matrix-vector and vector-vector operation
+gets broken into 8 parts that run one per core, and into 4 single float vectors
+that execute on the SSE SIMD units.
+
+Why is this
+implementation faster?  Recall that the L2 cache can old up to 128 image
+vectors' worth of data.  Since we're only aligning data for one gallery user at a
+time the working set of the problem fits into L2 cache, resulting in
+significantly higher image alignment speed than previous cpu implementations.
+
+In our proposed CPU implementation of the iterative alignment stage, we
+therefore perform parallelization in a manner similar to the matlab code:
+solving a single problem instance at a time, relying on the threaded version of
+MKL BLAS for both multi-core and vector parallelism.
+
 This section discusses how the iterative alignment problem can be implemented
-on CPU hardware.  The implementations presented in \cite{WagnerA2009-CVPR} and
-\cite{WagnerA2011-PAMI} were written using the OpenCV library.
+on CPU hardware.
 
-OpenCV wraps either the operating system's built-in BLAS library, or Math
-Kernel Library (MKL), Intel's BLAS implementation, and thus achieves vector
-(i.e. pixel)-level parallelism implicitly.  The authors of
-\cite{WagnerA2011-PAMI} used OpenMP to schedule instances of the alignment
-problem to the CPU cores in a 1:1 fashion.  With eight alignment problems for
-ten training users being solved concurrently, and with 20-40 images per
-training user, all of the required data fits into L3 cache.  This provided a
-significant, though sub-linear improvement over a single-core implementation.
+{\bf Comparison to Previous Implementations}
+The implementations presented in \cite{WagnerA2009-CVPR} and
+\cite{WagnerA2011-PAMI} were written using the OpenCV library.  OpenCV wraps
+either the operating system's built-in BLAS library, or Math Kernel Library
+(MKL), Intel's BLAS implementation, and thus achieves vector (i.e. pixel)-level
+parallelism implicitly.  The authors of \cite{WagnerA2011-PAMI} used OpenMP to
+schedule instances of the alignment problem to the CPU cores in a 1:1 fashion.
+With eight alignment problems for ten training users being solved concurrently,
+and with 20-40 images per training user, all of the required data fits into L3
+cache.  This provided a significant, though sub-linear improvement over a
+single-core implementation.
 
@@
 \end{enumerate}
-
-In the proposed CPU implementation of the iterative alignment stage, we
-therefore perform parallelization in a manner similar to the matlab code:
-solving a single problem instance at a time, relying on the threaded version of
-MKL BLAS for both multi-core and vector parallelism.  Why is this
-implementation faster?  Recall that the L2 cache can old up to 128 image
-vectors' worth of data.  Since we're only aligning data for 1 gallery user at a
-time the working set of the problem fits into L2 cache, resulting in
-significantly higher image alignment speed than previous cpu implementations.
 
@@
 
-In this section, we benchmark the performance of our algorithm on CPU and GPU
-platforms.  In order to show how our $\ell_1$ minimization algorithms scale
-with problem size, in Section \ref{sec:simulation} we will begin with
-benchmarks on synthetic data.  We will then progress in Sections
-\ref{sec:alignment_benchmark} and \ref{sec:recognition_benchmark} to
-demonstrating the speed of our implementations as applied to the alignment and
-recognition stages of the face recognition problem.
+In this section, we benchmark the performance of our parallel implementations
+of $\ell_1$ minimization on CPU and GPU platforms.  In order to show how our
+$\ell_1$ minimization algorithms scale with problem size, in Section
+\ref{sec:simulation} we will begin with benchmarks on synthetic data.  We will
+then progress in Sections \ref{sec:alignment_benchmark} and
+\ref{sec:recognition_benchmark} to demonstrating the speed of our
+implementations as applied to the alignment and recognition stages of the face
+recognition problem.
 

### 2011-03-16 main.tex
Working on standardizing terminology

@@
 
-\section{Hardware Parallelism} \label{sec:parallelism} 
-In this section we discuss the hardware architectures used in this paper. Aside
-from the $\ell_1$-min algorithm, our choice of the architectures plays a
-significant role in many practical decisions we make in Section
-\ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
-the system makes use of all the cores in as many CPUs are present on one
-motherboard.  For GPU implementation, the system makes the best use of a
-single many-core Nvidia GPU chip on a PCI card (GPUs are commonly shipped with
-a single GPU chip per PCI-express card).  Unless otherwise specified, all
-implementations utilize single precision floating point datatypes.  Since on-chip
-memory (cache) is often orders of magnitude faster than off-chip memory, the sizes
-and speeds of on-chip memory (cache) have big performance implications.  We give
-a brief overview of the caches that are available in our target architectures,
-and defer discussion of their performance implications to sections
-\ref{sec:alignment_implementation_cpu} and
+\section{Hardware Concurrency} \label{sec:parallelism}
+
+In this paper we will clearly distinguish between algorithm {\em parallelism},
+which is a property of an algorithm, and hardware {\em concurrency}, which is a
+property of a hardware architecture.  Parallelism in an algorithm provides the
+opportunity to perform computations concurrently on hardware.  Algorithms often
+exhibit multiple levels of parallelism, and hardware often provides multiple
+levels of concurrency.  In these terms, the primary contribution of this paper
+will be to determine the optimal mappings between the parallelism available in
+$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU architectures.
+
+In this section we discuss the levels of concurrency available in the hardware
+architectures considered in this paper, as well as other aspects of the
+hardware that are important for performance.  In particular, since on-chip
+memory (cache) is often orders of magnitude faster than off-chip memory, the
+sizes and speeds of on-chip memory (cache) have big performance implications.
+We give a brief overview of the caches that are available in our target
+architectures, and defer discussion of their performance implications to
+sections \ref{sec:alignment_implementation_cpu} and
 \ref{sec:alignment_implementation_gpu}.
 
-\subsection{CPU Hardware Parallelism} 
+We will primarily be comparing CPU and GPU architectures on a per-board basis:
+for CPU implementation we make use of all the cores in as many CPUs are present
+on one motherboard, and for GPU implementation we make use of a single
+many-core Nvidia GPU chip on a PCI card. GPUs are commonly shipped with a
+single GPU chip per PCI-express card.  Note that most engineering  workstations
+are capable of supporting several GPU cards.  Unless otherwise specified, all
+implementations utilize single precision floating point datatypes.
+
+\subsection{CPU Hardware Concurrency} 
+The main defining characteristics of contemporary multi-core CPU architectures
+are that they have two levels of concurrency, relatively large amounts of
+on-chip cache, and relatively high clock speeds.
+
 The baseline architecture for our experiments is a Linux workstation with two
@@
 
-There are thus two important levels of parallelism that need to be exploited to
-efficiently use a modern CPU: core-level parallelism and vector-level
-parallelism. CPU tools for exploiting these two levels of parallelism are
+There are thus two important levels of concurrency that need to be exploited to
+efficiently use a modern CPU: {\em core-level} concurrency and {\em vector-level}
+concurrency. CPU tools for exploiting these two levels of concurrency are
 relatively mature.  Heavily optimized Basic Linear Algebra Subprograms (BLAS) are
-commercially available that make good use of both levels of CPU parallelism.
+commercially available that make good use of both levels of CPU concurrency.
 For operations that do not fit well with the BLAS API, we 
-achieve both vector-level and core-level parallelism through a combination of
+achieve both vector-level and core-level concurrency through a combination of
 the OpenMP API \cite{dagum2002openmp} and the automatic vectorization 
@@
 % info anyway.
-\subsection{GPU Hardware Parallelism}
+\subsection{GPU Hardware Concurrency}
+The main defining characteristics of contemporary multi-core GPU architectures
+are that they have two (much wider) levels of concurrency, relatively small amounts of
+on-chip cache, and relatively low clock speeds.
+
 Compared to CPU architectures, GPU architectures dedicate less of their transistors
@@
 
-\section{Parallelization of the Face Alignment Stage}
+\section{Parallelism in the Face Alignment Stage}
 \label{sec:alignment}
-%{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
-This section provides a brief explanation of the iterative alignment stage
-of the recognition pipeline.  
-The inputs of the algorithm are the test image $f(x,y)$ stacked into an $M$-
+There are essentially two levels of parallelism available for exploitation in
+for iterative alignment in the face recognition application.  The primary goal
+of this section is to introduce these two levels of parallelism.  Additionally,
+in order for some important aspects of our implementations to be clear later,
+we need to establish how the iterative alignment stage works in some
+detail.
+
+The inputs of the iterative alignment stage are the test image $f(x,y)$ stacked into an $M$-
 dimensional vector $\f$, and the aligned and stacked training images $A_i$ for
@@
 and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
-once per test image.  The iterative alignment stage is summarized as Algorithm \eqref{alg:iterative_alignment}.
+once per test image.  The per-user iterative alignment algorithm is summarized as Algorithm \eqref{alg:iterative_alignment}.
 \begin{algorithm}[thb]
-\caption{\bf (Iterative Alignment Stage)} \label{alg:iterative_alignment}
+\caption{\bf (Per-User Iterative Alignment)} \label{alg:iterative_alignment}
 \begin{algorithmic}[1]
 \begin{small}
-\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
+\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n}\}$ for gallery subject $i$, the test image
 $\f\in\Re^M$ 
 %\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
-\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
-\STATE {\bf for} each gallery subject $i$, 
-\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
-\STATE \hspace{6mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
-\STATE \hspace{6mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
-\STATE \hspace{6mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
-\STATE \hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
+%\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
+%\STATE {\bf for} each gallery subject $i$, 
+\STATE \hspace{0mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
+\STATE \hspace{3mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
+\STATE \hspace{3mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
+\STATE \hspace{3mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
+\STATE \hspace{3mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
 \tau$; 
@@
 
-\subsection{Computational Complexity} Any discussion of the speed of an
+\subsection{Computational Complexity and Parallelism} Any discussion of the speed of an
 algorithm should discuss the comutational complexity of the algorithm.  For the
@@
 
-I summary, the alignment stage essentially two levels of available parallelism:
-at the higher level there are multiple alignent problems that are solved
-independently, or {\bf problem-level} parallelism.  At a lower level, the first-order linear algebraic operations
-exhibit parallelism within image operations, i.e. at a pixel level.  We term
-this {\bf pixel-level} parallelism.
+In summary, the alignment stage essentially two levels of available
+parallelism: at the higher level there are multiple iterative alignent problems
+that are solved independently, or {\bf problem-level} parallelism.  At a lower
+level, the first-order linear algebraic operations exhibit parallelism within
+image operations, i.e. at a pixel level.  We term this {\bf pixel-level}
+parallelism.
 
@@
 \label{sec:alignment_implementation_cpu}
-
 One of the main contributions of this paper is a more efficent mapping of the
 two levels of parallelism available in iterative face alignment to the two
-levels of concurrency available in atypical multicore CPU architecture.
+levels of concurrency available in a typical multicore CPU architecture.
 

### 2011-03-17 main.tex
Added some new figures temporarily at the bottom of the paper.

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_old}
+\caption{Old mapping of alignment parallelism onto cpu concurrency}
+\label{fig:alignment_mapping_old}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_cpu}
+\caption{Proposed mapping of alignment parallelism onto cpu concurrency}
+\label{fig:alignment_mapping_cpu}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
+\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
+\label{fig:alignment_mapping_gpu}
+\end{figure}
+
 {\small

### 2011-03-22 main.tex
added caption
revised the experimental section

@@
 \ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
-nicely onto the BLAS API (such as SGEMV, SAXPY {\bf? Victor, fill in here}), we
+nicely onto the BLAS API (such as SGEMV, SAXPY), we
 call Intel's MKL BLAS library, which internally takes advantage of the SSE
@@
 Even for operations that could have been implemented via multiple BLAS calls,
-this was found to have better performance due to reducede bandwidth and kernel
-call overhead {\bf Victor, verify or correct this}.
+this was found to have better performance due to reduced bandwidth and kernel
+call overhead {\bf Victor, verify or correct this.  I'm not sure, currently asking Mark}.
 
@@
 randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
-$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
+$10:1$ to $10:5$ with $d$ varying from 500 to 8000.
 The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
-Because the ground truth signal $\xx_0$ is known, the algorithm terminates
+Because the ground truth signal $\xx_0$ was known, the algorithm terminated
 when $norm(\bx-\bx_0) < tolerance$.
 The measurement vector was generated with the equation $\y_0 = A \x_0$.
-%We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
-We averaged the time taken for the ALM algorithm to solve 10 instances of a problem.
 
@@
 The x-axis represents the size of the $A$ matrix and the y-axis represents the
-average amount of time it to complete a single problem instance.  A can be seen
-in Figure \ref{fig:random_data}, the GPU implementation tends to be faster than
+average amount of time it to complete a single problem instance.  The GPU implementation tends to be faster than
 the CPU implementation at solving a single large problem, whereas the the CPU
-implementation is faster at solving a single small problem.  
+implementation is faster at solving a single small problem.  Figure \ref{fig:random_data2} shows the various sizes and tolerances at which our GPU implementation
+begins to run faster than the CPU implementation on synthetic data and will provide researchers 
+with a starting point of where to begin setting their tolerance levels for the ALM algorithm.
 \begin{figure}
@@
 \begin{center}
-\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_2}
+\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_3}
 \end{center}
-\caption{\bf CAPTION NEEDED}
+\caption{Relation between CPU and GPU runtime vs tolerance level}
 \label{fig:random_data2}

### 2011-03-23 main.tex
Converted to IJCB syle, which is really the cvpr style we started with.
Rewrote CPU alignment implementation section.  Not compiling.

--- 2011-03-17 main.tex
@@
 
-\usepackage{iccv}
+\usepackage{cvpr}
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
@@
 
-\def\iccvPaperID{257} % *** Enter the CVPR Paper ID here
+\def\cvprPaperID{257} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 % Pages are numbered in submission mode, and unnumbered in camera-ready
-\ificcvfinal\pagestyle{empty}\fi
+\ifcvprfinal\pagestyle{empty}\fi
 \begin{document}
@@
 %\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
-\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition
-%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}
-}
+\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
 
@@
 will be to determine the optimal mappings between the parallelism available in
-$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU architectures.
+$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU
+architectures.
 
@@
 
-With Hyperthreading enabled, each core can support two threads simultaneously.
+%With Hyperthreading enabled, each core can support two threads simultaneously.
 For floating point instructions, each core also has a vector processing unit
@@
 efficiently use a modern CPU: {\em core-level} concurrency and {\em vector-level}
-concurrency. CPU tools for exploiting these two levels of concurrency are
-relatively mature.  Heavily optimized Basic Linear Algebra Subprograms (BLAS) are
-commercially available that make good use of both levels of CPU concurrency.
-For operations that do not fit well with the BLAS API, we 
-achieve both vector-level and core-level concurrency through a combination of
-the OpenMP API \cite{dagum2002openmp} and the automatic vectorization 
-facilities of Intel's C++ compiler (ICC)\cite{dulong1999overview}.
-% DW: I'm really not sure these citations are even really relevant;
-% our target audience will just google icc openmp and get more up-to-date
-% info anyway.
+concurrency.
+
 \subsection{GPU Hardware Concurrency}
@@
 image operations, i.e. at a pixel level.  We term this {\bf pixel-level}
-parallelism.
+parallelism.  In the following two sections will will discuss how to
+exploit these two levels of parallelism on cpu and gpu architectures.
 
@@
 \label{sec:alignment_implementation_cpu}
-One of the main contributions of this paper is a more efficent mapping of the
-two levels of parallelism available in iterative face alignment to the two
-levels of concurrency available in a typical multicore CPU architecture.
+This section discusses how to efficently map the two levels of parallelism
+available in iterative face alignment to the two levels of concurrency
+available in a typical multicore CPU architecture.
 
-We argue that, in constrast with previous implementations, which mapped problem
-level parallelism onto core-level concurrency, it is more efficient to
+We will show that, in constrast with previous implementations, which mapped
+problem level parallelism onto core-level concurrency, it is more efficient to
 completely disregard problem-level parallelism, and instead map pixel-level
-parallelism onto both core-level and vector-level cpu concurrency.  In other words,
-the implementation we advocate performs alignment of the test image against the gallery
-images of a single user at a time.  Each matrix-vector and vector-vector operation
-gets broken into 8 parts that run one per core, and into 4 single float vectors
-that execute on the SSE SIMD units.
+parallelism onto both core-level and vector-level cpu concurrency.  In other
+words, the cpu implementation we advocate performs alignment of the test image
+against the gallery images of a single user at a time.  Each matrix-vector and
+vector-vector operation gets broken into 8 parts that run one per core, and
+into 4 single float vectors that execute on the SSE SIMD units.
+This mapping of the alignment problem onto the concurrency provided by the pair
+of CPUs is visualized in Figure \ref{fig:alignment_mapping_cpu}.
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_cpu}
+\caption{Proposed mapping of alignment parallelism onto cpu concurrency}
+\label{fig:alignment_mapping_cpu}
+\end{figure}
 
-Why is this
-implementation faster?  Recall that the L2 cache can old up to 128 image
-vectors' worth of data.  Since we're only aligning data for one gallery user at a
-time the working set of the problem fits into L2 cache, resulting in
-significantly higher image alignment speed than previous cpu implementations.
+Recall that the L2 cache can old up to 128 image vectors' worth of data.  Since
+we're only aligning data for one gallery user at a time the working set of the
+problem fits into L2 cache.  Due to the higher bandwidth of L2 cache relative
+to L3 cache, this results in significantly higher image alignment speed than
+previous cpu implementations, as will be seen and discussed further in section
+\ref{sec:alignment_benchmark}.
 
-In our proposed CPU implementation of the iterative alignment stage, we
-therefore perform parallelization in a manner similar to the matlab code:
-solving a single problem instance at a time, relying on the threaded version of
-MKL BLAS for both multi-core and vector parallelism.
+{\bf Implementation Details} In our implementation, most of the operations
+required by the alignment stage take advantage of Intel Math Kernel Library
+(MKL), a commercial implementation of the standard Basic Linear Algebra
+Subprograms (BLAS).  For image resampling, we make use of Intel's Integrated
+Performance Primitives (IPP) library.  Both of these libraries are heavily
+optimized for Intel multicore CPU's, and are able to make good use of both
+core-level and vector-level concurrency.  For operations that do not map well
+onto existing libraries, we achieve both vector-level and core-level
+concurrency through a combination of the OpenMP API \cite{dagum2002openmp} and
+the automatic vectorization facilities of Intel's C++ compiler
+(ICC)\cite{dulong1999overview}.
 
-This section discusses how the iterative alignment problem can be implemented
-on CPU hardware.
+{\bf Comparison to Previous Implementations} The implementation presented in
+\cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} schedules instances of the
+alignment problem to the cpu cores in a 1:1 fashion.  
+This mapping of the alignment problem onto the concurrency provided by the pair
+of CPUs is visualized in Figure \ref{fig:alignment_mapping_old}.
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_old}
+\caption{Old mapping of alignment parallelism onto cpu concurrency}
+\label{fig:alignment_mapping_old}
+\end{figure}
 
-{\bf Comparison to Previous Implementations}
-The implementations presented in \cite{WagnerA2009-CVPR} and
-\cite{WagnerA2011-PAMI} were written using the OpenCV library.  OpenCV wraps
-either the operating system's built-in BLAS library, or Math Kernel Library
-(MKL), Intel's BLAS implementation, and thus achieves vector (i.e. pixel)-level
-parallelism implicitly.  The authors of \cite{WagnerA2011-PAMI} used OpenMP to
-schedule instances of the alignment problem to the CPU cores in a 1:1 fashion.
-With eight alignment problems for ten training users being solved concurrently,
-and with 20-40 images per training user, all of the required data fits into L3
-cache.  This provided a significant, though sub-linear improvement over a
-single-core implementation.
+This implementation was written using the OpenCV library
+as a wrapper for MKL, and, like our proposed implementation, achieves vector-level concurrency implicitly.  
+With eight alignment problems (associated with 8 training users) being solved
+concurrently, and with 20-40 images per training user, all of the required data
+fits into L3 cache.  While the L3 cache has much higher bandwidth and lower latency
+than off-chip memory, this implementation is slower than the proposed implementation.
 
-During the course of development, many implementation strategies were tried.
-Listed in order of increasing performance based on a synthetic benchmark of
-the $\ell_1$ minimization step, they included:
-\begin{enumerate}
-\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
-\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
-\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
-\end{enumerate}
+%During the course of development, many implementation strategies were tried.
+%Listed in order of increasing performance based on a synthetic benchmark of
+%the $\ell_1$ minimization step, they included:
+%\begin{enumerate}
+%\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
+%\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
+%\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
+%\end{enumerate}
 
@@
 hardware will allow.
+
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
+\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
+\label{fig:alignment_mapping_gpu}
+\end{figure}
 
@@
 \end{center}
-\caption{\bf CAPTION NEEDED}
+\caption{Relation between CPU and GPU runtime vs tolerance level}
 \label{fig:random_data2}
@@
 
-\subsection{Face Alignment Stage Benchmark} \label{sec:alignment_benchmark}
+\subsection{Face Alignment Stage Benchmark} 
+\label{sec:alignment_benchmark}
 This section presents benchmarks of our CPU and GPU implementations of the
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_old}
-\caption{Old mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_old}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_cpu}
-\caption{Proposed mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_cpu}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
-\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
-\label{fig:alignment_mapping_gpu}
-\end{figure}
-
 {\small
@@
 }
-
-\end{document}

### 2011-03-23 main.tex
Fixed missing \end{document}
Updated figure

@@
 \label{sec:alignment_implementation_gpu}
+This section discusses how to efficently map the two levels of parallelism
+available in iterative face alignment to the two levels of concurrency
+available in the NVIDIA Fermi GPU architecture.
 
-This section discusses how to implement algorithm
-\eqref{alg:iterative_alignment} on the GPU.  Like for the CPU case, many GPU
-implementation strategies were explored during the course of our development,
-and are only mentioned briefly.  Listed in order of increasing performance
-based on a synthetic benchmark of the $\ell_1$ minimization step, they
-included:
+Our proposed GPU implementation for the iterative alignment stage consists of a
+single kernel that performs all of the test image alignment problems.  Each of
+instance of the alignment problem is assigned its own thread block, and the GPU
+hardware schedules as many of these thread blocks to run concurrently as the 
+hardware will allow.  
+This mapping of the alignment problem onto the concurrency provided by the pair
+of CPUs is visualized in Figure \ref{fig:alignment_mapping_gpu}.
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
+\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
+\label{fig:alignment_mapping_gpu}
+\end{figure}
+The full resolution test image, along with its numerical derivatives in both
+image directions, are stored in GPU texture memory. This  allows the
+recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
+versions of the test image and its derivatives, to exploit the GPU's special
+purpose hardware for bilinear interpolation.
+
+\subsection{Other Implementation Strategies}
+Several GPU implementation strategies were explored during the course of our
+development, and are only mentioned briefly.  Listed in order of increasing
+performance based on a synthetic benchmark of the $\ell_1$ minimization step,
+they included:
 \begin{enumerate}
@@
 % that they're done
-
-Our proposed GPU implementation for the iterative alignment stage consists of a
-single kernel that performs all of the test image alignment problems.  Each of
-instance of the alignment problem is assigned its own thread block, and the GPU
-hardware schedules as many of these thread blocks to run concurrently as the 
-hardware will allow.
-
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
-\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
-\label{fig:alignment_mapping_gpu}
-\end{figure}
-
-The full resolution test image, along with its numerical derivatives in both
-image directions, are stored in GPU texture memory. This  allows the
-recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
-versions of the test image and its derivatives, to exploit the GPU's special
-purpose hardware for bilinear interpolation.
 
@@
 units on all available CPU cores.
-{\bf TODO: Our current implementation links against the MKL
-library provided internally with matlab; we are in the process of updating the code to link directly against a newer version
-of MKL}
 
@@
 }
+
+\end{document}

### 2011-03-23 main.tex

@@
 \begin{center}
-\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_2}
+\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_3}
 \end{center}

### 2011-03-25 main.tex
updated section 6

@@
 
-\usepackage{cvpr}
+\usepackage{iccv}
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
@@
 
-\def\cvprPaperID{257} % *** Enter the CVPR Paper ID here
+\def\iccvPaperID{257} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 %\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
-\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
+\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition
+%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}
+}
 
@@
 will be to determine the optimal mappings between the parallelism available in
-$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU
-architectures.
+$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU architectures.
 
@@
 
-%With Hyperthreading enabled, each core can support two threads simultaneously.
+With Hyperthreading enabled, each core can support two threads simultaneously.
 For floating point instructions, each core also has a vector processing unit
@@
 efficiently use a modern CPU: {\em core-level} concurrency and {\em vector-level}
-concurrency.
-
+concurrency. CPU tools for exploiting these two levels of concurrency are
+relatively mature.  Heavily optimized Basic Linear Algebra Subprograms (BLAS) are
+commercially available that make good use of both levels of CPU concurrency.
+For operations that do not fit well with the BLAS API, we 
+achieve both vector-level and core-level concurrency through a combination of
+the OpenMP API \cite{dagum2002openmp} and the automatic vectorization 
+facilities of Intel's C++ compiler (ICC)\cite{dulong1999overview}.
+% DW: I'm really not sure these citations are even really relevant;
+% our target audience will just google icc openmp and get more up-to-date
+% info anyway.
 \subsection{GPU Hardware Concurrency}
@@
 image operations, i.e. at a pixel level.  We term this {\bf pixel-level}
-parallelism.  In the following two sections will will discuss how to
-exploit these two levels of parallelism on cpu and gpu architectures.
+parallelism.
 
@@
 \label{sec:alignment_implementation_cpu}
-This section discusses how to efficently map the two levels of parallelism
-available in iterative face alignment to the two levels of concurrency
-available in a typical multicore CPU architecture.
+One of the main contributions of this paper is a more efficent mapping of the
+two levels of parallelism available in iterative face alignment to the two
+levels of concurrency available in a typical multicore CPU architecture.
 
-We will show that, in constrast with previous implementations, which mapped
-problem level parallelism onto core-level concurrency, it is more efficient to
+We argue that, in constrast with previous implementations, which mapped problem
+level parallelism onto core-level concurrency, it is more efficient to
 completely disregard problem-level parallelism, and instead map pixel-level
-parallelism onto both core-level and vector-level cpu concurrency.  In other
-words, the cpu implementation we advocate performs alignment of the test image
-against the gallery images of a single user at a time.  Each matrix-vector and
-vector-vector operation gets broken into 8 parts that run one per core, and
-into 4 single float vectors that execute on the SSE SIMD units.
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_cpu}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_cpu}
-\caption{Proposed mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_cpu}
-\end{figure}
+parallelism onto both core-level and vector-level cpu concurrency.  In other words,
+the implementation we advocate performs alignment of the test image against the gallery
+images of a single user at a time.  Each matrix-vector and vector-vector operation
+gets broken into 8 parts that run one per core, and into 4 single float vectors
+that execute on the SSE SIMD units.
 
-Recall that the L2 cache can old up to 128 image vectors' worth of data.  Since
-we're only aligning data for one gallery user at a time the working set of the
-problem fits into L2 cache.  Due to the higher bandwidth of L2 cache relative
-to L3 cache, this results in significantly higher image alignment speed than
-previous cpu implementations, as will be seen and discussed further in section
-\ref{sec:alignment_benchmark}.
+Why is this
+implementation faster?  Recall that the L2 cache can old up to 128 image
+vectors' worth of data.  Since we're only aligning data for one gallery user at a
+time the working set of the problem fits into L2 cache, resulting in
+significantly higher image alignment speed than previous cpu implementations.
 
-{\bf Implementation Details} In our implementation, most of the operations
-required by the alignment stage take advantage of Intel Math Kernel Library
-(MKL), a commercial implementation of the standard Basic Linear Algebra
-Subprograms (BLAS).  For image resampling, we make use of Intel's Integrated
-Performance Primitives (IPP) library.  Both of these libraries are heavily
-optimized for Intel multicore CPU's, and are able to make good use of both
-core-level and vector-level concurrency.  For operations that do not map well
-onto existing libraries, we achieve both vector-level and core-level
-concurrency through a combination of the OpenMP API \cite{dagum2002openmp} and
-the automatic vectorization facilities of Intel's C++ compiler
-(ICC)\cite{dulong1999overview}.
+In our proposed CPU implementation of the iterative alignment stage, we
+therefore perform parallelization in a manner similar to the matlab code:
+solving a single problem instance at a time, relying on the threaded version of
+MKL BLAS for both multi-core and vector parallelism.
 
-{\bf Comparison to Previous Implementations} The implementation presented in
-\cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} schedules instances of the
-alignment problem to the cpu cores in a 1:1 fashion.  
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_old}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_old}
-\caption{Old mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_old}
-\end{figure}
+This section discusses how the iterative alignment problem can be implemented
+on CPU hardware.
 
-This implementation was written using the OpenCV library
-as a wrapper for MKL, and, like our proposed implementation, achieves vector-level concurrency implicitly.  
-With eight alignment problems (associated with 8 training users) being solved
-concurrently, and with 20-40 images per training user, all of the required data
-fits into L3 cache.  While the L3 cache has much higher bandwidth and lower latency
-than off-chip memory, this implementation is slower than the proposed implementation.
+{\bf Comparison to Previous Implementations}
+The implementations presented in \cite{WagnerA2009-CVPR} and
+\cite{WagnerA2011-PAMI} were written using the OpenCV library.  OpenCV wraps
+either the operating system's built-in BLAS library, or Math Kernel Library
+(MKL), Intel's BLAS implementation, and thus achieves vector (i.e. pixel)-level
+parallelism implicitly.  The authors of \cite{WagnerA2011-PAMI} used OpenMP to
+schedule instances of the alignment problem to the CPU cores in a 1:1 fashion.
+With eight alignment problems for ten training users being solved concurrently,
+and with 20-40 images per training user, all of the required data fits into L3
+cache.  This provided a significant, though sub-linear improvement over a
+single-core implementation.
 
-%During the course of development, many implementation strategies were tried.
-%Listed in order of increasing performance based on a synthetic benchmark of
-%the $\ell_1$ minimization step, they included:
-%\begin{enumerate}
-%\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
-%\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
-%\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
-%\end{enumerate}
+During the course of development, many implementation strategies were tried.
+Listed in order of increasing performance based on a synthetic benchmark of
+the $\ell_1$ minimization step, they included:
+\begin{enumerate}
+\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
+\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
+\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
+\end{enumerate}
 
@@
 \label{sec:alignment_implementation_gpu}
-This section discusses how to efficently map the two levels of parallelism
-available in iterative face alignment to the two levels of concurrency
-available in the NVIDIA Fermi GPU architecture.
 
-Our proposed GPU implementation for the iterative alignment stage consists of a
-single kernel that performs all of the test image alignment problems.  Each of
-instance of the alignment problem is assigned its own thread block, and the GPU
-hardware schedules as many of these thread blocks to run concurrently as the 
-hardware will allow.  
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_gpu}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
-\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
-\label{fig:alignment_mapping_gpu}
-\end{figure}
-The full resolution test image, along with its numerical derivatives in both
-image directions, are stored in GPU texture memory. This  allows the
-recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
-versions of the test image and its derivatives, to exploit the GPU's special
-purpose hardware for bilinear interpolation.
-
-\subsection{Other Implementation Strategies}
-Several GPU implementation strategies were explored during the course of our
-development, and are only mentioned briefly.  Listed in order of increasing
-performance based on a synthetic benchmark of the $\ell_1$ minimization step,
-they included:
+This section discusses how to implement algorithm
+\eqref{alg:iterative_alignment} on the GPU.  Like for the CPU case, many GPU
+implementation strategies were explored during the course of our development,
+and are only mentioned briefly.  Listed in order of increasing performance
+based on a synthetic benchmark of the $\ell_1$ minimization step, they
+included:
 \begin{enumerate}
@@
 % that they're done
+
+Our proposed GPU implementation for the iterative alignment stage consists of a
+single kernel that performs all of the test image alignment problems.  Each of
+instance of the alignment problem is assigned its own thread block, and the GPU
+hardware schedules as many of these thread blocks to run concurrently as the 
+hardware will allow.
+
+The full resolution test image, along with its numerical derivatives in both
+image directions, are stored in GPU texture memory. This  allows the
+recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
+versions of the test image and its derivatives, to exploit the GPU's special
+purpose hardware for bilinear interpolation.
 
@@
 units on all available CPU cores. 
+{\bf TODO: Our current implementation links against the MKL
+library provided internally with matlab; we are in the process of updating the code to link directly against a newer version
+of MKL}
 
@@
 The first experiment compares the performance of our two proposed implementations of 
-the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
-comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
+the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
 vector $\xx_0$.  
 The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
-randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
-$10:1$ to $10:5$ with $d$ varying from 500 to 8000.
-The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
+randomly generated from the standard normal and normalized with unit column norm.  The ratio of $d:n$ varied from
+$1:10$ to $5:10$ with $n$ varying from 500 to 8000.
+The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and also normalized with unit norm.
 Because the ground truth signal $\xx_0$ was known, the algorithm terminated
-when $norm(\bx-\bx_0) < tolerance$.
-The measurement vector was generated with the equation $\y_0 = A \x_0$.
+when $\|\bx-\bx_0\| < \tau$.
+The measurement vector was generated with the equation $\b_0 = A \x_0$.
 
@@
 
-\subsection{Face Alignment Stage Benchmark} 
-\label{sec:alignment_benchmark}
+\subsection{Face Alignment Stage Benchmark} \label{sec:alignment_benchmark}
 This section presents benchmarks of our CPU and GPU implementations of the

### 2011-03-29 main.tex
Added diagram showing caches

--- 2011-03-25 main.tex
@@
 by the algorithms, and is not an important performance consideration.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figures/caches}
+\end{center}
+\caption{Scale Drawing of the CPU and GPU cache sizes}
+\label{fig:caches}
+\end{figure}
 
@@
 The first experiment compares the performance of our two proposed implementations of 
-the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
+the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
+comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
 vector $\xx_0$.  
 The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
-randomly generated from the standard normal and normalized with unit column norm.  The ratio of $d:n$ varied from
-$1:10$ to $5:10$ with $n$ varying from 500 to 8000.
-The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and also normalized with unit norm.
+randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
+$10:1$ to $10:5$ with $d$ varying from 500 to 8000.
+The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
 Because the ground truth signal $\xx_0$ was known, the algorithm terminated
-when $\|\bx-\bx_0\| < \tau$.
-The measurement vector was generated with the equation $\b_0 = A \x_0$.
+when $norm(\bx-\bx_0) < tolerance$.
+The measurement vector was generated with the equation $\y_0 = A \x_0$.
 

### 2011-03-30 main.tex
fixed typsetting error
Added references to the caches diagram.

--- 2011-03-25 main.tex
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figures/caches}
+\end{center}
+\caption{Scale Drawing of the CPU and GPU cache sizes}
+\label{fig:caches}
+\end{figure}
+
 % CPU CACHE DISCUSSION
@@
 Overall, this means the algorithm has approximately 16MB of on-chip cache
-available.  This corresponds to 1024 image vectors.
+available.  This corresponds to 1024 image vectors.  A scale drawing
+of the caches of the processor we're targeting can be seen in Figure \ref{fig:caches}.
 
@@
 KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount
-of cache (1/23 as much as CPU L3) on the GPU is ballanced by a significantly
+of cache (1/23 as much as CPU L3) on the GPU is balanced by a significantly
 higher bandwidth between the processor chip and off-chip memory (DRAM) compared
-to the CPU.
+to the CPU.  A scale drawing of the caches available on the GPU can be seen in Figure \ref{fig:caches}.
 
@@
 when $\|\bx-\bx_0\| < \tau$.
-The measurement vector was generated with the equation $\b_0 = A \x_0$.
+The measurement vector was generated with the equation $\bb_0 = A \xx_0$.
 

### 2011-04-07 main_allen.tex
Partially done merging Allen's changes

--- 2011-03-25 main.tex
@@
 \cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  
-$\ell_1$-min refers to finding the minimum $\ell$-norm solution to an
+$\ell_1$-min refers to finding the minimum $\ell_1$-norm solution to an
 underdetermined linear system $\bb=A\xx$:
@@
 state-of-the-art $\ell_1$-min solution based on the classical framework of
-\emph{augmented Lagrange multiplier} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
+\emph{augmented Lagrangian methods} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
 
 While solving \eqref{eq:l1min_denoise} can achieve exceptional
-recognition accuracy on public databases, its success relies on the assumption
+recognition accuracy on public databases, its success also relies on the assumption
 that the query image is well aligned with the training images. In
@@
 (e.g., affine or homography) acting on the image domain, which aligns $\tilde\bb$
-with the training images from the $i$-th class $A_i$. The algorithm 
-extends Lucas-Kanade iterative alignment \cite{LucasB1981} to both to use the
-$\ell_1$ norm as a robust error function, and also simultaneously 
-to estimate the illumination model $A\x$.  
-The original solution \cite{WagnerA2009-CVPR} was implemented based 
-on standard interior-point
-methods, and its speed has been since improved by ALM in a more recent paper
-\cite{WagnerA2011-PAMI}. However, primarily due the high per-user cost
-of the alignment step, the recognition system was still not fast enough
-for real-time recognition against datasets of hundreds of users.
-In this paper, in addition to solving the generic
+with the training (frontal) images from the $i$-th class $A_i$. The algorithm 
+extends Lucas-Kanade iterative alignment \cite{LucasB1981} using the
+$\ell_1$-norm as a robust error function.
+
+Numerically, the original solution \cite{WagnerA2009-CVPR} was implemented based 
+on standard interior-point methods, and its speed has been since improved by ALM in a more recent paper
+\cite{WagnerA2011-PAMI}. However, due the high per-class cost
+of the alignment step, the recognition system still fails to achieve \emph{real-time recognition} against datasets of hundreds or thousands of users.
+In this paper, in addition to accelerating the generic
 $\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
-discuss how to properly accelerate the iterative face alignment step
+also discuss how to properly accelerate the iterative face alignment step
 \eqref{eq:l1min_alignment} on multi-core CPUs and GPUs.
+
+In this paper we will clearly distinguish between algorithm {\em parallelism},
+which is a property of an algorithm, and hardware {\em concurrency}, which is a
+property of a hardware architecture. Parallelism in an algorithm provides the
+opportunity to perform computations concurrently on hardware.  Algorithms often
+exhibit multiple levels of parallelism, and hardware often provides multiple
+levels of concurrency.  In these terms, the primary contribution of this paper
+will be to determine the optimal mappings between the parallelism available in
+$\ell_1$-min and the concurrency available in multi-core CPU and GPU
+architectures.
 
@@
 generic ALM $\ell_1$-min algorithm and the corresponding face recognition
-routines on massively parallel CPUs and GPUs.  To this end, the face
+routines on massively parallel CPUs and GPUs.  To this end, a new face
 recognition system has been fully implemented on a state-of-the-art
-workstation, as shown in Figure \ref{fig:pipeline}. 
+workstation for security and access control applications, as shown in Figure \ref{fig:pipeline}. 
 Since the parallelization of any algorithm
@@
 the architecture of our test system in detail in Section
-\ref{sec:parallelism}.  We will show that a highly accurate and robust face
-recognition system can be constructed to recognize face images from a 
-database of several hundred subjects quickly enough to be practical for
-access control applications. 
-Note that for applications targeting thousands of users or more, it would likely be
-necessary to distribute the work over a network of machines; this paper
+\ref{sec:parallelism}.
+Finally, note that for applications targeting thousands of users or more, it would likely be
+necessary to distribute the work over a network of computers, while this paper
 focuses on the (shared memory) parallelism available in a single machine.
-The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture
-will be made available for evaluation by other researchers.
+The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture,
+will be made available for peer evaluation.
 \begin{figure}
@@
 {\tiny \input{figures/pipeline_simplified.tex}}
-\caption{Overview of the face recognition pipeline.}
+\caption{The face recognition pipeline.}
 \label{fig:pipeline}
@@
 \subsection{Literature Review} 
-In this section, we briefly review the literature of $\ell_1$-min and existing
+We first briefly review the literature of $\ell_1$-min and existing
 works on its parallelization.  Traditionally, $\ell_1$-min (or the so-called
@@
 BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
-interior path
-%{\bf DW: Following the interior path is what made them faster than
-%traditional linear programming techniques that follow the edge of the feasible set;
-%does Homotopy actually go outside the feasible set?  Just want to check if this is
-%an accurate representation...}
-, which is often computationally expensive. To mitigate these
-issues, an approach called \emph{Homotopy} has been studied to accelerate the
+interior path via gradient descent or conjugate gradient methods, which are computationally expensive.
+To mitigate these issues, an approach called \emph{Homotopy} has been recently studied to accelerate the
 speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were
@@
 In Section \ref{sec:ALM}, we contend
-that ALM is a better choice among Homotopy and other existing $\ell_1$-min
+that ALM is a better choice among existing $\ell_1$-min
 solvers for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
 category of approximate $\ell_1$-min solutions called \emph{iterative
-shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}.
-IST algorithms mainly involves elementary operations such as vector 
+shrinkage-thresholding} (IST) methods \cite{WrightS2008,BeckA2009}.
+IST algorithms mainly involve elementary operations such as vector 
 algebra and matrix-vector multiplication. Therefore,
 when the dimension of the problem becomes high, IST-type algorithms are
-particularly suitable for systems in parallel-computing environments. In
+particularly suitable for systems with parallel-computing pipelines. In
 \cite{YangA2010-ICIP}, the authors further showed that, compared to other
@@
 parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia
-programmable GPUs.
+programmable GPUs, which is similar to the hardware setup of this paper.
 
@@
 Lagrange multipliers in an iterative fashion.  For $\ell_1$-min, the augmented
-Lagrange function is defined as: \begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
+Lagrangian is defined as: \begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
 \yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2, \end{equation}
@@
 satisfies the second-order sufficiency conditions for optimality
-\cite{BertsekasD2003}, then for a sufficiently large $mu$, the optimal
+\cite{BertsekasD2003}, then for a sufficiently large $\mu$, the optimal
 $\ell_1$-min solution also minimizes
@@
 
-In practice, the optimal values for the triplet $(\xx^*, \yy^*, \uu)$ are all
-unknown. Furthermore, it has been observed that if solving
+In practice, the optimal values for the triplet $(\xx^*, \yy^*, \mu)$ are all
+unknown. Furthermore, it has been observed that solving
 \eqref{eq:optimal-ALM} with a large initial value of $\mu$ tends to lead to
 slower convergence speed \cite{WrightS2008,YangA2010-ICIP}. In
-\cite{BertsekasD2003,YangJ2009}, an alternating procedure has been proposed to
+\cite{BertsekasD2003,YangJ2009}, an alternating procedure has been shown to
 iteratively update $\xx$ and $\yy$:
@@
 \end{equation}
-where $\left\{\mu_{k}\right\}$ is a monotonically increasing positive sequence.
-The iteration terminates when the estimates $(\xx, \yy)$ converges.
+where $\mu_{k}\rightarrow \infty$ is a monotonically increasing positive sequence.
+The iteration terminates when the estimates $(\xx, \yy)$ converge.
 
-Note that in the iterative procedure outlined in \eqref{eq:alm}, the second
+Note that in the iterative procedure \eqref{eq:alm}, the second
 step only involves vector algebra and matrix-vector multiplication. Therefore,
 the procedure is computationally efficient only if it is easier to minimize the
-augmented Lagrange function compared to solving the original problem
-\eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k}
-(\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise
+augmented Lagrangian $L_{\mu_k} (\xx,\yy_k)$ compared to solving the original problem
+\eqref{eq:l1min} directly. In fact, this problem can be solved element-wise
 iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009},
 whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
-Algorithm \eqref{alg:alm} summarizes the ALM $\ell_1$-min algorithm \footnote{While we
-present the ALM algorithm in the primal domain, there also
-exists an implementation ALM in the dual domain.}.
+Algorithm \ref{alg:alm} summarizes the generic ALM $\ell_1$-min algorithm.\footnote{For conciseness, we
+only present the ALM algorithm in the primal domain. There also
+exist implementations in the dual domain \cite{YangJ2009,YangA2010-ICIP}.}
 
  \begin{algorithm}[h]
-\caption{Augmented Lagrange Multiplier (ALM)}
-{\bf INPUT:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$, $\tau\leftarrow$ largest eigenvalue of $A^TA$, and a constant $\rho>1$.
+\caption{Augmented Lagrangian Method (ALM)}
+{\bf INPUT:} $\bb \in \Re^m$, $A=[A_1,\cdots, A_K] \in \Re^{m \times n}$, $\tau\leftarrow \max\mbox{eig}(A^TA)$, and constant $\rho>1$.
 \begin{algorithmic}[1]
-\WHILE{not converged ($k = 1,2,\ldots$)} \STATE $t_1 \leftarrow 1$, $\zz_1
-\leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ \WHILE{not
-converged ($l = 1,2,\ldots$)} \STATE $\uu_{l+1}  \leftarrow
-\mathrm{soft}\left(\zz_l - \frac{1}{\tau}A^T\left(A\zz_l - \bb
-- \frac{1}{\mu_k}\yy_k\right), \frac{1}{\mu_k\tau}\right)$
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$ \STATE $\zz_{l+1} \leftarrow \uu_{l+1}
-+ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ \ENDWHILE \STATE$
-\xx_{k+1} \leftarrow \uu_{l+1}$ \STATE $\yy_{k+1} \leftarrow
-\yy_k + \mu_k (\bb - A\xx_{k+1})$ \STATE $\mu_{k+1} \leftarrow
-\rho\cdot\mu_k$ \ENDWHILE 
+\WHILE{not converged ($k = 1,2,\ldots$)} 
+\STATE $t_1 \leftarrow 1$, $\zz_1 \leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ 
+\WHILE{not converged ($l = 1,2,\ldots$)} 
+\STATE $\uu_{l+1}  \leftarrow \mbox{shrink}(\zz_l - \frac{1}{\tau}A^T(A\zz_l - \bb - \frac{1}{\mu_k}\yy_k), \frac{1}{\mu_k\tau})$
+\STATE $t_{l+1} \leftarrow \frac{1}{2}( 1 + \sqrt{1+4t_l^2})$
+\STATE $\zz_{l+1} \leftarrow \uu_{l+1}+ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ 
+\ENDWHILE 
+\STATE $\xx_{k+1} \leftarrow \uu_{l+1}$ 
+\STATE $\yy_{k+1} \leftarrow \yy_k + \mu_k (\bb - A\xx_{k+1})$ 
+\STATE $\mu_{k+1} \leftarrow \rho\cdot\mu_k$ 
+\ENDWHILE 
 \end{algorithmic}
@@
 
-In this paper we will clearly distinguish between algorithm {\em parallelism},
-which is a property of an algorithm, and hardware {\em concurrency}, which is a
-property of a hardware architecture.  Parallelism in an algorithm provides the
-opportunity to perform computations concurrently on hardware.  Algorithms often
-exhibit multiple levels of parallelism, and hardware often provides multiple
-levels of concurrency.  In these terms, the primary contribution of this paper
-will be to determine the optimal mappings between the parallelism available in
-$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU
-architectures.
-
 In this section we discuss the levels of concurrency available in the hardware
@@
 memory (cache) is often orders of magnitude faster than off-chip memory, the
-sizes and speeds of on-chip memory (cache) have big performance implications.
+size and speed of on-chip memory (cache) have big performance implications.
 We give a brief overview of the caches that are available in our target
 architectures, and defer discussion of their performance implications to
-sections \ref{sec:alignment_implementation_cpu} and
-\ref{sec:alignment_implementation_gpu}.
+sections \ref{sec:alignment} and \ref{sec:recognition}.
 
-We will primarily be comparing CPU and GPU architectures on a per-board basis:
+We will primarily compare CPU and GPU architectures on a per-board basis:
 for CPU implementation we make use of all the cores in as many CPUs are present
@@
 \subsection{CPU Hardware Concurrency} 
+\label{sec:CPU-concurrency}
 The main defining characteristics of contemporary multi-core CPU architectures
 are that they have two levels of concurrency, relatively large amounts of
-on-chip cache, and relatively high clock speeds.
+on-chip cache, and relatively high clock speed.
 
-The baseline architecture for our experiments is a Linux workstation with two
+The baseline architecture for our experiment is a Linux workstation with two
 quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
@@
 %... so can fit 2**7  = 128 images.
-Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache.  
-Between the 8 cores, this is a total of 2\,MB L2 cache.
-For the low resolution images used in sparse representation based
-face recognition (on the order of $64\times64$ pixel images), the L2 cache
-can in principle hold up to 128 image vectors.
+Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache (\ie, a total of 2\,MB L2 cache available for the eight cores).
+Therefore, for the low-resolution images used in face recognition (say, on the order of $64\times64$ grayscale images), the L2 cache
+in principle can hold up to 16 images per core and 128 images in total.
 %16MB = 2**24 bytes = 16777216 bytes
@@
 %... so can fit 2**10  = 1024 images.
-Each processor further has 8MB of L3 cache that is shared by all four cores.
-Overall, this means the algorithm has approximately 16MB of on-chip cache
-available.  This corresponds to 1024 image vectors.
+Each processor further has 8\,MB of L3 cache that is shared by the four cores.
+Overall, the algorithm has approximately 16\,MB of L3 cache
+available for a dual-processor configuration, which corresponds to about 1024 images.
 
 %With Hyperthreading enabled, each core can support two threads simultaneously.
-For floating point instructions, each core also has a vector processing unit
-(SSE) capable of performing the same arithmetic operation on four single
-precision floating point values simultaneously.
+For floating-point instructions, each core also has a vector processing unit
+(SSE) capable of performing the same arithmetic operation on four single-precision floating point values simultaneously.
 
@@
 are that they have two (much wider) levels of concurrency, relatively small amounts of
-on-chip cache, and relatively low clock speeds.
+on-chip cache, and relatively low clock speed.
 
-Compared to CPU architectures, GPU architectures dedicate less of their transistors
+Compared to CPU architectures, modern GPU architectures dedicate less of their transistors
 to cache, and more of them for arithmetic logic. Our target GPU architecture
-is the Nvidia Fermi architecture.  CUDA based GPUs are comprised of several 
-streaming multiprocessors (SMP) (currently 14 to 16 for a high end GPU).
+is the Nvidia Fermi architecture (\eg, GTX 480 GPU used in this paper). CUDA based GPUs are comprised of several 
+\emph{streaming multiprocessors} (SMP) (currently 14 to 16 for a high end GPU).
 
@@
 groups of 32 threads called \emph{warps}.  These warps are then grouped into
-higher levels called \emph{thread blocks} and \emph{grids}.  It is impossible
-to communicate between SMPs during gpu program execution, so computational tasks
-are typically broken up into multiple gpu programs (kernels) that are executed
-serially.
+higher levels called \emph{thread blocks} and \emph{grids}.  Since SMPs are not designed
+to communicate between themselves during GPU program execution, tasks
+are typically broken up into multiple GPU programs (kernels) that are executed
+serially.\footnote{Please add a footnote to further explain.}
 
-% GPU CACHE DISCUSSION
-% for L2 cache
-% 768KB = 786432 bytes
-% 64*64*4 = 2**14 bytes = 16384 bytes per image
-% 48 images
-Each SMP contains its own L1-cache, which is divided between hardware managed
-and software managed ("shared") memory.  Additionally, all SMPs share a common
-L2-cache.  For our system, each SMP has 64 KB of L1 cache with a total of 896
+Each SMP contains its own L1 cache, which is divided between hardware managed
+and software managed (``shared'') memory.  Additionally, all SMPs share a common
+L2 cache.  For our system, each SMP has 64 KB of L1 cache with a total of 896
 KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount
@@
 
-The GPU has its own memory system, any data the GPU uses must first be
-transferred from CPU DRAM to GPU DRAM over PCI-Express.  For our application,
-this transfer overhead is amortized over a large amount of computation and is
-not a major concern.
-
-Compared to a typical CPU, a GPU has much higher peak computational bandwidth
-as well as data bandwith to DRAM.  However, since our application is primarily
+Compared to typical CPUs, GPUs have much higher peak computational bandwidth
+as well as data bandwidth to DRAM.  However, since our system primarily consists
 of matrix-vector operations and image resampling operations, it has a very low
-arithmetic intensity (i.e. we perform a very small number of arithmetic
+arithmetic intensity (\ie, we perform a very small number of arithmetic
 operations for every number that has to be loaded), and thus performance is
 typically constrained by memory (and cache) bandwidth.  At large problem sizes
-the CPU is limited by its relatively low bandwidth to DRAM compared to the gpu,
+the CPU is limited by its relatively low bandwidth to DRAM compared to the GPU,
 while at small problem sizes the GPU is limited by the relatively high latency
@@
 \label{sec:alignment}
-There are essentially two levels of parallelism available for exploitation in
-for iterative alignment in the face recognition application.  The primary goal
-of this section is to introduce these two levels of parallelism.  Additionally,
-in order for some important aspects of our implementations to be clear later,
-we need to establish how the iterative alignment stage works in some
-detail.
 
-The inputs of the iterative alignment stage are the test image $f(x,y)$ stacked into an $M$-
-dimensional vector $\f$, and the aligned and stacked training images $A_i$ for
-each training user $i$. 
-The algorithm repeatedly solves a linearized version of the minimization problem
-in \eqref{eq:l1min_alignment}.  This involves two computational steps:
-the computation of the linearization, and the solution of the minimization problem.
+Given the hardware concurrency in modern parallel computing architectures, next,
+we shall study the optimal mappings of the algorithm parallelism in face recognition to the multi-core CPUs and GPUs, respectively.
+In this section, we first focus on the face alignment stage (see Figure \ref{fig:pipeline}).
 
-The computation of the linearization is complicated by the need to avoid trivial 
-solutions where the algorithm
-progressively zooms in on a black pixel of the test image.
-This is achieved by minimizing the $\ell_1$ error between the illumination model $Ax$ and the cropped,
-resampled, and normalized test image $\tilde\bb$, rather than the un-normalized version as in 
-conventional Lucas-Kanade iterative alignment.
-This involves a bit of calculus:
+Face alignment \eqref{eq:l1min_alignment} estimates an image transformation $\tau$ that rectifies the query image $\tilde{\bb}$ with possible pose variation
+w.r.t. each training class $A_i$, which leads to the minimal sparsity in error $\ee$ after the alignment. Note that directly solving \eqref{eq:l1min_alignment} would require
+simultaneously estimating $\xx$, $\ee$, and $\tau$ typically on a nonlinear manifold, which renders the problem extremely difficult to solve exactly with many local minima.
+
+To mitigate the problem, if the system can obtain to a good initial guess of the transformation $\tau$ (\eg, provided by a good face detector),
+the optimal solution for $\tau$ can be sought iteratively by a sequence of linearized approximations $j=1, 2, \cdots$ as follows:
 \begin{equation}
-\frac{\partial}{\partial \tau_p} \frac{\bb}{||\bb||} 
-= \frac{1}{||\bb||} \frac{\partial \bb}{\partial \tau_p} - 
-\frac{\bb}{||\bb||^3} \bb^T \frac{\partial \bb}{\partial \tau_p}
+\min_{\xx, \ee, \Delta \tau_j}\|\ee\|_1\quad \subj\quad \tilde\bb\circ\tau_j +  J_j\Delta \tau_j = A_i\xx + \ee,
 \end{equation}
-If the warping $\tau$ is constrained to be a similarity transform,
-it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
-reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
-and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
-once per test image.  The per-user iterative alignment algorithm is summarized as Algorithm \eqref{alg:iterative_alignment}.
-\begin{algorithm}[thb]
-\caption{\bf (Per-User Iterative Alignment)} \label{alg:iterative_alignment}
+where $J_j = \nabla_{\tau_j}(\tilde\bb\circ\tau_j)$ is the Jacobian, and $\Delta \tau$ is a step update to $\tau$.
+Denote $\bb_j = \tilde\bb\circ\tau_j$, $B_j = [A_i, -J_j]$ and $\ww^T = [\xx^T, \Delta \tau^T]$, then the update $\Delta \tau$ can be computed by solving the following linear program:
+\begin{equation}
+\min_{\ww, \ee}\|\ee\|_1\quad \subj\quad \bb_j = B_j\ww + \ee.
+\label{eq:alignment-linearization}
+\end{equation}
+
+The per-class alignment algorithm via ALM is summarized in Algorithm \ref{alg:alignment_alm}.
+In terms of computational complexity, one significant term is the computation of the matrix psudo-inverse $B_j^\dagger$, whose computational cost is bounded by $O(n_i^2m + n_i^3)$.
+For per-class face alignment, since both $A_i$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$ for all $i=1, \cdots, K$ classes), we find that the computation time is still dominated by the iterative computation of matrix-vector multiplications in the inner loop.
+\begin{algorithm}[ht!]
+\caption{\bf (Face Alignment via ALM)} \label{alg:alignment_alm}
+{\bf Input:} $\bb$, $A_i$, $\x_0 = \mathbf{0}$, $\tau_0$, and $J_0$.
 \begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n}\}$ for gallery subject $i$, the test image
-$\f\in\Re^M$ 
-%\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
-%\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
-%\STATE {\bf for} each gallery subject $i$, 
-\STATE \hspace{0mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
-\STATE \hspace{3mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
-\STATE \hspace{3mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
-\STATE \hspace{3mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
-\STATE \hspace{3mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$; 
-\end{small}
+\WHILE{not converged ($j = 1,2,\ldots$)}
+\STATE Update $\bb_j \leftarrow \frac{\bb\circ \tau_{j-1}}{\|\bb\circ \tau_{j-1}\|}$; $B_j= [A_i, -J_{j-1}]$ and corresponding $(B_j^\dagger)^T$
+\STATE Initialize $\ww_0 = \mathbf 0$, $\yy_0 = \mathbf 0$
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $\uu_0\leftarrow \ww_{k-1}$; $\zz_0\leftarrow \ee_{k-1}$
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\zz_l \leftarrow \mbox{shrink}\left(\bb_j - B_j\uu_{l-1} + \frac{\yy_{k}}{\mu_{k-1}}, \frac{1}{\mu_{k-1}}\right)$
+\STATE $\uu_l \leftarrow (B_j^\dagger)^T \left(\bb_j - \zz_{l} + \frac{\yy_{k-1}}{\mu_{k-1}} \right) $
+\ENDWHILE
+\STATE $\ww_k \leftarrow \uu_l$; $\ee_k \leftarrow \zz_l$
+\STATE $\yy_{k} \leftarrow \yy_{k-1} + \mu_{k-1} (\bb_j - B_j\ww_{k} - \e_{k})$
+\STATE $\mu_{k} \leftarrow \rho\mu_{k-1}$
+\ENDWHILE
+\STATE Update $\ee_j$, $\tau_j$, and $J_j$
+\ENDWHILE
 \end{algorithmic}
+{\bf Output:} $\tau^*\leftarrow \tau_j, \e^*\leftarrow \ee_j$
 \end{algorithm}
 
-The solution of the minimization problem in Algorithm
-\eqref{alg:iterative_alignment} Step 7 can be solved via a
-variant of ALM, which is summarized as Algorithm \eqref{alg:alignment_alm}.
-$\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of
-parameter $\mu$, we take the same strategy as in \cite{YangJ2009-pp} and set
-$\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
-
-\begin{algorithm}[h]
-\caption{\bf (Augmented Lagrange Multiplier Method For Alignment)}
-\begin{algorithmic}[1]
-\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
-\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
-\ENDWHILE
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
-\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{algorithmic}
-\label{alg:alignment_alm}
-\end{algorithm}
-%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-
-\subsection{Computational Complexity and Parallelism} Any discussion of the speed of an
-algorithm should discuss the comutational complexity of the algorithm.  For the
-alignment stage ALM algorithm \eqref{alg:alignment_alm}, this is rather simple.
-The Matrix-vector computions are $O(mn)$, and the vector-vector operations are
-$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
-very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
-computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
-A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
-Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
-of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
-Fortunately, $A^\dagger$ only needs to be computed once per instance of the
-$\ell_1$ minimization problem.
-
-Therefore, for problems with large $n$, we expect the computation of the
-pseudoinverse to dominate.  However, for the problem sizes needed in face
-recognition problems, we find that the computation time is dominated by the
-bandwidth required by repeated computation of matrix-vector and matrix-matrix
-multiplications in the inner loop.
-
-In summary, the alignment stage essentially two levels of available
-parallelism: at the higher level there are multiple iterative alignent problems
-that are solved independently, or {\bf problem-level} parallelism.  At a lower
+In summary, the alignment stage essentially contains two levels of available
+parallelism. At the higher level, there are per-class alignment problems
+that are solved independently, or \emph{problem-level parallelism}.  At a lower
 level, the first-order linear algebraic operations exhibit parallelism within
-image operations, i.e. at a pixel level.  We term this {\bf pixel-level}
-parallelism.  In the following two sections will will discuss how to
-exploit these two levels of parallelism on cpu and gpu architectures.
+image operations, \ie, at the pixel level.  We call this \emph{pixel-level
+parallelism}.  In the following two sections will will discuss how to
+exploit these two levels of parallelism on different CPU and GPU architectures.
 
@@
 \label{sec:alignment_implementation_cpu}
-This section discusses how to efficently map the two levels of parallelism
-available in iterative face alignment to the two levels of concurrency
-available in a typical multicore CPU architecture.
 
-We will show that, in constrast with previous implementations, which mapped
-problem level parallelism onto core-level concurrency, it is more efficient to
-completely disregard problem-level parallelism, and instead map pixel-level
-parallelism onto both core-level and vector-level cpu concurrency.  In other
-words, the cpu implementation we advocate performs alignment of the test image
-against the gallery images of a single user at a time.  Each matrix-vector and
-vector-vector operation gets broken into 8 parts that run one per core, and
-into 4 single float vectors that execute on the SSE SIMD units.
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_cpu}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_cpu}
-\caption{Proposed mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_cpu}
-\end{figure}
-
-Recall that the L2 cache can old up to 128 image vectors' worth of data.  Since
-we're only aligning data for one gallery user at a time the working set of the
-problem fits into L2 cache.  Due to the higher bandwidth of L2 cache relative
-to L3 cache, this results in significantly higher image alignment speed than
-previous cpu implementations, as will be seen and discussed further in section
-\ref{sec:alignment_benchmark}.
-
-{\bf Implementation Details} In our implementation, most of the operations
-required by the alignment stage take advantage of Intel Math Kernel Library
-(MKL), a commercial implementation of the standard Basic Linear Algebra
-Subprograms (BLAS).  For image resampling, we make use of Intel's Integrated
-Performance Primitives (IPP) library.  Both of these libraries are heavily
-optimized for Intel multicore CPU's, and are able to make good use of both
-core-level and vector-level concurrency.  For operations that do not map well
-onto existing libraries, we achieve both vector-level and core-level
-concurrency through a combination of the OpenMP API \cite{dagum2002openmp} and
-the automatic vectorization facilities of Intel's C++ compiler
-(ICC)\cite{dulong1999overview}.
-
-{\bf Comparison to Previous Implementations} The implementation presented in
-\cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} schedules instances of the
-alignment problem to the cpu cores in a 1:1 fashion.  
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_old}.
+The optimal strategy to implement Algorithm \ref{alg:alignment_alm} on a multicore CPU must take into account
+the trade-off between the available memory/cache sizes at multiple hierarchies and their bandwidth based on a particular
+CPU architecture. Figure \ref{fig:alignment_mapping_old} shows the the sizes of L2 and L3 caches of the Intel E5530 CPU, 
+which is a good example of modern multicore CPU architectures. 
 \begin{figure}
@@
 \includegraphics[width=3.4in]{figures/alignment_mapping_old}
-\caption{Old mapping of alignment parallelism onto cpu concurrency}
+\caption{Available L2 and L3 caches on the Intel E5530 multicore CPU architecture with dual CPU configuration.}
 \label{fig:alignment_mapping_old}
@@
 
-This implementation was written using the OpenCV library
-as a wrapper for MKL, and, like our proposed implementation, achieves vector-level concurrency implicitly.  
-With eight alignment problems (associated with 8 training users) being solved
-concurrently, and with 20-40 images per training user, all of the required data
-fits into L3 cache.  While the L3 cache has much higher bandwidth and lower latency
-than off-chip memory, this implementation is slower than the proposed implementation.
+As we have calculated in Section \ref{sec:CPU-concurrency}, the L2 cache in each core is only able to store about 16 images.
+For image alignment problem, the number of training images per class is typically larger than 16.
+Hence, if \eqref{eq:alignment-linearization} would be solved in parallel on the eight cores, it would require L3 cache to hold data up to four classes at a time, and the L2 cache should retrieve subsets of the data frequently during the iterative process to execute Algorithm \ref{alg:alignment_alm}.
+Therefore, the bottleneck in this strategy becomes the bandwidth between the L2 cache and L3 cache.
+The implementation of some previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this mapping strategy.
 
-%During the course of development, many implementation strategies were tried.
-%Listed in order of increasing performance based on a synthetic benchmark of
-%the $\ell_1$ minimization step, they included:
-%\begin{enumerate}
-%\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
-%\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
-%\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
-%\end{enumerate}
-
-%\subsection{Tuning the GEMV operation}
-%When matrix-vector operations are called sequentially on the same data,
-%the operation can be blocked so that each thread operates on the same block of data for each operation.
-%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
-%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
-%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
-
-%\subsection{Tuning the Soft-Thresholding Operator}
-%...Make sure there are no branches
-%...Make sure there are no function calls
-%...Make sure the loop gets automatically vectorized
-%...Block the computation into work for different cpu cores
+In this paper, we show that for CPU parallelism, a more traditional implementation for pixel-level parallelism in fact outperforms problem-level parallelism using eight separate cores.
+In other words, the CPU implementation performs alignment of a query image against the training images of single or dual users at a time on the dual processors, depending on the setting of the compiler.
+More specifically, in our implementation, most of the operations take advantage of Intel Math Kernel Library (MKL), a commercial implementation of the standard Basic Linear Algebra Subprograms (BLAS).
+For the image resampling step, we make use of the Intel Integrated Performance Primitives (IPP) library.
+Both MKL and IPP are optimized for Intel multicore CPUs, and are able to automatically utilize both core-level and vector-level concurrency.
+For other operations that are not optimized by Intel in-house libraries, we achieve both vector-level and core-level concurrency through a combination of the Open MP API \cite{dagum2002openmp} and the automatic vectorization facilities of the Intel C++ compiler (ICC) \cite{dulong1999overview}.
+We shall compare the performance of the two parallelism methods in Section \ref{sec:experiment}.
 
@@
 \label{sec:alignment_implementation_gpu}
-This section discusses how to efficently map the two levels of parallelism
-available in iterative face alignment to the two levels of concurrency
-available in the NVIDIA Fermi GPU architecture.
+
+Similarly, we also consider the trade-off between the available cache sizes and their bandwidth based on the Nvidia CUDA GPU architecture chosen for the paper.
+Figure \ref{fig:alignment_mapping_gpu} illustrates the multicore architecture and the L1/L2 caches of the Nvidia GTX 480 GPU.
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
+\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
+\label{fig:alignment_mapping_gpu}
+\end{figure}
 
@@
 hardware will allow.
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_gpu}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
-\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
-\label{fig:alignment_mapping_gpu}
-\end{figure}
-The full resolution test image, along with its numerical derivatives in both
+
+In addition, the full resolution query image, along with its numerical derivatives in both
 image directions, are stored in GPU texture memory. This  allows the
-recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
+computation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
 versions of the test image and its derivatives, to exploit the GPU's special
@@
 
-\subsection{Other Implementation Strategies}
-Several GPU implementation strategies were explored during the course of our
-development, and are only mentioned briefly.  Listed in order of increasing
-performance based on a synthetic benchmark of the $\ell_1$ minimization step,
-they included:
-\begin{enumerate}
-\item Solving a single problem instance at a time using Cublas\footnote{Nvidia provided BLAS implementation} (302 ms)
-\item Pure CUDA C implementation solving one problem per SMP using cuda streams\footnote{Streams is a CUDA feature that allows multiple
-kernels to be run concurrently} (70 ms)
-\item Pure CUDA C implementation solving one problem per SMP using one large kernel call (36 ms)
-\end{enumerate}
-
-We attribute the low performance of Cublas in this benchmark to the relatively
-low problem size. Cublas is intended to take advantage of the resources of the
-entire gpu to solve a single large problem, and our problem size is too small
-to hide the memory latency.  Effects of this will be seen again in the
-benchmarks in section \ref{sec:simulation}.
-
-One possible implementation of the iterative alignment stage would be to
-perform the optimization for all instances simultaneously, but to perform each
-linear algebraic operation in a separate gpu kernel call.  The primary
-advantage of this implementation strategy is that every kernel would be
-performing a very simple operation requiring fewer registers per thread.  The
-registers in a GPU are shared between all executing threads thus, barring other
-hardware limitations, the fewer registers each thread requires, the more
-threads can be scheduled to run simultaneously.  The primary downside of this
-approach is that it forces synchronization between all problem instances.  It will
-also have increased kernel launch overhead and potentially less efficient use of
-the GPU's L2 cache.
-% DREW Idea... in hindsight, maybe we should still try many kernel launches. 
-% instances can still exit early, and can still update a global variable signaling
-% that they're done
-
-% Should really verify somehow that we're maxing out GPU bandwidth
+In one experiment, we have exploited several possible GPU implementation strategies, which include
+solving single alignment problems one at a time using CUBLAS (302\,ms); solving one problem
+per SMP using CUDA streams (70\,ms); and solving one problem per SMP using one large kernel call
+ (36\,ms). The inferior performance of CUBLAS may be attributed to the relatively small problem
+ size. In Section \ref{sec:experiment}, we will further exploit the difference between CPU and GPU
+ via synthetic and real image data.
 
@@
 \label{sec:recognition}
-{\bf AY: Add an algorithm here describe the recognition block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
-{\bf VS: Why did we modify the algorithm for the recognition stage? }
-
-Below is the algorithm for the face recognition stage in Figure \ref{fig:pipeline}, which
-solves the $\ell_1$ minimization problem posed in \eqref{eq:l1min_denoise}.
-
+After the face alignment stage is complete, a single recognition algorithm will be called
+to estimate a sparse representation of the query image w.r.t. all the training classes, 
+as shown in \eqref{eq:l1min_denoise}. In \cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI},
+the recognition algorithm via $\ell_1$-min has been extensively studied. The algorithm itself via ALM
+is summarized in Algorithm \ref{alg:alm_rec}.
+In this section, we mainly focus our discussion on its pixel-level parallelism on the multicore CPUs and GPUs. Then in Section \ref{sec:experiment}, we benchmark
+the performance of the two architectures and further demonstrate the improvement in both accuracy and speed
+compared to the existing literature.
 \begin{algorithm}[t]
-\caption{\bf (Face Recognition via ALM)}
+\caption{\bf (Face Recognition via ALM)} \label{alg:alm_rec} 
 \begin{algorithmic}[1]
@@
 \WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}\left(\bb - A\x_k +
-\frac{1}{\mu}\y_k, \frac{1}{\mu}\right)$;
+\STATE $\e_{k+1} = \textup{shrink}(\bb - A\x_k +
+\frac{1}{\mu}\y_k, \frac{1}{\mu})$;
 \STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
 \WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
-\frac{1}{\gamma}A^T\left(\bb - A\z_l - \e_{k+1} +
-\frac{1}{\mu}\y_k\right), \frac{1}{\mu\gamma}\right)$;
-\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
-\sqrt{1+4t_l^2}\right)$;
+\STATE $\w_{l+1} \leftarrow \textup{shrink}(\z_l +
+\frac{1}{\gamma}A^T(\bb - A\z_l - \e_{k+1} +
+\frac{1}{\mu}\y_k), \frac{1}{\mu\gamma})$;
+\STATE $t_{l+1} \leftarrow \frac{1}{2}( 1 +
+\sqrt{1+4t_l^2})$;
 \STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
@@
 \end{algorithmic}
-\label{alg:alm_rec} 
 \end{algorithm}
 
-\subsection{CPU Implementation} Our CPU implementation of Algorithm
-\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
-nicely onto the BLAS API (such as SGEMV, SAXPY), we
-call Intel's MKL BLAS library, which internally takes advantage of the SSE
-units on all available CPU cores.
+For CPU implementation, {\bf AY: Drew, please fill in one or two paragraphs}.
 
-For operations that do not map nicely onto the BLAS API, such as the shrinkage
-operator, {\bf Argue that these parts are optimized, and/or demonstrate that
-performance of these parts is insignificant}.
+For GPU implementation, {\bf AY: Drew, please fill in one or two paragraphs}.
 
-\subsection{GPU Implementation} Our GPU implementation of Algorithm
-\ref{alg:alm_rec} is similar to the CPU implementation described in the
-previous section, in that only one problem is solved at a time, and each linear
-algebraic operation is parallelized separately.
-
-For most of the operations, calls to MKL CPU BLAS library map directly onto
-similar calls in Nvidia's GPU BLAS library (CUBLAS).  The CUBLAS library is
-designed to utilize all of the SM's on the GPU.
-
-Operations that do not map well onto the BLAS API (mostly combined
-vector-vector operations) were implemented as gpu kernels (in C for CUDA).
-Even for operations that could have been implemented via multiple BLAS calls,
-this was found to have better performance due to reduced bandwidth and kernel
-call overhead {\bf Victor, verify or correct this.  I'm not sure, currently asking Mark}.
-
-In order to avoid expensive data transfer across the bus connecting the GPU
-card and the CPU motherboard (the PCI express bus), all of the data is
-transferred to GPU DRAM once, and all non-trivial tasks are performed on the
-GPU on data stored in GPU DRAM.  The sole exception to the above implementation
-strategy is that the pseudoinverse of A is still computed on the CPU via the
-LAPACK implementation that is part of Intel's MKL. This operation need only be
-performed (and results uploaded to GPU DRAM) once before GPU computation
-starts. {\bf we should prove justification for A pseudoinverse being cheap,
-either here or perhaps later in a pie chart}
-
-\section{Experiments}
-
+\section{Experiment}
+\label{sec:experiment}
 In this section, we benchmark the performance of our parallel implementations
@@
 makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Unless otherwise specified, all implementations utilize single
-precision floating point datatypes.
+chip).  Unless otherwise specified, all implementations utilize single-precision floating point datatypes.
 
@@
 The first experiment compares the performance of our two proposed implementations of 
-the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
+the general $\ell_1$-min (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
 vector $\xx_0$.  
@@
 \end{center}
-\caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
+\caption{Speed vs Dimensions of $A$.}
 \label{fig:random_data}

### 2011-04-11 main.tex
Quick chanegs

@@
 state-of-the-art $\ell_1$-min solution based on the classical framework of
-\emph{augmented Lagrange multiplier} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
+\emph{augmented Lagrangian methods} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
 
 While solving \eqref{eq:l1min_denoise} can achieve exceptional
-recognition accuracy on public databases, its success relies on the assumption
+recognition accuracy on public databases, its success also relies on the assumption
 that the query image is well aligned with the training images. In
 \cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align
-a query image to one subject class. The face alignment algorithm in turn solves
+a query image to each subject class individually. The face alignment algorithm solves
 the following minimization problem:
@@
 where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations
-(e.g., affine or homography) acting on the image domain, which aligns $\tilde\bb$
-with the training images from the $i$-th class $A_i$. The algorithm 
-extends Lucas-Kanade iterative alignment \cite{LucasB1981} to both to use the
-$\ell_1$ norm as a robust error function, and also simultaneously 
-to estimate the illumination model $A\x$.  
-The original solution \cite{WagnerA2009-CVPR} was implemented based 
-on standard interior-point
-methods, and its speed has been since improved by ALM in a more recent paper
-\cite{WagnerA2011-PAMI}. However, primarily due the high per-user cost
-of the alignment step, the recognition system was still not fast enough
-for real-time recognition against datasets of hundreds of users.
-In this paper, in addition to solving the generic
+(e.g., affine or homography) acting on the image domain.  The sequence
+$\tau_i$ converges to a transformation that aligns the test image $\tilde\bb$
+with the training images from the $i$-th class $A_i$. In other words, the algorithm 
+extends Lucas-Kanade iterative alignment \cite{LucasB1981} to use the
+$\ell_1$-norm as a robust error function, while simultaneously estimating
+the coefficients of the linear illumination model $A\x$.
+
+Numerically, the original solution \cite{WagnerA2009-CVPR} was implemented based 
+on standard interior-point methods, and its speed has been since improved by ALM in a more recent paper
+\cite{WagnerA2011-PAMI}. However, due the high per-class cost
+of the alignment step, the recognition system still fails to achieve \emph{real-time recognition} against datasets of hundreds or thousands of users.
+In this paper, in addition to accelerating the generic
 $\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
-discuss how to properly accelerate the iterative face alignment step
+also discuss how to properly accelerate the iterative face alignment step
 \eqref{eq:l1min_alignment} on multi-core CPUs and GPUs.
+
+In this paper we will clearly distinguish between algorithm {\em parallelism},
+which is a property of an algorithm, and hardware {\em concurrency}, which is a
+property of a hardware architecture. Parallelism in an algorithm provides the
+opportunity to perform computations concurrently on hardware.  Algorithms often
+exhibit multiple levels of parallelism, and hardware often provides multiple
+levels of concurrency.  In these terms, the primary contribution of this paper
+will be to determine the optimal mappings between the parallelism available in
+$\ell_1$-min and the concurrency available in multi-core CPU and GPU
+architectures.
 
@@
 generic ALM $\ell_1$-min algorithm and the corresponding face recognition
-routines on massively parallel CPUs and GPUs.  To this end, the face
-recognition system has been fully implemented on a state-of-the-art
+routines on massively parallel CPUs and GPUs.  To this end, a new face
+recognition system for security and access control applications has been fully implemented on a state-of-the-art
 workstation, as shown in Figure \ref{fig:pipeline}. 
@@
 the architecture of our test system in detail in Section
-\ref{sec:parallelism}.  We will show that a highly accurate and robust face
-recognition system can be constructed to recognize face images from a 
-database of several hundred subjects quickly enough to be practical for
-access control applications. 
-Note that for applications targeting thousands of users or more, it would likely be
-necessary to distribute the work over a network of machines; this paper
-focuses on the (shared memory) parallelism available in a single machine.
-The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture
-will be made available for evaluation by other researchers.
+\ref{sec:concurrency}.
+Finally, note that for applications targeting thousands of users or more, it would likely be
+necessary to distribute the work over a network of computers, while this paper
+focuses on the parallelism available in a single (shared memory) machine.
+The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture,
+will be made available for peer evaluation.
 \begin{figure}
@@
 In Section \ref{sec:ALM}, we contend
-that ALM is a better choice among Homotopy and other existing $\ell_1$-min
-solvers for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
+that ALM is a better choice for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
 category of approximate $\ell_1$-min solutions called \emph{iterative
-shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}.
-IST algorithms mainly involves elementary operations such as vector 
+shrinkage-thresholding} (IST) methods \cite{WrightS2008,BeckA2009}.
+IST algorithms mainly involve elementary operations such as vector 
 algebra and matrix-vector multiplication. Therefore,
 when the dimension of the problem becomes high, IST-type algorithms are
-particularly suitable for systems in parallel-computing environments. In
+particularly suitable for hardware systems with a high degree of concurrency. In
 \cite{YangA2010-ICIP}, the authors further showed that, compared to other
@@
 parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia
-programmable GPUs.
+CUDA programmable GPU architecture, which is also used in this paper.
 
@@
 
-\section{Hardware Concurrency} \label{sec:parallelism}
-
-In this paper we will clearly distinguish between algorithm {\em parallelism},
-which is a property of an algorithm, and hardware {\em concurrency}, which is a
-property of a hardware architecture.  Parallelism in an algorithm provides the
-opportunity to perform computations concurrently on hardware.  Algorithms often
-exhibit multiple levels of parallelism, and hardware often provides multiple
-levels of concurrency.  In these terms, the primary contribution of this paper
-will be to determine the optimal mappings between the parallelism available in
-$\ell_1$-Minimization and the concurrency available in multi-core CPU and GPU
-architectures.
+\section{Hardware Concurrency} \label{sec:concurrency}
 
@@
 \subsection{CPU Hardware Concurrency} 
+\label{sec:CPU-concurrency}
 The main defining characteristics of contemporary multi-core CPU architectures
@@
 Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache.
-Between the 8 cores, this is a total of 2\,MB L2 cache.
-For the low resolution images used in sparse representation based
-face recognition (on the order of $64\times64$ pixel images), the L2 cache
-can in principle hold up to 128 image vectors.
 %16MB = 2**24 bytes = 16777216 bytes
@@
 %... so can fit 2**10  = 1024 images.
-Each processor further has 8MB of L3 cache that is shared by all four cores.
-Overall, this means the algorithm has approximately 16MB of on-chip cache
-available.  This corresponds to 1024 image vectors.  A scale drawing
-of the caches of the processor we're targeting can be seen in Figure \ref{fig:caches}.
+Each processor further has 8\,MB of L3 cache that is shared by the four cores.
+Overall, the algorithm has approximately 16\,MB of L3 cache
+available for a dual-processor configuration.
 
-%With Hyperthreading enabled, each core can support two threads simultaneously.
-For floating point instructions, each core also has a vector processing unit
-(SSE) capable of performing the same arithmetic operation on four single
-precision floating point values simultaneously.
+For floating-point instructions, each core also has a vector processing unit
+(SSE) capable of performing the same arithmetic operation on four single-precision 
+floating point values simultaneously \footnote{By the publication date, the vector
+units of typical high-end CPUS will double.}.
 
@@
 are that they have two (much wider) levels of concurrency, relatively small amounts of
-on-chip cache, and relatively low clock speeds.
+on-chip cache, and relatively low clock speeds.  Whereas most of the transistors on
+a typical CPU are dedicated either to cache or to hardware that enables higher
+clock speeds (such as branch prediction, out-of-order instruction execution, etc...),
+most of the transistors on a GPU are dedicated to arithmetic logic units.
+An explanation of the CUDA programming model at a useful level of completeness would take 
+more space than is available here, so we will instead frame our discussion in terms of 
+hardware capabilities, and refer the reader to the NVIDIA C for CUDA Programming Guide.
 
-Compared to CPU architectures, GPU architectures dedicate less of their transistors
-to cache, and more of them for arithmetic logic. Our target GPU architecture
-is the Nvidia Fermi architecture.  CUDA based GPUs are comprised of several 
-streaming multiprocessors (SMP) (currently 14 to 16 for a high end GPU).
+Our target GPU architecture
+is the Nvidia Fermi architecture (\eg, the GTX 480 GPU used in this paper). 
+CUDA based GPUs are comprised of several \emph{streaming multiprocessors} (SMP) 
+(currently 14 to 16 for a high end GPU).  Each SMP is capable of executing up to 64
+floating point operations concurrently.
 
-The GPU executes programs as SIMT (Single Instruction Multiple Threads) in
-groups of 32 threads called \emph{warps}.  These warps are then grouped into
-higher levels called \emph{thread blocks} and \emph{grids}.  It is impossible
-to communicate between SMPs during gpu program execution, so computational tasks
-are typically broken up into multiple gpu programs (kernels) that are executed
-serially.
+%The GPU uses a SPMD (Single Program Multiple Data) programming model: at any given
+%point in time, all threads are running the same gpu program, or \emph{kernel}, but
+%they all operate on different data.  During kernel execution, threads can only communicate
+%(or synchronize with each other) within groups of threads called blocks.  Each block
+%is further subdivided into groups of 32 threads called \emph{warps}. 
+%Typically, an SMP will hold several thread blocks at a time, although it "only" has enough 
+%hardware to schedule
+%2 warps to perform a floating point operation at a time (the remaining warps are either
+%inactive, or waiting for data to arrive).  Threads within a block are able to synchronize
+%(wait at certain points of a program), which gives them a chance to communicate via a
+%relatively small amount of shared cache.
 
@@
 and software managed ("shared") memory.  Additionally, all SMPs share a common
-L2-cache.  For our system, each SMP has 64 KB of L1 cache with a total of 896
+L2-cache.  For our system, each SMP has 64 KB of L1 cache for a total of 896
 KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount

### 2011-04-17 main.tex
Still merging...

@@
 There are thus two important levels of concurrency that need to be exploited to
-efficiently use a modern CPU: {\em core-level} concurrency and {\em vector-level}
+efficiently use a modern CPU: {\em core-level} concurrency and {\em sse-level}
 concurrency.
@@
 more space than is available here, so we will instead frame our discussion in terms of 
-hardware capabilities, and refer the reader to the NVIDIA C for CUDA Programming Guide.
+hardware capabilities\footnote{A discussion of the CUDA programming model can be found
+in the NVIDIA C for CUDA Programming Guide}.
 
-Our target GPU architecture
-is the Nvidia Fermi architecture (\eg, the GTX 480 GPU used in this paper). 
-CUDA based GPUs are comprised of several \emph{streaming multiprocessors} (SMP) 
-(currently 14 to 16 for a high end GPU).  Each SMP is capable of executing up to 64
-floating point operations concurrently.
-
-%The GPU uses a SPMD (Single Program Multiple Data) programming model: at any given
-%point in time, all threads are running the same gpu program, or \emph{kernel}, but
-%they all operate on different data.  During kernel execution, threads can only communicate
-%(or synchronize with each other) within groups of threads called blocks.  Each block
-%is further subdivided into groups of 32 threads called \emph{warps}. 
-%Typically, an SMP will hold several thread blocks at a time, although it "only" has enough 
-%hardware to schedule
-%2 warps to perform a floating point operation at a time (the remaining warps are either
-%inactive, or waiting for data to arrive).  Threads within a block are able to synchronize
-%(wait at certain points of a program), which gives them a chance to communicate via a
-%relatively small amount of shared cache.
+Our target GPU architecture is the Nvidia Fermi architecture (\eg, the GTX 480
+GPU used in this paper).  CUDA based GPUs are comprised of several
+\emph{streaming multiprocessors} (SMP), each of which is roughly analogous to a CPU core.
+For Fermi architecture GPU's, there are up to 16 SMP's, and each SMP is capable of executing up to 64 single precision 
+floating point operations
+concurrently\footnote{In CUDA terms, the warp width is 32, and the floating pipeline can
+issue up to two warps simultaneously}.
 
@@
 and software managed ("shared") memory.  Additionally, all SMPs share a common
-L2-cache.  For our system, each SMP has 64 KB of L1 cache for a total of 896
-KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount
+L2-cache.  For our system, each SMP has 64 KB of L1 cache, and all SMP's share
+768KB of L2 cache. 
+%The cache hierarchy 
+The relatively small amount
 of cache (1/23 as much as CPU L3) on the GPU is balanced by a significantly
@@
 to the CPU.  A scale drawing of the caches available on the GPU can be seen in Figure \ref{fig:caches}.
-
-The GPU has its own memory system, any data the GPU uses must first be
+The GPU has its own memory system, and any data the GPU uses must first be
 transferred from CPU DRAM to GPU DRAM over PCI-Express.  For our application,
-this transfer overhead is amortized over a large amount of computation and is
+this transfer overhead can be amortized over a large amount of computation and is
 not a major concern.
 
-Compared to a typical CPU, a GPU has much higher peak computational bandwidth
-as well as data bandwith to DRAM.  However, since our application is primarily
-of matrix-vector operations and image resampling operations, it has a very low
-arithmetic intensity (i.e. we perform a very small number of arithmetic
-operations for every number that has to be loaded), and thus performance is
-typically constrained by memory (and cache) bandwidth.  At large problem sizes
-the CPU is limited by its relatively low bandwidth to DRAM compared to the gpu,
-while at small problem sizes the GPU is limited by the relatively high latency
-of its access to DRAM.
+While the programming model for
+the SIMD units of each SMP is somewhat more flexible than on the CPU,
+leveraging the flexibility typically comes at the cost of reduced parallelism
+\footnote{In CUDA terms, code branches that cause warp divergence usually
+result in serialization of the different groups of threads}.  Therefore, for
+the purposes of comparing hardware architectures, the SIMD units on the GPU are
+roughly analogous to the CPU SIMD units. Thus, in summary, the GPU hardware 
+provides two levels of concurrency: {\em SMP-level}
+concurrency and {\em warp-level} concurrency.  Note that the GPU provides more
+concurrency than a cpu at both levels (14 SMP's vs. 8 cores) and (64 wide vs. 4 wide SIMD
+units).
 
@@
 \label{sec:alignment}
-There are essentially two levels of parallelism available for exploitation in
-for iterative alignment in the face recognition application.  The primary goal
-of this section is to introduce these two levels of parallelism.  Additionally,
-in order for some important aspects of our implementations to be clear later,
-we need to establish how the iterative alignment stage works in some
-detail.
 
-The inputs of the iterative alignment stage are the test image $f(x,y)$ stacked into an $M$-
-dimensional vector $\f$, and the aligned and stacked training images $A_i$ for
-each training user $i$. 
-The algorithm repeatedly solves a linearized version of the minimization problem
-in \eqref{eq:l1min_alignment}.  This involves two computational steps:
-the computation of the linearization, and the solution of the minimization problem.
+Given the hardware concurrency in modern parallel computing architectures, next,
+we shall study the optimal mappings of the algorithm parallelism in face recognition to the multi-core CPUs and GPUs, respectively.
+In this section, we first focus on the face alignment stage (see Figure \ref{fig:pipeline}).
 
-The computation of the linearization is complicated by the need to avoid trivial 
-solutions where the algorithm
-progressively zooms in on a black pixel of the test image.
-This is achieved by minimizing the $\ell_1$ error between the illumination model $Ax$ and the cropped,
-resampled, and normalized test image $\tilde\bb$, rather than the un-normalized version as in 
-conventional Lucas-Kanade iterative alignment.
-This involves a bit of calculus:
+Face alignment \eqref{eq:l1min_alignment} estimates an image transformation $\tau$ that rectifies the query image $\tilde{\bb}$ with possible pose variation
+w.r.t. each training class $A_i$, which leads to the minimal sparsity in error $\ee$ after the alignment. Note that directly solving \eqref{eq:l1min_alignment} would require
+simultaneously estimating $\xx$, $\ee$, and $\tau$ typically on a nonlinear manifold, which renders the problem extremely difficult to solve exactly with many local minima.
+
+To mitigate the problem, if the system can obtain to a good initial guess of the transformation $\tau$ (\eg, provided by a good face detector),
+the optimal solution for $\tau$ can be sought iteratively by a sequence of linearized approximations $j=1, 2, \cdots$ as follows:
 \begin{equation}
-\frac{\partial}{\partial \tau_p} \frac{\bb}{||\bb||} 
-= \frac{1}{||\bb||} \frac{\partial \bb}{\partial \tau_p} - 
-\frac{\bb}{||\bb||^3} \bb^T \frac{\partial \bb}{\partial \tau_p}
+\min_{\xx, \ee, \Delta \tau_j}\|\ee\|_1\quad \subj\quad \tilde\bb\circ\tau_j +  J_j\Delta \tau_j = A_i\xx + \ee,
 \end{equation}
@@
 completely disregard problem-level parallelism, and instead map pixel-level
-parallelism onto both core-level and vector-level cpu concurrency.  In other
+parallelism onto both core-level and sse-level cpu concurrency.  In other
 words, the cpu implementation we advocate performs alignment of the test image
@@
 optimized for Intel multicore CPU's, and are able to make good use of both
-core-level and vector-level concurrency.  For operations that do not map well
-onto existing libraries, we achieve both vector-level and core-level
+core-level and sse-level concurrency.  For operations that do not map well
+onto existing libraries, we achieve both sse-level and core-level
 concurrency through a combination of the OpenMP API \cite{dagum2002openmp} and
@@
 This implementation was written using the OpenCV library
-as a wrapper for MKL, and, like our proposed implementation, achieves vector-level concurrency implicitly.  
+as a wrapper for MKL, and, like our proposed implementation, achieves sse-level concurrency implicitly.  
 With eight alignment problems (associated with 8 training users) being solved
@@
 than off-chip memory, this implementation is slower than the proposed implementation.
-
-%During the course of development, many implementation strategies were tried.
-%Listed in order of increasing performance based on a synthetic benchmark of
-%the $\ell_1$ minimization step, they included:
-%\begin{enumerate}
-%\item A pure C++ implementation threaded and vectorized with OpenMP and icpc (200 ms).
-%\item An OpenCV (with MKL backend) implementation threaded with OpenMP similar to \cite{WagnerA2011-PAMI} (190ms.)
-%\item A straightforward MATLAB implementation (which uses threaded MKL internally) (60 ms.)
-%\end{enumerate}
-
-%\subsection{Tuning the GEMV operation}
-%When matrix-vector operations are called sequentially on the same data,
-%the operation can be blocked so that each thread operates on the same block of data for each operation.
-%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
-%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
-%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!
-
-%\subsection{Tuning the Soft-Thresholding Operator}
-%...Make sure there are no branches
-%...Make sure there are no function calls
-%...Make sure the loop gets automatically vectorized
-%...Block the computation into work for different cpu cores
 
@@
 the GPU's L2 cache.
-% DREW Idea... in hindsight, maybe we should still try many kernel launches. 
-% instances can still exit early, and can still update a global variable signaling
-% that they're done
-
-% Should really verify somehow that we're maxing out GPU bandwidth
 
@@
 
+Compared to a typical CPU, a GPU has much higher peak computational bandwidth
+as well as data bandwith to DRAM.  However, since our application is primarily
+of matrix-vector operations and image resampling operations, it has a very low
+arithmetic intensity (i.e. we perform a very small number of arithmetic
+operations for every number that has to be loaded), and thus performance is
+typically constrained by memory (and cache) bandwidth.  At large problem sizes
+the CPU is limited by its relatively low bandwidth to DRAM compared to the gpu,
+while at small problem sizes the GPU is limited by the relatively high latency
+of its access to DRAM.
+
 {\small

### 2011-04-17 main_allen.tex
Still merging...

@@
 \cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align
-a query image to one subject class. The face alignment algorithm in turn solves
+a query image to each subject class individually. The face alignment algorithm solves
 the following minimization problem:
@@
 where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations
-(e.g., affine or homography) acting on the image domain, which aligns $\tilde\bb$
-with the training (frontal) images from the $i$-th class $A_i$. The algorithm 
-extends Lucas-Kanade iterative alignment \cite{LucasB1981} using the
-$\ell_1$-norm as a robust error function.
+(e.g., affine or homography) acting on the image domain.  The sequence
+$\tau_i$ converges to a transformation that aligns the test image $\tilde\bb$
+with the training images from the $i$-th class $A_i$. In other words, the algorithm 
+extends Lucas-Kanade iterative alignment \cite{LucasB1981} to use the
+$\ell_1$-norm as a robust error function, while simultaneously estimating
+the coefficients of the linear illumination model $A\x$.
 
@@
 routines on massively parallel CPUs and GPUs.  To this end, a new face
-recognition system has been fully implemented on a state-of-the-art
-workstation for security and access control applications, as shown in Figure \ref{fig:pipeline}. 
+recognition system for security and access control applications has been fully implemented on a state-of-the-art
+workstation, as shown in Figure \ref{fig:pipeline}. 
 Since the parallelization of any algorithm
@@
 the architecture of our test system in detail in Section
-\ref{sec:parallelism}.
+\ref{sec:concurrency}.
 Finally, note that for applications targeting thousands of users or more, it would likely be
 necessary to distribute the work over a network of computers, while this paper
-focuses on the (shared memory) parallelism available in a single machine.
+focuses on the parallelism available in a single (shared memory) machine.
 The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture,
@@
 In Section \ref{sec:ALM}, we contend
-that ALM is a better choice among existing $\ell_1$-min
-solvers for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
+that ALM is a better choice for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
 category of approximate $\ell_1$-min solutions called \emph{iterative
@@
 when the dimension of the problem becomes high, IST-type algorithms are
-particularly suitable for systems with parallel-computing pipelines. In
+particularly suitable for hardware systems with a high degree of concurrency. In
 \cite{YangA2010-ICIP}, the authors further showed that, compared to other
@@
 parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia
-programmable GPUs, which is similar to the hardware setup of this paper.
+CUDA programmable GPU architecture, which is also used in this paper.
 
@@
 
-\section{Hardware Concurrency} \label{sec:parallelism}
+\section{Hardware Concurrency} \label{sec:concurrency}
 
@@
 memory (cache) is often orders of magnitude faster than off-chip memory, the
-size and speed of on-chip memory (cache) have big performance implications.
+sizes and speeds of on-chip memory (cache) have big performance implications.
 We give a brief overview of the caches that are available in our target
 architectures, and defer discussion of their performance implications to
-sections \ref{sec:alignment} and \ref{sec:recognition}.
+sections \ref{sec:alignment_implementation_cpu} and
+\ref{sec:alignment_implementation_gpu}.
 
-We will primarily compare CPU and GPU architectures on a per-board basis:
+We will primarily be comparing CPU and GPU architectures on a per-board basis:
 for CPU implementation we make use of all the cores in as many CPUs are present
@@
 are that they have two levels of concurrency, relatively large amounts of
-on-chip cache, and relatively high clock speed.
+on-chip cache, and relatively high clock speeds.
 
-The baseline architecture for our experiment is a Linux workstation with two
+The baseline architecture for our experiments is a Linux workstation with two
 quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
@@
 %... so can fit 2**7  = 128 images.
-Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache (\ie, a total of 2\,MB L2 cache available for the eight cores).
-Therefore, for the low-resolution images used in face recognition (say, on the order of $64\times64$ grayscale images), the L2 cache
-in principle can hold up to 16 images per core and 128 images in total.
+Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache.
 %16MB = 2**24 bytes = 16777216 bytes
@@
 Overall, the algorithm has approximately 16\,MB of L3 cache
-available for a dual-processor configuration, which corresponds to about 1024 images.
+available for a dual-processor configuration.
 
-%With Hyperthreading enabled, each core can support two threads simultaneously.
 For floating-point instructions, each core also has a vector processing unit
-(SSE) capable of performing the same arithmetic operation on four single-precision floating point values simultaneously.
+(SSE) capable of performing the same arithmetic operation on four single-precision 
+floating point values simultaneously \footnote{By the publication date, the vector
+units of typical high-end CPUS will double.}.
 
 There are thus two important levels of concurrency that need to be exploited to
-efficiently use a modern CPU: {\em core-level} concurrency and {\em vector-level}
+efficiently use a modern CPU: {\em core-level} concurrency and {\em sse-level}
 concurrency.
@@
 are that they have two (much wider) levels of concurrency, relatively small amounts of
-on-chip cache, and relatively low clock speed.
+on-chip cache, and relatively low clock speeds.  Whereas most of the transistors on
+a typical CPU are dedicated either to cache or to hardware that enables higher
+clock speeds (such as branch prediction, out-of-order instruction execution, etc...),
+most of the transistors on a GPU are dedicated to arithmetic logic units.
+An explanation of the CUDA programming model at a useful level of completeness would take 
+more space than is available here, so we will instead frame our discussion in terms of 
+hardware capabilities\footnote{A discussion of the CUDA programming model can be found
+in the NVIDIA C for CUDA Programming Guide}.
 
-Compared to CPU architectures, modern GPU architectures dedicate less of their transistors
-to cache, and more of them for arithmetic logic. Our target GPU architecture
-is the Nvidia Fermi architecture (\eg, GTX 480 GPU used in this paper). CUDA based GPUs are comprised of several 
-\emph{streaming multiprocessors} (SMP) (currently 14 to 16 for a high end GPU).
+Our target GPU architecture is the Nvidia Fermi architecture (\eg, the GTX 480
+GPU used in this paper).  CUDA based GPUs are comprised of several
+\emph{streaming multiprocessors} (SMP), each of which is roughly analogous to a CPU core.
+For Fermi architecture GPU's, there are up to 16 SMP's, and each SMP is capable of executing up to 64 single precision 
+floating point operations
+concurrently\footnote{In CUDA terms, the warp width is 32, and the floating pipeline can
+issue up to two warps simultaneously}.
 
-The GPU executes programs as SIMT (Single Instruction Multiple Threads) in
-groups of 32 threads called \emph{warps}.  These warps are then grouped into
-higher levels called \emph{thread blocks} and \emph{grids}.  Since SMPs are not designed
-to communicate between themselves during GPU program execution, tasks
-are typically broken up into multiple GPU programs (kernels) that are executed
-serially.\footnote{Please add a footnote to further explain.}
+% GPU CACHE DISCUSSION
+% for L2 cache
+% 768KB = 786432 bytes
+% 64*64*4 = 2**14 bytes = 16384 bytes per image
+% 48 images
+Each SMP contains its own L1-cache, which is divided between hardware managed
+and software managed ("shared") memory.  Additionally, all SMPs share a common
+L2-cache.  For our system, each SMP has 64 KB of L1 cache, and all SMP's share
+768KB of L2 cache. 
+%The cache hierarchy 
+The relatively small amount
+of cache (1/23 as much as CPU L3) on the GPU is balanced by a significantly
+higher bandwidth between the processor chip and off-chip memory (DRAM) compared
+to the CPU.  A scale drawing of the caches available on the GPU can be seen in Figure \ref{fig:caches}.
+The GPU has its own memory system, and any data the GPU uses must first be
+transferred from CPU DRAM to GPU DRAM over PCI-Express.  For our application,
+this transfer overhead can be amortized over a large amount of computation and is
+not a major concern.
 
-Each SMP contains its own L1 cache, which is divided between hardware managed
-and software managed (``shared'') memory.  Additionally, all SMPs share a common
-L2 cache.  For our system, each SMP has 64 KB of L1 cache with a total of 896
-KB of L1 cache/shared memory and 768KB of L2 cache. The relatively small amount
-of cache (1/23 as much as CPU L3) on the GPU is ballanced by a significantly
-higher bandwidth between the processor chip and off-chip memory (DRAM) compared
-to the CPU.
-
-Compared to typical CPUs, GPUs have much higher peak computational bandwidth
-as well as data bandwidth to DRAM.  However, since our system primarily consists
-of matrix-vector operations and image resampling operations, it has a very low
-arithmetic intensity (\ie, we perform a very small number of arithmetic
-operations for every number that has to be loaded), and thus performance is
-typically constrained by memory (and cache) bandwidth.  At large problem sizes
-the CPU is limited by its relatively low bandwidth to DRAM compared to the GPU,
-while at small problem sizes the GPU is limited by the relatively high latency
-of its access to DRAM.
+While the programming model for
+the SIMD units of each SMP is somewhat more flexible than on the CPU,
+leveraging the flexibility typically comes at the cost of reduced parallelism
+\footnote{In CUDA terms, code branches that cause warp divergence usually
+result in serialization of the different groups of threads}.  Therefore, for
+the purposes of comparing hardware architectures, the SIMD units on the GPU are
+roughly analogous to the CPU SIMD units. Thus, in summary, the GPU hardware 
+provides two levels of concurrency: {\em SMP-level}
+concurrency and {\em warp-level} concurrency.  Note that the GPU provides more
+concurrency than a cpu at both levels (14 SMP's vs. 8 cores) and (64 wide vs. 4 wide SIMD
+units).
 

### 2011-04-17 main.tex
More mergning... need to print out...

@@
 \end{equation}
-If the warping $\tau$ is constrained to be a similarity transform,
-it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
-reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
-and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
-once per test image.  The per-user iterative alignment algorithm is summarized as Algorithm \eqref{alg:iterative_alignment}.
-\begin{algorithm}[thb]
-\caption{\bf (Per-User Iterative Alignment)} \label{alg:iterative_alignment}
+where $J_j = \nabla_{\tau_j}(\tilde\bb\circ\tau_j)$ is the Jacobian, and
+$\Delta \tau$ is a step update to $\tau$. Denote $\bb_j =
+\tilde\bb\circ\tau_j$, $B_j = [A_i, -J_j]$ and $\ww^T = [\xx^T, \Delta
+\tau^T]$, then the update $\Delta \tau$ can be computed by solving the
+following linear program:
+\begin{equation}
+\min_{\ww, \ee}\|\ee\|_1\quad \subj\quad \bb_j = B_j\ww + \ee.
+\label{eq:alignment-linearization}
+\end{equation}
+%If the warping $\tau$ is constrained to be a similarity transform,
+%it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
+%reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
+%and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
+%once per test image.
+
+The per-class alignment algorithm via ALM is summarized in Algorithm
+\ref{alg:alignment_alm}. In terms of computational complexity, one significant
+term is the computation of the matrix psudo-inverse $B_j^\dagger$, whose
+computational cost is bounded by $O(n_i^2m + n_i^3)$. For per-class face
+alignment, since both $A_i$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$
+for all $i=1, \cdots, K$ classes), we find that the computation time is still
+dominated by the iterative computation of matrix-vector multiplications in the
+inner loop.
+\begin{algorithm}[ht!]
+\caption{\bf (Face Alignment via ALM)} \label{alg:alignment_alm}
+{\bf Input:} $\bb$, $A_i$, $\x_0 = \mathbf{0}$, $\tau_0$, and $J_0$.
 \begin{algorithmic}[1]
-\begin{small}
-\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n}\}$ for gallery subject $i$, the test image
-$\f\in\Re^M$ 
-%\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
-%\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
-%\STATE {\bf for} each gallery subject $i$, 
-\STATE \hspace{0mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
-\STATE \hspace{3mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
-\STATE \hspace{3mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
-\STATE \hspace{3mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
-\STATE \hspace{3mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
-\tau$; 
-\end{small}
+\WHILE{not converged ($j = 1,2,\ldots$)}
+\STATE Update $\bb_j \leftarrow \frac{\bb\circ \tau_{j-1}}{\|\bb\circ \tau_{j-1}\|}$; $B_j= [A_i, -J_{j-1}]$ and corresponding $(B_j^\dagger)^T$
+\STATE Initialize $\ww_0 = \mathbf 0$, $\yy_0 = \mathbf 0$
+\WHILE{not converged ($k = 1,2,\ldots$)}
+\STATE $\uu_0\leftarrow \ww_{k-1}$; $\zz_0\leftarrow \ee_{k-1}$
+\WHILE{not converged ($l = 1,2,\ldots$)}
+\STATE $\zz_l \leftarrow \mbox{shrink}\left(\bb_j - B_j\uu_{l-1} + \frac{\yy_{k}}{\mu_{k-1}}, \frac{1}{\mu_{k-1}}\right)$
+\STATE $\uu_l \leftarrow (B_j^\dagger)^T \left(\bb_j - \zz_{l} + \frac{\yy_{k-1}}{\mu_{k-1}} \right) $
+\ENDWHILE
+\STATE $\ww_k \leftarrow \uu_l$; $\ee_k \leftarrow \zz_l$
+\STATE $\yy_{k} \leftarrow \yy_{k-1} + \mu_{k-1} (\bb_j - B_j\ww_{k} - \e_{k})$
+\STATE $\mu_{k} \leftarrow \rho\mu_{k-1}$
+\ENDWHILE
+\STATE Update $\ee_j$, $\tau_j$, and $J_j$
+\ENDWHILE
 \end{algorithmic}
+{\bf Output:} $\tau^*\leftarrow \tau_j, \e^*\leftarrow \ee_j$
 \end{algorithm}
 
-The solution of the minimization problem in Algorithm
-\eqref{alg:iterative_alignment} Step 7 can be solved via a
-variant of ALM, which is summarized as Algorithm \eqref{alg:alignment_alm}.
-$\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of
-parameter $\mu$, we take the same strategy as in \cite{YangJ2009-pp} and set
-$\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
-
-\begin{algorithm}[h]
-\caption{\bf (Augmented Lagrange Multiplier Method For Alignment)}
-\begin{algorithmic}[1]
-\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
-$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
-\WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
-\WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
-\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
-\ENDWHILE
-\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
-\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
-\ENDWHILE \STATE
-{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
-\end{algorithmic}
-\label{alg:alignment_alm}
-\end{algorithm}
-%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
-
-\subsection{Computational Complexity and Parallelism} Any discussion of the speed of an
-algorithm should discuss the comutational complexity of the algorithm.  For the
-alignment stage ALM algorithm \eqref{alg:alignment_alm}, this is rather simple.
-The Matrix-vector computions are $O(mn)$, and the vector-vector operations are
-$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
-very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
-computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
-A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
-Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
-of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
-Fortunately, $A^\dagger$ only needs to be computed once per instance of the
-$\ell_1$ minimization problem.
-
-Therefore, for problems with large $n$, we expect the computation of the
-pseudoinverse to dominate.  However, for the problem sizes needed in face
-recognition problems, we find that the computation time is dominated by the
-bandwidth required by repeated computation of matrix-vector and matrix-matrix
-multiplications in the inner loop.
-
-In summary, the alignment stage essentially two levels of available
-parallelism: at the higher level there are multiple iterative alignent problems
-that are solved independently, or {\bf problem-level} parallelism.  At a lower
+In summary, the alignment stage essentially contains two levels of available
+parallelism. At the higher level, there are per-class alignment problems
+that are solved independently, or \emph{problem-level parallelism}.  At a lower
 level, the first-order linear algebraic operations exhibit parallelism within
-image operations, i.e. at a pixel level.  We term this {\bf pixel-level}
-parallelism.  In the following two sections will will discuss how to
-exploit these two levels of parallelism on cpu and gpu architectures.
+image operations, \ie, at the pixel level.  We call this \emph{pixel-level
+parallelism}.  In the following two sections will will discuss how to
+exploit these two levels of parallelism on different CPU and GPU architectures.
 
@@
 \label{sec:recognition}
-{\bf AY: Add an algorithm here describe the recognition block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
-{\bf VS: Why did we modify the algorithm for the recognition stage? }
-
-Below is the algorithm for the face recognition stage in Figure \ref{fig:pipeline}, which
-solves the $\ell_1$ minimization problem posed in \eqref{eq:l1min_denoise}.
-
+After the face alignment stage is complete, a single recognition algorithm will
+be called to estimate a sparse representation of the query image w.r.t. all the
+training classes, as shown in \eqref{eq:l1min_denoise}. In
+\cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI}, the recognition
+algorithm via $\ell_1$-min has been extensively studied. The algorithm itself
+via ALM is summarized in Algorithm \ref{alg:alm_rec}.  In this section, we
+mainly focus our discussion on its pixel-level parallelism on the multicore
+CPUs and GPUs. Then in Section \ref{sec:experiment}, we benchmark the
+performance of the two architectures and further demonstrate the improvement in
+both accuracy and speed compared to the existing literature.
 \begin{algorithm}[t]
-\caption{\bf (Face Recognition via ALM)}
+\caption{\bf (Face Recognition via ALM)} \label{alg:alm_rec} 
 \begin{algorithmic}[1]
@@
 \end{algorithmic}
-\label{alg:alm_rec} 
 \end{algorithm}
@@
 \section{Experiments}
-
+\label{sec:experiment}
 In this section, we benchmark the performance of our parallel implementations
@@
 makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Unless otherwise specified, all implementations utilize single
-precision floating point datatypes.
+chip).  Unless otherwise specified, all implementations utilize single-precision 
+floating point datatypes.
 
@@
 
-The first experiment compares the performance of our two proposed implementations of 
-the general $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
-vector $\xx_0$.  
-The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
-randomly generated from the standard normal and normalized with unit column norm.  The ratio of $d:n$ varied from
-$1:10$ to $5:10$ with $n$ varying from 500 to 8000.
-The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and also normalized with unit norm.
-Because the ground truth signal $\xx_0$ was known, the algorithm terminated
-when $\|\bx-\bx_0\| < \tau$.
-The measurement vector was generated with the equation $\bb_0 = A \xx_0$.
+The first experiment compares the performance of our two proposed
+implementations of the general $\ell_1$-min (Algorithm \eqref{alg:alm}) given a
+generic random matrix $A$ and a random sparse vector $\xx_0$.  The measurement
+$d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each
+entry randomly generated from the standard normal and normalized with unit
+column norm.  The ratio of $d:n$ varied from $1:10$ to $5:10$ with $n$ varying
+from 500 to 8000.  The ground truth signals, $\xx_0$ had a sparsity rate of 0.1
+with elements generated using the standard normal and also normalized with unit
+norm.  Because the ground truth signal $\xx_0$ was known, the algorithm
+terminated when $\|\bx-\bx_0\| < \tau$.  The measurement vector was generated
+with the equation $\bb_0 = A \xx_0$.
 
@@
 The x-axis represents the size of the $A$ matrix and the y-axis represents the
-average amount of time it to complete a single problem instance.  The GPU implementation tends to be faster than
-the CPU implementation at solving a single large problem, whereas the the CPU
-implementation is faster at solving a single small problem.  Figure \ref{fig:random_data2} shows the various sizes and tolerances at which our GPU implementation
-begins to run faster than the CPU implementation on synthetic data and will provide researchers 
-with a starting point of where to begin setting their tolerance levels for the ALM algorithm.
+average amount of time it to complete a single problem instance.  The GPU
+implementation tends to be faster than the CPU implementation at solving a
+single large problem, whereas the the CPU implementation is faster at solving a
+single small problem.  Figure \ref{fig:random_data2} shows the various sizes
+and tolerances at which our GPU implementation begins to run faster than the
+CPU implementation on synthetic data and will provide researchers with a
+starting point of where to begin setting their tolerance levels for the ALM
+algorithm.
 \begin{figure}

### 2011-04-17 main_allen.tex
More mergning... need to print out...

@@
 \end{equation}
-where $J_j = \nabla_{\tau_j}(\tilde\bb\circ\tau_j)$ is the Jacobian, and $\Delta \tau$ is a step update to $\tau$.
-Denote $\bb_j = \tilde\bb\circ\tau_j$, $B_j = [A_i, -J_j]$ and $\ww^T = [\xx^T, \Delta \tau^T]$, then the update $\Delta \tau$ can be computed by solving the following linear program:
+where $J_j = \nabla_{\tau_j}(\tilde\bb\circ\tau_j)$ is the Jacobian, and
+$\Delta \tau$ is a step update to $\tau$. Denote $\bb_j =
+\tilde\bb\circ\tau_j$, $B_j = [A_i, -J_j]$ and $\ww^T = [\xx^T, \Delta
+\tau^T]$, then the update $\Delta \tau$ can be computed by solving the
+following linear program:
 \begin{equation}
@@
 \end{equation}
+%If the warping $\tau$ is constrained to be a similarity transform,
+%it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
+%reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
+%and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
+%once per test image.
 
-The per-class alignment algorithm via ALM is summarized in Algorithm \ref{alg:alignment_alm}.
-In terms of computational complexity, one significant term is the computation of the matrix psudo-inverse $B_j^\dagger$, whose computational cost is bounded by $O(n_i^2m + n_i^3)$.
-For per-class face alignment, since both $A_i$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$ for all $i=1, \cdots, K$ classes), we find that the computation time is still dominated by the iterative computation of matrix-vector multiplications in the inner loop.
+The per-class alignment algorithm via ALM is summarized in Algorithm
+\ref{alg:alignment_alm}. In terms of computational complexity, one significant
+term is the computation of the matrix psudo-inverse $B_j^\dagger$, whose
+computational cost is bounded by $O(n_i^2m + n_i^3)$. For per-class face
+alignment, since both $A_i$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$
+for all $i=1, \cdots, K$ classes), we find that the computation time is still
+dominated by the iterative computation of matrix-vector multiplications in the
+inner loop.
 \begin{algorithm}[ht!]
@@
 
-As we have calculated in Section \ref{sec:CPU-concurrency}, the L2 cache in each core is only able to store about 16 images.
-For image alignment problem, the number of training images per class is typically larger than 16.
-Hence, if \eqref{eq:alignment-linearization} would be solved in parallel on the eight cores, it would require L3 cache to hold data up to four classes at a time, and the L2 cache should retrieve subsets of the data frequently during the iterative process to execute Algorithm \ref{alg:alignment_alm}.
-Therefore, the bottleneck in this strategy becomes the bandwidth between the L2 cache and L3 cache.
-The implementation of some previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this mapping strategy.
+As we have calculated in Section \ref{sec:CPU-concurrency}, the L2 cache in
+each core is only able to store about 16 images. For image alignment problem,
+the number of training images per class is typically larger than 16. Hence, if
+\eqref{eq:alignment-linearization} would be solved in parallel on the eight
+cores, it would require L3 cache to hold data up to four classes at a time, and
+the L2 cache should retrieve subsets of the data frequently during the
+iterative process to execute Algorithm \ref{alg:alignment_alm}. Therefore, the
+bottleneck in this strategy becomes the bandwidth between the L2 cache and L3
+cache. The implementation of some previous works
+\cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this mapping
+strategy.
 
-In this paper, we show that for CPU parallelism, a more traditional implementation for pixel-level parallelism in fact outperforms problem-level parallelism using eight separate cores.
-In other words, the CPU implementation performs alignment of a query image against the training images of single or dual users at a time on the dual processors, depending on the setting of the compiler.
-More specifically, in our implementation, most of the operations take advantage of Intel Math Kernel Library (MKL), a commercial implementation of the standard Basic Linear Algebra Subprograms (BLAS).
-For the image resampling step, we make use of the Intel Integrated Performance Primitives (IPP) library.
-Both MKL and IPP are optimized for Intel multicore CPUs, and are able to automatically utilize both core-level and vector-level concurrency.
-For other operations that are not optimized by Intel in-house libraries, we achieve both vector-level and core-level concurrency through a combination of the Open MP API \cite{dagum2002openmp} and the automatic vectorization facilities of the Intel C++ compiler (ICC) \cite{dulong1999overview}.
-We shall compare the performance of the two parallelism methods in Section \ref{sec:experiment}.
+In this paper, we show that for CPU parallelism, a more traditional
+implementation for pixel-level parallelism in fact outperforms problem-level
+parallelism using eight separate cores. In other words, the CPU implementation
+performs alignment of a query image against the training images of single or
+dual users at a time on the dual processors, depending on the setting of the
+compiler. More specifically, in our implementation, most of the operations take
+advantage of Intel Math Kernel Library (MKL), a commercial implementation of
+the standard Basic Linear Algebra Subprograms (BLAS). For the image resampling
+step, we make use of the Intel Integrated Performance Primitives (IPP) library.
+Both MKL and IPP are optimized for Intel multicore CPUs, and are able to
+automatically utilize both core-level and vector-level concurrency. For other
+operations that are not optimized by Intel in-house libraries, we achieve both
+vector-level and core-level concurrency through a combination of the Open MP
+API \cite{dagum2002openmp} and the automatic vectorization facilities of the
+Intel C++ compiler (ICC) \cite{dulong1999overview}. We shall compare the
+performance of the two parallelism methods in Section \ref{sec:experiment}.
 
@@
 
-Similarly, we also consider the trade-off between the available cache sizes and their bandwidth based on the Nvidia CUDA GPU architecture chosen for the paper.
-Figure \ref{fig:alignment_mapping_gpu} illustrates the multicore architecture and the L1/L2 caches of the Nvidia GTX 480 GPU.
+Similarly, we also consider the trade-off between the available cache sizes and
+their bandwidth based on the Nvidia CUDA GPU architecture chosen for the paper.
+Figure \ref{fig:alignment_mapping_gpu} illustrates the multicore architecture
+and the L1/L2 caches of the Nvidia GTX 480 GPU.
 \begin{figure}
@@
 \label{sec:recognition}
-After the face alignment stage is complete, a single recognition algorithm will be called
-to estimate a sparse representation of the query image w.r.t. all the training classes, 
-as shown in \eqref{eq:l1min_denoise}. In \cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI},
-the recognition algorithm via $\ell_1$-min has been extensively studied. The algorithm itself via ALM
-is summarized in Algorithm \ref{alg:alm_rec}.
-In this section, we mainly focus our discussion on its pixel-level parallelism on the multicore CPUs and GPUs. Then in Section \ref{sec:experiment}, we benchmark
-the performance of the two architectures and further demonstrate the improvement in both accuracy and speed
-compared to the existing literature.
+After the face alignment stage is complete, a single recognition algorithm will
+be called to estimate a sparse representation of the query image w.r.t. all the
+training classes, as shown in \eqref{eq:l1min_denoise}. In
+\cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI}, the recognition
+algorithm via $\ell_1$-min has been extensively studied. The algorithm itself
+via ALM is summarized in Algorithm \ref{alg:alm_rec}.  In this section, we
+mainly focus our discussion on its pixel-level parallelism on the multicore
+CPUs and GPUs. Then in Section \ref{sec:experiment}, we benchmark the
+performance of the two architectures and further demonstrate the improvement in
+both accuracy and speed compared to the existing literature.
 \begin{algorithm}[t]
@@
 
-For CPU implementation, {\bf AY: Drew, please fill in one or two paragraphs}.
+\subsection{CPU Implementation} Our CPU implementation of Algorithm
+\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
+nicely onto the BLAS API (such as SGEMV, SAXPY), we
+call Intel's MKL BLAS library, which internally takes advantage of the SSE
+units on all available CPU cores.
 
-For GPU implementation, {\bf AY: Drew, please fill in one or two paragraphs}.
+For operations that do not map nicely onto the BLAS API, such as the shrinkage
+operator, {\bf Argue that these parts are optimized, and/or demonstrate that
+performance of these parts is insignificant}.
 
-\section{Experiment}
+\subsection{GPU Implementation} Our GPU implementation of Algorithm
+\ref{alg:alm_rec} is similar to the CPU implementation described in the
+previous section, in that only one problem is solved at a time, and each linear
+algebraic operation is parallelized separately.
+
+For most of the operations, calls to MKL CPU BLAS library map directly onto
+similar calls in Nvidia's GPU BLAS library (CUBLAS).  The CUBLAS library is
+designed to utilize all of the SM's on the GPU.
+
+Operations that do not map well onto the BLAS API (mostly combined
+vector-vector operations) were implemented as gpu kernels (in C for CUDA).
+Even for operations that could have been implemented via multiple BLAS calls,
+this was found to have better performance due to reduced bandwidth and kernel
+call overhead {\bf Victor, verify or correct this.  I'm not sure, currently asking Mark}.
+
+In order to avoid expensive data transfer across the bus connecting the GPU
+card and the CPU motherboard (the PCI express bus), all of the data is
+transferred to GPU DRAM once, and all non-trivial tasks are performed on the
+GPU on data stored in GPU DRAM.  The sole exception to the above implementation
+strategy is that the pseudoinverse of A is still computed on the CPU via the
+LAPACK implementation that is part of Intel's MKL. This operation need only be
+performed (and results uploaded to GPU DRAM) once before GPU computation
+starts. {\bf we should prove justification for A pseudoinverse being cheap,
+either here or perhaps later in a pie chart}
+
+\section{Experiments}
 \label{sec:experiment}
@@
 makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Unless otherwise specified, all implementations utilize single-precision floating point datatypes.
+chip).  Unless otherwise specified, all implementations utilize single-precision 
+floating point datatypes.
 
@@
 
-The first experiment compares the performance of our two proposed implementations of 
-the general $\ell_1$-min (Algorithm \eqref{alg:alm}) given a generic random matrix $A$ and a random sparse
-vector $\xx_0$.  
-The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
-randomly generated from the standard normal and normalized with unit column norm.  The ratio of $d:n$ varied from
-$1:10$ to $5:10$ with $n$ varying from 500 to 8000.
-The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and also normalized with unit norm.
-Because the ground truth signal $\xx_0$ was known, the algorithm terminated
-when $\|\bx-\bx_0\| < \tau$.
-The measurement vector was generated with the equation $\bb_0 = A \xx_0$.
+The first experiment compares the performance of our two proposed
+implementations of the general $\ell_1$-min (Algorithm \eqref{alg:alm}) given a
+generic random matrix $A$ and a random sparse vector $\xx_0$.  The measurement
+$d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each
+entry randomly generated from the standard normal and normalized with unit
+column norm.  The ratio of $d:n$ varied from $1:10$ to $5:10$ with $n$ varying
+from 500 to 8000.  The ground truth signals, $\xx_0$ had a sparsity rate of 0.1
+with elements generated using the standard normal and also normalized with unit
+norm.  Because the ground truth signal $\xx_0$ was known, the algorithm
+terminated when $\|\bx-\bx_0\| < \tau$.  The measurement vector was generated
+with the equation $\bb_0 = A \xx_0$.
 

### 2011-04-17 main_allen.tex
Fixed some writing in CPU and GPU alignment implementation sections.

@@
 
-The optimal strategy to implement Algorithm \ref{alg:alignment_alm} on a multicore CPU must take into account
-the trade-off between the available memory/cache sizes at multiple hierarchies and their bandwidth based on a particular
-CPU architecture. Figure \ref{fig:alignment_mapping_old} shows the the sizes of L2 and L3 caches of the Intel E5530 CPU, 
-which is a good example of modern multicore CPU architectures. 
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_old}
-\caption{Available L2 and L3 caches on the Intel E5530 multicore CPU architecture with dual CPU configuration.}
-\label{fig:alignment_mapping_old}
-\end{figure}
+Optimal implementation of Algorithm \ref{alg:alignment_alm} on a multicore CPU must take 
+into account the properties of the cache hierarchy. In general, cache that is 
+closer to the core (i.e. L1 cache) has higher bandwidth, but smaller size compared to cache
+that is farther from the core (L3 cache).  For reference, Figure \ref{fig:caches} 
+shows the the sizes of L2 and L3 caches of the Intel E5530 CPU, 
+which is a representative example of a modern multicore CPU.
 
-As we have calculated in Section \ref{sec:CPU-concurrency}, the L2 cache in
-each core is only able to store about 16 images. For image alignment problem,
-the number of training images per class is typically larger than 16. Hence, if
-\eqref{eq:alignment-linearization} would be solved in parallel on the eight
-cores, it would require L3 cache to hold data up to four classes at a time, and
-the L2 cache should retrieve subsets of the data frequently during the
-iterative process to execute Algorithm \ref{alg:alignment_alm}. Therefore, the
-bottleneck in this strategy becomes the bandwidth between the L2 cache and L3
-cache. The implementation of some previous works
-\cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this mapping
-strategy.
+For the E5530, the L2 cache in each core is only able to store about 16 images.
+For image alignment problem, the number of training images per class is
+typically larger than 16.  Therefore, if eight instances of
+\eqref{eq:alignment-linearization} were be executed in parallel on the eight
+cores, there would be an insufficient amount of L1 or L2 cache to hold $A$ and
+$A^\dagger$ during the iterative process to execute Algorithm
+\ref{alg:alignment_alm}. Therefore, the bottleneck in this strategy becomes the
+bandwidth between the L2 cache and L3 cache. The implementation of some
+previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this
+mapping strategy.
 
-In this paper, we show that for CPU parallelism, a more traditional
-implementation for pixel-level parallelism in fact outperforms problem-level
-parallelism using eight separate cores. In other words, the CPU implementation
-performs alignment of a query image against the training images of single or
-dual users at a time on the dual processors, depending on the setting of the
-compiler. More specifically, in our implementation, most of the operations take
+In this paper, we show that for CPU parallelism of face alignment, a higher
+performance implementation can be achieved by {\em not} taking advantage of
+problem-level parallelism at all, and instead map the pixel-level parallelism
+onto both the core-level and the sse-level concurrency provided by the cpu.  In
+other words, the proposed CPU implementation performs alignment of a query
+image against the training images of a single user at a time using all cores on
+both processors.
+This mapping of the problem is illustrated in Figure \ref{fig:alignment_mapping_cpu}.
+
+More specifically, in our implementation, most of the operations take
 advantage of Intel Math Kernel Library (MKL), a commercial implementation of
@@
 \label{sec:alignment_implementation_gpu}
+While on the CPU, algorithm performance is highly dependent on effective
+cache usage, on the GPU cache is relatively unimportant for ALM based
+$\ell_1$-min.  As can be seen in Figure \ref{fig:caches}, the GPU has 
+a very small amount of cache compared to the CPU.  While it might be possible
+to fit a single instance of the alignment problem (with a slightly reduced problem size)
+into L2 cache, this would not be an efficient use of the GPU's resources.
+The strength of the GPU's memory architecture for our purposes it's ability to sustain a very
+high bandwidth to DRAM.  A higher overall bandwidth can be achieved if 
+many memory transactions are performed simultaneously, than can be achieved
+with a large number of smaller transactions that are just large enough to fill
+the L2 cache.
 
-Similarly, we also consider the trade-off between the available cache sizes and
-their bandwidth based on the Nvidia CUDA GPU architecture chosen for the paper.
-Figure \ref{fig:alignment_mapping_gpu} illustrates the multicore architecture
-and the L1/L2 caches of the Nvidia GTX 480 GPU.
+Therefore, whereas on the CPU we advocate performing alignment for one training
+class at a time, on the GPU we advocate performing alignment for many classes
+concurrently.  \footnote{In CUDA terminology, Our proposed alignment stage
+implementation consists of a single kernel that performs alignment for all
+subject classes.  Each instance of the alignment problem is assigned its own
+thread block, and the GPU hardware schedules as many thread blocks to run
+concurrently as the hardware will allow.} The number of subject classes that
+actually get aligned concurrently is chosen by the gpu hardware. In our
+experiments the hardware is able to schedule 4-5 classes per SMP to run
+concurrently. This mapping of the problem-level parallelism onto the SMP-level
+concurrency is illustrated in Figure \ref{fig:alignment_mapping_gpu}.
+
 \begin{figure}
@@
 
-Our proposed GPU implementation for the iterative alignment stage consists of a
-single kernel that performs all of the test image alignment problems.  Each of
-instance of the alignment problem is assigned its own thread block, and the GPU
-hardware schedules as many of these thread blocks to run concurrently as the 
-hardware will allow.
-
 In addition, the full resolution query image, along with its numerical derivatives in both
@@
 
-In one experiment, we have exploited several possible GPU implementation strategies, which include
-solving single alignment problems one at a time using CUBLAS (302\,ms); solving one problem
-per SMP using CUDA streams (70\,ms); and solving one problem per SMP using one large kernel call
- (36\,ms). The inferior performance of CUBLAS may be attributed to the relatively small problem
- size. In Section \ref{sec:experiment}, we will further exploit the difference between CPU and GPU
- via synthetic and real image data.
+In one experiment, we have exploited several possible GPU implementation
+strategies, which include solving single alignment problems one at a time using
+CUBLAS (302\,ms); solving one problem per SMP using CUDA streams (70\,ms); and
+solving one problem per SMP using one large kernel call (36\,ms). The inferior
+performance of CUBLAS may be attributed to the relatively small problem size.
+In general, the libraries provided by Nvidia are optimized for problem sizes
+that are significantly larger than our individual alignment problems.  In
+Section \ref{sec:experiment}, we show the performance of our CPU and GPU
+implementations running on synthetic and real image data.
 
@@
 vector-vector operations) were implemented as gpu kernels (in C for CUDA).
-Even for operations that could have been implemented via multiple BLAS calls,
-this was found to have better performance due to reduced bandwidth and kernel
-call overhead {\bf Victor, verify or correct this.  I'm not sure, currently asking Mark}.
+Even for some operations that could have been implemented via multiple BLAS
+calls, performance improvements were achieved by combining multiple
+vector-vector operations into a single kernel, due to reduced bandwidth and
+kernel call overhead.
 
@@
 performed (and results uploaded to GPU DRAM) once before GPU computation
-starts. {\bf we should prove justification for A pseudoinverse being cheap,
-either here or perhaps later in a pie chart}
+starts. Since the $A$ is very tall, it is only necessary to invert a matrix
+of size $n \times n$, which executes very quickly.
 

### 2011-05-19 main.tex
added new plot for accuracy vs resolution

@@
 As can be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly diminishing
-returns as the number of kept users is increased beyond ten.  
+returns as the number of kept users is increased beyond 20.  
 \begin{figure}
@@
 \label{sec:pipeline_benchmark}
-The fourth experiment compares the recognition rate of the face recognition algorithm at different
-image resolutions.  We fixed the number of users to be $enter here$ which was shown in figure  $enter here$  have 95\% probability
-to have the true user in the training set, inner loop tolerance to 1e-2 and outer loop tolerance to 1e-3.
+The fourth experiment compares the recognition rate of the face recognition algorithm at different image resolutions.
+We tested 5 square resolutions {32, 48, 64, 96, 128} with no image perturbations.
+We fixed the number of users to 20 which was shown in figure \ref{fig:user_alignment_rank_plot} have approximately 95\% probability to have the true user in the training set, inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.
+
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/accuracyVsResolution}
+\caption{Accuracy vs Image Resolution}
+\label{fig:accuracy_vs_resolution}
+\end{figure}
+
+The results shown in figure \ref{fig:accuracy_vs_resolution} indicate that
+image resolutions of 64x64 provides the best recognition rate for our pipeline.
+One possible reason why the higher resolutions have a lower recognition rate is that the higher resolution images may contain more noise.
 

### 2011-05-20 main.tex

@@
 \label{sec:alignment_implementation_cpu}
-This section discusses how to efficently map the two levels of parallelism
-available in iterative face alignment to the two levels of concurrency
-available in a typical multicore CPU architecture.
 
-We will show that, in constrast with previous implementations, which mapped
-problem level parallelism onto core-level concurrency, it is more efficient to
-completely disregard problem-level parallelism, and instead map pixel-level
-parallelism onto both core-level and sse-level cpu concurrency.  In other
-words, the cpu implementation we advocate performs alignment of the test image
-against the gallery images of a single user at a time.  Each matrix-vector and
-vector-vector operation gets broken into 8 parts that run one per core, and
-into 4 single float vectors that execute on the SSE SIMD units.
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_cpu}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_cpu}
-\caption{Proposed mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_cpu}
-\end{figure}
+Optimal implementation of Algorithm \ref{alg:alignment_alm} on a multicore CPU must take 
+into account the properties of the cache hierarchy. In general, cache that is 
+closer to the core (i.e. L1 cache) has higher bandwidth, but smaller size compared to cache
+that is farther from the core (L3 cache).  For reference, Figure \ref{fig:caches} 
+shows the the sizes of L2 and L3 caches of the Intel E5530 CPU, 
+which is a representative example of a modern multicore CPU.
 
-Recall that the L2 cache can old up to 128 image vectors' worth of data.  Since
-we're only aligning data for one gallery user at a time the working set of the
-problem fits into L2 cache.  Due to the higher bandwidth of L2 cache relative
-to L3 cache, this results in significantly higher image alignment speed than
-previous cpu implementations, as will be seen and discussed further in section
-\ref{sec:alignment_benchmark}.
+For the E5530, the L2 cache in each core is only able to store about 16 images.
+For image alignment problem, the number of training images per class is
+typically larger than 16.  Therefore, if eight instances of
+\eqref{eq:alignment-linearization} were be executed in parallel on the eight
+cores, there would be an insufficient amount of L1 or L2 cache to hold $A$ and
+$A^\dagger$ during the iterative process to execute Algorithm
+\ref{alg:alignment_alm}. Therefore, the bottleneck in this strategy becomes the
+bandwidth between the L2 cache and L3 cache. The implementation of some
+previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this
+mapping strategy.
 
-{\bf Implementation Details} In our implementation, most of the operations
-required by the alignment stage take advantage of Intel Math Kernel Library
-(MKL), a commercial implementation of the standard Basic Linear Algebra
-Subprograms (BLAS).  For image resampling, we make use of Intel's Integrated
-Performance Primitives (IPP) library.  Both of these libraries are heavily
-optimized for Intel multicore CPU's, and are able to make good use of both
-core-level and sse-level concurrency.  For operations that do not map well
-onto existing libraries, we achieve both sse-level and core-level
-concurrency through a combination of the OpenMP API \cite{dagum2002openmp} and
-the automatic vectorization facilities of Intel's C++ compiler
-(ICC)\cite{dulong1999overview}.
+In this paper, we show that for CPU parallelism of face alignment, a higher
+performance implementation can be achieved by {\em not} taking advantage of
+problem-level parallelism at all, and instead map the pixel-level parallelism
+onto both the core-level and the sse-level concurrency provided by the cpu.  In
+other words, the proposed CPU implementation performs alignment of a query
+image against the training images of a single user at a time using all cores on
+both processors.
+This mapping of the problem is illustrated in Figure \ref{fig:alignment_mapping_cpu}.
 
-{\bf Comparison to Previous Implementations} The implementation presented in
-\cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} schedules instances of the
-alignment problem to the cpu cores in a 1:1 fashion.  
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_old}.
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_old}
-\caption{Old mapping of alignment parallelism onto cpu concurrency}
-\label{fig:alignment_mapping_old}
-\end{figure}
-
-This implementation was written using the OpenCV library
-as a wrapper for MKL, and, like our proposed implementation, achieves sse-level concurrency implicitly.  
-With eight alignment problems (associated with 8 training users) being solved
-concurrently, and with 20-40 images per training user, all of the required data
-fits into L3 cache.  While the L3 cache has much higher bandwidth and lower latency
-than off-chip memory, this implementation is slower than the proposed implementation.
+More specifically, in our implementation, most of the operations take
+advantage of Intel Math Kernel Library (MKL), a commercial implementation of
+the standard Basic Linear Algebra Subprograms (BLAS). For the image resampling
+step, we make use of the Intel Integrated Performance Primitives (IPP) library.
+Both MKL and IPP are optimized for Intel multicore CPUs, and are able to
+automatically utilize both core-level and vector-level concurrency. For other
+operations that are not optimized by Intel in-house libraries, we achieve both
+vector-level and core-level concurrency through a combination of the Open MP
+API \cite{dagum2002openmp} and the automatic vectorization facilities of the
+Intel C++ compiler (ICC) \cite{dulong1999overview}. We shall compare the
+performance of the two parallelism methods in Section \ref{sec:experiment}.
 
@@
 \label{sec:alignment_implementation_gpu}
-This section discusses how to efficently map the two levels of parallelism
-available in iterative face alignment to the two levels of concurrency
-available in the NVIDIA Fermi GPU architecture.
+While on the CPU, algorithm performance is highly dependent on effective
+cache usage, on the GPU cache is relatively unimportant for ALM based
+$\ell_1$-min.  As can be seen in Figure \ref{fig:caches}, the GPU has 
+a very small amount of cache compared to the CPU.  While it might be possible
+to fit a single instance of the alignment problem (with a slightly reduced problem size)
+into L2 cache, this would not be an efficient use of the GPU's resources.
+The strength of the GPU's memory architecture for our purposes it's ability to sustain a very
+high bandwidth to DRAM.  A higher overall bandwidth can be achieved if 
+many memory transactions are performed simultaneously, than can be achieved
+with a large number of smaller transactions that are just large enough to fill
+the L2 cache.
 
-Our proposed GPU implementation for the iterative alignment stage consists of a
-single kernel that performs all of the test image alignment problems.  Each of
-instance of the alignment problem is assigned its own thread block, and the GPU
-hardware schedules as many of these thread blocks to run concurrently as the 
-hardware will allow.  
-This mapping of the alignment problem onto the concurrency provided by the pair
-of CPUs is visualized in Figure \ref{fig:alignment_mapping_gpu}.
+Therefore, whereas on the CPU we advocate performing alignment for one training
+class at a time, on the GPU we advocate performing alignment for many classes
+concurrently.  \footnote{In CUDA terminology, Our proposed alignment stage
+implementation consists of a single kernel that performs alignment for all
+subject classes.  Each instance of the alignment problem is assigned its own
+thread block, and the GPU hardware schedules as many thread blocks to run
+concurrently as the hardware will allow.} The number of subject classes that
+actually get aligned concurrently is chosen by the gpu hardware. In our
+experiments the hardware is able to schedule 4-5 classes per SMP to run
+concurrently. This mapping of the problem-level parallelism onto the SMP-level
+concurrency is illustrated in Figure \ref{fig:alignment_mapping_gpu}.
+
 \begin{figure}
@@
 \end{figure}
-The full resolution test image, along with its numerical derivatives in both
+
+In addition, the full resolution query image, along with its numerical derivatives in both
 image directions, are stored in GPU texture memory. This  allows the
-recomputation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
+computation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
 versions of the test image and its derivatives, to exploit the GPU's special
@@
 
-\subsection{Other Implementation Strategies}
-Several GPU implementation strategies were explored during the course of our
-development, and are only mentioned briefly.  Listed in order of increasing
-performance based on a synthetic benchmark of the $\ell_1$ minimization step,
-they included:
-\begin{enumerate}
-\item Solving a single problem instance at a time using Cublas\footnote{Nvidia provided BLAS implementation} (302 ms)
-\item Pure CUDA C implementation solving one problem per SMP using cuda streams\footnote{Streams is a CUDA feature that allows multiple
-kernels to be run concurrently} (70 ms)
-\item Pure CUDA C implementation solving one problem per SMP using one large kernel call (36 ms)
-\end{enumerate}
-
-We attribute the low performance of Cublas in this benchmark to the relatively
-low problem size. Cublas is intended to take advantage of the resources of the
-entire gpu to solve a single large problem, and our problem size is too small
-to hide the memory latency.  Effects of this will be seen again in the
-benchmarks in section \ref{sec:simulation}.
-
-One possible implementation of the iterative alignment stage would be to
-perform the optimization for all instances simultaneously, but to perform each
-linear algebraic operation in a separate gpu kernel call.  The primary
-advantage of this implementation strategy is that every kernel would be
-performing a very simple operation requiring fewer registers per thread.  The
-registers in a GPU are shared between all executing threads thus, barring other
-hardware limitations, the fewer registers each thread requires, the more
-threads can be scheduled to run simultaneously.  The primary downside of this
-approach is that it forces synchronization between all problem instances.  It will
-also have increased kernel launch overhead and potentially less efficient use of
-the GPU's L2 cache.
+In one experiment, we have exploited several possible GPU implementation
+strategies, which include solving single alignment problems one at a time using
+CUBLAS (302\,ms); solving one problem per SMP using CUDA streams (70\,ms); and
+solving one problem per SMP using one large kernel call (36\,ms). The inferior
+performance of CUBLAS may be attributed to the relatively small problem size.
+In general, the libraries provided by Nvidia are optimized for problem sizes
+that are significantly larger than our individual alignment problems.  In
+Section \ref{sec:experiment}, we show the performance of our CPU and GPU
+implementations running on synthetic and real image data.
 
@@
 
-Compared to a typical CPU, a GPU has much higher peak computational bandwidth
-as well as data bandwith to DRAM.  However, since our application is primarily
-of matrix-vector operations and image resampling operations, it has a very low
-arithmetic intensity (i.e. we perform a very small number of arithmetic
-operations for every number that has to be loaded), and thus performance is
-typically constrained by memory (and cache) bandwidth.  At large problem sizes
-the CPU is limited by its relatively low bandwidth to DRAM compared to the gpu,
-while at small problem sizes the GPU is limited by the relatively high latency
-of its access to DRAM.
-
 {\small

### 2011-05-20 main.tex
Some wording fixes in intro

@@
 \cite{WagnerA2011-PAMI}. However, due the high per-class cost
-of the alignment step, the recognition system still fails to achieve \emph{real-time recognition} against datasets of hundreds or thousands of users.
+of the alignment step, the recognition system still fails to achieve real-time recognition against datasets of hundreds or thousands of users.
 In this paper, in addition to accelerating the generic
@@
 
-In this paper we will clearly distinguish between algorithm {\em parallelism},
+In this discussion we will clearly distinguish between algorithm {\em parallelism},
 which is a property of an algorithm, and hardware {\em concurrency}, which is a
@@
 generic ALM $\ell_1$-min algorithm and the corresponding face recognition
-routines on massively parallel CPUs and GPUs.  To this end, a new face
-recognition system for security and access control applications has been fully implemented on a state-of-the-art
-workstation, as shown in Figure \ref{fig:pipeline}. 
+routines on massively parallel CPUs and GPUs.  To this end, a face
+recognition system for security and access control applications has been implemented, 
+and is diagrammed in Figure \ref{fig:pipeline}. This pipeline follows the work of \cite{WagnerA2011-PAMI}, but
+the fast implementations presented in this paper are novel.
 Since the parallelization of any algorithm
@@
 \ref{sec:concurrency}.
-Finally, note that for applications targeting thousands of users or more, it would likely be
-necessary to distribute the work over a network of computers, while this paper
-focuses on the parallelism available in a single (shared memory) machine.
-The C++ source code of our implementations, which target Nvidia's CUDA GPU architecture,
-will be made available for peer evaluation.
+Finally, note that 
+while this paper focuses on the parallelism available in a single (shared memory) machine,
+applications targeting thousands of users or more, will likely require computation be
+distributed over a cluster or network of computers, and this is not addressed here.
+The source code of our implementations, which target Nvidia's CUDA GPU architecture,
+as well as the x86-64 architecture, will be made available for peer evaluation.
 \begin{figure}
@@
 computationally expensive operations such as matrix-matrix multiplication and
-the solution of least-squares problems with varying $A$ matrices. 
-In Section \ref{sec:ALM}, we contend
-that ALM is a better choice for implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
+the solution of least-squares problems with varying $A$
+matrices\footnote{During the face alignment algorithm presented later, The $A$
+matrix changes between alignment iterations but not within the L1 solver}.  In
+Section \ref{sec:ALM}, we contend that ALM is a better choice for
+implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
 category of approximate $\ell_1$-min solutions called \emph{iterative
-shrinkage-thresholding} (IST) methods \cite{WrightS2008,BeckA2009}.
-IST algorithms mainly involve elementary operations such as vector 
-algebra and matrix-vector multiplication. Therefore,
-when the dimension of the problem becomes high, IST-type algorithms are
-particularly suitable for hardware systems with a high degree of concurrency. In
-\cite{YangA2010-ICIP}, the authors further showed that, compared to other
-IST-type algorithms, ALM is able to achieve good estimation accuracy comparable
-to BP and Homotopy, while at the same time significantly improve the speed
-compared to these exact $\ell_1$-min solutions. Due to these facts, in this
-paper, we choose ALM as the core algorithm to implement the complete parallel
-face recognition pipeline.
+shrinkage-thresholding} (IST) methods \cite{WrightS2008,BeckA2009}.  IST
+algorithms mainly involve elementary operations such as vector algebra and
+matrix-vector multiplication. Therefore, when the dimension of the problem
+becomes high, IST-type algorithms are particularly suitable for hardware
+systems with a high degree of concurrency. In \cite{YangA2010-ICIP}, the
+authors showed that ALM is able to significantly improve the solver speed,
+while still achieving estimation accuracy competitive with exact $\ell_1$-min
+solutions, such as BP and Homotopy methods, as well as other IST-type
+algorithms.  Therefore, in this paper, we choose ALM as the core algorithm to
+implement the complete parallel face recognition pipeline.
 

### 2011-05-20 main.tex
removed what used to be figure 5

@@
 single large problem, whereas the the CPU implementation is faster at solving a
-single small problem.  Figure \ref{fig:random_data2} shows the various sizes
-and tolerances at which our GPU implementation begins to run faster than the
-CPU implementation on synthetic data and will provide researchers with a
-starting point of where to begin setting their tolerance levels for the ALM
-algorithm.
+single small problem.  
 \begin{figure}
@@
 \label{fig:random_data}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{results/random_data/size_vs_speed_crossover_ratio_3}
-\end{center}
-\caption{Relation between CPU and GPU runtime vs tolerance level}
-\label{fig:random_data2}
 \end{figure}

### 2011-05-20 figures/user_alignment_rank_histogram.m
fixed a few figures.  they are printed to eps from matlab

@@ -2,7 +2,7 @@ load('sorted_user_ranks.dat');
 
 h = figure;
 [a, x] = hist(sorted_user_ranks, 100);
-a = a / sum(a);
+a = a / sum(a)*100;
 bar(x,a);
 
 b = [a(1)];
@@ -13,12 +13,21 @@ b = b;
 plot(x+1,b);
 
 xlabel('Number of users in reduced training set');
-ylabel('Probability that user lies in reduced training set');
+%ylabel('Probability that user lies in reduced training set');
+ylabel('Probability');
 
-set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
-set (gcf,'Units','Inches');                     %   Using units of inches
-pos=get(gcf,'position');                        %   Save position values
-set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
-set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
+% set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
+% set (gcf,'Units','Inches');                     %   Using units of inches
+% pos=get(gcf,'position');                        %   Save position values
+% set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
+% set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
 
-plot2svg(['user_alignment_rank_plot.svg'], h, 'png');
+set (h,'windowstyle','normal');               %   Window must be undocked for the following
+set (h,'Units','Inches');                     %   Using units of inches
+pos=get(h,'position');                        %   Save position values
+set (h,'position',[pos(1),pos(2),pos(1)+3.75,pos(2)]);       %   and set plot area to 5 by 3.75
+set(h, 'PaperPosition', [0 0 3.75 3]);
+
+
+print(h, 'user_alignment_rank_plot.eps', '-depsc2');
+%plot2svg(['user_alignment_rank_plot.svg'], h, 'png');

### 2011-05-20 results/accuracy/analyze_data.m

@@ -1,0 +1,281 @@
+function [blasAc, cudaAc] = analyze_data(imsize1)
+imsize = 64;
+saveQ = true;
+
+if exist('imsize1', 'var')
+    imsize = imsize1;
+    saveQ = false;
+end
+
+dir  = ['size' num2str(imsize) 'x' num2str(imsize) '/no_perturb/'];
+%dir  = ['size' num2str(imsize) 'x' num2str(imsize) '/1_1_3/'];
+blas = importdata([dir 'blas_recognition_output_keepers_20.txt']);
+cuda = importdata([dir 'cuda_recognition_output_keepers_20.txt']);
+
+
+labels = {'upload time', 'solve time', 'user names', 'threshold', 'tol', ...
+    'tol_int', 'maxIterInner', 'maxIterOuter', 'm', 'n', 'k', ...
+    'solve_nIter', 'imagesPerUser', 'SCI', 'trueUserRank', ...
+    'user_names (in order)', 'user residual values', ...
+    'user names', 'x'};
+
+rows = size(blas.textdata,1);
+true_user   = zeros(rows, 1);
+for i=1:rows
+    true_user(i) = str2double(blas.textdata{i,3});
+end
+
+up_time     = [blas.data(:,1), cuda.data(:,1)];
+solve_time  = [blas.data(:,2), cuda.data(:,2)];
+threshold   = mean([blas.data(:,4), cuda.data(:,4)]);
+tolerance   = mean([blas.data(:,5), cuda.data(:,5)]);
+tol_int     = mean([blas.data(:,6), cuda.data(:,6)]);
+maxIterInner = mean([blas.data(:,7), cuda.data(:,7)]);
+maxIterOuter = mean([blas.data(:,8), cuda.data(:,8)]);
+m           = mean([blas.data(:,9), cuda.data(:,9)]);
+n           = mean([blas.data(:,10), cuda.data(:,10)]);
+k           = mean([blas.data(:,11), cuda.data(:,11)]);
+iter        = [blas.data(:,12), cuda.data(:,12)];
+imagesPerUser = mean([blas.data(:,13), cuda.data(:,13)]);
+SCI         = [blas.data(:,14), cuda.data(:,14)];
+trueUserRank = [blas.data(:,15), cuda.data(:,15)];
+
+start_row = 16;
[…]

### 2011-05-20 results/accuracy/plotAccVsRes.m

@@ -1,0 +1,29 @@
+function plotAccVsRes()
+%%
+res = [32 48 64 80 96 128];
+acc = zeros(numel(res),2);
+
+for i=1:numel(res)
+    resolution = res(i);
+    [b1,c1] = analyze_data(resolution);
+    acc(i,1) = b1;
+    acc(i,2) = c1;
+end
+%%
+close all;
+h = figure;
+plot(res, acc(:,1) * 100 , '--o', res, acc(:,2) * 100, 'x-');
+legend('CPU', 'GPU', 'Location', 'best');
+xlabel('Resolution');
+ylabel('Accuracy %');
+ylim([0 100]);
+
+set (h,'windowstyle','normal');               %   Window must be undocked for the following
+set (h,'Units','Inches');                     %   Using units of inches
+pos=get(h,'position');                        %   Save position values
+set (h,'position',[pos(1),pos(2),pos(1)+3.75,pos(2)]);       %   and set plot area to 5 by 3.75
+set(h, 'PaperPosition', [0 0 3.75 3]);
+print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
+print(h, 'accuracyVsResolution.eps', '-depsc2');
+
+end

### 2011-05-22 main.tex
updated figure 4.  code to change it is in results/random_data/bench_PALM1_paper1.m deleted bad experiments

@@
 entry randomly generated from the standard normal and normalized with unit
-column norm.  The ratio of $d:n$ varied from $1:10$ to $5:10$ with $n$ varying
-from 500 to 8000.  The ground truth signals, $\xx_0$ had a sparsity rate of 0.1
+column norm.  The ratio of $d/n$ =$1/2$ with $n$ varying
+from 1000 to 8000.  The ground truth signals, $\xx_0$ had a sparsity rate of 10\% of d
 with elements generated using the standard normal and also normalized with unit
 norm.  Because the ground truth signal $\xx_0$ was known, the algorithm
-terminated when $\|\bx-\bx_0\| < \tau$.  The measurement vector was generated
+terminated when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement vector was generated
 with the equation $\bb_0 = A \xx_0$.
@@
 \begin{center}
-\includegraphics[width=3.5in]{results/random_data/time_vs_matrix_size_constant_tol}
+\includegraphics[width=3.5in]{figures/time_vs_matrix_size_constant_tol}
 \end{center}

### 2011-05-22 results/random_data/bench_PALM_chart.m

@@ -19,7 +19,7 @@ level_times = [];
 i = 0;
 
 for ratio=2:4
-    name = [type '_' num2str(n) '_' num2str(ratio) '_'];
+    name = [type '_' num2str(ratio) '_' num2str(n) '_'];
     
     gpuName = [name 'GPU.mat'];
     gpuData = load(gpuName);
@@ -43,7 +43,7 @@ for ratio=2:4
             
             count = 0;
             
-            for m=-1:0.5:1
+            for m=-0.5:0.5:0.5
                 count = count + 1;
                 xpoint  = find(squeeze(gpuData.averageRunTime(j,k,:,2,2)) + m/2 > squeeze(blasData.averageRunTime(j,k,:,3,2))==0,1);
                 if isempty(xpoint);
@@ -92,22 +92,24 @@ for ratio=unique(ratios)'
         end
         
         plotX = [plotX; xpoint];
-        plotY = [plotY; unique_tols(j) unique_tols(j) unique_tols(j) unique_tols(j) unique_tols(j)];
+        plotY = [plotY; unique_tols(j) unique_tols(j) unique_tols(j)]; % unique_tols(j) unique_tols(j)];
     end
         
     semilogy(plotX, plotY, '-o');
     set(gca, 'ydir', 'reverse');
+    ylabel(['Outer tolerance level']);
     xlabel('A matrix');
     xlabel(['A matrix (m x ' num2str(ratio/10) 'm)']);
     xlim([min(min(plotX))-500 max(max(plotX))+500]);
-    legend('GPU slower by 1 second','GPU slower by 1/2 second', 'GPU = CPU speed', 'GPU faster by 1/2 second', 'GPU faster by 1 second');
+    %legend('GPU slower by 1 second','GPU slower by 1/2 second', 'GPU = CPU speed', 'GPU faster by 1/2 second', 'GPU faster by 1 second');
+    legend('GPU slower by 1/2 second', 'GPU = CPU speed', 'GPU faster by 1/2 second');
     %axis([]);
     
     %hold off;
     set (h,'windowstyle','normal');               %   Window must be undocked for the following
     set (h,'Units','Inches');                     %   Using units of inches
     pos=get(h,'position');                        %   Save position values
-    set (h,'position',[pos(1),pos(2),12,6]);       %   and set plot area to 5 by 3.75
+    set (h,'position',[pos(1),pos(2),11,6]);       %   and set plot area to 5 by 3.75
[…]

### 2011-05-22 results/random_data/bench_PALM_paper1.m

@@ -15,8 +15,8 @@ tol_ints = [1e-1 1e-2 1e-3 1e-4 1e-5];
 
 h = figure;
 
-for i=2:4
-    name = [type '_' num2str(n) '_' num2str(i) '_'];
+for i=5
+    name = [type '_' num2str(i) '_' num2str(n) '_'];
     
     gpuName = [name 'GPU.mat'];
     
@@ -34,9 +34,9 @@ for i=2:4
     gpuASize       = (gpuDimsPowers.^2 * i / 10 ) * 4 / (2^20);
     blasASize      = (blasDimsPowers.^2 * i / 10 ) * 4 / (2^20);
     
-    subplot(1, 3, i-1);
-    plotA = squeeze(gpuData.averageRunTime(1,5,:,2,2))';
-    plotC = squeeze(blasData.averageRunTime(1,5,:,3,2))';
+    %subplot(1, 3, i-1);
+    plotA = squeeze(gpuData.averageRunTime(3,3,:,2,2))';
+    plotC = squeeze(blasData.averageRunTime(3,3,:,3,2))';
     %plot(gpuASize(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
     plot(gpuDimsPowers(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
     hold on
@@ -44,7 +44,7 @@ for i=2:4
     %title(['Tolerance ' num2str(tols(1)) ', Inner Tolerance ' num2str(tol_ints(5))]);
     ylabel('Elapsed time (s)');
     %axis([0 100 0 30])
-    xlabel(['A matrix (m x ' num2str(i/10) 'm)']);
+    xlabel(['A matrix (' num2str(i/10) 'm x m)']);
     %xlabel(['A matrix dimension (m x 0.' num2str(i) 'm)']);
     hold off
     legend('gpu', 'cpu', 'Location', 'NorthWest');
@@ -54,11 +54,21 @@ end
 
 i = 1;
 
-set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
-set (gcf,'Units','Inches');                     %   Using units of inches
-pos=get(gcf,'position');                        %   Save position values
-set (gcf,'position',[pos(1),pos(2),12,6]);       %   and set plot area to 5 by 3.75
-set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
 
-plot2svg(['time_vs_matrix_size_constant_tol.svg'], h, 'png');
[…]

### 2011-05-23 results/random_data/bench_PALM_paper1.m
added some comments to describe the different indices in the large array

@@ -35,6 +35,12 @@ for i=5
     blasASize      = (blasDimsPowers.^2 * i / 10 ) * 4 / (2^20);
     
     %subplot(1, 3, i-1);
+    % index 1 is outer tolerance
+    % index 2 is inner tolerance
+    % index 3 is different resolutions
+    % index 4 is (2) GPU (3) BLAS
+    % index 5 is (1) time to upload (2) time to solve (3) time to free memory
+
     plotA = squeeze(gpuData.averageRunTime(3,3,:,2,2))';
     plotC = squeeze(blasData.averageRunTime(3,3,:,3,2))';
     %plot(gpuASize(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');

### 2011-05-23 main.tex
Working on cache figures
Added abstract

@@
 \usepackage{algorithmic}
+\usepackage{subfigure}
 
@@
 
-\begin{abstract}
-
-\end{abstract}
+\begin{abstract} Recently a family of promising face recognition algorithms
+that are based on sparse representation and $\ell_1$ minimization have been
+been developed.  These algorithms have not yet seen commercial application,
+largely due to their greater computational requirements compared to more
+traditional algorithms.  This paper studies techniques for leveraging the
+massive parallelism available in GPU and multi-core CPU hardware to accelerate
+augmented Lagrangian method (ALM) based $\ell_1$ minimization solvers.  The
+proposed implementations are significantly faster than than previously
+published implementations.  Despite the low arithmetic intensity of ALM
+solvers, the GPU can still outperform the CPU by a wide margin for large
+problem sizes, and when solving many small problems simultaneously.  The
+proposed CPU implementation remains faster for solving a single moderately
+sized problem.  \end{abstract}
 
@@
 \begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figures/caches}
-\end{center}
-\caption{Scale Drawing of the CPU and GPU cache sizes}
-\label{fig:caches}
+\centering
+\subfigure[The larger algorithm data structures]
+{
+	\includegraphics[scale=1.2]{figures/arrays}
+	\label{fig:arrays}
+}
+\subfigure[The caches on a E5530 CPU]
+{
+	\includegraphics[scale=1.2]{figures/cpu_caches}
+	\label{fig:cpu_caches}
+}
+\subfigure[The caches on a GTX480 GPU]
+{
+	\includegraphics[scale=1.2]{figures/gpu_caches}
+	\label{fig:gpu_caches}
+}
+\caption{Scale drawings for the visual comparison of the algorithm
+working set to available caches}
 \end{figure}

### 2011-05-24 main.tex
Lots of changes.  Especially in experiments section.

@@
 authors showed that ALM is able to significantly improve the solver speed,
-while still achieving estimation accuracy competitive with exact $\ell_1$-min
-solutions, such as BP and Homotopy methods, as well as other IST-type
-algorithms.  Therefore, in this paper, we choose ALM as the core algorithm to
-implement the complete parallel face recognition pipeline.
+while achieving estimation accuracy competitive with exact $\ell_1$-min
+solutions such as BP and Homotopy methods, and other IST-type
+algorithms.  Therefore, we choose ALM as the core algorithm for
+implementation of the parallel face recognition pipeline.
 
@@
 architectures, and defer discussion of their performance implications to
-sections \ref{sec:alignment_implementation_cpu} and
+Sections \ref{sec:alignment_implementation_cpu} and
 \ref{sec:alignment_implementation_gpu}.
@@
 many-core Nvidia GPU chip on a PCI card. GPUs are commonly shipped with a
-single GPU chip per PCI-express card.  Note that most engineering  workstations
-are capable of supporting several GPU cards.  Unless otherwise specified, all
-implementations utilize single precision floating point datatypes.
+single GPU chip per PCI-express card.  Note that while typical engineering
+workstations can only support a single high-end GPU due to their high current
+draw, purpose-built servers can support multiple high-end GPU's.  Unless
+otherwise specified, all implementations utilize single precision floating
+point datatypes.
 
@@
 }
+\label{fig:caches}
 \caption{Scale drawings for the visual comparison of the algorithm
@@
 higher bandwidth between the processor chip and off-chip memory (DRAM) compared
-to the CPU.  A scale drawing of the caches available on the GPU can be seen in Figure \ref{fig:caches}.
+to the CPU.  A scale drawing of the caches available on the GPU can be seen in Figure \ref{fig:gpu_caches}.
 The GPU has its own memory system, and any data the GPU uses must first be
@@
 the SIMD units of each SMP is somewhat more flexible than on the CPU,
-leveraging the flexibility typically comes at the cost of reduced parallelism
+leveraging the flexibility typically comes at the cost of reduced concurrency
 \footnote{In CUDA terms, code branches that cause warp divergence usually
@@
 concurrency and {\em warp-level} concurrency.  Note that the GPU provides more
-concurrency than a cpu at both levels (14 SMP's vs. 8 cores) and (64 wide vs. 4 wide SIMD
+concurrency than a CPU at both levels (14 SMP's vs. 8 cores) and (64 wide vs. 4 wide SIMD
 units).
@@
 \ref{alg:alignment_alm}. In terms of computational complexity, one significant
-term is the computation of the matrix psudo-inverse $B_j^\dagger$, whose
+term is the computation of the matrix pseudo-inverse $B_j^\dagger$, whose
 computational cost is bounded by $O(n_i^2m + n_i^3)$. For per-class face
@@
 
-Optimal implementation of Algorithm \ref{alg:alignment_alm} on a multicore CPU must take 
+Optimal implementation of Algorithm \ref{alg:alignment_alm} on a multi-core CPU must take 
 into account the properties of the cache hierarchy. In general, cache that is 
 closer to the core (i.e. L1 cache) has higher bandwidth, but smaller size compared to cache
-that is farther from the core (L3 cache).  For reference, Figure \ref{fig:caches} 
+that is farther from the core (L3 cache).  For reference, Figure \ref{fig:cpu_caches} 
 shows the the sizes of L2 and L3 caches of the Intel E5530 CPU, 
-which is a representative example of a modern multicore CPU.
+which is a representative example of a modern multi-core CPU.
 
@@
 performance implementation can be achieved by {\em not} taking advantage of
-problem-level parallelism at all, and instead map the pixel-level parallelism
-onto both the core-level and the sse-level concurrency provided by the cpu.  In
+problem-level parallelism at all, and instead mapping the pixel-level parallelism
+onto both the core-level and the sse-level concurrency provided by the CPU.  In
 other words, the proposed CPU implementation performs alignment of a query
@@
 both processors.
-This mapping of the problem is illustrated in Figure \ref{fig:alignment_mapping_cpu}.
 
@@
 step, we make use of the Intel Integrated Performance Primitives (IPP) library.
-Both MKL and IPP are optimized for Intel multicore CPUs, and are able to
+Both MKL and IPP are optimized for Intel multi-core CPUs, and are able to
 automatically utilize both core-level and vector-level concurrency. For other
-operations that are not optimized by Intel in-house libraries, we achieve both
+operations that are not optimized by Intel in-house libraries, such as the shrinkage
+operator in step 7 of Algorithm \ref{alg:alignment_alm}, we achieve both
 vector-level and core-level concurrency through a combination of the Open MP
 API \cite{dagum2002openmp} and the automatic vectorization facilities of the
-Intel C++ compiler (ICC) \cite{dulong1999overview}. We shall compare the
-performance of the two parallelism methods in Section \ref{sec:experiment}.
+Intel C++ compiler (ICC) \cite{dulong1999overview}. 
+%We shall compare the performance of the two parallelism methods in Section \ref{sec:experiment}.
 
@@
 concurrently as the hardware will allow.} The number of subject classes that
-actually get aligned concurrently is chosen by the gpu hardware. In our
+actually get aligned concurrently is chosen by the GPU hardware. In our
 experiments the hardware is able to schedule 4-5 classes per SMP to run
@@
 \includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
-\caption{Proposed mapping of alignment parallelism onto gpu concurrency}
+\caption{Proposed mapping of alignment parallelism onto GPU concurrency}
 \label{fig:alignment_mapping_gpu}
@@
 
-In addition, the full resolution query image, along with its numerical derivatives in both
+The full resolution query image, along with its numerical derivatives in both
 image directions, are stored in GPU texture memory. This  allows the
@@
 
-In one experiment, we have exploited several possible GPU implementation
-strategies, which include solving single alignment problems one at a time using
-CUBLAS (302\,ms); solving one problem per SMP using CUDA streams (70\,ms); and
-solving one problem per SMP using one large kernel call (36\,ms). The inferior
-performance of CUBLAS may be attributed to the relatively small problem size.
-In general, the libraries provided by Nvidia are optimized for problem sizes
-that are significantly larger than our individual alignment problems.  In
-Section \ref{sec:experiment}, we show the performance of our CPU and GPU
+Solving one alignment problem per SMP necessitated the implementation of a
+custom CUDA matrix inversion routine.  This was done using Gaussian elimination
+with partial pivoting.  To achieve precision comparable to the single precision
+LAPACK routines in intel's MKL with this simplistic algorithm, we use double
+precision.  Since the matrix to be inverted is only $n \times n$, the cost of
+the inversion is dwarfed by the cost of other steps.  Similarly, while a
+commercial library might perform reductions on long vectors (i.e. sum or
+$\ell_1$ norm of an image) using a tree structure to improve precision without
+the cost of double precision math, we simply use a double precision temporary
+variable for accumulation.
+
+In Section \ref{sec:experiment}, we show the performance of our CPU and GPU
 implementations running on synthetic and real image data.
@@
 \label{sec:recognition}
-After the face alignment stage is complete, a single recognition algorithm will
-be called to estimate a sparse representation of the query image w.r.t. all the
-training classes, as shown in \eqref{eq:l1min_denoise}. In
-\cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI}, the recognition
-algorithm via $\ell_1$-min has been extensively studied. The algorithm itself
-via ALM is summarized in Algorithm \ref{alg:alm_rec}.  In this section, we
-mainly focus our discussion on its pixel-level parallelism on the multicore
-CPUs and GPUs. Then in Section \ref{sec:experiment}, we benchmark the
-performance of the two architectures and further demonstrate the improvement in
-both accuracy and speed compared to the existing literature.
+After the face alignment stage is complete, a recognition stage consisting of a
+single $\ell_1$ minimization is performed to represent the aligned testing
+image as a sparse combination of aligned training images from the 20 users with lowest alignment
+residual, as shown in \eqref{eq:l1min_denoise}. 
+The coefficients $\x$ are then used to compute error residuals that are
+used for classification.  This recognition algorithm has been extensively studied in 
+\cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI}.
+
+The $ell_1$ minimization is performed via the ALM algorithm shown in 
+Algorithm \ref{alg:alm_rec}.  In this stage of the algorithm, there is no problem-level
+parallelism to exploit, so this section will discuss how to exploit exploit pixel-level parallelism
+on CPU and GPU hardware.
+Then in Section \ref{sec:experiment}, we will benchmark the
+performance of the two architectures and further demonstrate the speed
+gains achieved by our proposed implementations over previously published implementations.
+
 \begin{algorithm}[t]
@@
 
-\subsection{CPU Implementation} Our CPU implementation of Algorithm
-\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
-nicely onto the BLAS API (such as SGEMV, SAXPY), we
-call Intel's MKL BLAS library, which internally takes advantage of the SSE
-units on all available CPU cores.
+\subsection{Recognition Stage Implementation} Compared to problem-level
+parallelism, pixel-level parallelism in Algorithm \ref{alg:alm_rec} is
+relatively easy to exploit using existing BLAS libraries.  For this reason,
+both the CPU and GPU implementation of the recognition stage is very similar to
+the CPU implementation of the $\ell_1$ minimization advocated in Section
+\ref{sec:alignment_implementation_cpu}.  On the CPU, most of the operations map
+well onto MKL BLAS calls, and the operations that do not can be easily
+parallelized using OpenMP and auto-vectorization.
 
-For operations that do not map nicely onto the BLAS API, such as the shrinkage
-operator, {\bf Argue that these parts are optimized, and/or demonstrate that
-performance of these parts is insignificant}.
-
-\subsection{GPU Implementation} Our GPU implementation of Algorithm
-\ref{alg:alm_rec} is similar to the CPU implementation described in the
-previous section, in that only one problem is solved at a time, and each linear
-algebraic operation is parallelized separately.
-
-For most of the operations, calls to MKL CPU BLAS library map directly onto
-similar calls in Nvidia's GPU BLAS library (CUBLAS).  The CUBLAS library is
-designed to utilize all of the SM's on the GPU.
-
-Operations that do not map well onto the BLAS API (mostly combined
-vector-vector operations) were implemented as gpu kernels (in C for CUDA).
-Even for some operations that could have been implemented via multiple BLAS
-calls, performance improvements were achieved by combining multiple
+On the GPU, most of the operations map well onto similar calls in Nvidia's GPU
+BLAS library (CUBLAS) which, like MKL, can take advantage of both levels of
+concurrency available in the hardware architecture.  Operations that do not map
+well onto the CUBLAS API were implemented as GPU kernels (in C for CUDA).
+Additionally, some operations that could have been implemented via multiple
+BLAS calls, performance improvements were achieved by combining multiple
 vector-vector operations into a single kernel, due to reduced bandwidth and
@@
 transferred to GPU DRAM once, and all non-trivial tasks are performed on the
-GPU on data stored in GPU DRAM.  The sole exception to the above implementation
-strategy is that the pseudoinverse of A is still computed on the CPU via the
-LAPACK implementation that is part of Intel's MKL. This operation need only be
-performed (and results uploaded to GPU DRAM) once before GPU computation
-starts. Since the $A$ is very tall, it is only necessary to invert a matrix
-of size $n \times n$, which executes very quickly.
+GPU on data stored in GPU DRAM.  All temporary variables are stored in GPU DRAM
+during solver execution, with the sole exception of the pseudoinverse of A,
+which is computed on the CPU via Intel's LAPACK implementation, and then
+uploaded to GPU DRAM.
 
-\section{Experiments}
-\label{sec:experiment}
-In this section, we benchmark the performance of our parallel implementations
-of $\ell_1$ minimization on CPU and GPU platforms.  In order to show how our
-$\ell_1$ minimization algorithms scale with problem size, in Section
-\ref{sec:simulation} we will begin with benchmarks on synthetic data.  We will
+\section{Experiments} \label{sec:experiment} In this section, we benchmark the
+performance of our parallel implementations of $\ell_1$ minimization on CPU and
+GPU platforms.  In order to show how our $\ell_1$ minimization algorithms scale
+with problem size, in Section \ref{sec:simulation} we will begin with
+benchmarks for the general $\ell_1$-min problem on synthetic data.  We will
 then progress in Sections \ref{sec:alignment_benchmark} and
-\ref{sec:recognition_benchmark} to demonstrating the speed of our
+\ref{sec:recognition_benchmark} to demonstrating the speed and accuracy of our
 implementations as applied to the alignment and recognition stages of the face
@@
 per-board basis:  For CPU implementations, the benchmark makes use of all of
-the cores in as many cpu's are present.  For GPU implementations, the benchmark
+the cores in as many CPU's are present.  For GPU implementations, the benchmark
 makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Unless otherwise specified, all implementations utilize single-precision 
-floating point datatypes.
+chip).  Aside from the exceptions already mentioned in Section
+\ref{sec:alignment_implementation_gpu}, all math operations are performed in
+single precision.
 
@@
 entry randomly generated from the standard normal and normalized with unit
-column norm.  The ratio of $d/n$ =$1/2$ with $n$ varying
-from 1000 to 8000.  The ground truth signals, $\xx_0$ had a sparsity rate of 10\% of d
-with elements generated using the standard normal and also normalized with unit
+column norm.  The ratio of $d/n$ =$1/2$ with $n$ varying from 1000 to 8000.
+The ground truth signals, $\xx_0$ had a sparsity rate of 10\% of d with
+elements generated using the standard normal and also normalized with unit
 norm.  Because the ground truth signal $\xx_0$ was known, the algorithm
-terminated when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement vector was generated
-with the equation $\bb_0 = A \xx_0$.
+terminated when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement
+vector was generated with the equation $\bb_0 = A \xx_0$.
 
@@
 implementation tends to be faster than the CPU implementation at solving a
-single large problem, whereas the the CPU implementation is faster at solving a
+single large problem, whereas the CPU implementation is faster at solving a
 single small problem.  
@@
 alignment stage of the face recognition pipeline described in Algorithm
-\ref{alg:iterative_alignment}.  The iterative alignment seeks a similarity
+\ref{alg:iterative_alignment}.
+
+First, in order to test our theory that the GPU performs best solving many
+$\ell_1$ problems-per concurrently, we benchmarked three implementations of the
+alignment $ell_1$ minimization with $A$ of size $5120 \times 32$, and a fixed
+$50$ inner loop iterations for each of $50$ outer loop iterations. The runtime
+on a GTX480 GPU was averaged over a large number of trials, which were run
+sequentially or concurrently depending on the implementation.
+%The results are shown in Table \ref{tbl:ubench}.
+The results are shown below:
+%\begin{table}[h] 
+\begin{tabular}{|l|c|}
+\hline
+Sequential solver using CUBLAS & 302\,ms \\
+\hline
+One problem per SMP using CUDA streams 
+\footnote{The streams implementation was limited significantly by a 
+cap on the number of concurrent streams in the current version of CUDA}
+& 70\,ms \\
+\hline
+One problem per SMP using single kernel & {\bf 36\,ms} \\
+\hline
+%\label{tbl:ubench}
+\end{tabular}
+%\end{table}
+Our proposed parallelization of the $\ell_1$ minimization used in the alignment
+stage is {\em eight times} faster than an implementation solving a single
+problem at a time using the stock BLAS libraries.
+
+For experiments on face data we use subsets of the CMU Multi-PIE Face Database.
+For gallery images we use frontal images from session 1, which contains 20
+images per user taken under different illuminations. For test images we use
+images from session 2.  The training images are prepared as follows:
+The iterative alignment stage seeks a similarity
 transformation between the coordinate frame of the "camera" resolution test
 image and the coordinate frame of the "canonical" resolution images which the
-images are compared.  We use a $64$ pixel wide square window in the canonical
+images are compared.  We use a $64 \times 64$ pixel window in the canonical
 frame.  The training images are aligned by manually clicking the two outer eye
@@
 that for well aligned images, the canonical window contains exclusively pixels
-from the subject's face.
+from the subject's face.  We assume the face detector is accurate, so for the test
+images we use clicked eye corners for test image alignment initialization\footnote{Previous
+publications have shown that iterative alignment improves recognition accuracy substantially
+even with clicked eye corners}.
 
-For experiments on face data we use subsets of the CMU Multi-PIE Face Database.
-For gallery images we use frontal images from session 1, which contains 20
-images per user taken under different illuminations. For test images we use
-images from session 2.
-
-In Figure \ref{fig:alignment_stage_runtime} we report the time it takes
-to align a single test image against each of the users in the gallery.
-We vary the size of the gallery to show how implementation scales in the
-number of gallery users.
+In Figure \ref{fig:alignment_stage_runtime} we report the time it takes fort
+our proposed CPU and GPU algorithms to align a single test image against each
+of the users in the gallery.  We vary the size of the gallery to show how
+implementation scales in the number of gallery users.
 \begin{figure}
@@
 \end{figure}
+Similar to the results in Figure \ref{fig:random_data},
+there is a crossover point above which the GPU out-performs the CPU.  Since
+the CPU algorithm solves problems sequentially, there is a linear trend.  
+The tiers seen in the GPU profile result from the GPU hardware scheduling alignment
+problems in concurrent batches, but the overall trend is linear as expected.
 
-Previous publications used a hard threshold on the number of subjects kept for the recognition
-stage as a compromise between recognition speed and potential recognition rate.
-In order to better motivate 
-a choice for a threshold of the number of users kept for the recognition stage,
-we varied the number of users kept for the recognition stage and estimated the probability
-of the test user being among the set of training users kept for the recognition stage.
-As can be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly diminishing
-returns as the number of kept users is increased beyond 20.  
+The following experiment fulfills the dual purpose of verifying the accuracy of
+the alignment stage, and motivating a threshold on how many users are kept for
+the recognition stage.  Setting this threshold establishes a compromise between
+recognition speed and potential recognition rate.  To explore the affect of the
+threshold on potential recognition rate, we varied the number of users kept for
+the recognition stage and estimated the probability of the test user being
+among the set of training users kept for the recognition stage.  This curve
+bounds above the achievable recognition rate of the overall pipeline.  As can
+be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly
+diminishing returns as the number of kept users is increased beyond 20; a small
+fraction of the gallery.
 \begin{figure} 
@@
 \includegraphics[width=3.4in]{figures/user_alignment_rank_plot}
-\caption{Probability of the test user being kept for the recognition stage vs. Number of traning
-users kept for the recognition stage}
-\label{fig:user_alignment_rank_plot}
-\end{figure}
+\caption{Probability of the test user being kept for the recognition stage vs.
+Number of training users kept for the recognition stage}
+\label{fig:user_alignment_rank_plot} \end{figure}
 
-\subsection{Face Recognition Stage Benchmarks}
-\label{sec:recognition_benchmark}
-
-\subsection{Face Recognition Pipeline Benchmarks}
-\label{sec:pipeline_benchmark}
-The fourth experiment compares the recognition rate of the face recognition algorithm at different image resolutions.
-We tested 5 square resolutions {32, 48, 64, 96, 128} with no image perturbations.
-We fixed the number of users to 20 which was shown in figure \ref{fig:user_alignment_rank_plot} have approximately 95\% probability to have the true user in the training set, inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.
-
+Finally, we perform an experiment verifying the recognition accuracy of the
+overall pipeline.  Since would expect higher recognition rates but longer
+execution time for higher resolution images, we compare the recognition rate of
+the overall pipeline with the face recognition stage run at 5 different
+canonical image resolutions: $32\times32, 48 \times 48, 64 \times 64, 96 \times
+96, \textrm{and} 128 \times 128$.  As motivated by Figure
+\ref{fig:user_alignment_rank_plot}, we fix the number of gallery users to 20,
+giving approximately 95\% probability to have the true user in the training
+set.  
+%The inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.  
 \begin{figure}
@@
 \end{figure}
-
-The results shown in figure \ref{fig:accuracy_vs_resolution} indicate that
-image resolutions of 64x64 provides the best recognition rate for our pipeline.
-One possible reason why the higher resolutions have a lower recognition rate is that the higher resolution images may contain more noise.
-
-Our final experiment is to compare the speeds of CPU and GPU implementations entire pipeline in Figure \ref{fig:pipeline}.
+As can be seen in Figure \ref{fig:accuracy_vs_resolution}, at the optimal
+resolution, which happens to match the alignment stage resolution, this bound
+is very nearly achieved by the GPU implementation.  For significantly lower
+resolutions, the accuracy drops off significantly.  Of the two main stages of
+the face recognition pipeline, the iterative alignment stage is by far the most
+expensive; the recognition stage requires less than one second.
 
 \section{Conclusion}
+We have shown that our proposed parallelization is {\em eight times} faster
+than an implementation solving a single problem at a time using the stock
+CUBLAS library. Furthermore, the proposed mapping of the parallelism of the
+face alignment problem onto the concurrency provided in modern CPU and GPU
+architectures results in dramatic improvements in recognition speed over
+previously reported implementations of the same pipeline.  Whereas earlier
+implementations required over a second to align a test image against just one
+of the 250 users in Multi-PIE session 1, the proposed CPU implementation can
+perform the same task in only 45 seconds, and the GPU implementation can do it
+in {\bf under 10 seconds}!.  In principle, the parallelization techniques
+presented here make it feasible to implement an access control system for
+hundreds of users on a handful of purpose-build multi-GPU workstations.
+
+%We argue that in addition to vendor supplied libraries
+%that attempt to leverage both levels of concurrency of the GPU, there is a need
+%for standard libraries that operate at the SMP level of concurrency.
 
@@
 \end{document}
+
+% To do list:
+%Add experiment timing for one instance per core on cpu?
+%Unify plot styles
+%Interpret 2 percent accuracy difference
+%Update speed comparisons and diagrams to GTX 580?
+%Better citation for previously reported speed?
+%Figure tweaks from Allen

### 2011-05-25 figures/user_alignment_rank_histogram.m
Made figure styles more uniform

@@ -1,6 +1,6 @@
 load('sorted_user_ranks.dat');
 
-h = figure;
+h = figure(6);
 [a, x] = hist(sorted_user_ranks, 100);
 a = a / sum(a)*100;
 bar(x,a);
@@ -10,24 +10,17 @@ for i=2:numel(a)
     b(i) = a(i) + b(i-1);
 end
 b = b;
-plot(x+1,b);
+plot(x+1,b,'k.-');
 
 xlabel('Number of users in reduced training set');
 %ylabel('Probability that user lies in reduced training set');
 ylabel('Probability');
 
-% set (gcf,'windowstyle','normal');               %   Window must be undocked for the following
-% set (gcf,'Units','Inches');                     %   Using units of inches
-% pos=get(gcf,'position');                        %   Save position values
-% set (gcf,'position',[pos(1),pos(2),12,10]);       %   and set plot area to 5 by 3.75
-% set(gcf,'paperpositionmode','auto');            %   Set output to WYSIWYG for current screen settings
-
 set (h,'windowstyle','normal');               %   Window must be undocked for the following
 set (h,'Units','Inches');                     %   Using units of inches
-pos=get(h,'position');                        %   Save position values
-set (h,'position',[pos(1),pos(2),pos(1)+3.75,pos(2)]);       %   and set plot area to 5 by 3.75
-set(h, 'PaperPosition', [0 0 3.75 3]);
+windowSize = [0 0 3.75 2];
+set(h,'position', windowSize);
+set(h, 'PaperPosition', windowSize);
 
-
-print(h, 'user_alignment_rank_plot.eps', '-depsc2');
+print(h, 'user_alignment_rank_plot.eps', '-deps2');
 %plot2svg(['user_alignment_rank_plot.svg'], h, 'png');

### 2011-05-25 results/accuracy/plotAccVsRes.m

@@ -11,19 +11,19 @@ for i=1:numel(res)
 end
 %%
 close all;
-h = figure;
-plot(res, acc(:,1) * 100 , '--o', res, acc(:,2) * 100, 'x-');
-legend('CPU', 'GPU', 'Location', 'best');
-xlabel('Resolution');
-ylabel('Accuracy %');
+h = figure(7);
+plot( res, acc(:,2) * 100, 'kx-', res, acc(:,1) * 100 , 'ko-');
+legend('GPU', 'CPU', 'Location', 'SouthEast');
+xlabel('Window width (pixels)');
+ylabel('Recognition Rate (%)');
 ylim([0 100]);
 
 set (h,'windowstyle','normal');               %   Window must be undocked for the following
 set (h,'Units','Inches');                     %   Using units of inches
-pos=get(h,'position');                        %   Save position values
-set (h,'position',[pos(1),pos(2),pos(1)+3.75,pos(2)]);       %   and set plot area to 5 by 3.75
-set(h, 'PaperPosition', [0 0 3.75 3]);
+windowSize = [0 0 3.75 2];
+set(h,'position', windowSize);
+set(h, 'PaperPosition', windowSize);
+
 print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
-print(h, 'accuracyVsResolution.eps', '-depsc2');
 
 end

### 2011-05-25 results/random_data/bench_PALM_paper1.m

@@ -13,7 +13,7 @@ ratio = 1:5;
 tols = [1e-1 1e-2 1e-3 1e-4];
 tol_ints = [1e-1 1e-2 1e-3 1e-4 1e-5];
 
-h = figure;
+h = figure(4);
 
 for i=5
     name = [type '_' num2str(i) '_' num2str(n) '_'];
@@ -44,17 +44,16 @@ for i=5
     plotA = squeeze(gpuData.averageRunTime(3,3,:,2,2))';
     plotC = squeeze(blasData.averageRunTime(3,3,:,3,2))';
     %plot(gpuASize(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
-    plot(gpuDimsPowers(1:gpuDimLen), plotA(1:gpuDimLen), 'bx-');
+    plot(gpuDimsPowers(1:gpuDimLen), plotA(1:gpuDimLen), 'kx-');
     hold on
-    plot(blasDimsPowers(1:blasDimLen), plotC(1:blasDimLen), 'g+-');
+    plot(blasDimsPowers(1:blasDimLen), plotC(1:blasDimLen), 'ko-');
     %title(['Tolerance ' num2str(tols(1)) ', Inner Tolerance ' num2str(tol_ints(5))]);
     ylabel('Elapsed time (s)');
     %axis([0 100 0 30])
-    xlabel(['A matrix (' num2str(i/10) 'm x m)']);
+    xlabel(['m (A is size ' num2str(i/10) 'm x m)']);
     %xlabel(['A matrix dimension (m x 0.' num2str(i) 'm)']);
     hold off
-    legend('gpu', 'cpu', 'Location', 'NorthWest');
-    
+    legend('GPU', 'CPU', 'Location', 'NorthWest');
     
 end
 
@@ -64,11 +63,15 @@ i = 1;
 set (h,'windowstyle','normal');               %   Window must be undocked for the following
 set (h,'Units','Inches');                     %   Using units of inches
 pos=get(h,'position');                        %   Save position values
-set (h,'position',[pos(1),pos(2),pos(1)+3.75,pos(2)]);       %   and set plot area to 5 by 3.75
-set(h, 'PaperPosition', [0 0 3.75 3]);
+windowSize = [0 0 3.75 2];
+%set(h,'position', [pos(1),pos(2),pos(1),pos(2)]+windowSize);
+set(h,'position', windowSize);
[…]

### 2011-05-27 main.tex
Fixed a caption
Most of Allen's changes applied

@@
 
-\def\cvprPaperID{257} % *** Enter the CVPR Paper ID here
+\def\cvprPaperID{184} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 
-\begin{abstract} Recently a family of promising face recognition algorithms
-that are based on sparse representation and $\ell_1$ minimization have been
-been developed.  These algorithms have not yet seen commercial application,
-largely due to their greater computational requirements compared to more
-traditional algorithms.  This paper studies techniques for leveraging the
-massive parallelism available in GPU and multi-core CPU hardware to accelerate
-augmented Lagrangian method (ALM) based $\ell_1$ minimization solvers.  The
-proposed implementations are significantly faster than than previously
-published implementations.  Despite the low arithmetic intensity of ALM
-solvers, the GPU can still outperform the CPU by a wide margin for large
-problem sizes, and when solving many small problems simultaneously.  The
-proposed CPU implementation remains faster for solving a single moderately
-sized problem.  \end{abstract}
+\begin{abstract} 
+Recently a family of promising face recognition algorithms that are based on sparse representation and $\ell_1$ minimization have been been developed.
+These algorithms have not yet seen commercial application, largely due to their greater computational requirements compared to more traditional algorithms.
+This paper studies techniques for leveraging the massive parallelism available in GPU and multi-core CPU hardware to accelerate augmented Lagrangian method (ALM) based $\ell_1$ minimization solvers.
+The proposed implementations are significantly faster than than previously published implementations.
+Despite the low arithmetic intensity of ALM solvers, the GPU can still outperform the CPU by a wide margin for large problem sizes, and when solving many small problems simultaneously.
+The proposed CPU implementation remains faster for solving a single moderately sized problem.
+\end{abstract}
 
 \section{Introduction} 
-$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
+$\ell_1$ minimization ($\ell_1$-min) has received much attention in recent
 years due to important applications in compressive sensing
@@
 For the completeness of the paper, in this section, we briefly describe the ALM
-algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze
-its complexity. Parallelization of ALM for face alignment and recognition will
+algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP}
+ %and analyze its complexity
+. 
+Parallelization of ALM for face alignment and recognition will
 be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition},
@@
 architectures considered in this paper, as well as other aspects of the
-hardware that are important for performance.  In particular, since on-chip
-memory (cache) is often orders of magnitude faster than off-chip memory, the
-sizes and speeds of on-chip memory (cache) have big performance implications.
-We give a brief overview of the caches that are available in our target
-architectures, and defer discussion of their performance implications to
-Sections \ref{sec:alignment_implementation_cpu} and
+hardware that are important for performance.  In particular, since caches
+(regions of on-chip memory) are often orders of magnitude faster than off-chip
+memory, their sizes and speeds have big performance implications.  We give a
+brief overview of the caches that are available in our target architectures,
+and defer discussion of their performance implications to Sections
+\ref{sec:alignment_implementation_cpu} and
 \ref{sec:alignment_implementation_gpu}.
@@
 on one motherboard, and for GPU implementation we make use of a single
-many-core Nvidia GPU chip on a PCI card. GPUs are commonly shipped with a
-single GPU chip per PCI-express card.  Note that while typical engineering
-workstations can only support a single high-end GPU due to their high current
-draw, purpose-built servers can support multiple high-end GPU's.  Unless
-otherwise specified, all implementations utilize single precision floating
-point datatypes.
+many-core Nvidia GPU chip on a PCI card. 
+%GPUs are commonly shipped with a
+%single GPU chip per PCI-express card.  Note that while typical engineering
+%workstations can only support a single high-end GPU due to their high current
+%draw, purpose-built servers can support multiple high-end GPU's.  
+Unless otherwise specified, all implementations utilize single precision
+floating point datatypes.
 
@@
 are that they have two levels of concurrency, relatively large amounts of
-on-chip cache, and relatively high clock speeds.
+cache, and relatively high clock speeds.
 
@@
 \label{fig:caches}
-\caption{Scale drawings for the visual comparison of the algorithm
-working set to available caches}
+\caption{A visual comparison of the algorithm working set to the CPU and GPU
+caches.  Note: Although the aspect ratio could not be preserved, all arrays and
+caches are drawn so that the area is proportional to the size of the data. }
 \end{figure}
@@
 %... so can fit 2**7  = 128 images.
-Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache.
+Each core has a private 32\,KiB L1 data cache and a private 256\,KiB L2 cache.
 %16MB = 2**24 bytes = 16777216 bytes
@@
 %... so can fit 2**10  = 1024 images.
-Each processor further has 8\,MB of L3 cache that is shared by the four cores.
-Overall, the algorithm has approximately 16\,MB of L3 cache
+Each processor further has 8\,MiB of L3 cache that is shared by the four cores.
+Overall, the algorithm has approximately 16\,MiB of L3 cache
 available for a dual-processor configuration.
@@
 (SSE) capable of performing the same arithmetic operation on four single-precision 
-floating point values simultaneously \footnote{By the publication date, the vector
-units of typical high-end CPUS will double.}.
-
+floating point values simultaneously.
 There are thus two important levels of concurrency that need to be exploited to
@@
 are that they have two (much wider) levels of concurrency, relatively small amounts of
-on-chip cache, and relatively low clock speeds.  Whereas most of the transistors on
+ cache, and relatively low clock speeds.  Whereas most of the transistors on
 a typical CPU are dedicated either to cache or to hardware that enables higher
@@
 and software managed ("shared") memory.  Additionally, all SMPs share a common
-L2-cache.  For our system, each SMP has 64 KB of L1 cache, and all SMP's share
+L2-cache.  For our system, each SMP has 64 KiB of L1 cache, and all SMP's share
 768KB of L2 cache. 
@@
 provides two levels of concurrency: {\em SMP-level}
-concurrency and {\em warp-level} concurrency.  Note that the GPU provides more
+concurrency and {\em thread-level} concurrency.  Note that the GPU provides more
 concurrency than a CPU at both levels (14 SMP's vs. 8 cores) and (64 wide vs. 4 wide SIMD
@@
 
-Given the hardware concurrency in modern parallel computing architectures, next,
-we shall study the optimal mappings of the algorithm parallelism in face recognition to the multi-core CPUs and GPUs, respectively.
-In this section, we first focus on the face alignment stage (see Figure \ref{fig:pipeline}).
+Given the hardware concurrency in modern parallel computing architectures,
+next, we shall study the optimal mappings of the algorithm parallelism in face
+recognition to the multi-core CPUs and GPUs, respectively.  In this section, we
+first focus on the face alignment stage (see Figure \ref{fig:pipeline}).
 
-Face alignment \eqref{eq:l1min_alignment} estimates an image transformation $\tau$ that rectifies the query image $\tilde{\bb}$ with possible pose variation
-w.r.t. each training class $A_i$, which leads to the minimal sparsity in error $\ee$ after the alignment. Note that directly solving \eqref{eq:l1min_alignment} would require
-simultaneously estimating $\xx$, $\ee$, and $\tau$ typically on a nonlinear manifold, which renders the problem extremely difficult to solve exactly with many local minima.
+Face alignment \eqref{eq:l1min_alignment} estimates an image transformation
+$\tau$ that rectifies the query image $\tilde{\bb}$ with possible pose
+variation w.r.t. each training class $A_i$, which leads to the minimal sparsity
+in error $\ee$ after the alignment. Note that directly solving
+\eqref{eq:l1min_alignment} would require simultaneously estimating $\xx$,
+$\ee$, and $\tau$ typically on a nonlinear manifold, which renders the problem
+extremely difficult to solve exactly with many local minima.
 
-To mitigate the problem, if the system can obtain to a good initial guess of the transformation $\tau$ (\eg, provided by a good face detector),
-the optimal solution for $\tau$ can be sought iteratively by a sequence of linearized approximations $j=1, 2, \cdots$ as follows:
+To mitigate the problem, if the system can obtain a good initial guess of the
+transformation $\tau$ (\eg, provided by a good face detector), the optimal
+solution for $\tau$ can be sought iteratively by linearization at each $j$th
+step as follows:
 \begin{equation}
@@
 %once per test image.  
-
 The per-class alignment algorithm via ALM is summarized in Algorithm
-\ref{alg:alignment_alm}. In terms of computational complexity, one significant
-term is the computation of the matrix pseudo-inverse $B_j^\dagger$, whose
-computational cost is bounded by $O(n_i^2m + n_i^3)$. For per-class face
-alignment, since both $A_i$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$
-for all $i=1, \cdots, K$ classes), we find that the computation time is still
-dominated by the iterative computation of matrix-vector multiplications in the
-inner loop.
+\ref{alg:alignment_alm}. 
+%In terms of computational complexity, one significant
+%term is the computation of the matrix pseudo-inverse $B_j^\dagger$, whose
+%computational cost is bounded by $O(n_i^2m + n_i^3)$. However, for per-class face
+%alignment, since both $B_j$ and $J_{j}$ are very tall matrices (\ie, $m\gg n_i$
+%for all $i=1, \cdots, K$ classes), we find that the computation cost is still
+%dominated by the frequently computed matrix-vector multiplications in the
+%inner loop.
 \begin{algorithm}[ht!]
@@
 \end{algorithmic}
-{\bf Output:} $\tau^*\leftarrow \tau_j, \e^*\leftarrow \ee_j$
+{\bf Output:} $\tau_i^*\leftarrow \tau_j, \e_i^*\leftarrow \ee_j$
 \end{algorithm}
@@
 \eqref{eq:alignment-linearization} were be executed in parallel on the eight
-cores, there would be an insufficient amount of L1 or L2 cache to hold $A$ and
-$A^\dagger$ during the iterative process to execute Algorithm
+cores, there would be an insufficient amount of L1 or L2 cache to hold $B_j$ and
+$B_j^\dagger$ during the iterative process to execute Algorithm
 \ref{alg:alignment_alm}. Therefore, the bottleneck in this strategy becomes the
@@
 problem-level parallelism at all, and instead mapping the pixel-level parallelism
-onto both the core-level and the sse-level concurrency provided by the CPU.  In
+onto both the core-level and the SSE-level concurrency provided by the CPU.  In
 other words, the proposed CPU implementation performs alignment of a query
@@
 Solving one alignment problem per SMP necessitated the implementation of a
-custom CUDA matrix inversion routine.  This was done using Gaussian elimination
-with partial pivoting.  To achieve precision comparable to the single precision
-LAPACK routines in intel's MKL with this simplistic algorithm, we use double
-precision.  Since the matrix to be inverted is only $n \times n$, the cost of
-the inversion is dwarfed by the cost of other steps.  Similarly, while a
-commercial library might perform reductions on long vectors (i.e. sum or
-$\ell_1$ norm of an image) using a tree structure to improve precision without
-the cost of double precision math, we simply use a double precision temporary
-variable for accumulation.
-
-In Section \ref{sec:experiment}, we show the performance of our CPU and GPU
-implementations running on synthetic and real image data.
+custom CUDA routine for computing $B_j^\dagger = (B_j^TB_j)^-1 B^T = G^-1 B_j^T$. $G$
+is inverted using Gaussian elimination with partial pivoting.  To achieve
+precision comparable to the single precision LAPACK routines in Intel's MKL
+with this simplistic algorithm, we use double precision.  Since $G$ is only $n
+\times n$ and $B_j^\dagger$ is only computed once per optimization problem, the
+cost of the inversion is dwarfed by the cost of other steps.  
+Additionally, reductions on long vectors (i.e. sum or $\ell_1$ norm
+of an image) are performed using a double precision temporary variable for accumulation.
+%In Section \ref{sec:experiment}, we show the performance of our CPU and GPU
+%implementations running on synthetic and real image data.
 
@@
 \label{sec:recognition}
-After the face alignment stage is complete, a recognition stage consisting of a
-single $\ell_1$ minimization is performed to represent the aligned testing
-image as a sparse combination of aligned training images from the 20 users with lowest alignment
-residual, as shown in \eqref{eq:l1min_denoise}. 
-The coefficients $\x$ are then used to compute error residuals that are
-used for classification.  This recognition algorithm has been extensively studied in 
-\cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI}.
-
-The $ell_1$ minimization is performed via the ALM algorithm shown in 
-Algorithm \ref{alg:alm_rec}.  In this stage of the algorithm, there is no problem-level
-parallelism to exploit, so this section will discuss how to exploit exploit pixel-level parallelism
-on CPU and GPU hardware.
+After the face alignment stage is complete, all but the 20 users with lowest
+$||\e_i||_1$ are thrown out, $\bb$ and $A_i$ are re-sampled to a common alignment
+using $\tau_i$, and $A_i$ are concatenated into a new matrix $A$.
+A sparse representation of $\bb$ w.r.t $A$ is then recovered
+by solving a single $\ell_1$-min problem, as shown
+in \eqref{eq:l1min_denoise}, using Algorithm \ref{alg:alm_rec}.  The
+coefficients $\x$ are then used to compute error residuals that are used for
+classification.  
+%This recognition algorithm has
+%been extensively studied in
+%\cite{WrightJ2009-PAMI,YangA2010-ICIP,WagnerA2011-PAMI}.
+In this stage of the algorithm, there is no problem-level
+parallelism to exploit, so this section will discuss how to exploit pixel-level parallelism
+on both CPU and GPU hardware.
 Then in Section \ref{sec:experiment}, we will benchmark the
@@
 both the CPU and GPU implementation of the recognition stage is very similar to
-the CPU implementation of the $\ell_1$ minimization advocated in Section
+the CPU implementation of the $\ell_1$-min advocated in Section
 \ref{sec:alignment_implementation_cpu}.  On the CPU, most of the operations map
@@
 GPU on data stored in GPU DRAM.  All temporary variables are stored in GPU DRAM
-during solver execution, with the sole exception of the pseudoinverse of A,
+during solver execution, with the sole exception of the pseudo-inverse of $B_j$,
 which is computed on the CPU via Intel's LAPACK implementation, and then
@@
 \section{Experiments} \label{sec:experiment} In this section, we benchmark the
-performance of our parallel implementations of $\ell_1$ minimization on CPU and
-GPU platforms.  In order to show how our $\ell_1$ minimization algorithms scale
-with problem size, in Section \ref{sec:simulation} we will begin with
-benchmarks for the general $\ell_1$-min problem on synthetic data.  We will
-then progress in Sections \ref{sec:alignment_benchmark} and
-\ref{sec:recognition_benchmark} to demonstrating the speed and accuracy of our
+performance of our parallel implementations of $\ell_1$-min on CPU and GPU
+platforms.  In order to show how our $\ell_1$-min algorithms scale with problem
+size, in Section \ref{sec:simulation} we will begin with benchmarks for the
+general $\ell_1$-min problem on synthetic data.  We will then progress in
+Section \ref{sec:benchmark} to demonstrating the speed and accuracy of our
 implementations as applied to the alignment and recognition stages of the face
@@
 generic random matrix $A$ and a random sparse vector $\xx_0$.  The measurement
-$d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each
+$d \times n$ matrix $A$ is generated as a random Gaussian matrix, with each
 entry randomly generated from the standard normal and normalized with unit
-column norm.  The ratio of $d/n$ =$1/2$ with $n$ varying from 1000 to 8000.
-The ground truth signals, $\xx_0$ had a sparsity rate of 10\% of d with
-elements generated using the standard normal and also normalized with unit
-norm.  Because the ground truth signal $\xx_0$ was known, the algorithm
-terminated when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement
-vector was generated with the equation $\bb_0 = A \xx_0$.
+column norm.  The ratio of $d/n$ is fixed at $1/2$ with $n$ varying from 1000 to 8000.
+The ground truth signals, $\xx_0$ have a sparsity rate of 10\% of d with
+elements sampled from a normal distribution and normalized to unit $\ell_2$ vector
+norm.  Because the ground truth signal $\xx_0$ is known, the algorithm
+terminats when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement
+vector is generated with the equation $\bb_0 = A \xx_0$.
 
@@
 \end{center}
-\caption{Speed vs Dimensions of $A$.}
+\caption{Runtime vs. Dimensions of $A$.}
 \label{fig:random_data}
@@
 \subsection{Face Alignment Stage Benchmark} 
-\label{sec:alignment_benchmark}
+\label{sec:benchmark}
 This section presents benchmarks of our CPU and GPU implementations of the
-alignment stage of the face recognition pipeline described in Algorithm
-\ref{alg:iterative_alignment}.
+alignment stage of the face recognition pipeline, as shown in Algorithm
+\ref{alg:iterative_alignment}, as well as the recognition stage of the
+face recognition pipeline, as seen in Algorithm \ref{alg:alm_rec}.
 
 First, in order to test our theory that the GPU performs best solving many
-$\ell_1$ problems-per concurrently, we benchmarked three implementations of the
-alignment $ell_1$ minimization with $A$ of size $5120 \times 32$, and a fixed
+$\ell_1$ problems-per concurrently, we benchmark three implementations of the
+alignment $ell_1$-min with $A$ of size $5120 \times 32$, and a fixed
 $50$ inner loop iterations for each of $50$ outer loop iterations. The runtime
-on a GTX480 GPU was averaged over a large number of trials, which were run
+on a GTX480 GPU is averaged over a large number of trials, which are run
 sequentially or concurrently depending on the implementation.
-%The results are shown in Table \ref{tbl:ubench}.
-The results are shown below:
-%\begin{table}[h] 
+The results are shown in Table \ref{tbl:ubench}.
+%The results are shown below:
+\begin{table}[h] 
 \begin{tabular}{|l|c|}
@@
 One problem per SMP using CUDA streams 
-\footnote{The streams implementation was limited significantly by a 
+\footnote{The streams implementation is limited significantly by a 
 cap on the number of concurrent streams in the current version of CUDA}
@@
 One problem per SMP using single kernel & {\bf 36\,ms} \\
-\hline
-%\label{tbl:ubench}
+%\hline
+\label{tbl:ubench}
 \end{tabular}
-%\end{table}
-Our proposed parallelization of the $\ell_1$ minimization used in the alignment
+\end{table}
+Our proposed parallelization of the $\ell_1$-min used in the alignment
 stage is {\em eight times} faster than an implementation solving a single
@@
 recognition speed and potential recognition rate.  To explore the affect of the
-threshold on potential recognition rate, we varied the number of users kept for
-the recognition stage and estimated the probability of the test user being
+threshold on potential recognition rate, we vary the number of users kept for
+the recognition stage and estimate the probability of the test user being
 among the set of training users kept for the recognition stage.  This curve

### 2011-05-27 main.tex
Changes while on the phone with Allen

@@
 the solution of least-squares problems with varying $A$
-matrices\footnote{During the face alignment algorithm presented later, The $A$
-matrix changes between alignment iterations but not within the L1 solver}.  In
+matrices.\footnote{During the face alignment algorithm presented later, The $A$
+matrix changes between alignment iterations but not within the L1 solver}  In
 Section \ref{sec:ALM}, we contend that ALM is a better choice for
@@
 whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
-Algorithm \ref{alg:alm} summarizes the generic ALM $\ell_1$-min algorithm.\footnote{For conciseness, we
+Algorithm \ref{alg:alm} summarizes the generic ALM $\ell_1$-min algorithm. \footnote{For conciseness, we
 only present the ALM algorithm in the primal domain. There also
@@
 more space than is available here, so we will instead frame our discussion in terms of 
-hardware capabilities\footnote{A discussion of the CUDA programming model can be found
-in the NVIDIA C for CUDA Programming Guide}.
+hardware capabilities. \footnote{A discussion of the CUDA programming model can be found
+in the NVIDIA C for CUDA Programming Guide}
 
@@
 floating point operations
-concurrently\footnote{In CUDA terms, the warp width is 32, and the floating pipeline can
-issue up to two warps simultaneously}.
+concurrently. \footnote{In CUDA terms, the warp width is 32, and the floating pipeline can
+issue up to two warps simultaneously}
 
@@
 leveraging the flexibility typically comes at the cost of reduced concurrency
-\footnote{In CUDA terms, code branches that cause warp divergence usually
-result in serialization of the different groups of threads}.  Therefore, for
+.\footnote{In CUDA terms, code branches that cause warp divergence usually
+result in serialization of the different groups of threads}  Therefore, for
 the purposes of comparing hardware architectures, the SIMD units on the GPU are
@@
 The per-class alignment algorithm via ALM is summarized in Algorithm
-\ref{alg:alignment_alm}. 
+\ref{alg:iterative_alignment}. 
 %In terms of computational complexity, one significant
@@
 \begin{algorithm}[ht!]
-\caption{\bf (Face Alignment via ALM)} \label{alg:alignment_alm}
+\caption{\bf (Face Alignment via ALM)} \label{alg:iterative_alignment}
 {\bf Input:} $\bb$, $A_i$, $\x_0 = \mathbf{0}$, $\tau_0$, and $J_0$.
@@
 \STATE $\zz_l \leftarrow \mbox{shrink}\left(\bb_j - B_j\uu_{l-1} + \frac{\yy_{k}}{\mu_{k-1}}, \frac{1}{\mu_{k-1}}\right)$
-\STATE $\uu_l \leftarrow (B_j^\dagger)^T \left(\bb_j - \zz_{l} + \frac{\yy_{k-1}}{\mu_{k-1}} \right) $
+\STATE $\uu_l \leftarrow B_j^\dagger \left(\bb_j - \zz_{l} + \frac{\yy_{k-1}}{\mu_{k-1}} \right) $
 \ENDWHILE
@@
 
-Optimal implementation of Algorithm \ref{alg:alignment_alm} on a multi-core CPU must take 
+Optimal implementation of Algorithm \ref{alg:iterative_alignment} on a multi-core CPU must take 
 into account the properties of the cache hierarchy. In general, cache that is 
@@
 $B_j^\dagger$ during the iterative process to execute Algorithm
-\ref{alg:alignment_alm}. Therefore, the bottleneck in this strategy becomes the
+\ref{alg:iterative_alignment}. Therefore, the bottleneck in this strategy becomes the
 bandwidth between the L2 cache and L3 cache. The implementation of some
@@
 operations that are not optimized by Intel in-house libraries, such as the shrinkage
-operator in step 7 of Algorithm \ref{alg:alignment_alm}, we achieve both
+operator in step 7 of Algorithm \ref{alg:iterative_alignment}, we achieve both
 vector-level and core-level concurrency through a combination of the Open MP
@@
 Solving one alignment problem per SMP necessitated the implementation of a
-custom CUDA routine for computing $B_j^\dagger = (B_j^TB_j)^-1 B^T = G^-1 B_j^T$. $G$
+custom CUDA routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$. $G$
 is inverted using Gaussian elimination with partial pivoting.  To achieve
@@
 transferred to GPU DRAM once, and all non-trivial tasks are performed on the
-GPU on data stored in GPU DRAM.  All temporary variables are stored in GPU DRAM
-during solver execution, with the sole exception of the pseudo-inverse of $B_j$,
-which is computed on the CPU via Intel's LAPACK implementation, and then
-uploaded to GPU DRAM.
+GPU on data stored in GPU DRAM.  
+All non-scalar temporary variables are stored in GPU DRAM
+during solver execution. 
+%with the sole exception of the pseudo-inverse of $B_j$,
+%which is computed on the CPU via Intel's LAPACK implementation, and then
+%uploaded to GPU DRAM.
 
@@
 \end{center}
-\caption{Runtime vs. Dimensions of $A$.}
+\caption{Comparison of $\ell_1$-min runtime vs. dimensions of $A$ on random data.}
 \label{fig:random_data}
@@
 
-\subsection{Face Alignment Stage Benchmark} 
+\subsection{Face Recognition Pipeline Benchmark} 
 \label{sec:benchmark}
@@
 $\ell_1$ problems-per concurrently, we benchmark three implementations of the
-alignment $ell_1$-min with $A$ of size $5120 \times 32$, and a fixed
+alignment $\ell_1$-min with $A$ of size $5120 \times 32$, and a fixed
 $50$ inner loop iterations for each of $50$ outer loop iterations. The runtime
@@
 The results are shown in Table \ref{tbl:ubench}.
+Our proposed parallelization of the $\ell_1$-min used in the alignment
+stage is {\em eight times} faster than an implementation solving a single
+problem at a time using the stock BLAS libraries.
 %The results are shown below:
-\begin{table}[h] 
+\begin{table}[h!]
+\caption{Benchmark of implementations of alignment $\ell_1$-min.}
 \begin{tabular}{|l|c|}
@@
 \hline
-One problem per SMP using CUDA streams 
-\footnote{The streams implementation is limited significantly by a 
-cap on the number of concurrent streams in the current version of CUDA}
-& 70\,ms \\
+One problem per SMP using CUDA streams & 70\,ms  \\
 \hline
 One problem per SMP using single kernel & {\bf 36\,ms} \\
-%\hline
+\hline
+\end{tabular}
 \label{tbl:ubench}
-\end{tabular}
 \end{table}
-Our proposed parallelization of the $\ell_1$-min used in the alignment
-stage is {\em eight times} faster than an implementation solving a single
-problem at a time using the stock BLAS libraries.
+%\footnote{The streams implementation is limited significantly by a 
+%cap on the number of concurrent streams in the current version of CUDA}
 
@@
 from the subject's face.  We assume the face detector is accurate, so for the test
-images we use clicked eye corners for test image alignment initialization\footnote{Previous
+images we use clicked eye corners for test image alignment initialization.\footnote{Previous
 publications have shown that iterative alignment improves recognition accuracy substantially
-even with clicked eye corners}.
+even with clicked eye corners}
 
@@
 \includegraphics[width=3.4in]{figures/alignment_runtime_graph}
-\caption{Alignment Stage Runtime vs. Size of Training Database}
+\caption{Alignment stage runtime vs. size of training database.}
 \label{fig:alignment_stage_runtime}
@@
 \caption{Probability of the test user being kept for the recognition stage vs.
-Number of training users kept for the recognition stage}
+number of training users kept for the recognition stage.}
 \label{fig:user_alignment_rank_plot} \end{figure}
@@
 \includegraphics[width=3.4in]{figures/accuracyVsResolution}
-\caption{Accuracy vs Image Resolution}
+\caption{Recognition rate of the full pipeline vs. face window resolution.}
 \label{fig:accuracy_vs_resolution}

### 2011-05-28 main.tex
Finished first pass through Allen's notes. Still have critical TODO items

@@
 $\ell_1$ problems-per concurrently, we benchmark three implementations of the
-alignment $\ell_1$-min with $A$ of size $5120 \times 32$, and a fixed
+alignment $\ell_1$-min solver with $A$ of size $5120 \times 32$, and a fixed
 $50$ inner loop iterations for each of $50$ outer loop iterations. The runtime
@@
 problem at a time using the stock BLAS libraries.
+\footnote{Note that the streams implementation is limited significantly by a 
+cap on the number of concurrent streams in the current version of CUDA}
 %The results are shown below:
@@
 \hline
-One problem per SMP using single kernel & {\bf 36\,ms} \\
+Four problems per SMP using single kernel & {\bf 36\,ms} \\
 \hline
@@
 \end{table}
-%\footnote{The streams implementation is limited significantly by a 
-%cap on the number of concurrent streams in the current version of CUDA}
 
-For experiments on face data we use subsets of the CMU Multi-PIE Face Database.
-For gallery images we use frontal images from session 1, which contains 20
-images per user taken under different illuminations. For test images we use
-images from session 2.  The training images are prepared as follows:
-The iterative alignment stage seeks a similarity
-transformation between the coordinate frame of the "camera" resolution test
-image and the coordinate frame of the "canonical" resolution images which the
-images are compared.  We use a $64 \times 64$ pixel window in the canonical
-frame.  The training images are aligned by manually clicking the two outer eye
-corners in the camera frame, and then warping the images via a similarity
-transformation such that the eye corners lie at a pair of canonical eye
-locations in the canonical frame.  The canonical eye corners are chosen such
-that for well aligned images, the canonical window contains exclusively pixels
-from the subject's face.  We assume the face detector is accurate, so for the test
-images we use clicked eye corners for test image alignment initialization.\footnote{Previous
-publications have shown that iterative alignment improves recognition accuracy substantially
-even with clicked eye corners}
+Using an implementation motivated by the previous result, we now benchmark our
+iterative alignment implementation on real face data.  For experiments on face
+data we use subsets of the CMU Multi-PIE Face Database.  For gallery images we
+use frontal images from session 1, which contains 20 images per user taken
+under different illuminations. For test images we use images from session 2.
+The training images are prepared as follows: The iterative alignment stage
+seeks a similarity transformation between the coordinate frame of the "camera"
+resolution test image and the coordinate frame of the "canonical" resolution
+images which the images are compared.  We use a $64 \times 64$ pixel window in
+the canonical frame.  The training images are aligned by manually clicking the
+two outer eye corners in the camera frame, and then warping the images via a
+similarity transformation such that the eye corners lie at a pair of canonical
+eye locations in the canonical frame.  The canonical eye corners are chosen
+such that for well aligned images, the canonical window contains exclusively
+pixels from the subject's face.  We assume the face detector is accurate, so
+for the test images we use clicked eye corners for test image alignment
+initialization.\footnote{Previous publications have shown that iterative
+alignment improves recognition accuracy substantially even with clicked eye
+corners}
 
-In Figure \ref{fig:alignment_stage_runtime} we report the time it takes fort
-our proposed CPU and GPU algorithms to align a single test image against each
-of the users in the gallery.  We vary the size of the gallery to show how
-implementation scales in the number of gallery users.
+Figure \ref{fig:alignment_stage_runtime} shows the average run time of the CPU and GPU implementations to align
+a query image against all the subject classes separately. We vary the number of
+subject classes to show how the algorithms scale. 
 \begin{figure}
@@
 Finally, we perform an experiment verifying the recognition accuracy of the
-overall pipeline.  Since would expect higher recognition rates but longer
+overall pipeline.  Since one would expect higher recognition rates but longer
 execution time for higher resolution images, we compare the recognition rate of

### 2011-05-28 results/accuracy/analyze_data.m
Working on extracting speed plot.

@@ -11,9 +11,12 @@ dir  = ['size' num2str(imsize) 'x' num2str(imsize) '/no_perturb/'];
 %dir  = ['size' num2str(imsize) 'x' num2str(imsize) '/1_1_3/'];
 blas = importdata([dir 'blas_recognition_output_keepers_20.txt']);
 cuda = importdata([dir 'cuda_recognition_output_keepers_20.txt']);
+%save([dir 'blas_recognition_output_keepers_20.mat'],blas);
+%save([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
+blas = load([dir 'blas_recognition_output_keepers_20.mat'],blas);
+blas = load([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
 
-
-labels = {'upload time', 'solve time', 'user names', 'threshold', 'tol', ...
+%%labels = {'upload time', 'solve time', 'user names', 'threshold', 'tol', ...
     'tol_int', 'maxIterInner', 'maxIterOuter', 'm', 'n', 'k', ...
     'solve_nIter', 'imagesPerUser', 'SCI', 'trueUserRank', ...
     'user_names (in order)', 'user residual values', ...
@@ -27,7 +30,7 @@ end
 
 up_time     = [blas.data(:,1), cuda.data(:,1)];
 solve_time  = [blas.data(:,2), cuda.data(:,2)];
-threshold   = mean([blas.data(:,4), cuda.data(:,4)]);
+%threshold   = mean([blas.data(:,4), cuda.data(:,4)]);
 tolerance   = mean([blas.data(:,5), cuda.data(:,5)]);
 tol_int     = mean([blas.data(:,6), cuda.data(:,6)]);
 maxIterInner = mean([blas.data(:,7), cuda.data(:,7)]);
@@ -56,226 +59,226 @@ start_row = end_row+1;
 end_row = start_row+n-1;
 x = {blas.data(:,start_row:end_row),cuda.data(:,start_row:end_row)};
 
-%%  plot some data
+%%%  plot some data
 
-subplot(3,1,1)
-plot(up_time(:,1), 'b');
-hold on
-plot(up_time(:,2), 'g');
-hold off
-title('upload time');
-xlabel('index');
-ylabel('time (s)');
[…]

### 2011-05-28 results/accuracy/plotSpeedVsRes.m

@@ -1,0 +1,29 @@
+function plotAccVsRes()
+
+res = [32 48 64 80 96 128];
+acc = zeros(numel(res),2);
+
+for i=1:numel(res)
+    resolution = res(i);
+    [b1,c1] = analyze_data(resolution);
+    acc(i,1) = b1;
+    acc(i,2) = c1;
+end
+
+%close all;
+%h = figure(7);
+%plot( res, acc(:,2) * 100, 'kx-', res, acc(:,1) * 100 , 'ko-');
+%legend('GPU', 'CPU', 'Location', 'SouthEast');
+%xlabel('Window width (pixels)');
+%ylabel('Recognition Rate (%)');
+%ylim([0 100]);
+
+%set (h,'windowstyle','normal');               %   Window must be undocked for the following
+%set (h,'Units','Inches');                     %   Using units of inches
+%windowSize = [0 0 3.75 2];
+%set(h,'position', windowSize);
+%set(h, 'PaperPosition', windowSize);
+
+%print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
+
+end

### 2011-05-29 results/accuracy/analyze_data.m
plots

@@ -13,8 +13,8 @@ blas = importdata([dir 'blas_recognition_output_keepers_20.txt']);
 cuda = importdata([dir 'cuda_recognition_output_keepers_20.txt']);
 %save([dir 'blas_recognition_output_keepers_20.mat'],blas);
 %save([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
-blas = load([dir 'blas_recognition_output_keepers_20.mat'],blas);
-blas = load([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
+%blas = load([dir 'blas_recognition_output_keepers_20.mat'],blas);
+%cuda = load([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
 
 %%labels = {'upload time', 'solve time', 'user names', 'threshold', 'tol', ...
     'tol_int', 'maxIterInner', 'maxIterOuter', 'm', 'n', 'k', ...

### 2011-05-29 results/accuracy/plotAccVsRes.m

@@ -26,4 +26,4 @@ set(h, 'PaperPosition', windowSize);
 
 print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
 
-end
+%end

### 2011-05-29 results/accuracy/plotSpeedVsRes.m

@@ -26,4 +26,4 @@ end
 
 %print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
 
-end
+%end

### 2011-05-29 results/accuracy/analyze_data.m
plots

@@ -14,9 +14,9 @@ cuda = importdata([dir 'cuda_recognition_output_keepers_20.txt']);
 %save([dir 'blas_recognition_output_keepers_20.mat'],blas);
 %save([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
 %blas = load([dir 'blas_recognition_output_keepers_20.mat'],blas);
-%cuda = load([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
+%blas = load([dir 'cuda_recognition_output_keepers_20.mat'],cuda);
 
-%%labels = {'upload time', 'solve time', 'user names', 'threshold', 'tol', ...
+labels = {'upload time', 'solve time', 'user names', 'threshold', 'tol', ...
     'tol_int', 'maxIterInner', 'maxIterOuter', 'm', 'n', 'k', ...
     'solve_nIter', 'imagesPerUser', 'SCI', 'trueUserRank', ...
     'user_names (in order)', 'user residual values', ...

### 2011-05-29 results/accuracy/analyze_data.m
plots

@@ -112,8 +112,8 @@ cuda_keep1 = true_user < 250;
 
 %sci_threshold = 0.17;
 
-%blas_keep2 = SCI(:,1) > sci_threshold;
-%cuda_keep2 = SCI(:,2) > sci_threshold;
+blas_keep2 = SCI(:,1) > sci_threshold;
+cuda_keep2 = SCI(:,2) > sci_threshold;
 
 blas_rank = trueUserRank(blas_keep1 & blas_keep2,1);
 cuda_rank = trueUserRank(cuda_keep1 & cuda_keep2,2);
@@ -217,8 +217,8 @@ rejected = sum(cuda_keep3 == 0);
 blas_keep1 = true_user <= 250;
 cuda_keep1 = true_user <= 250;
 
-%blas_keep2 = SCI(:,1) > sci_threshold;
-%cuda_keep2 = SCI(:,2) > sci_threshold;
+blas_keep2 = SCI(:,1) > sci_threshold;
+cuda_keep2 = SCI(:,2) > sci_threshold;
 %blas_keep3 = users_residuals{1}(:,1) < residual_threshold;
 %cuda_keep3 = users_residuals{2}(:,1) < residual_threshold;
 

### 2011-05-29 results/accuracy/extract_times.sh
Recognitoin stage runtime should work in matlab.

@@ -1,0 +1,25 @@
+#!/bin/bash
+for y in blas cuda
+do
+	rm -f ${y}times
+	touch ${y}times
+	for x in 32 48 64 80 96 128
+	do
+		cat size${x}x${x}/no_perturb/${y}_recognition_output_keepers_20.txt | cut -d' ' -f8 > temptimes;
+		paste -d' ' ${y}times temptimes > temptimes2
+		mv temptimes2 ${y}times
+	done
+done
+
+mfile=recognition_times.m
+rm -f $mfile
+echo "res=[32 48 64 80 96 128];" >> $mfile
+for y in blas cuda
+do
+	echo "${y}times=[" >> $mfile
+	cat ${y}times >> $mfile
+	echo "];">> $mfile 
+	echo "${y}times = mean(${y}times,1);" >> $mfile
+done
+
+rm -f temptimes cudatimes blastimes

### 2011-05-29 results/accuracy/plotAccVsRes.m

@@ -26,4 +26,4 @@ set(h, 'PaperPosition', windowSize);
 
 print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
 
-%end
+end

### 2011-05-29 results/accuracy/plotSpeedVsRes.m

@@ -1,29 +1,16 @@
-function plotAccVsRes()
+recognition_times % Run extract_times.sh to generate this file.
 
-res = [32 48 64 80 96 128];
-acc = zeros(numel(res),2);
+close all;
+h = figure(7);
+plot( res, cudatimes, 'kx-', res, blastimes, 'ko-');
+legend('GPU', 'CPU', 'Location', 'SouthEast');
+xlabel('Window width (pixels)');
+ylabel('Elapsed time (s)');
 
-for i=1:numel(res)
-    resolution = res(i);
-    [b1,c1] = analyze_data(resolution);
-    acc(i,1) = b1;
-    acc(i,2) = c1;
-end
+set (h,'windowstyle','normal');               
+set (h,'Units','Inches');                    
+windowSize = [0 0 3.75 2];
+set(h,'position', windowSize);
+set(h, 'PaperPosition', windowSize);
+print(h, '../../figures/speedVsResolution.eps', '-deps2');
 
-%close all;
-%h = figure(7);
-%plot( res, acc(:,2) * 100, 'kx-', res, acc(:,1) * 100 , 'ko-');
-%legend('GPU', 'CPU', 'Location', 'SouthEast');
-%xlabel('Window width (pixels)');
-%ylabel('Recognition Rate (%)');
-%ylim([0 100]);
-
-%set (h,'windowstyle','normal');               %   Window must be undocked for the following
-%set (h,'Units','Inches');                     %   Using units of inches
-%windowSize = [0 0 3.75 2];
-%set(h,'position', windowSize);
-%set(h, 'PaperPosition', windowSize);
-
-%print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
-
-%end

### 2011-05-29 results/accuracy/recognition_times.m

@@ -1,0 +1,2713 @@
+res=[32 48 64 80 96 128];
+blastimes=[
+ 0.142199 0.182301 0.33289 0.320231 0.250251 0.340394
+ 0.0966094 0.336167 0.279986 0.25431 0.389691 0.470237
+ 0.180092 0.209197 0.206542 0.220751 0.257808 1.02017
+ 0.235881 0.468272 0.21033 0.293386 0.404377 0.841716
+ 0.459029 0.466585 0.219512 0.244405 0.607816 0.922753
+ 0.366286 0.278246 0.416003 0.361309 0.386845 0.689421
+ 0.0963254 0.276924 0.394689 0.22374 0.18971 0.606888
+ 0.107564 0.187594 0.651839 0.204195 0.606492 0.451667
+ 0.152562 0.556813 0.367029 0.231086 0.66808 0.891359
+ 0.111311 0.365451 0.557755 0.167097 0.19954 0.290176
+ 0.0818299 0.665589 0.79132 0.243515 0.54785 0.363848
+ 0.13422 0.145826 0.207274 0.274385 1.15226 1.09175
+ 0.153948 0.141769 0.221627 0.222811 0.282943 0.481207
+ 0.128116 0.48511 0.165649 0.184896 0.293558 0.365582
+ 0.161931 0.173731 0.218378 0.304281 0.301603 0.411524
+ 0.171077 0.475349 0.22882 0.30139 0.260238 0.805285
+ 0.139828 0.583685 0.198339 0.329591 0.282227 0.446354
+ 0.131941 0.503648 0.266116 0.19022 0.697944 0.669807
+ 0.367452 0.216821 0.218725 0.239172 0.302052 0.385372
+ 0.317629 0.19673 0.274175 0.246163 0.230664 0.409808
+ 0.244311 0.181225 0.192825 0.277889 0.411446 1.03337
+ 0.208915 0.16918 0.415602 0.255712 0.694796 0.987522
+ 0.556288 0.134825 0.452023 0.277563 0.754809 0.860671
+ 0.100176 0.138005 0.205762 0.194752 0.518453 0.880802
+ 0.131393 0.180384 0.189868 0.200622 0.814892 0.44913
+ 0.107987 0.154148 0.242897 0.243 1.13824 0.618187
+ 0.102612 0.16858 0.135345 0.185524 0.232397 0.465844
+ 0.0769558 0.230183 0.191888 0.246655 0.298242 0.359311
[…]

### 2011-05-29 main.tex
unified todo list
Added recognition stage runtime plot, fixed legend

@@
 
-The following experiment fulfills the dual purpose of verifying the accuracy of
-the alignment stage, and motivating a threshold on how many users are kept for
-the recognition stage.  Setting this threshold establishes a compromise between
-recognition speed and potential recognition rate.  To explore the affect of the
-threshold on potential recognition rate, we vary the number of users kept for
-the recognition stage and estimate the probability of the test user being
-among the set of training users kept for the recognition stage.  This curve
-bounds above the achievable recognition rate of the overall pipeline.  As can
-be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly
-diminishing returns as the number of kept users is increased beyond 20; a small
-fraction of the gallery.
+%The following experiment fulfills the dual purpose of verifying the accuracy of
+%the alignment stage, and motivating a threshold on how many users are kept for
+%the recognition stage.  Setting this threshold establishes a compromise between
+%recognition speed and potential recognition rate.  To explore the affect of the
+%threshold on potential recognition rate, we vary the number of users kept for
+%the recognition stage and estimate the probability of the test user being
+%among the set of training users kept for the recognition stage.  This curve
+%bounds above the achievable recognition rate of the overall pipeline.  As can
+%be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly
+%diminishing returns as the number of kept users is increased beyond 20; a small
+%fraction of the gallery.
+%\begin{figure} 
+%\centering
+%\includegraphics[width=3.4in]{figures/user_alignment_rank_plot}
+%\caption{Probability of the test user being kept for the recognition stage vs.
+%number of training users kept for the recognition stage.}
+%\label{fig:user_alignment_rank_plot} \end{figure}
+
 \begin{figure} 
 \centering
-\includegraphics[width=3.4in]{figures/user_alignment_rank_plot}
-\caption{Probability of the test user being kept for the recognition stage vs.
-number of training users kept for the recognition stage.}
-\label{fig:user_alignment_rank_plot} \end{figure}
+\includegraphics[width=3.4in]{figures/speedVsResolution}
+\caption{Recognition stage runtime vs. size of training database.}
+\label{fig:recognition_stage_runtime} \end{figure}
 
@@
 \end{document}
-
-% To do list:
-%Add experiment timing for one instance per core on cpu?
-%Unify plot styles
-%Interpret 2 percent accuracy difference
-%Update speed comparisons and diagrams to GTX 580?
-%Better citation for previously reported speed?
-%Figure tweaks from Allen

### 2011-05-29 main.tex
Fixed writing to account for new figure.

@@
 
+The next experiment compares the speed of the GPU and CPU implementations of
+the recognition stage.  It was determined that keeping 20 gallery users is
+sufficient to ensure that the correct user is kept for the recognition stage
+with 95\% probability.  Since recognition failures are typically caused by a
+poor alignment, we find that keeping more users for the recognition stage does
+not necessarily improve recognition rate.
+%The inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.  
+In Figure \ref{fig:recognition_stage_runtime} we show recognition stage runtime
+for canonical images of size: $32\times32$, $48 \times 48$, $64 \times 64$, $96 \times
+96$, and $128 \times 128$.  For all image resolutions the problem size
+is sufficiently small that the CPU is significantly faster than the GPU.  Note
+also that for both implementations, the recognition stage takes much less time
+than the alignment stage.
 \begin{figure} 
@@
 
-Finally, we perform an experiment verifying the recognition accuracy of the
-overall pipeline.  Since one would expect higher recognition rates but longer
-execution time for higher resolution images, we compare the recognition rate of
-the overall pipeline with the face recognition stage run at 5 different
-canonical image resolutions: $32\times32, 48 \times 48, 64 \times 64, 96 \times
-96, \textrm{and} 128 \times 128$.  As motivated by Figure
-\ref{fig:user_alignment_rank_plot}, we fix the number of gallery users to 20,
-giving approximately 95\% probability to have the true user in the training
-set.  
-%The inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.  
 \begin{figure}
@@
 \end{figure}
-As can be seen in Figure \ref{fig:accuracy_vs_resolution}, at the optimal
-resolution, which happens to match the alignment stage resolution, this bound
-is very nearly achieved by the GPU implementation.  For significantly lower
-resolutions, the accuracy drops off significantly.  Of the two main stages of
-the face recognition pipeline, the iterative alignment stage is by far the most
-expensive; the recognition stage requires less than one second.
+Finally, we perform an experiment verifying the recognition accuracy of the
+overall pipeline.  As can be seen in Figure \ref{fig:accuracy_vs_resolution},
+at the optimal resolution, which happens to match the alignment stage
+resolution, the GPU implementation reaches 95\% recognition rate, the max
+achievable given the alignment results.  For significantly lower resolutions,
+the accuracy drops off significantly.
 

### 2011-05-31 main.tex
Merged in latest corrections from Allen.

@@
 % Pages are numbered in submission mode, and unnumbered in camera-ready
+
+\hyphenation{under-determined}
 \ifcvprfinal\pagestyle{empty}\fi
@@
 \title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
+%\title{Parallelization of Fast $\ell_1$-Minimization Towards Real-Time Face Recognition}
 
@@
 
-\begin{abstract} 
-Recently a family of promising face recognition algorithms that are based on sparse representation and $\ell_1$ minimization have been been developed.
-These algorithms have not yet seen commercial application, largely due to their greater computational requirements compared to more traditional algorithms.
-This paper studies techniques for leveraging the massive parallelism available in GPU and multi-core CPU hardware to accelerate augmented Lagrangian method (ALM) based $\ell_1$ minimization solvers.
-The proposed implementations are significantly faster than than previously published implementations.
-Despite the low arithmetic intensity of ALM solvers, the GPU can still outperform the CPU by a wide margin for large problem sizes, and when solving many small problems simultaneously.
-The proposed CPU implementation remains faster for solving a single moderately sized problem.
-\end{abstract}
+\begin{abstract} Recently a family of promising face recognition algorithms
+based on sparse representation and $\ell_1$-minimization ($\ell_1$-min) have been
+developed.  These algorithms have not yet seen commercial application, 
+largely due to higher computational cost compared to other traditional
+algorithms. This paper studies techniques for leveraging the 
+massive parallelism available in GPU and CPU hardware to accelerate
+$\ell_1$-min based on augmented Lagrangian method (ALM) solvers. We demonstrate
+through extensive simulation and real-world experiment that the proposed
+algorithms significantly speed up the sparsity recovery problem compared to the
+state of the art.  Despite the low arithmetic intensity of ALM solvers, the GPU
+can still outperform the CPU by a wide margin for large problem sizes, and when
+solving many small problems simultaneously. The CPU implementation remains
+faster for solving a single moderately sized problem. The source code of the
+algorithms will be made available for peer evaluation.  \end{abstract}
 
 \section{Introduction} 
-$\ell_1$ minimization ($\ell_1$-min) has received much attention in recent
+$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
 years due to important applications in compressive sensing
 \cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  
-$\ell_1$-min refers to finding the minimum $\ell_1$-norm solution to an
+The problem refers to finding the minimum $\ell_1$-norm solution to an
 underdetermined linear system $\bb=A\xx$:
@@
 $\ell_1$-norm as a robust error function, while simultaneously estimating
-the coefficients of the linear illumination model $A\x$.
+the coefficients of the linear illumination model $A_i\x$.
 
@@
 
-In this discussion we will clearly distinguish between algorithm {\em parallelism},
-which is a property of an algorithm, and hardware {\em concurrency}, which is a
+To clarify discussion, we clearly distinguish between {\em parallelism},
+which is a property of an algorithm, and {\em concurrency}, which is a
 property of a hardware architecture. Parallelism in an algorithm provides the
@@
 distributed over a cluster or network of computers, and this is not addressed here.
-The source code of our implementations, which target Nvidia's CUDA GPU architecture,
+The source code of our implementations, which target Nvidia's CUDA GPU architecture
 as well as the x86-64 architecture, will be made available for peer evaluation.
@@
 \subsection{Literature Review} 
-We first briefly review the literature of $\ell_1$-min and existing
-works on its parallelization.  Traditionally, $\ell_1$-min (or the so-called
+Traditionally, $\ell_1$-min (a.k.a.
 basis pursuit (BP)) has been formulated as a linear program
@@
 computationally expensive operations such as matrix-matrix multiplication and
-the solution of least-squares problems with varying $A$
-matrices.\footnote{During the face alignment algorithm presented later, The $A$
-matrix changes between alignment iterations but not within the L1 solver}  In
-Section \ref{sec:ALM}, we contend that ALM is a better choice for
+linear least-squares problems with varying $A$
+matrices.
+%\footnote{During the face alignment algorithm presented later, The $A$ matrix changes between alignment iterations but not within the L1 solver}  
+In Section \ref{sec:ALM}, we contend that ALM is a better choice for
 implementation on many-core CPUs and GPUs. The ALM algorithm belongs to a
@@
 algorithms. Therefore, we choose ALM as the core algorithm for
-implementation of the parallel face recognition pipeline.
+implementation of $\ell_1$-min in the parallel face recognition pipeline.
 
@@
 
-\section{Augmented Lagrange Multiplier Method}
+\section{Augmented Lagrangian Method}
 \label{sec:ALM}
 
-For the completeness of the paper, in this section, we briefly describe the ALM
-algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP}
- %and analyze its complexity
-. 
-Parallelization of ALM for face alignment and recognition will
-be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition},
-respectively.
-
-Lagrange multiplier methods have been frequently used in convex programming to
-eliminate equality constraints via adding a significant penalty term to the
+In this section, we briefly describe the ALM algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP}
+and analyze its complexity. Lagrange multipliers have been frequently used in convex programming to
+eliminate equality constraints via adding a penalty term to the
 cost function for infeasible points. ALM methods differ from other
 penalty-based approaches by simultaneously estimating the optimal solution and
-Lagrange multipliers in an iterative fashion.  For $\ell_1$-min, the augmented
-Lagrangian is defined as: \begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
-\yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2, \end{equation}
+Lagrange multipliers in an iterative fashion.  For $\ell_1$-min \eqref{eq:l1min}, the augmented
+Lagrange function is defined as: 
+\begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
+\yy^T(\bb - A\xx) + \frac{\mu}{2} \| \bb-A\xx \|_2^2, 
+\end{equation}
 where $\mu > 0$ is a constant that penalizes infeasibility and $\yy$ is a
-vector of lagrange multipliers.
+vector of Lagrange multipliers.
 
-In Lagrange Multiplier Theory, if there exists a Lagrangian $y^*$ that
-satisfies the second-order sufficiency conditions for optimality
-\cite{BertsekasD2003}, then for a sufficiently large $\mu$, the optimal
+In Lagrange Multiplier Theory \cite{BertsekasD2003}, if there exists a Lagrangian $y^*$ that
+satisfies the second-order sufficiency conditions for optimality, then for a sufficiently large $\mu$, the optimal
 $\ell_1$-min solution also minimizes
@@
 \end{equation}
-where $\mu_{k}\rightarrow \infty$ is a monotonically increasing positive sequence.
-The iteration terminates when the estimates $(\xx, \yy)$ converge.
+where $\mu_{k}\rightarrow \infty$ is monotonically increasing.
+The iteration terminates when the estimates $(\xx_k, \yy_k)$ converge.
 
@@
 \bibliographystyle{ieee}
-\bibliography{faces}
+\bibliography{faces-short}
 }

### 2011-05-31 results/accuracy/plotAccVsRes.m
Unfunctionized a .m file to make debugging easier.

@@ -1,4 +1,3 @@
-function plotAccVsRes()
 %%
 res = [32 48 64 80 96 128];
 acc = zeros(numel(res),2);
@@ -26,4 +25,3 @@ set(h, 'PaperPosition', windowSize);
 
 print(h, '../../figures/accuracyVsResolution.eps', '-depsc2');
 
-end

### 2011-05-31 figures/alignment_runtime_graph.m
Working on alignment stage runtime plot.

@@ -1,0 +1,29 @@
+
+% For MATLAB, you might need to manually remove the first row if there is a header.
+cpu1=load('alignment_runtime_graph_reference_cpu.dat');
+cpu2=load('alignment_runtime_graph_reference_cpu_percore.dat');
+gtx480=load('alignment_runtime_graph_reference_gtx480.dat');
+gtx580=load('alignment_runtime_graph_reference_gtx580.dat');
+
+userCount = cpu1(:,1);
+
+cpu1 = cpu1(:,2);
+cpu2 = cpu2(:,2);
+gtx480 = gtx480(:,2);
+gtx580 = gtx580(:,2);
+
+close all;
+h = figure(7);
+plot( userCount, gtx480, 'kx-' , userCount, gtx580, 'ko-' , userCount, cpu1, 'ks-' , userCount, cpu2, 'kd-');
+legend('gtx480', 'gtx580', '2 x E5530, single instance', '2 x E5530, 1 instance per core');
+xlabel('Number of training users');
+ylabel('Elapsed time (s)');
+
+set (h,'windowstyle','normal');               
+set (h,'Units','Inches');                    
+%windowSize = [0 0 3.75 2];
+windowSize = [0 0 20 15];
+set(h,'position', windowSize);
+set(h, 'PaperPosition', windowSize);
+print(h, 'alignment_runtime_graph.eps', '-deps2');
+

### 2011-06-01 figures/alignment_runtime_graph.m
Alignment stage benchmark

@@ -14,16 +14,18 @@ gtx580 = gtx580(:,2);
 
 close all;
 h = figure(7);
-plot( userCount, gtx480, 'kx-' , userCount, gtx580, 'ko-' , userCount, cpu1, 'ks-' , userCount, cpu2, 'kd-');
-legend('gtx480', 'gtx580', '2 x E5530, single instance', '2 x E5530, 1 instance per core');
+plot( userCount, gtx480, 'k-.', userCount, gtx580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
+legend('gtx480 GPU', 'gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core', 'Location','NorthWest');
 xlabel('Number of training users');
 ylabel('Elapsed time (s)');
 
 set (h,'windowstyle','normal');               
 set (h,'Units','Inches');                    
-%windowSize = [0 0 3.75 2];
-windowSize = [0 0 20 15];
+windowSize = [0 0 3.75 2];
+windowSize = windowSize*4;
+%windowSize = [0 0 20 15];
 set(h,'position', windowSize);
 set(h, 'PaperPosition', windowSize);
 print(h, 'alignment_runtime_graph.eps', '-deps2');
+print(h, 'alignment_runtime_graph.png', '-dpng');
 

### 2011-06-03 main.tex
Re-wrote abstract and part of intro.

@@
 massive parallelism available in GPU and CPU hardware to accelerate
-$\ell_1$-min based on augmented Lagrangian method (ALM) solvers. We demonstrate
-through extensive simulation and real-world experiment that the proposed
-algorithms significantly speed up the sparsity recovery problem compared to the
-state of the art.  Despite the low arithmetic intensity of ALM solvers, the GPU
-can still outperform the CPU by a wide margin for large problem sizes, and when
-solving many small problems simultaneously. The CPU implementation remains
-faster for solving a single moderately sized problem. The source code of the
+$\ell_1$-min based on augmented Lagrangian method (ALM) solvers. 
+Due to the low arithmetic intensity of ALM solvers, both
+architectures are bandwidth limited.  
+with large problems running faster on the GPU, and small problems running faster
+on the CPU.  
+On both architectures, the proposed implementations dramatically out-perform
+nieve library-based implementations. 
+At the problem sizes most important for $\ell_1$-min based
+face recognition, both architectures achieve comparable performance.
+The source code of the
 algorithms will be made available for peer evaluation.  \end{abstract}
@@
 distributed over a cluster or network of computers, and this is not addressed here.
-The source code of our implementations, which target Nvidia's CUDA GPU architecture
+The source code of our implementations, which target NVIDIA's CUDA GPU architecture
 as well as the x86-64 architecture, will be made available for peer evaluation.
@@
 Moreau-Yosida regularization. In \cite{MurphyM2010}, Murphy et al. presented
-parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia
-CUDA programmable GPU architecture, which is also used in this paper.
+parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on 
+the same GPU architecture addressed in this paper.
 
@@
 
-In this section, we briefly describe the ALM algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP}
-and analyze its complexity. Lagrange multipliers have been frequently used in convex programming to
-eliminate equality constraints via adding a penalty term to the
-cost function for infeasible points. ALM methods differ from other
-penalty-based approaches by simultaneously estimating the optimal solution and
-Lagrange multipliers in an iterative fashion.  For $\ell_1$-min \eqref{eq:l1min}, the augmented
-Lagrange function is defined as: 
+In this section, we briefly describe the ALM algorithm for $\ell_1$-min
+\eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze its complexity. Lagrange
+multipliers have been frequently used in convex programming to eliminate
+equality constraints via adding a penalty term to the cost function for
+infeasible points. ALM methods differ from other penalty-based approaches by
+simultaneously estimating the optimal solution and Lagrange multipliers in an
+iterative fashion.  For $\ell_1$-min \eqref{eq:l1min}, the augmented Lagrange
+function is defined as: 
 \begin{equation} L_\mu(\xx,\yy) = \|\xx\|_1 +
@@
 (regions of on-chip memory) are often orders of magnitude faster than off-chip
-memory, their sizes and speeds have big performance implications.  We give a
+memory, their sizes and speeds have a dramatic effect on performance.  We give a
 brief overview of the caches that are available in our target architectures,
-and defer discussion of their performance implications to Sections
+and defer discussion of their performance effects to Sections
 \ref{sec:alignment_implementation_cpu} and
@@
 
-We will primarily be comparing CPU and GPU architectures on a per-board basis:
-for CPU implementation we make use of all the cores in as many CPUs are present
-on one motherboard, and for GPU implementation we make use of a single
-many-core Nvidia GPU chip on a PCI card. 
-%GPUs are commonly shipped with a
-%single GPU chip per PCI-express card.  Note that while typical engineering
-%workstations can only support a single high-end GPU due to their high current
-%draw, purpose-built servers can support multiple high-end GPU's.  
+Our discussion and experiments will focus on the most common hardware
+configuration for engineering workstations: a motherboard with two quad-core
+processors, and a PCI card with a single high-end GPU.  
+Recognition involving more than a few hundred users with contemporary hardware
+would require a server(or cluster) configuration with an expandable 
+number of processors.  We will not address efficient parallelization for 
+these systems, which may have additional challenges associated with
+their non-shared memory model.\footnote{Non-shared memory configurations
+include CPU blade servers, multi-GPU servers, and clusters}.
 Unless otherwise specified, all implementations utilize single precision
 floating point datatypes.  
+% WORKING HERE
 
@@
 
-Our target GPU architecture is the Nvidia Fermi architecture (\eg, the GTX 480
+Our target GPU architecture is the NVIDIA Fermi architecture (\eg, the GTX 480
 GPU used in this paper).  CUDA based GPUs are comprised of several
@@
 
-On the GPU, most of the operations map well onto similar calls in Nvidia's GPU
+On the GPU, most of the operations map well onto similar calls in NVIDIA's GPU
 BLAS library (CUBLAS) which, like MKL, can take advantage of both levels of

### 2011-06-04 main.tex
Finished modifying paper to fit new result

@@
 \cite{WagnerA2011-PAMI}. However, due the high per-class cost
-of the alignment step, the recognition system still fails to achieve real-time recognition against datasets of hundreds or thousands of users.
-In this paper, in addition to accelerating the generic
+of the alignment step, the recognition system still fails to achieve real-time
+recognition against datasets of hundreds or thousands of users.  In this paper,
+in addition to accelerating the generic
 $\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
@@
 
-Our discussion and experiments will focus on the most common hardware
+Our discussion and experiments will address the most common hardware
 configuration for engineering workstations: a motherboard with two quad-core
@@
 For image alignment problem, the number of training images per class is
-typically larger than 16.  Therefore, if eight instances of
-\eqref{eq:alignment-linearization} were be executed in parallel on the eight
-cores, there would be an insufficient amount of L1 or L2 cache to hold $B_j$ and
-$B_j^\dagger$ during the iterative process to execute Algorithm
-\ref{alg:iterative_alignment}. Therefore, the bottleneck in this strategy becomes the
-bandwidth between the L2 cache and L3 cache. The implementation of some
-previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} is indeed based on this
-mapping strategy.
+typically larger than 16.  
+In this paper we consider two mappings of the parallelism available in the alignment
+stage to the concurrency available on the CPU.  We argue that the latter
+implementation should achieve higher performance, and later verify that this is the case.
 
-In this paper, we show that for CPU parallelism of face alignment, a higher
-performance implementation can be achieved by {\em not} taking advantage of
-problem-level parallelism at all, and instead mapping the pixel-level parallelism
+The first implementation disregards 
+problem-level parallelism, and instead mapping the pixel-level parallelism
 onto both the core-level and the SSE-level concurrency provided by the CPU.  In
-other words, the proposed CPU implementation performs alignment of a query
+other words, alignment is performed of the query
 image against the training images of a single user at a time using all cores on
-both processors.
+both processors.  The advantage of this implementation is that all of the variables
+for the inner loop, most notably $B_j$ and $B_j^\dagger$, fit in L2 cache.
 
-More specifically, in our implementation, most of the operations take
+The second implementation maps problem-level parallelism onto core-level
+concurrency, and maps pixel-level parallelism onto SSE-level concurrency.  In
+other words, eight instances of \eqref{eq:alignment-linearization} are executed
+in parallel with each problem running on one of the eight cores.  The advantage
+of this implementation is that since the cores are operating on different
+alignment problem instances, no data is shared between cores, and therefore no
+synchronization is necessary between the cores.  Furthermore, even with eight
+problem instances running concurrently, the local variables for the inner
+loop still fit in L3 cache.  Thanks to the sequential data access patterns of
+the solver, prefetching significantly reduces the bandwidth difference between
+L2 and L3 cache.  For these reasons, we expect this implementation to out-perform
+the previous implementation.
+
+%The implementation of some
+%previous works \cite{WagnerA2009-CVPR,WagnerA2011-PAMI} are indeed based on the latter
+%parallelization strategy.
+
+%64 byte cache line-size per clock.  L3 cache running at lower clock than cpu.  Two reads and a write per clock.
+%4 floats at a time in SSE.
+
+In both implementations, most of the operations take
 advantage of Intel Math Kernel Library (MKL), a commercial implementation of
@@
 Both MKL and IPP are optimized for Intel multi-core CPUs, and are able to
-automatically utilize both core-level and vector-level concurrency. For other
+automatically utilize both core-level and vector-level concurrency (for the first implementation),
+but are also available in single-threaded versions (for the second implementation). For 
 operations that are not optimized by Intel in-house libraries, such as the shrinkage
-operator in step 7 of Algorithm \ref{alg:iterative_alignment}, we achieve both
-vector-level and core-level concurrency through a combination of the Open MP
-API \cite{dagum2002openmp} and the automatic vectorization facilities of the
-Intel C++ compiler (ICC) \cite{dulong1999overview}. 
+operator in step 7 of Algorithm \ref{alg:iterative_alignment}, we achieve 
+vector-level concurrency via
+the automatic vectorization facilities of the Intel C++ compiler (ICC) \cite{dulong1999overview}.
+To achieve core-level concurrency we make use of the Open MP API. \cite{dagum2002openmp} 
 %We shall compare the performance of the two parallelism methods in Section \ref{sec:experiment}.
@@
 into L2 cache, this would not be an efficient use of the GPU's resources.
-The strength of the GPU's memory architecture for our purposes it's ability to sustain a very
-high bandwidth to DRAM.  A higher overall bandwidth can be achieved if 
-many memory transactions are performed simultaneously, than can be achieved
-with a large number of smaller transactions that are just large enough to fill
-the L2 cache.
+The strength of the GPU's memory architecture for our purposes is it's ability to sustain a very
+high bandwidth to DRAM.  This bandwidth is achieved by having a very large number of
+threads issuing memory requests simultaneously.  Solving many alignment problem instances
+concurrently increases the number of threads that can be used.
 
-Therefore, whereas on the CPU we advocate performing alignment for one training
-class at a time, on the GPU we advocate performing alignment for many classes
-concurrently.  \footnote{In CUDA terminology, Our proposed alignment stage
+Therefore, our recommended GPU implementation is strikingly similar in spirit
+to our recommended CPU implementation: on the CPU we use the cores to solve
+multiple instances concurrently, and on the GPU we use multiple SMP's for the 
+same purpose.
+\footnote{In CUDA terminology, Our proposed alignment stage
 implementation consists of a single kernel that performs alignment for all
@@
 concurrently as the hardware will allow.} The number of subject classes that
-actually get aligned concurrently is chosen by the GPU hardware. In our
-experiments the hardware is able to schedule 4-5 classes per SMP to run
-concurrently. This mapping of the problem-level parallelism onto the SMP-level
-concurrency is illustrated in Figure \ref{fig:alignment_mapping_gpu}.
+actually get aligned concurrently is ultimately chosen by the GPU hardware, 
+but can be influenced by manual tuning of the code. We have empirically determined
+that performance is highest with 5-7 subject classes aligned simultaneously on each SMP.
+%This mapping of the problem-level parallelism onto the SMP-level
+%concurrency is illustrated in Figure \ref{fig:alignment_mapping_gpu}.
+%\begin{figure}
+%\centering
+%\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
+%\caption{Proposed mapping of alignment parallelism onto GPU concurrency}
+%\label{fig:alignment_mapping_gpu}
+%\end{figure}
 
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/alignment_mapping_gpu}
-\caption{Proposed mapping of alignment parallelism onto GPU concurrency}
-\label{fig:alignment_mapping_gpu}
-\end{figure}
-
+Several other aspects of the implementation merit discussion.  
 The full resolution query image, along with its numerical derivatives in both
@@
 
-Solving one alignment problem per SMP necessitated the implementation of a
+Solving one alignment problem per SMP necessitates the implementation of a
 custom CUDA routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$. $G$
@@
 \subsection{Recognition Stage Implementation} Compared to problem-level
-parallelism, pixel-level parallelism in Algorithm \ref{alg:alm_rec} is
-relatively easy to exploit using existing BLAS libraries.  For this reason,
-both the CPU and GPU implementation of the recognition stage is very similar to
-the CPU implementation of the $\ell_1$-min advocated in Section
-\ref{sec:alignment_implementation_cpu}.  On the CPU, most of the operations map
-well onto MKL BLAS calls, and the operations that do not can be easily
+parallelism, the exclusively pixel-level parallelism in Algorithm \ref{alg:alm_rec} is
+relatively straightforward to exploit:  
+On the CPU, most of the operations map
+well onto MKL BLAS calls, and operations that do not can be easily
 parallelized using OpenMP and auto-vectorization.
@@
 recognition problem.
-
-To generate a meaningful comparison between different hardware architectures,
-in all cases we compare the performance of hardware architectures on a
-per-board basis:  For CPU implementations, the benchmark makes use of all of
-the cores in as many CPU's are present.  For GPU implementations, the benchmark
-makes the best use of the entire GPU chip (most GPU boards have a single GPU
-chip).  Aside from the exceptions already mentioned in Section
+%To generate a meaningful comparison between different hardware architectures,
+%in all cases we compare the performance of hardware architectures on a
+%per-board basis:  For CPU implementations, the benchmark makes use of all of
+%the cores in as many CPU's are present.  For GPU implementations, the benchmark
+%makes the best use of the entire GPU chip (most GPU boards have a single GPU
+%chip).  
+Aside from the exceptions already mentioned in Section
 \ref{sec:alignment_implementation_gpu}, all math operations are performed in
@@
 
-The first experiment compares the performance of our two proposed
-implementations of the general $\ell_1$-min (Algorithm \eqref{alg:alm}) given a
+The first experiment compares the performance of our proposed CPU and GPU
+implementations of the general $\ell_1$-min (Algorithm \eqref{alg:alm}), given a
 generic random matrix $A$ and a random sparse vector $\xx_0$.  The measurement
@@
 norm.  Because the ground truth signal $\xx_0$ is known, the algorithm
-terminats when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement
+terminates when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement
 vector is generated with the equation $\bb_0 = A \xx_0$.
@@
 single large problem, whereas the CPU implementation is faster at solving a
-single small problem.  
+single small problem.  The transition between the two regimes occurs where
+the problem size begins to exceed the size of the CPU L3 cache. 
 \begin{figure}
@@
 \end{figure}
-Similar to the results in Figure \ref{fig:random_data},
-there is a crossover point above which the GPU out-performs the CPU.  Since
-the CPU algorithm solves problems sequentially, there is a linear trend.  
 The tiers seen in the GPU profile result from the GPU hardware scheduling alignment
 problems in concurrent batches, but the overall trend is linear as expected.
+The CPU implementation that solves one problem instance per core out-performs
+the single-instance-at-a-time implementation by a wide margin, and is even competitive
+with the GPU implementation.\footnote{Note that the GPU implementation was tuned
+to be fast for the full problem set; tuning affects the locations of the tiers,
+so the lower corners of the tiers are most relevant.}
 
@@
 \section{Conclusion}
-We have shown that our proposed parallelization is {\em eight times} faster
-than an implementation solving a single problem at a time using the stock
-CUBLAS library. Furthermore, the proposed mapping of the parallelism of the
-face alignment problem onto the concurrency provided in modern CPU and GPU
-architectures results in dramatic improvements in recognition speed over
-previously reported implementations of the same pipeline.  Whereas earlier
-implementations required over a second to align a test image against just one
-of the 250 users in Multi-PIE session 1, the proposed CPU implementation can
-perform the same task in only 45 seconds, and the GPU implementation can do it
-in {\bf under 10 seconds}!.  In principle, the parallelization techniques
-presented here make it feasible to implement an access control system for
-hundreds of users on a handful of purpose-build multi-GPU workstations.
+We have demonstrated that on both CPU and GPU algorithms, parallelizations of
+ALM that solve multiple face alignment problems concurrently are significantly
+faster than implementations that rely purely on vendor-supplied BLAS libraries.
+Furthermore, thanks to a combination of faster hardware, and more efficient use
+of the hardware, we have demonstrated dramatic improvements in recognition
+speed over previously reported implementations of the same pipeline.  Whereas
+earlier implementations required over a second to align a test image against
+just one of the 250 users in Multi-PIE session 1, the proposed 
+implementations can perform the same task in 
+{\bf under 10 seconds}!.  In principle, the
+parallelization techniques presented here make it feasible to implement an
+access control system for hundreds of users on a handful of purpose-build
+multi-GPU workstations.  As CPU manufacturers 
+increase the number of cores and vector widths, and as GPU manufacturers increase the amount
+of cache, both architectures are rapidly converging towards an architectural balance
+that increasingly favorable for $ell_1$-min based face recognition.
 

### 2011-06-04 figures/alignment_runtime_graph.m

@@ -5,24 +5,37 @@ cpu2=load('alignment_runtime_graph_reference_cpu_percore.dat');
 gtx480=load('alignment_runtime_graph_reference_gtx480.dat');
 gtx580=load('alignment_runtime_graph_reference_gtx580.dat');
 
-userCount = cpu1(:,1);
+%userCount = cpu1(:,1);
+%cpu1 = cpu1(:,2);
+%cpu2 = cpu2(:,2);
+%gtx480 = gtx480(:,2);
+%gtx580 = gtx580(:,2);
 
-cpu1 = cpu1(:,2);
-cpu2 = cpu2(:,2);
-gtx480 = gtx480(:,2);
-gtx580 = gtx580(:,2);
-
+userCount = cpu1(1:2:end,1);
+cpu1 = cpu1(1:2:end,2);
+cpu2 = cpu2(1:2:end,2);
+gtx480 = gtx480(1:2:end,2);
+gtx580 = gtx580(1:2:end,2);
 close all;
 h = figure(7);
-plot( userCount, gtx480, 'k-.', userCount, gtx580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
-legend('gtx480 GPU', 'gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core', 'Location','NorthWest');
+
+%plot( userCount, gtx480, 'k-.', userCount, gtx580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
+%legend('gtx480 GPU', 'gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core', 'Location','NorthWest');
+
+%plot( userCount, gtx580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
+plot( userCount, gtx580, 'kx-' , userCount, cpu1, 'k*-' , userCount, cpu2, 'ko-');
+legend('gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core','Location', 'NorthWest');
+
 xlabel('Number of training users');
 ylabel('Elapsed time (s)');
 
 set (h,'windowstyle','normal');               
-set (h,'Units','Inches');                    
+%set (h,'Units','Inches'); % matlab
+set (h,'Units','inches'); % octave
 windowSize = [0 0 3.75 2];
[…]

### 2011-06-05 main.tex
Applied fixes from Dan.  Sent to Allen.

@@
 On both architectures, the proposed implementations dramatically out-perform
-nieve library-based implementations. 
+naive library-based implementations. 
 At the problem sizes most important for $\ell_1$-min based
@@
 image-based face recognition as a sparse representation problem
-\cite{WrightJ2009-PAMI}.  If we stack the training images of $K$ subject
-classes $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$ into the
-columns of a matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, given a new
-query image in the same vector form $\bb\in\Re^d$, \emph{sparsity-based
+\cite{WrightJ2009-PAMI}.  If we stack the pixels of the training images of $K$ subject
+classes into the columns of matrices $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$, combine
+the matrices into a larger matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, and arrange the pixels of a new
+query image into a vector $\bb\in\Re^d$, \emph{sparsity-based
 classification} (SBC) solves the following minimization problem:
@@
 image, and the dominant nonzero coefficients of $\xx$ reveal the membership of
-$\bb$ based on the training image labels associated with $A$.  Since $A$ ideally
+$\bb$ based on the training image labels associated with $A$.  Since $A$
 contains many images per user taken under different illuminations, 
-$A\x$ also acts as a linear illumination model for the test image $\bb$.
+$A_i\x_i$ acts as a linear illumination model for the test image $\bb$.
 
@@
 program, conventional algorithms such as interior-point methods
-\cite{ChenS2001-SIAM,TibshiraniR1996} are known to suffer from poor scalability
-for large-scale applications such as face recognition. Recently, a number of
+\cite{ChenS2001-SIAM,TibshiraniR1996} scale poorly
+for applications such as face recognition. Recently, a number of
 accelerated algorithms have been proposed that explicitly take advantage of
@@
 that the query image is well aligned with the training images. In
-\cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align
-a query image to each subject class individually. The face alignment algorithm solves
+\cite{WagnerA2009-CVPR}, the framework of SBC was extended to align
+a query image to each subject class individually. The face alignment algorithm iteratively solves
 the following minimization problem:
@@
 of the alignment step, the recognition system still fails to achieve real-time
-recognition against datasets of hundreds or thousands of users.  In this paper,
+recognition against datasets of hundreds or thousands of subjects.  In this paper,
 in addition to accelerating the generic
 $\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
-also discuss how to properly accelerate the iterative face alignment step
+also discuss how to efficiently accelerate the iterative face alignment step
 \eqref{eq:l1min_alignment} on multi-core CPUs and GPUs.
@@
 while this paper focuses on the parallelism available in a single (shared memory) machine,
-applications targeting thousands of users or more, will likely require computation be
+applications targeting thousands of subjects or more, will likely require computation be
 distributed over a cluster or network of computers, and this is not addressed here.
-The source code of our implementations, which target NVIDIA's CUDA GPU architecture
-as well as the x86-64 architecture, will be made available for peer evaluation.
+The source code of our implementations, which target NVIDIA Fermi GPUs
+as well as Intel x86-64 CPUs, will be made available for peer evaluation.
 \begin{figure}
@@
 processors, and a PCI card with a single high-end GPU.  
-Recognition involving more than a few hundred users with contemporary hardware
+Recognition involving more than a few hundred subjects with contemporary hardware
 would require a server(or cluster) configuration with an expandable 
@@
 most of the transistors on a GPU are dedicated to arithmetic logic units.
-An explanation of the CUDA programming model at a useful level of completeness would take 
-more space than is available here, so we will instead frame our discussion in terms of 
+For our GPU implementations we target NVIDIA Fermi GPU architecture (\eg, the GTX 480
+GPU used in this paper).
+An explanation of the GPU programming model (CUDA) at a useful level of completeness would take more space than is available here, so we will instead frame our discussion in terms of
 hardware capabilities. \footnote{A discussion of the CUDA programming model can be found
 in the NVIDIA C for CUDA Programming Guide}
-
-Our target GPU architecture is the NVIDIA Fermi architecture (\eg, the GTX 480
-GPU used in this paper).  CUDA based GPUs are comprised of several
+CUDA programmable GPUs are comprised of several
 \emph{streaming multiprocessors} (SMP), each of which is roughly analogous to a CPU core.
@@
 in error $\ee$ after the alignment. Note that directly solving
-\eqref{eq:l1min_alignment} would require simultaneously estimating $\xx$,
-$\ee$, and $\tau$ typically on a nonlinear manifold, which renders the problem
-extremely difficult to solve exactly with many local minima.
-
-To mitigate the problem, if the system can obtain a good initial guess of the
-transformation $\tau$ (\eg, provided by a good face detector), the optimal
+\eqref{eq:l1min_alignment} is impossible since it is a non-convex problem
+and may exhibit local minima.
+However, given a good initial guess of the
+transformation $\tau$ (\eg, provided by an accurate face detector), the optimal
 solution for $\tau$ can be sought iteratively by linearization at each $j$th
-step as follows:
+step:
 \begin{equation}
@@
 image operations, \ie, at the pixel level.  We call this \emph{pixel-level
-parallelism}.  In the following two sections will will discuss how to
-exploit these two levels of parallelism on different CPU and GPU architectures.
+parallelism}.  The following two sections discuss methods for exploiting
+these two levels of parallelism on CPU and GPU architectures.
 
@@
 typically larger than 16.  
-In this paper we consider two mappings of the parallelism available in the alignment
-stage to the concurrency available on the CPU.  We argue that the latter
-implementation should achieve higher performance, and later verify that this is the case.
+In this paper we compare two mappings of the parallelism available in the alignment
+stage to the concurrency available on the CPU: a naive implementation that is purely 
+library-based, and the optimized version we advocate.
 
-The first implementation disregards 
-problem-level parallelism, and instead mapping the pixel-level parallelism
+The naive implementation disregards 
+problem-level parallelism, and instead maps pixel-level parallelism
 onto both the core-level and the SSE-level concurrency provided by the CPU.  In
@@
 image against the training images of a single user at a time using all cores on
-both processors.  The advantage of this implementation is that all of the variables
+both processors.  The potential advantage of this implementation is that all of the variables
 for the inner loop, most notably $B_j$ and $B_j^\dagger$, fit in L2 cache.
 
-The second implementation maps problem-level parallelism onto core-level
+The optimized implementation maps problem-level parallelism onto core-level
 concurrency, and maps pixel-level parallelism onto SSE-level concurrency.  In
@@
 problem instances running concurrently, the local variables for the inner
-loop still fit in L3 cache.  Thanks to the sequential data access patterns of
-the solver, prefetching significantly reduces the bandwidth difference between
-L2 and L3 cache.  For these reasons, we expect this implementation to out-perform
+loop still fit in L3 cache.  Because of the sequential data access patterns of
+the solver, the CPU hardware is able to use the full L3 cache bandwidth, which is 
+high enough to make the program CPU limited, rather than memory limited.  
+For these reasons, we expect this implementation to out-perform
 the previous implementation.
@@
 high bandwidth to DRAM.  This bandwidth is achieved by having a very large number of
-threads issuing memory requests simultaneously.  Solving many alignment problem instances
+threads issuing interleaved memory requests.  Solving many alignment problem instances
 concurrently increases the number of threads that can be used.
@@
 same purpose.
-\footnote{In CUDA terminology, Our proposed alignment stage
+\footnote{In CUDA terms, our proposed alignment stage
 implementation consists of a single kernel that performs alignment for all
@@
 concurrently as the hardware will allow.} The number of subject classes that
-actually get aligned concurrently is ultimately chosen by the GPU hardware, 
-but can be influenced by manual tuning of the code. We have empirically determined
+are actually scheduled to run concurrently is chosen by the GPU hardware, 
+but can be indirectly influenced by manual tuning of the code.\footnote{In CUDA terms, it is the kernel launch
+configuration that is being tuned.} We have empirically determined
 that performance is highest with 5-7 subject classes aligned simultaneously on each SMP.
@@
 Several other aspects of the implementation merit discussion.  
-The full resolution query image, along with its numerical derivatives in both
-image directions, are stored in GPU texture memory. This  allows the
-computation of $b(\tau)$ and $J(\tau)$, which essentially consist of warped
-versions of the test image and its derivatives, to exploit the GPU's special
-purpose hardware for bilinear interpolation.
-
-Solving one alignment problem per SMP necessitates the implementation of a
-custom CUDA routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$. $G$
+First of all, we take advantage of the GPU's special hardware dedicated to bilinear interpolation
+for the computation of $b(\tau)$ and $J(\tau)$, which essentially consist of resampled
+versions of the test image and its derivatives.
+Second of all, since there are no standard CUDA libraries that work at the SMP level, we use a
+custom routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$. $G$
 is inverted using Gaussian elimination with partial pivoting.  To achieve
@@
 cost of the inversion is dwarfed by the cost of other steps.  
-Additionally, reductions on long vectors (i.e. sum or $\ell_1$ norm
-of an image) are performed using a double precision temporary variable for accumulation.
+Similarly, reductions (i.e. sum or $\ell_1$ norm) on image-sized vectors are also
+performed in double precision.
+%The lack of a atomic floating point addition on the GPU
+%significantly reduces the perfomance of this step of the algorithm.
 %In Section \ref{sec:experiment}, we show the performance of our CPU and GPU
@@
 \label{sec:recognition}
-After the face alignment stage is complete, all but the 20 users with lowest
-$||\e_i||_1$ are thrown out, $\bb$ and $A_i$ are re-sampled to a common alignment
+After the face alignment stage is complete, the 20 subjects with lowest
+$||\e_i||_1$ are selected for recognition, $\bb$ and $A_i$ are re-sampled to a common alignment
 using $\tau_i$, and $A_i$ are concatenated into a new matrix $A$.
@@
 concurrency available in the hardware architecture.  Operations that do not map
-well onto the CUBLAS API were implemented as GPU kernels (in C for CUDA).
+well onto the CUBLAS API were implemented directly in CUDA code.
 Additionally, some operations that could have been implemented via multiple
@@
 GPU on data stored in GPU DRAM.  
-All non-scalar temporary variables are stored in GPU DRAM
-during solver execution. 
 %with the sole exception of the pseudo-inverse of $B_j$,
@@
 %chip).
-Aside from the exceptions already mentioned in Section
-\ref{sec:alignment_implementation_gpu}, all math operations are performed in
-single precision.
 
@@
 The first experiment compares the performance of our proposed CPU and GPU
-implementations of the general $\ell_1$-min (Algorithm \eqref{alg:alm}), given a
-generic random matrix $A$ and a random sparse vector $\xx_0$.  The measurement
-$d \times n$ matrix $A$ is generated as a random Gaussian matrix, with each
-entry randomly generated from the standard normal and normalized with unit
+implementations of the general $\ell_1$-min solver (Algorithm \eqref{alg:alm}).
+The $d \times n$ measurement matrix $A$ is a random Gaussian matrix, with each
+entry generated from the standard normal distribution and normalized to unit
 column norm.  The ratio of $d/n$ is fixed at $1/2$ with $n$ varying from 1000 to 8000.
-The ground truth signals, $\xx_0$ have a sparsity rate of 10\% of d with
-elements sampled from a normal distribution and normalized to unit $\ell_2$ vector
+The ground truth signal, $\xx_0$ has a sparsity rate of 10\% of d with
+elements sampled from a normal distribution and normalized to unit column
 norm.  Because the ground truth signal $\xx_0$ is known, the algorithm
@@
 
-First, in order to test our theory that the GPU performs best solving many
-$\ell_1$ problems-per concurrently, we benchmark three implementations of the
+First, in order to measure the impact of solving many
+$\ell_1$ problems-per concurrently on the GPU, we benchmark three implementations of the
 alignment $\ell_1$-min solver with $A$ of size $5120 \times 32$, and a fixed
@@
 \end{figure}
-The tiers seen in the GPU profile result from the GPU hardware scheduling alignment
+The tiers seen in the GPU profile result from the GPU hardware scheduling the computation of alignment
 problems in concurrent batches, but the overall trend is linear as expected.
@@
 %The following experiment fulfills the dual purpose of verifying the accuracy of
-%the alignment stage, and motivating a threshold on how many users are kept for
+%the alignment stage, and motivating a threshold on how many subjects are kept for
 %the recognition stage.  Setting this threshold establishes a compromise between
 %recognition speed and potential recognition rate.  To explore the affect of the
-%threshold on potential recognition rate, we vary the number of users kept for
+%threshold on potential recognition rate, we vary the number of subjects kept for
 %the recognition stage and estimate the probability of the test user being
-%among the set of training users kept for the recognition stage.  This curve
+%among the set of subjects kept for the recognition stage.  This curve
 %bounds above the achievable recognition rate of the overall pipeline.  As can
 %be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly
-%diminishing returns as the number of kept users is increased beyond 20; a small
+%diminishing returns as the number of kept subjects is increased beyond 20; a small
 %fraction of the gallery.
@@
 %\caption{Probability of the test user being kept for the recognition stage vs.
-%number of training users kept for the recognition stage.}
+%number of subjects kept for the recognition stage.}
 %\label{fig:user_alignment_rank_plot} \end{figure}
@@
 The next experiment compares the speed of the GPU and CPU implementations of
-the recognition stage.  It was determined that keeping 20 gallery users is
+the recognition stage.  It was determined that keeping 20 gallery subjects is
 sufficient to ensure that the correct user is kept for the recognition stage
 with 95\% probability.  Since recognition failures are typically caused by a
-poor alignment, we find that keeping more users for the recognition stage does
+poor alignment, we find that keeping more subjects for the recognition stage does
 not necessarily improve recognition rate.
@@
 earlier implementations required over a second to align a test image against
-just one of the 250 users in Multi-PIE session 1, the proposed 
+just one of the 250 subjects in Multi-PIE session 1, the proposed 
 implementations can perform the same task in 
@@
 parallelization techniques presented here make it feasible to implement an
-access control system for hundreds of users on a handful of purpose-build
+access control system for hundreds of subjects on a handful of purpose-build
 multi-GPU workstations.  As CPU manufacturers 

### 2011-06-05 figures/alignment_runtime_graph.m
New figure 4, no markers.

@@ -19,12 +19,12 @@ gtx580 = gtx580(1:2:end,2);
 close all;
 h = figure(7);
 
-%plot( userCount, gtx480, 'k-.', userCount, gtx580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
-%legend('gtx480 GPU', 'gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core', 'Location','NorthWest');
+plot( userCount, gtx480, 'k-', userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
+legend('GPU', 'CPU, library threading', 'CPU, manual threading', 'Location','NorthWest');
 
-%plot( userCount, gtx580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
-plot( userCount, gtx580, 'kx-' , userCount, cpu1, 'k*-' , userCount, cpu2, 'ko-');
-legend('gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core','Location', 'NorthWest');
+%plot( userCount, GTX 580, 'k-' , userCount, cpu1, 'k:' , userCount, cpu2, 'k--');
+%plot( userCount, GTX 580, 'kx-' , userCount, cpu1, 'k*-' , userCount, cpu2, 'ko-');
+%legend('gtx580 GPU', '2 x E5530 CPU, single instance', '2 x E5530 CPU, 1 instance per core','Location', 'NorthWest');
 
 xlabel('Number of training users');
 ylabel('Elapsed time (s)');
@@ -33,8 +33,8 @@ set (h,'windowstyle','normal');
 %set (h,'Units','Inches'); % matlab
 set (h,'Units','inches'); % octave
 windowSize = [0 0 3.75 2];
-windowSize = windowSize*0.8;
-%windowSize = windowSize*1.5;
+%windowSize = windowSize*0.8;
+windowSize = windowSize*1.2;
 %windowSize = windowSize*4;
 %windowSize = [0 0 20 15];
 set(h,'position', windowSize);

### 2011-06-05 main.tex
Merge changes from Allen

@@
 $\ell_1$-min based on augmented Lagrangian method (ALM) solvers. 
-Due to the low arithmetic intensity of ALM solvers, both
-architectures are bandwidth limited.  
-with large problems running faster on the GPU, and small problems running faster
+Due to the low arithmetic intensity of ALM solvers, implementations on both
+architectures are primarily limited by memory bandwidth,  
+with large problems running faster on the GPU and small problems running faster
 on the CPU.  
-On both architectures, the proposed implementations dramatically out-perform
+On both architectures, the proposed implementations significantly outperform
 naive library-based implementations. 
@@
 algorithms will be made available for peer evaluation.  \end{abstract}
+\vspace{-0.1in}
 
 \section{Introduction} 
+\vspace{-0.05in}
 $\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
@@
 image, and the dominant nonzero coefficients of $\xx$ reveal the membership of
-$\bb$ based on the training image labels associated with $A$.  Since $A$
-contains many images per user taken under different illuminations, 
-$A_i\x_i$ acts as a linear illumination model for the test image $\bb$.
+$\bb$ based on the training image labels associated with $A$. 
+%Since $A$ contains many images per user taken under different illuminations, 
+%$A_i\x_i$ acts as a linear illumination model for the test image $\bb$.
 
@@
 
-While solving \eqref{eq:l1min_denoise} can achieve exceptional
-recognition accuracy on public databases, its success also relies on the assumption
-that the query image is well aligned with the training images. In
-\cite{WagnerA2009-CVPR}, the framework of SBC was extended to align
-a query image to each subject class individually. The face alignment algorithm iteratively solves
-the following minimization problem:
+In addition, the accuracy of the solution given by \eqref{eq:l1min_denoise} relies on the assumption
+that the query image is well aligned with the training images.
+In \cite{WagnerA2009-CVPR}, SBC was extended to align
+a query image to each subject class individually. The alignment algorithm solves
+the following problem:
 \begin{equation}
@@
 
-Numerically, the original solution \cite{WagnerA2009-CVPR} was implemented based 
-on standard interior-point methods, and its speed has been since improved by ALM in a more recent paper
-\cite{WagnerA2011-PAMI}. However, due the high per-class cost
+The $ell_1$ solvers in the original system prototype were based 
+on standard interior-point methods \cite{WagnerA2009-CVPR}. Their speed has been improved by ALM in a more recent paper \cite{WagnerA2011-PAMI}. However, due the high per-class cost
 of the alignment step, the recognition system still fails to achieve real-time
@@
 {\tiny \input{figures/pipeline_simplified.tex}}
-\caption{The face recognition pipeline.}
+\caption{The face recognition pipeline.\vspace{-0.1in}}
 \label{fig:pipeline}
@@
 
-\subsection{Literature Review} 
+\subsection{Literature Review} \vspace{-0.05in}
 Traditionally, $\ell_1$-min (a.k.a.
@@
 \label{sec:ALM}
-
+\vspace{-0.05in}
 In this section, we briefly describe the ALM algorithm for $\ell_1$-min
@@
 \caption{Augmented Lagrangian Method (ALM)}
+\small
 {\bf INPUT:} $\bb \in \Re^m$, $A=[A_1,\cdots, A_K] \in \Re^{m \times n}$, $\tau\leftarrow \max\mbox{eig}(A^TA)$, and constant $\rho>1$.
@@
 
+\vspace{-0.05in}
 \section{Hardware Concurrency} \label{sec:concurrency}
-
+\vspace{-0.05in}
 In this section we discuss the levels of concurrency available in the hardware
@@
 Recognition involving more than a few hundred subjects with contemporary hardware
-would require a server(or cluster) configuration with an expandable 
+would require a server (or cluster) configuration with an expandable 
 number of processors.  We will not address efficient parallelization for 
 these systems, which may have additional challenges associated with
-their non-shared memory model.\footnote{Non-shared memory configurations
-include CPU blade servers, multi-GPU servers, and clusters}.
+their non-shared memory model.
+%\footnote{Non-shared memory configurations include CPU blade servers, multi-GPU servers, and clusters}.
 Unless otherwise specified, all implementations utilize single precision
@@
 \subsection{CPU Hardware Concurrency}
+\vspace{-0.05in}
 \label{sec:CPU-concurrency}
@@
 are that they have two levels of concurrency, relatively large amounts of
-cache, and relatively high clock speeds.
-
-The baseline architecture for our experiments is a Linux workstation with two
+cache, and relatively high clock speeds. The baseline architecture for our experiments 
+is a Linux workstation with two
 quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
 has its own memory interface, and is directly interfaced to half of the RAM
-installed in the machine.  The amount of RAM installed exceeds the amount used
+installed in the machine.  The amount of RAM exceeds the amount used
 by the algorithms, and is not an important performance consideration.
@@
 caches.  Note: Although the aspect ratio could not be preserved, all arrays and
-caches are drawn so that the area is proportional to the size of the data. }
+caches are drawn so that the area is proportional to the size of the data. } \vspace{-0.1in}
 \end{figure}
@@
 There are thus two important levels of concurrency that need to be exploited to
-efficiently use a modern CPU: {\em core-level} concurrency and {\em sse-level}
+efficiently use a modern CPU: {\em core-level} concurrency and {\em SSE-level}
 concurrency.
 
-\subsection{GPU Hardware Concurrency}
+\subsection{GPU Hardware Concurrency} \vspace{-0.05in}
 The main defining characteristics of contemporary multi-core GPU architectures
@@
 An explanation of the GPU programming model (CUDA) at a useful level of completeness would take more space than is available here, so we will instead frame our discussion in terms of
-hardware capabilities. \footnote{A discussion of the CUDA programming model can be found
-in the NVIDIA C for CUDA Programming Guide}
+hardware capabilities. 
+%\footnote{A discussion of the CUDA programming model can be found in the NVIDIA C for CUDA Programming Guide}
 CUDA programmable GPUs are comprised of several
@@
 the SIMD units of each SMP is somewhat more flexible than on the CPU,
-leveraging the flexibility typically comes at the cost of reduced concurrency
-.\footnote{In CUDA terms, code branches that cause warp divergence usually
+leveraging the flexibility typically comes at the cost of reduced concurrency.
+\footnote{In CUDA terms, code branches that cause warp divergence usually
 result in serialization of the different groups of threads}  Therefore, for
@@
 
-\section{Parallelism in the Face Alignment Stage}
+\section{Parallelism in the Face Alignment Stage}\vspace{-0.05in}
 \label{sec:alignment}
@@
 \caption{\bf (Face Alignment via ALM)} \label{alg:iterative_alignment}
+\small
 {\bf Input:} $\bb$, $A_i$, $\x_0 = \mathbf{0}$, $\tau_0$, and $J_0$.
@@
 
-\subsection{CPU Implementation} 
+\subsection{CPU Implementation} \vspace{-0.05in}
 \label{sec:alignment_implementation_cpu}
@@
 
-\subsection{GPU Implementation}
+\subsection{GPU Implementation} \vspace{-0.05in}
 \label{sec:alignment_implementation_gpu}
@@
 
-\section{Parallelization of the Face Recognition Stage}
+\section{Parallelization of the Face Recognition Stage} \vspace{-0.05in}
 \label{sec:recognition}
@@
 
-\subsection{Recognition Stage Implementation} Compared to problem-level
+\subsection{Recognition Stage Implementation} \vspace{-0.05in}
+Compared to problem-level
 parallelism, the exclusively pixel-level parallelism in Algorithm \ref{alg:alm_rec} is
@@
 
-\section{Experiments} \label{sec:experiment} In this section, we benchmark the
+\section{Experiments} \vspace{-0.05in}
+\label{sec:experiment} 
+In this section, we benchmark the
 performance of our parallel implementations of $\ell_1$-min on CPU and GPU
@@
 \begin{center}
-\includegraphics[width=3.5in]{figures/time_vs_matrix_size_constant_tol}
-\end{center}
-\caption{Comparison of $\ell_1$-min runtime vs. dimensions of $A$ on random data.}
+\includegraphics[width=2.5in]{figures/time_vs_matrix_size_constant_tol}
+\end{center} \vspace{-0.1in}
+\caption{\small Comparison of $\ell_1$-min runtime vs. dimensions of $A$ on random data.} \vspace{-0.1in}
 \label{fig:random_data}
@@
 
-\subsection{Face Recognition Pipeline Benchmark} 
+\subsection{Face Recognition Pipeline Benchmark} \vspace{-0.05in}
 \label{sec:benchmark}
 This section presents benchmarks of our CPU and GPU implementations of the
-alignment stage of the face recognition pipeline, as shown in Algorithm
-\ref{alg:iterative_alignment}, as well as the recognition stage of the
-face recognition pipeline, as seen in Algorithm \ref{alg:alm_rec}.
+alignment stage of the face recognition pipeline (i.e., Algorithm
+\ref{alg:iterative_alignment}), as well as the recognition stage of the
+face recognition pipeline (i.e., Algorithm \ref{alg:alm_rec}).
 
@@
 problem at a time using the stock BLAS libraries.
-\footnote{Note that the streams implementation is limited significantly by a 
+\footnote{The streams implementation is limited significantly by a 
 cap on the number of concurrent streams in the current version of CUDA}
 %The results are shown below:
-\begin{table}[h!]
-\caption{Benchmark of implementations of alignment $\ell_1$-min.}
+\begin{table}[t!]
+\caption{\small Benchmark of implementations of alignment $\ell_1$-min.}
+\small
 \begin{tabular}{|l|c|}
@@
 \hline
-\end{tabular}
+\end{tabular} \vspace{-0.1in}
 \label{tbl:ubench}
@@
 the canonical frame. The training images are aligned by manually clicking the
-two outer eye corners in the camera frame, and then warping the images via a
+two outer eye corners in the camera frame, and then warped via a
 similarity transformation such that the eye corners lie at a pair of canonical
-eye locations in the canonical frame.  The canonical eye corners are chosen
-such that for well aligned images, the canonical window contains exclusively
-pixels from the subject's face.  We assume the face detector is accurate, so
-for the test images we use clicked eye corners for test image alignment
-initialization.\footnote{Previous publications have shown that iterative
-alignment improves recognition accuracy substantially even with clicked eye
-corners}
+eye locations in the canonical frame.
 
 Figure \ref{fig:alignment_stage_runtime} shows the average run time of the CPU and GPU implementations to align
-a query image against all the subject classes separately. We vary the number of
-subject classes to show how the algorithms scale. 
-\begin{figure}
+a query image against all the subject classes separately. We vary the total number of
+subject classes to show how the algorithms scale. The plateaus seen in the GPU curve
+result from the GPU hardware scheduling the computation of alignment
+problems in concurrent batches, but the overall trend is linear as expected.
+The manually threaded CPU implementation slightly outperforms the GPU implementation,
+and it beats the \emph{naive}
+library threaded implementation by a wide margin.
+Compared to previously published results, the new implementation can align the query
+image with 250 subject classes in under 10 seconds, while the best previous effort
+required 1.7 seconds per subject class (i.e. 253 seconds for 250 subjects).
+
+\begin{figure}[t!]
 \centering
-\includegraphics[width=3.4in]{figures/alignment_runtime_graph}
-\caption{Alignment stage runtime vs. size of training database.}
+\includegraphics[width=3in]{figures/alignment_runtime_graph}\vspace{0.05in}
+\caption{Alignment stage runtime vs. size of training database.} \vspace{-0.1in}
 \label{fig:alignment_stage_runtime}
 \end{figure}
-The tiers seen in the GPU profile result from the GPU hardware scheduling the computation of alignment
-problems in concurrent batches, but the overall trend is linear as expected.
-The CPU implementation that solves one problem instance per core out-performs
-the single-instance-at-a-time implementation by a wide margin, and is even competitive
-with the GPU implementation.\footnote{Note that the GPU implementation was tuned
-to be fast for the full problem set; tuning affects the locations of the tiers,
-so the lower corners of the tiers are most relevant.}
-
-%The following experiment fulfills the dual purpose of verifying the accuracy of
-%the alignment stage, and motivating a threshold on how many subjects are kept for
-%the recognition stage.  Setting this threshold establishes a compromise between
-%recognition speed and potential recognition rate.  To explore the affect of the
-%threshold on potential recognition rate, we vary the number of subjects kept for
-%the recognition stage and estimate the probability of the test user being
-%among the set of subjects kept for the recognition stage.  This curve
-%bounds above the achievable recognition rate of the overall pipeline.  As can
-%be seen in Figure \ref{fig:user_alignment_rank_plot}, there are rapidly
-%diminishing returns as the number of kept subjects is increased beyond 20; a small
-%fraction of the gallery.
-%\begin{figure} 
-%\centering
-%\includegraphics[width=3.4in]{figures/user_alignment_rank_plot}
-%\caption{Probability of the test user being kept for the recognition stage vs.
-%number of subjects kept for the recognition stage.}
-%\label{fig:user_alignment_rank_plot} \end{figure}
 
@@
 not necessarily improve recognition rate.
-%The inner loop tolerance to 1e-5 and outer loop tolerance to 1e-2.  
-In Figure \ref{fig:recognition_stage_runtime} we show recognition stage runtime
+
+Figure \ref{fig:recognition_stage_runtime} shows the recognition stage runtime
 for canonical images of size: $32\times32$, $48 \times 48$, $64 \times 64$, $96 \times
-96$, and $128 \times 128$.  For all image resolutions the problem size
+96$, and $128 \times 128$.  For all image resolutions, the problem size
 is sufficiently small that the CPU is significantly faster than the GPU.  Note
@@
 than the alignment stage.
-\begin{figure} 
+\begin{figure}[t!]
 \centering
-\includegraphics[width=3.4in]{figures/speedVsResolution}
-\caption{Recognition stage runtime vs. size of training database.}
+\includegraphics[width=2.5in]{figures/speedVsResolution} \vspace{-0.05in}
+\caption{Recognition stage runtime vs. face window resolution}\vspace{-0.1in}
 \label{fig:recognition_stage_runtime} \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics[width=3.4in]{figures/accuracyVsResolution}
-\caption{Recognition rate of the full pipeline vs. face window resolution.}
-\label{fig:accuracy_vs_resolution}
-\end{figure}
 Finally, we perform an experiment verifying the recognition accuracy of the
-overall pipeline.  As can be seen in Figure \ref{fig:accuracy_vs_resolution},
+overall pipeline.  As shown in Figure \ref{fig:accuracy_vs_resolution},
 at the optimal resolution, which happens to match the alignment stage
@@
 achievable given the alignment results.  For significantly lower resolutions,
-the accuracy drops off significantly.
+the accuracy drops off significantly.  The slight difference in CPU vs. GPU
+accuracy may be a result of numerical precision differences in our matrix
+inversion and vector reduction routines.
+\begin{figure}[t!]
+\centering
+\includegraphics[width=2.5in]{figures/accuracyVsResolution} \vspace{-0.05in}
+\caption{Recognition rate of the full pipeline vs. face window resolution.} \vspace{-0.1in}
+\label{fig:accuracy_vs_resolution}
+\end{figure}
 
-\section{Conclusion}
+\section{Conclusion}\vspace{-0.05in}
 We have demonstrated that on both CPU and GPU algorithms, parallelizations of
@@
 of the hardware, we have demonstrated dramatic improvements in recognition
-speed over previously reported implementations of the same pipeline.  Whereas
-earlier implementations required over a second to align a test image against
-just one of the 250 subjects in Multi-PIE session 1, the proposed 
-implementations can perform the same task in 
-{\bf under 10 seconds}!.  In principle, the
-parallelization techniques presented here make it feasible to implement an
-access control system for hundreds of subjects on a handful of purpose-build
-multi-GPU workstations.  As CPU manufacturers 
-increase the number of cores and vector widths, and as GPU manufacturers increase the amount
+speed over previously reported implementations.
+As CPU manufacturers increase the number of cores and vector widths, and as GPU 
+manufacturers increase the amount
 of cache, both architectures are rapidly converging towards an architectural balance
-that increasingly favorable for $ell_1$-min based face recognition.
+that increasingly favorable for $\ell_1$-min based face recognition.
+% of the same pipeline.  Whereas
+%earlier implementations required over a second to align a test image against
+%just one of the 250 subjects in Multi-PIE session 1, the proposed 
+%implementations can perform the same task in 
+%{\bf under 10 seconds}!.  
+%The parallelization techniques presented here make it feasible to implement an
+%access control system for hundreds of subjects on a handful of purpose-build
+%multi-GPU workstations.
 

### 2011-06-05 old.tex
Added Allen's changes.  Submitting this version to be safe.

--- 2011-06-05 main.tex
@@
 $\ell_1$-min based on augmented Lagrangian method (ALM) solvers. 
-Due to the low arithmetic intensity of ALM solvers, implementations on both
-architectures are primarily limited by memory bandwidth,  
-with large problems running faster on the GPU and small problems running faster
-on the CPU.  
-On both architectures, the proposed implementations significantly outperform
-naive library-based implementations. 
+For very large problems the GPU is faster to higher memory bandwidth, while 
+for problems that fit in the larger CPU L3 cache, the CPU is faster.
 At the problem sizes most important for $\ell_1$-min based
 face recognition, both architectures achieve comparable performance.
+On both architectures, the proposed implementations significantly outperform
+naive library-based implementations, as well as published results. 
 The source code of the
@@
 library threaded implementation by a wide margin.
-Compared to previously published results, the new implementation can align the query
-image with 250 subject classes in under 10 seconds, while the best previous effort
-required 1.7 seconds per subject class (i.e. 253 seconds for 250 subjects).
+The new implementation can align the query
+image in $\approx 40$ ms per subject, 
+while the fastest previously published result \cite{WagnerA2011-PAMI}
+required $\approx 600$ ms seconds per subject in a similar setting.
 

### 2011-06-05 main.tex
Fixed some awkward sentences, make fig. 4 bigger.

@@
 %\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
-\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
+\title{\vspace{-0.2in} Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
 %\title{Parallelization of Fast $\ell_1$-Minimization Towards Real-Time Face Recognition}
@@
 $\ell_1$-min based on augmented Lagrangian method (ALM) solvers. 
-Due to the low arithmetic intensity of ALM solvers, implementations on both
-architectures are primarily limited by memory bandwidth,  
-with large problems running faster on the GPU and small problems running faster
-on the CPU.  
+For very large problems, the GPU is faster due to higher memory bandwidth, while 
+for problems that fit in the larger CPU L3 cache, the CPU is faster.
 On both architectures, the proposed implementations significantly outperform
-naive library-based implementations. 
-At the problem sizes most important for $\ell_1$-min based
-face recognition, both architectures achieve comparable performance.
+naive library-based implementations, as well as previous systems. 
 The source code of the
@@
 \cite{LorisI2009,YangA2010-ICIP}. We investigate parallelization of a
-state-of-the-art $\ell_1$-min solution based on the classical framework of
+state-of-the-art $\ell_1$-min solution based on
 \emph{augmented Lagrangian methods} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.
@@
 
-The $ell_1$ solvers in the original system prototype were based 
-on standard interior-point methods \cite{WagnerA2009-CVPR}. Their speed has been improved by ALM in a more recent paper \cite{WagnerA2011-PAMI}. However, due the high per-class cost
-of the alignment step, the recognition system still fails to achieve real-time
-recognition against datasets of hundreds or thousands of subjects.  In this paper,
+Although ALM has been applied to improve the speed of a prototype face recognition system \cite{WagnerA2011-PAMI},
+due to the high per-class cost
+of the alignment step, the system still fails to achieve \emph{real-time}
+performance against datasets of hundreds or thousands of subjects.  In this paper,
 in addition to accelerating the generic
 $\ell_1$-min objectives \eqref{eq:l1min} and \eqref{eq:l1min_denoise}, we 
-also discuss how to efficiently accelerate the iterative face alignment step
+also discuss how to efficiently accelerate the face alignment step
 \eqref{eq:l1min_alignment} on multi-core CPUs and GPUs.
@@
 while this paper focuses on the parallelism available in a single (shared memory) machine,
-applications targeting thousands of subjects or more, will likely require computation be
-distributed over a cluster or network of computers, and this is not addressed here.
+applications targeting thousands of subjects or more will likely require computation to be
+distributed over a cluster or network of computers. These systems are not addressed here.
 The source code of our implementations, which target NVIDIA Fermi GPUs
@@
 
-One of the drawbacks of most interior-point methods for $\ell_1$-min (including
-BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
+One of the drawbacks of most interior-point methods for $\ell_1$-min is that they require the solution sequence to follow an
 interior path via gradient descent or conjugate gradient methods, which are computationally expensive.
 To mitigate these issues, an approach called \emph{Homotopy} has been recently studied to accelerate the
-speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were
-first studied in the context of LASSO \cite{OsborneM2000}, which inspired a
-solution to the \emph{forward stagewise linear regression} problem called LARS
-\cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms
-for BP in \cite{MalioutovD2005,DonohoD2006}.
-
+speed of $\ell_1$-min \cite{OsborneM2000,EfronB2004,MalioutovD2005,DonohoD2006}.
 Although Homotopy can be shown to exactly estimate BP when the solution is
@@
 matrices.
+
+%Homotopy methods for recovering sparse signals were
+%first studied in the context of LASSO \cite{OsborneM2000}, which inspired a
+%solution to the \emph{forward stagewise linear regression} problem called LARS
+%\cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms
+%for BP in \cite{MalioutovD2005,DonohoD2006}.
+
 %\footnote{During the face alignment algorithm presented later, The $A$ matrix changes between alignment iterations but not within the L1 solver}  
@@
 authors showed that ALM is able to significantly improve the solver speed,
-while achieving estimation accuracy competitive with exact $\ell_1$-min
-solutions such as BP and Homotopy methods, and other IST-type
-algorithms. Therefore, we choose ALM as the core algorithm for
+while achieving estimation accuracy competitive with other $\ell_1$-min
+solutions. Therefore, we choose ALM as the core algorithm for
 implementation of $\ell_1$-min in the parallel face recognition pipeline.
@@
 \end{equation}
-where $\mu_{k}\rightarrow \infty$ is monotonically increasing.
+where $\mu_{k}\rightarrow \infty$ increases monotonically.
 The iteration terminates when the estimates $(\xx_k, \yy_k)$ converge.
@@
 most of the transistors on a GPU are dedicated to arithmetic logic units.
-For our GPU implementations we target NVIDIA Fermi GPU architecture (\eg, the GTX 480
+For our GPU implementations, we target NVIDIA Fermi GPU architecture (\eg, the GTX 480
 GPU used in this paper).
@@
 the SIMD units of each SMP is somewhat more flexible than on the CPU,
-leveraging the flexibility typically comes at the cost of reduced concurrency.
-\footnote{In CUDA terms, code branches that cause warp divergence usually
+leveraging the flexibility typically comes at the cost of reduced concurrency.\footnote{
+In CUDA terms, code branches that cause warp divergence usually
 result in serialization of the different groups of threads}  Therefore, for
@@
 in error $\ee$ after the alignment. Note that directly solving
-\eqref{eq:l1min_alignment} is impossible since it is a non-convex problem
+\eqref{eq:l1min_alignment} is inefficient since it is a non-convex problem
 and may exhibit local minima.
@@
 typically larger than 16.  
-In this paper we compare two mappings of the parallelism available in the alignment
+In this paper, we compare two mappings of the parallelism available in the alignment
 stage to the concurrency available on the CPU: a naive implementation that is purely 
-library-based, and the optimized version we advocate.
+based on stock libraries and compilers, and a manually optimized version we advocate.
 
@@
 onto both the core-level and the SSE-level concurrency provided by the CPU.  In
-other words, alignment is performed of the query
-image against the training images of a single user at a time using all cores on
-both processors.  The potential advantage of this implementation is that all of the variables
+other words, alignment of the query image is performed
+against a single subject at a time using all available cores on
+the processors.  The potential advantage of this implementation is that all of the variables
 for the inner loop, most notably $B_j$ and $B_j^\dagger$, fit in L2 cache.
 
-The optimized implementation maps problem-level parallelism onto core-level
+In contrast, our manually optimized implementation maps problem-level parallelism onto core-level
 concurrency, and maps pixel-level parallelism onto SSE-level concurrency.  In
@@
 high enough to make the program CPU limited, rather than memory limited.  
-For these reasons, we expect this implementation to out-perform
-the previous implementation.
+For these reasons, this implementation outperforms the previous naive solution.
 
@@
 While on the CPU, algorithm performance is highly dependent on effective
-cache usage, on the GPU cache is relatively unimportant for ALM based
+cache usage, cache is relatively unimportant on the GPU for ALM based
 $\ell_1$-min.  As can be seen in Figure \ref{fig:caches}, the GPU has 
@@
 into L2 cache, this would not be an efficient use of the GPU's resources.
-The strength of the GPU's memory architecture for our purposes is it's ability to sustain a very
+The strength of the GPU's memory architecture for our purposes is its ability to sustain a very
 high bandwidth to DRAM.  This bandwidth is achieved by having a very large number of
@@
 multiple instances concurrently, and on the GPU we use multiple SMP's for the 
-same purpose.
-\footnote{In CUDA terms, our proposed alignment stage
+same purpose.\footnote{In CUDA terms, our proposed alignment stage
 implementation consists of a single kernel that performs alignment for all
@@
 are actually scheduled to run concurrently is chosen by the GPU hardware, 
-but can be indirectly influenced by manual tuning of the code.\footnote{In CUDA terms, it is the kernel launch
-configuration that is being tuned.} We have empirically determined
+but can be indirectly influenced by manual tuning of the code (i.e., the kernel launch
+configuration in CUDA terms). We have empirically determined
 that performance is highest with 5-7 subject classes aligned simultaneously on each SMP.
@@
 versions of the test image and its derivatives.
-Second of all, since there are no standard CUDA libraries that work at the SMP level, we use a
-custom routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$. $G$
-is inverted using Gaussian elimination with partial pivoting.  To achieve
+Second, since there are no standard CUDA libraries that work at the SMP level, we use a
+custom routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$, with $G$
+inverted using Gaussian elimination with partial pivoting.  To achieve
 precision comparable to the single precision LAPACK routines in Intel's MKL
@@
 cost of the inversion is dwarfed by the cost of other steps.  
-Similarly, reductions (i.e. sum or $\ell_1$ norm) on image-sized vectors are also
+Similarly, sums and dot-products of large vectors are also
 performed in double precision.
@@
 \label{sec:recognition}
-After the face alignment stage is complete, the 20 subjects with lowest
-$||\e_i||_1$ are selected for recognition, $\bb$ and $A_i$ are re-sampled to a common alignment
+After the face alignment stage is complete, the 20 subject classes with lowest
+alignment error are selected for recognition, $\bb$ and $A_i$ are re-sampled to a common alignment
 using $\tau_i$, and $A_i$ are concatenated into a new matrix $A$.
@@
 terminates when $\|\bx-\bx_0\| < \tau$ with $\tau=10^{-3}$.  The measurement
-vector is generated with the equation $\bb_0 = A \xx_0$.
+vector is generated by $\bb = A \xx_0$.
 
 The results of this benchmark can be found in Figure \ref{fig:random_data}.
-The x-axis represents the size of the $A$ matrix and the y-axis represents the
-average amount of time it to complete a single problem instance.  The GPU
+The $x$-axis represents the size of the $A$ matrix and the $y$-axis represents the
+average amount of time to complete a single problem instance.  The GPU
 implementation tends to be faster than the CPU implementation at solving a
 single large problem, whereas the CPU implementation is faster at solving a
-single small problem.  The transition between the two regimes occurs where
+single small problem.  The transition between the two regimes occurs (at 2000) where
 the problem size begins to exceed the size of the CPU L3 cache. 
@@
 \label{sec:benchmark}
-This section presents benchmarks of our CPU and GPU implementations of the
-alignment stage of the face recognition pipeline (i.e., Algorithm
-\ref{alg:iterative_alignment}), as well as the recognition stage of the
-face recognition pipeline (i.e., Algorithm \ref{alg:alm_rec}).
+This section presents benchmarks of the CPU and GPU implementations of the
+alignment stage (i.e., Algorithm
+\ref{alg:iterative_alignment}) as well as the recognition stage (i.e., Algorithm \ref{alg:alm_rec}) of the
+face recognition pipeline.
 
@@
 data we use subsets of the CMU Multi-PIE Face Database.  For gallery images we
-use frontal images from session 1, which contains 20 images per user taken
+use frontal images from session 1, which contains 20 images per subject taken
 under different illuminations. For test images we use images from session 2.
 The training images are prepared as follows: The iterative alignment stage
-seeks a similarity transformation between the coordinate frame of the "camera"
-resolution test image and the coordinate frame of the "canonical" resolution
-images which the images are compared. We use a $64 \times 64$ pixel window in
-the canonical frame. The training images are aligned by manually clicking the
-two outer eye corners in the camera frame, and then warped via a
-similarity transformation such that the eye corners lie at a pair of canonical
-eye locations in the canonical frame.
+seeks a similarity transformation between the coordinate frame of the
+full-resolution test image and a ``canonical'' coordinate frame in which images
+are compared.  The training images are aligned by applying a similarity
+transormation that maps two manually clicked outer eye corners to fixed
+locations in the canonical frame.  A $64 \times 64$
+pixel window in the canonical frame is used for resampling.
 
@@
 \centering
-\includegraphics[width=3in]{figures/alignment_runtime_graph}\vspace{0.05in}
+\includegraphics[width=3.3in]{figures/alignment_runtime_graph}\vspace{0.05in}
 \caption{Alignment stage runtime vs. size of training database.} \vspace{-0.1in}
@@
 the recognition stage.  It was determined that keeping 20 gallery subjects is
-sufficient to ensure that the correct user is kept for the recognition stage
+sufficient to ensure that the correct subject is kept for the recognition stage
 with 95\% probability.  Since recognition failures are typically caused by a
@@
 resolution, the GPU implementation reaches 95\% recognition rate, the max
-achievable given the alignment results.  For significantly lower resolutions,
+achievable given the alignment selection process.  For significantly lower resolutions,
 the accuracy drops off significantly.  The slight difference in CPU vs. GPU

### 2011-06-05 main.tex
Applied fixes from parents, and Ray

@@
 image-based face recognition as a sparse representation problem
-\cite{WrightJ2009-PAMI}.  If we stack the pixels of the training images of $K$ subject
+\cite{WrightJ2009-PAMI}.  If we stack the $d$ pixels of the training images of $K$ subject
 classes into the columns of matrices $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$, combine
@@
 
-Finally, we present extensive benchmark to compare the performance of the
+Finally, we present extensive benchmarks to compare the performance of the
 generic ALM $\ell_1$-min algorithm and the corresponding face recognition
@@
 configuration for engineering workstations: a motherboard with two quad-core
-processors, and a PCI card with a single high-end GPU.  
+processors and a PCIe card with a single high-end GPU.  
 Recognition involving more than a few hundred subjects with contemporary hardware
@@
 CUDA programmable GPUs are comprised of several
-\emph{streaming multiprocessors} (SMP), each of which is roughly analogous to a CPU core.
-For Fermi architecture GPU's, there are up to 16 SMP's, and each SMP is capable of executing up to 64 single precision 
+\emph{streaming multiprocessors} (SM), each of which is roughly analogous to a CPU core.
+For Fermi architecture GPUs there are up to 16 SMs, and each SM is capable of executing up to 64 single precision 
 floating point operations
@@
 % 48 images
-Each SMP contains its own L1-cache, which is divided between hardware managed
-and software managed ("shared") memory.  Additionally, all SMPs share a common
-L2-cache.  For our system, each SMP has 64 KiB of L1 cache, and all SMP's share
+Each SM contains its own L1-cache, which is divided between hardware managed
+and software managed (``shared") memory.  Additionally, all SMs share a common
+L2-cache.  For our system, each SM has 64 KiB of L1 cache, and all SMs share
 768KB of L2 cache. 
@@
 The GPU has its own memory system, and any data the GPU uses must first be
-transferred from CPU DRAM to GPU DRAM over PCI-Express.  For our application,
+transferred from CPU DRAM to GPU DRAM over PCIe.  For our application,
 this transfer overhead can be amortized over a large amount of computation and is
@@
 While the programming model for
-the SIMD units of each SMP is somewhat more flexible than on the CPU,
+the SIMD units of each SM is somewhat more flexible than on the CPU,
 leveraging the flexibility typically comes at the cost of reduced concurrency.\footnote{
@@
 roughly analogous to the CPU SIMD units. Thus, in summary, the GPU hardware 
-provides two levels of concurrency: {\em SMP-level}
+provides two levels of concurrency: {\em SM-level}
 concurrency and {\em thread-level} concurrency.  Note that the GPU provides more
-concurrency than a CPU at both levels (14 SMP's vs. 8 cores) and (64 wide vs. 4 wide SIMD
+concurrency than a CPU at both levels (14 SMs vs. 8 cores) and (64 wide vs. 4 wide SIMD
 units).
@@
 
-Given the hardware concurrency in modern parallel computing architectures,
-next, we shall study the optimal mappings of the algorithm parallelism in face
-recognition to the multi-core CPUs and GPUs, respectively.  In this section, we
-first focus on the face alignment stage (see Figure \ref{fig:pipeline}).
+We will next discuss the parallelism available in the face recognition
+pipeline, and propose techniques for exploiting this parallelism on the GPU and CPU hardware 
+described in the previous section.  In this section we first focus on the
+face alignment stage (see Figure \ref{fig:pipeline}).
 
@@
 and may exhibit local minima.
-However, given a good initial guess of the
+However, given a good initial estimate of the
 transformation $\tau$ (\eg, provided by an accurate face detector), the optimal
@@
 cache usage, cache is relatively unimportant on the GPU for ALM based
-$\ell_1$-min.  As can be seen in Figure \ref{fig:caches}, the GPU has 
+$\ell_1$-min.  As can be seen in Figure \ref{fig:gpu_caches}, the GPU has 
 a very small amount of cache compared to the CPU.  While it might be possible
@@
 to our recommended CPU implementation: on the CPU we use the cores to solve
-multiple instances concurrently, and on the GPU we use multiple SMP's for the 
+multiple instances concurrently, and on the GPU we use multiple SMs for the 
 same purpose.\footnote{In CUDA terms, our proposed alignment stage
@@
 configuration in CUDA terms). We have empirically determined
-that performance is highest with 5-7 subject classes aligned simultaneously on each SMP.
-%This mapping of the problem-level parallelism onto the SMP-level
+that performance is highest with 5-7 subject classes aligned simultaneously on each SM.
+%This mapping of the problem-level parallelism onto the SM-level
 %concurrency is illustrated in Figure \ref{fig:alignment_mapping_gpu}.
@@
 versions of the test image and its derivatives.
-Second, since there are no standard CUDA libraries that work at the SMP level, we use a
+Second, since there are no standard CUDA libraries that work at the SM level, we use a
 custom routine for computing $B_j^\dagger = (B_j^TB_j)^{-1} B^T = G^{-1} B_j^T$, with $G$
@@
 in \eqref{eq:l1min_denoise}, using Algorithm \ref{alg:alm_rec}.  The
-coefficients $\x$ are then used to compute error residuals that are used for
+coefficients in $\x$ are then used to compute error residuals that are used for
 classification.  
@@
 \hline
-One problem per SMP using CUDA streams & 70\,ms  \\
+One problem per SM using CUDA streams & 70\,ms  \\
 \hline
-Four problems per SMP using single kernel & {\bf 36\,ms} \\
+Four problems per SM using single kernel & {\bf 36\,ms} \\
 \hline
@@
 The manually threaded CPU implementation slightly outperforms the GPU implementation,
-and it beats the \emph{naive}
+and it surpasses the \emph{naive}
 library threaded implementation by a wide margin.
@@
 faster than implementations that rely purely on vendor-supplied BLAS libraries.
-Furthermore, thanks to a combination of faster hardware, and more efficient use
+Furthermore, thanks to a combination of faster hardware and more efficient use
 of the hardware, we have demonstrated dramatic improvements in recognition
@@
 %that attempt to leverage both levels of concurrency of the GPU, there is a need
-%for standard libraries that operate at the SMP level of concurrency.
+%for standard libraries that operate at the SM level of concurrency.
 

### 2011-06-05 main.tex
Applied Hossein's corrections

@@
 image-based face recognition as a sparse representation problem
-\cite{WrightJ2009-PAMI}.  If we stack the $d$ pixels of the training images of $K$ subject
-classes into the columns of matrices $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$, combine
-the matrices into a larger matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, and arrange the pixels of a new
-query image into a vector $\bb\in\Re^d$, \emph{sparsity-based
+\cite{WrightJ2009-PAMI}.  If we stack the $m$ pixels of the $n_i$ training images of $K$ subject
+classes into the columns of matrices $(A_1\in\Re^{m\times n_1}, \cdots, A_K\in\Re^{m\times n_K})$, combine
+the matrices into a larger matrix $A = [A_1, \cdots, A_K]\in\Re^{m\times n}$, and arrange the pixels of a new
+query image into a vector $\bb\in\Re^m$, \emph{sparsity-based
 classification} (SBC) solves the following minimization problem:
@@
 
-In Lagrange Multiplier Theory \cite{BertsekasD2003}, if there exists a Lagrangian $y^*$ that
+In Lagrange Multiplier Theory \cite{BertsekasD2003}, if there exists a multiplier $\y^*$ that
 satisfies the second-order sufficiency conditions for optimality, then for a sufficiently large $\mu$, the optimal
@@
 exist implementations in the dual domain \cite{YangJ2009,YangA2010-ICIP}.}
+The soft-thresholding algorithm involves the element-wise operator $\shrink(x) = \sign(x)\max(|x|,0)$.
 
@@
 \WHILE{not converged ($l = 1,2,\ldots$)} 
-\STATE $\uu_{l+1}  \leftarrow \mbox{shrink}(\zz_l - \frac{1}{\tau}A^T(A\zz_l - \bb - \frac{1}{\mu_k}\yy_k), \frac{1}{\mu_k\tau})$
+\STATE $\uu_{l+1}  \leftarrow \shrink(\zz_l - \frac{1}{\tau}A^T(A\zz_l - \bb - \frac{1}{\mu_k}\yy_k), \frac{1}{\mu_k\tau})$
 \STATE $t_{l+1} \leftarrow \frac{1}{2}( 1 + \sqrt{1+4t_l^2})$
@@
 \WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\zz_l \leftarrow \mbox{shrink}\left(\bb_j - B_j\uu_{l-1} + \frac{\yy_{k}}{\mu_{k-1}}, \frac{1}{\mu_{k-1}}\right)$
+\STATE $\zz_l \leftarrow \shrink\left(\bb_j - B_j\uu_{l-1} + \frac{\yy_{k}}{\mu_{k-1}}, \frac{1}{\mu_{k-1}}\right)$
 \STATE $\uu_l \leftarrow B_j^\dagger \left(\bb_j - \zz_{l} + \frac{\yy_{k-1}}{\mu_{k-1}} \right) $
@@
 but are also available in single-threaded versions (for the second implementation). For 
-operations that are not optimized by Intel in-house libraries, such as the shrinkage
+operations that are not optimized by Intel in-house libraries, such as the soft thresholding
 operator in step 7 of Algorithm \ref{alg:iterative_alignment}, we achieve 
@@
 \WHILE{not converged ($k = 1,2,\ldots$)}
-\STATE $\e_{k+1} = \textup{shrink}(\bb - A\x_k +
+\STATE $\e_{k+1} = \shrink(\bb - A\x_k +
 \frac{1}{\mu}\y_k, \frac{1}{\mu})$;
@@
 \WHILE{not converged ($l = 1,2,\ldots$)}
-\STATE $\w_{l+1} \leftarrow \textup{shrink}(\z_l +
+\STATE $\w_{l+1} \leftarrow \shrink(\z_l +
 \frac{1}{\gamma}A^T(\bb - A\z_l - \e_{k+1} +
@@
 implementations of the general $\ell_1$-min solver (Algorithm \eqref{alg:alm}).
-The $d \times n$ measurement matrix $A$ is a random Gaussian matrix, with each
+The $m \times n$ measurement matrix $A$ is a random Gaussian matrix, with each
 entry generated from the standard normal distribution and normalized to unit
-column norm.  The ratio of $d/n$ is fixed at $1/2$ with $n$ varying from 1000 to 8000.
-The ground truth signal, $\xx_0$ has a sparsity rate of 10\% of d with
+column norm.  The ratio of $m/n$ is fixed at $1/2$ with $n$ varying from 1000 to 8000.
+The ground truth signal, $\xx_0$ has a sparsity rate of 10\% of $m$ with
 elements sampled from a normal distribution and normalized to unit column

### 2011-06-06 main.tex
Tweaked figures, applied changes from Hossein.

@@
 \section{Introduction} 
-\vspace{-0.05in}
+\vspace{-0.06in}
 $\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
@@
 
-\subsection{Literature Review} \vspace{-0.05in}
+\subsection{Literature Review} \vspace{-0.06in}
 Traditionally, $\ell_1$-min (a.k.a.
@@
 \label{sec:ALM}
-\vspace{-0.05in}
+\vspace{-0.06in}
 In this section, we briefly describe the ALM algorithm for $\ell_1$-min
@@
 
-\vspace{-0.05in}
+\vspace{-0.06in}
 \section{Hardware Concurrency} \label{sec:concurrency}
-\vspace{-0.05in}
+\vspace{-0.06in}
 In this section we discuss the levels of concurrency available in the hardware
@@
 \subsection{CPU Hardware Concurrency}
-\vspace{-0.05in}
+\vspace{-0.06in}
 \label{sec:CPU-concurrency}
@@
 
-\subsection{GPU Hardware Concurrency} \vspace{-0.05in}
+\subsection{GPU Hardware Concurrency} \vspace{-0.06in}
 The main defining characteristics of contemporary multi-core GPU architectures
@@
 
-\section{Parallelism in the Face Alignment Stage}\vspace{-0.05in}
+\section{Parallelism in the Face Alignment Stage}\vspace{-0.06in}
 \label{sec:alignment}
@@
 
-\subsection{CPU Implementation} \vspace{-0.05in}
+\subsection{CPU Implementation} \vspace{-0.06in}
 \label{sec:alignment_implementation_cpu}
@@
 
-\subsection{GPU Implementation} \vspace{-0.05in}
+\subsection{GPU Implementation} \vspace{-0.06in}
 \label{sec:alignment_implementation_gpu}
@@
 
-\section{Parallelization of the Face Recognition Stage} \vspace{-0.05in}
+\section{Parallelization of the Face Recognition Stage} \vspace{-0.06in}
 \label{sec:recognition}
@@
 
-\subsection{Recognition Stage Implementation} \vspace{-0.05in}
+\subsection{Recognition Stage Implementation} \vspace{-0.06in}
 Compared to problem-level
@@
 
-\section{Experiments} \vspace{-0.05in}
+\section{Experiments} \vspace{-0.06in}
 \label{sec:experiment} 
@@
 
-\subsection{Face Recognition Pipeline Benchmark} \vspace{-0.05in}
+\subsection{Face Recognition Pipeline Benchmark} \vspace{-0.06in}
 \label{sec:benchmark}
@@
 \centering
-\includegraphics[width=3.3in]{figures/alignment_runtime_graph}\vspace{0.05in}
+\includegraphics[width=3.3in]{figures/alignment_runtime_graph}\vspace{0.06in}
 \caption{Alignment stage runtime vs. size of training database.} \vspace{-0.1in}
@@
 \centering
-\includegraphics[width=2.5in]{figures/speedVsResolution} \vspace{-0.05in}
+\includegraphics[width=2.5in]{figures/speedVsResolution} \vspace{-0.06in}
 \caption{Recognition stage runtime vs. face window resolution}\vspace{-0.1in}
@@
 \centering
-\includegraphics[width=2.5in]{figures/accuracyVsResolution} \vspace{-0.05in}
+\includegraphics[width=2.5in]{figures/accuracyVsResolution} \vspace{-0.06in}
 \caption{Recognition rate of the full pipeline vs. face window resolution.} \vspace{-0.1in}
@@
 
-\section{Conclusion}\vspace{-0.05in}
+\section{Conclusion}\vspace{-0.06in}
 We have demonstrated that on both CPU and GPU algorithms, parallelizations of
@@
 
+\vspace{-0.06in}
 {\small

### 2011-06-06 figures/alignment_runtime_graph.m

@@ -5,17 +5,17 @@ cpu2=load('alignment_runtime_graph_reference_cpu_percore.dat');
 gtx480=load('alignment_runtime_graph_reference_gtx480.dat');
 gtx580=load('alignment_runtime_graph_reference_gtx580.dat');
 
-%userCount = cpu1(:,1);
-%cpu1 = cpu1(:,2);
-%cpu2 = cpu2(:,2);
-%gtx480 = gtx480(:,2);
-%gtx580 = gtx580(:,2);
+userCount = cpu1(:,1);
+cpu1 = cpu1(:,2);
+cpu2 = cpu2(:,2);
+gtx480 = gtx480(:,2);
+gtx580 = gtx580(:,2);
+%userCount = cpu1(1:2:end,1);
+%cpu1 = cpu1(1:2:end,2);
+%cpu2 = cpu2(1:2:end,2);
+%gtx480 = gtx480(1:2:end,2);
+%gtx580 = gtx580(1:2:end,2);
 
-userCount = cpu1(1:2:end,1);
-cpu1 = cpu1(1:2:end,2);
-cpu2 = cpu2(1:2:end,2);
-gtx480 = gtx480(1:2:end,2);
-gtx580 = gtx580(1:2:end,2);
 close all;
 h = figure(7);
 

### 2011-06-08 main.tex
Explained crossover point.  Giving up on re-runing experiments.
Added authors for submission as tech report.

@@
 
-% \cvprfinalcopy % *** Uncomment this line for the final submission
+\cvprfinalcopy % *** Uncomment this line for the final submission
 
@@
 
-%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
-%\title{Efficient Parallelization of Sparse Representation for Face Recognition}
-%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
-%\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
-\title{\vspace{-0.2in} Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
-%\title{Parallelization of Fast $\ell_1$-Minimization Towards Real-Time Face Recognition}
+\title{\vspace{-0.75in} Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
 
-\author{First Author\\
-Institution1\\
-Institution1 address\\
-{\tt\small firstauthor@i1.org}
 % For a paper whose authors are all at the same institution,
@@
 % To save space, use either the email address or home page, not both
+
+\author{Andrew Wagner\\
+University of Illinois at Urbana-Champaign\\
+{\small awagner@illinois.edu}
 \and
-Second Author\\
-Institution2\\
-First line of institution2 address\\
-{\small\url{http://www.author.org/~second}}
+Victor Shia\\
+University of California Berkeley\\
+{\small vshia@eecs.berkeley.edu}
+\and
+Allen Yang\\
+{\small yang@eecs.berkeley.edu}
+\and
+Mark Murphy\\
+{\small mjmurphy@eecs.berkeley.edu}
+\and
+Yi Ma\\
+University of Illinois at Urbana-Champaign\\
+{\small yima@illinois.edu}
 }
@@
 single large problem, whereas the CPU implementation is faster at solving a
-single small problem.  The transition between the two regimes occurs (at 2000) where
-the problem size begins to exceed the size of the CPU L3 cache. 
+single small problem.  The transition between the two regimes occurs where the size of
+$A$ reaches $2000 \times \1000 \times 4 = 8$ MB, i.e. the size of the CPU L3 cache. 
 \begin{figure}

### 2011-08-18 main.tex
Typesets with ACM... need to polish.

@@
-\documentclass[10pt,twocolumn,letterpaper]{article}
-% ICCV Rules:
-% 8 page limit (6 free + 2 paid)
-% 10MB pdf file size limit
-% 30MB supplementary materials file size limit (PDF or ZIP)
+\documentclass[preprint]{sigplanconf}
 
-\usepackage{cvpr}
+%\usepackage{cvpr}
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
@@
 \usepackage{subfigure}
-
-% Needed for getting formulas in dia diagrams.
 \usepackage{pstricks}
 \usepackage{tikz}
-
 \usepackage{mydefs}
 
-\cvprfinalcopy % *** Uncomment this line for the final submission
-
-\def\cvprPaperID{184} % *** Enter the CVPR Paper ID here
-\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
-
-% Pages are numbered in submission mode, and unnumbered in camera-ready
+% A few things in cvpr style file but not in ACM...
+\def\eg{\emph{e.g.\ }} 
+\def\ie{\emph{i.e.\ }}
 
 \hyphenation{under-determined}
-\ifcvprfinal\pagestyle{empty}\fi
 \begin{document}
 
-\title{\vspace{-0.75in} Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
+\conferenceinfo{PPoPP '12}{Feb. 29, New Orleans.} 
+\copyrightyear{2012} 
+\copyrightdata{[to be supplied]}
 
-% For a paper whose authors are all at the same institution,
-% omit the following lines up until the closing ``}''.
-% Additional authors and addresses can be added with ``\and'',
-% just like the second author.
-% To save space, use either the email address or home page, not both
+\titlebanner{REVIEW COPY; PLEASE TO NOT REDISTRIBUTE}        % These are ignored unless
+%\preprintfooter{short description of paper}   % 'preprint' option specified.
 
-\author{Andrew Wagner\\
-University of Illinois at Urbana-Champaign\\
-{\small awagner@illinois.edu}
-\and
-Victor Shia\\
-University of California Berkeley\\
-{\small vshia@eecs.berkeley.edu}
-\and
-Allen Yang\\
-{\small yang@eecs.berkeley.edu}
-\and
-Mark Murphy\\
-{\small mjmurphy@eecs.berkeley.edu}
-\and
-Yi Ma\\
-University of Illinois at Urbana-Champaign\\
-{\small yima@illinois.edu}
-}
+\title{Parallelization of Fast $\ell_1$-Minimization for Face Recognition}
+
+%\authorinfo{Name1}
+           %{Affiliation1}
+           %{Email1}
+%\authorinfo{Name2\and Name3}
+           %{Affiliation2/3}
+           %{Email2/3}
+\authorinfo{Andrew Wagner\and Yi Ma}
+			{University of Illinois at Urbana-Champaign}
+			{\{awagner,yima\}@illinois.edu}
+\authorinfo{Victor Shia\and Allen Yang\and Mark Murphy}
+			{University of California Berkeley}
+			{\{vshia,yang,mjmurphy\}@eecs.berkeley.edu}
+%\authorinfo{Allen Yang}
+			%{University of California Berkeley}
+			%{yang@eecs.berkeley.edu}
+%\authorinfo{Mark Murphy}
+			%{University of California Berkeley}
+			%{mjmurphy@eecs.berkeley.edu}
+%\authorinfo{Yi Ma}
+			%{University of Illinois at Urbana-Champaign}
+			%{yima@illinois.edu}
 
@@
 
-\vspace{-0.06in}
-{\small
-\bibliographystyle{ieee}
+\bibliographystyle{abbrvnat}
 \bibliography{faces-short}
-}
 

### 2011-08-19 main.tex
Added figure illustrating transformation geometry, applied some comments from reviewer.

@@
 
-%\authorinfo{Name1}
-           %{Affiliation1}
-           %{Email1}
-%\authorinfo{Name2\and Name3}
-           %{Affiliation2/3}
-           %{Email2/3}
-\authorinfo{Andrew Wagner\and Yi Ma}
-			{University of Illinois at Urbana-Champaign}
-			{\{awagner,yima\}@illinois.edu}
-\authorinfo{Victor Shia\and Allen Yang\and Mark Murphy}
-			{University of California Berkeley}
-			{\{vshia,yang,mjmurphy\}@eecs.berkeley.edu}
-%\authorinfo{Allen Yang}
+\authorinfo{Name1}
+		   {Affiliation1}
+		   {Email1}
+\authorinfo{Name2\and Name3}
+		   {Affiliation2/3}
+		   {Email2/3}
+%\authorinfo{Andrew Wagner\and Yi Ma}
+			%{University of Illinois at Urbana-Champaign}
+			%{\{awagner,yima\}@illinois.edu}
+%\authorinfo{Victor Shia\and Allen Yang\and Mark Murphy}
 			%{University of California Berkeley}
-			%{yang@eecs.berkeley.edu}
-%\authorinfo{Mark Murphy}
-			%{University of California Berkeley}
-			%{mjmurphy@eecs.berkeley.edu}
-%\authorinfo{Yi Ma}
-			%{University of Illinois at Urbana-Champaign}
-			%{yima@illinois.edu}
+			%{\{vshia,mjmurphy,yang\}@eecs.berkeley.edu}
 
@@
 algorithms will be made available for peer evaluation.  \end{abstract}
-\vspace{-0.1in}
 
 \section{Introduction} 
-\vspace{-0.06in}
 $\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
@@
 $\bb$ based on the training image labels associated with $A$.  
-%Since $A$ contains many images per user taken under different illuminations, 
-%$A_i\x_i$ acts as a linear illumination model for the test image $\bb$.
+{\bf Several
+classifiers based on $\xx$ have been proposed that all have nearly identical 
+performance.  In this paper classification is performed using the following residual:
+\begin{equation}
+i^* = \argmin{i} || \bb - \ee - A_i \xx_i||_2
+\label{eq:classification}
+\end{equation}
+}
+% changed
 
@@
 {\tiny \input{figures/pipeline_simplified.tex}}
-\caption{The face recognition pipeline.\vspace{-0.1in}}
+\caption{The face recognition pipeline.}
 \label{fig:pipeline}
@@
 
-\subsection{Literature Review} \vspace{-0.06in}
+\subsection{Literature Review} 
 Traditionally, $\ell_1$-min (a.k.a.
@@
 \label{sec:ALM}
-\vspace{-0.06in}
 In this section, we briefly describe the ALM algorithm for $\ell_1$-min
@@
 
-\vspace{-0.06in}
 \section{Hardware Concurrency} \label{sec:concurrency}
-\vspace{-0.06in}
 In this section we discuss the levels of concurrency available in the hardware
@@
 floating point datatypes.
-% WORKING HERE
 
 \subsection{CPU Hardware Concurrency}
-\vspace{-0.06in}
 \label{sec:CPU-concurrency}
@@
 caches.  Note: Although the aspect ratio could not be preserved, all arrays and
-caches are drawn so that the area is proportional to the size of the data. } \vspace{-0.1in}
+caches are drawn so that the area is proportional to the size of the data. } 
 \end{figure}
@@
 
-\subsection{GPU Hardware Concurrency} \vspace{-0.06in}
+\subsection{GPU Hardware Concurrency} 
 The main defining characteristics of contemporary multi-core GPU architectures
@@
 
-\section{Parallelism in the Face Alignment Stage}\vspace{-0.06in}
+\section{Parallelism in the Face Alignment Stage}
 \label{sec:alignment}
@@
 
+{\bf
 Face alignment \eqref{eq:l1min_alignment} estimates an image transformation
-$\tau$ that rectifies the query image $\tilde{\bb}$ with possible pose
-variation w.r.t. each training class $A_i$, which leads to the minimal sparsity
-in error $\ee$ after the alignment. Note that directly solving
-\eqref{eq:l1min_alignment} is inefficient since it is a non-convex problem
-and may exhibit local minima.
-However, given a good initial estimate of the
-transformation $\tau$ (\eg, provided by an accurate face detector), the optimal
-solution for $\tau$ can be sought iteratively by linearization at each $j$th
-step:
+$\tau$ that maps pixels in a canonical frame, where the face is well aligned,
+back to pixels in the original image.  
+This is illustrated in Figure \ref{fig:transformation}.
+\begin{figure}
+\centering
+\includegraphics[width=3.4in]{figures/transformation}
+\caption{The simililarity transformation $\tau$ maps pixels in a canonical frame
+where the image is aligned back to pixel locations in the original-sized images.}
+\label{fig:transformation}
+\end{figure}
+The training images are prepared as
+follows: canonical outer-eye corner locations are defined in the canonical
+frame, and the outer-eye corners are manually selected in the training images.
+The training images are warped/resampled by a similarity transformation
+(rotation, translation, and scale) between the two reference frames that is
+computed from the eye corners.
+The canonical eye corners are also illustrated in Figure \ref{fig:transformation}
+
+During iterative alignment, the transformation $\tau$ between the original test
+image and the canonical frame is initialized via a face detector, and then updated
+iteratively.  A different $\tau_i$ is computed separately for each training class $A_i$.  
+$\ell_1$ minimization is performed on canonical-sized images that
+have been stacked into the columns of matrices.  
+Directly solving \eqref{eq:l1min_alignment} in its original formulation is difficult since 
+it is a non-convex problem and may exhibit local minima.
+Minimization is therefore performed iteratively, alternating between updating
+a linearized version of the problem at each $j$th step, and solving the linearized problem:
 \begin{equation}
@@
 \end{equation}
+}
 %If the warping $\tau$ is constrained to be a similarity transform,
@@
 
-\subsection{CPU Implementation} \vspace{-0.06in}
+\subsection{CPU Implementation} 
 \label{sec:alignment_implementation_cpu}
@@
 
-\subsection{GPU Implementation} \vspace{-0.06in}
+\subsection{GPU Implementation} 
 \label{sec:alignment_implementation_gpu}
@@
 
-\section{Parallelization of the Face Recognition Stage} \vspace{-0.06in}
+\section{Parallelization of the Face Recognition Stage} 
 \label{sec:recognition}
@@
 
-\subsection{Recognition Stage Implementation} \vspace{-0.06in}
+\subsection{Recognition Stage Implementation} 
 Compared to problem-level
@@
 
-\section{Experiments} \vspace{-0.06in}
+\section{Experiments} 
 \label{sec:experiment} 
@@
 The results of this benchmark can be found in Figure \ref{fig:random_data}.
-The $x$-axis represents the size of the $A$ matrix and the $y$-axis represents the
-average amount of time to complete a single problem instance.  The GPU
+The $x$-axis represents the size of the $A$ matrix and the $y$-axis represents
+the average amount of time to complete a single problem instance.  The GPU
 implementation tends to be faster than the CPU implementation at solving a
 single large problem, whereas the CPU implementation is faster at solving a
-single small problem.  The transition between the two regimes occurs where the size of
-$A$ reaches $2000 \times \1000 \times 4 = 8$ MB, i.e. the size of the CPU L3 cache. 
+single small problem.  {\bf The change in slope of the CPU curve is caused by a
+transition from the CPU being cache-bandwidth-bound to memory-bandwidth-bound.
+Since there are other data structures besides $A$, The transition between the
+two regimes occurs slightly before the size of $A$ reaches $2000 \times \1000
+\times 4 = 8$ MB, i.e. the size of the CPU L3 cache. }
+%changed
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=2.5in]{figures/time_vs_matrix_size_constant_tol}
-\end{center} \vspace{-0.1in}
-\caption{\small Comparison of $\ell_1$-min runtime vs. dimensions of $A$ on random data.} \vspace{-0.1in}
+\includegraphics[width=3.35in]{figures/time_vs_matrix_size_constant_tol}
+\end{center} 
+\caption{\small Comparison of $\ell_1$-min runtime vs. dimensions of $A$ on random data.  {\bf The crossover
+occurs near where the size of the problem exceeds the size of the CPU L3 cache.}} 
 \label{fig:random_data}
@@
 
-\subsection{Face Recognition Pipeline Benchmark} \vspace{-0.06in}
+\subsection{Face Recognition Pipeline Benchmark} 
 \label{sec:benchmark}
@@
 \small
+\centering
 \begin{tabular}{|l|c|}
@@
 \hline
-\end{tabular} \vspace{-0.1in}
+\end{tabular} 
 \label{tbl:ubench}
@@
 \centering
-\includegraphics[width=3.3in]{figures/alignment_runtime_graph}\vspace{0.06in}
-\caption{Alignment stage runtime vs. size of training database.} \vspace{-0.1in}
+\includegraphics[width=3.35in]{figures/alignment_runtime_graph}
+\caption{Alignment stage runtime vs. size of training database.} 
 \label{fig:alignment_stage_runtime}
@@
 \centering
-\includegraphics[width=2.5in]{figures/speedVsResolution} \vspace{-0.06in}
-\caption{Recognition stage runtime vs. face window resolution}\vspace{-0.1in}
+\includegraphics[width=3.35in]{figures/speedVsResolution} 
+\caption{Recognition stage runtime vs. face window resolution}
 \label{fig:recognition_stage_runtime} \end{figure}
@@
 \centering
-\includegraphics[width=2.5in]{figures/accuracyVsResolution} \vspace{-0.06in}
-\caption{Recognition rate of the full pipeline vs. face window resolution.} \vspace{-0.1in}
+\includegraphics[width=3.35in]{figures/accuracyVsResolution} 
+\caption{Recognition rate of the full pipeline vs. face window resolution.} 
 \label{fig:accuracy_vs_resolution}
@@
 
-\section{Conclusion}\vspace{-0.06in}
+\section{Conclusion}
 We have demonstrated that on both CPU and GPU algorithms, parallelizations of
